% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{article}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage[margin=1in]{geometry}
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\usepackage{xcolor}
\definecolor{aliceblue}{HTML}{F0F8FF}
\definecolor{antiquewhite}{HTML}{FAEBD7}
\definecolor{aqua}{HTML}{00FFFF}
\definecolor{aquamarine}{HTML}{7FFFD4}
\definecolor{azure}{HTML}{F0FFFF}
\definecolor{beige}{HTML}{F5F5DC}
\definecolor{bisque}{HTML}{FFE4C4}
\definecolor{black}{HTML}{000000}
\definecolor{blanchedalmond}{HTML}{FFEBCD}
\definecolor{blue}{HTML}{0000FF}
\definecolor{blueviolet}{HTML}{8A2BE2}
\definecolor{brown}{HTML}{A52A2A}
\definecolor{burlywood}{HTML}{DEB887}
\definecolor{cadetblue}{HTML}{5F9EA0}
\definecolor{chartreuse}{HTML}{7FFF00}
\definecolor{chocolate}{HTML}{D2691E}
\definecolor{coral}{HTML}{FF7F50}
\definecolor{cornflowerblue}{HTML}{6495ED}
\definecolor{cornsilk}{HTML}{FFF8DC}
\definecolor{crimson}{HTML}{DC143C}
\definecolor{cyan}{HTML}{00FFFF}
\definecolor{darkblue}{HTML}{00008B}
\definecolor{darkcyan}{HTML}{008B8B}
\definecolor{darkgoldenrod}{HTML}{B8860B}
\definecolor{darkgray}{HTML}{A9A9A9}
\definecolor{darkgreen}{HTML}{006400}
\definecolor{darkgrey}{HTML}{A9A9A9}
\definecolor{darkkhaki}{HTML}{BDB76B}
\definecolor{darkmagenta}{HTML}{8B008B}
\definecolor{darkolivegreen}{HTML}{556B2F}
\definecolor{darkorange}{HTML}{FF8C00}
\definecolor{darkorchid}{HTML}{9932CC}
\definecolor{darkred}{HTML}{8B0000}
\definecolor{darksalmon}{HTML}{E9967A}
\definecolor{darkseagreen}{HTML}{8FBC8F}
\definecolor{darkslateblue}{HTML}{483D8B}
\definecolor{darkslategray}{HTML}{2F4F4F}
\definecolor{darkslategrey}{HTML}{2F4F4F}
\definecolor{darkturquoise}{HTML}{00CED1}
\definecolor{darkviolet}{HTML}{9400D3}
\definecolor{deeppink}{HTML}{FF1493}
\definecolor{deepskyblue}{HTML}{00BFFF}
\definecolor{dimgray}{HTML}{696969}
\definecolor{dimgrey}{HTML}{696969}
\definecolor{dodgerblue}{HTML}{1E90FF}
\definecolor{firebrick}{HTML}{B22222}
\definecolor{floralwhite}{HTML}{FFFAF0}
\definecolor{forestgreen}{HTML}{228B22}
\definecolor{fuchsia}{HTML}{FF00FF}
\definecolor{gainsboro}{HTML}{DCDCDC}
\definecolor{ghostwhite}{HTML}{F8F8FF}
\definecolor{gold}{HTML}{FFD700}
\definecolor{goldenrod}{HTML}{DAA520}
\definecolor{gray}{HTML}{808080}
\definecolor{green}{HTML}{008000}
\definecolor{greenyellow}{HTML}{ADFF2F}
\definecolor{grey}{HTML}{808080}
\definecolor{honeydew}{HTML}{F0FFF0}
\definecolor{hotpink}{HTML}{FF69B4}
\definecolor{indianred}{HTML}{CD5C5C}
\definecolor{indigo}{HTML}{4B0082}
\definecolor{ivory}{HTML}{FFFFF0}
\definecolor{khaki}{HTML}{F0E68C}
\definecolor{lavender}{HTML}{E6E6FA}
\definecolor{lavenderblush}{HTML}{FFF0F5}
\definecolor{lawngreen}{HTML}{7CFC00}
\definecolor{lemonchiffon}{HTML}{FFFACD}
\definecolor{lightblue}{HTML}{ADD8E6}
\definecolor{lightcoral}{HTML}{F08080}
\definecolor{lightcyan}{HTML}{E0FFFF}
\definecolor{lightgoldenrodyellow}{HTML}{FAFAD2}
\definecolor{lightgray}{HTML}{D3D3D3}
\definecolor{lightgreen}{HTML}{90EE90}
\definecolor{lightgrey}{HTML}{D3D3D3}
\definecolor{lightpink}{HTML}{FFB6C1}
\definecolor{lightsalmon}{HTML}{FFA07A}
\definecolor{lightseagreen}{HTML}{20B2AA}
\definecolor{lightskyblue}{HTML}{87CEFA}
\definecolor{lightslategray}{HTML}{778899}
\definecolor{lightslategrey}{HTML}{778899}
\definecolor{lightsteelblue}{HTML}{B0C4DE}
\definecolor{lightyellow}{HTML}{FFFFE0}
\definecolor{lime}{HTML}{00FF00}
\definecolor{limegreen}{HTML}{32CD32}
\definecolor{linen}{HTML}{FAF0E6}
\definecolor{magenta}{HTML}{FF00FF}
\definecolor{maroon}{HTML}{800000}
\definecolor{mediumaquamarine}{HTML}{66CDAA}
\definecolor{mediumblue}{HTML}{0000CD}
\definecolor{mediumorchid}{HTML}{BA55D3}
\definecolor{mediumpurple}{HTML}{9370DB}
\definecolor{mediumseagreen}{HTML}{3CB371}
\definecolor{mediumslateblue}{HTML}{7B68EE}
\definecolor{mediumspringgreen}{HTML}{00FA9A}
\definecolor{mediumturquoise}{HTML}{48D1CC}
\definecolor{mediumvioletred}{HTML}{C71585}
\definecolor{midnightblue}{HTML}{191970}
\definecolor{mintcream}{HTML}{F5FFFA}
\definecolor{mistyrose}{HTML}{FFE4E1}
\definecolor{moccasin}{HTML}{FFE4B5}
\definecolor{navajowhite}{HTML}{FFDEAD}
\definecolor{navy}{HTML}{000080}
\definecolor{oldlace}{HTML}{FDF5E6}
\definecolor{olive}{HTML}{808000}
\definecolor{olivedrab}{HTML}{6B8E23}
\definecolor{orange}{HTML}{FFA500}
\definecolor{orangered}{HTML}{FF4500}
\definecolor{orchid}{HTML}{DA70D6}
\definecolor{palegoldenrod}{HTML}{EEE8AA}
\definecolor{palegreen}{HTML}{98FB98}
\definecolor{paleturquoise}{HTML}{AFEEEE}
\definecolor{palevioletred}{HTML}{DB7093}
\definecolor{papayawhip}{HTML}{FFEFD5}
\definecolor{peachpuff}{HTML}{FFDAB9}
\definecolor{peru}{HTML}{CD853F}
\definecolor{pink}{HTML}{FFC0CB}
\definecolor{plum}{HTML}{DDA0DD}
\definecolor{powderblue}{HTML}{B0E0E6}
\definecolor{purple}{HTML}{800080}
\definecolor{red}{HTML}{FF0000}
\definecolor{rosybrown}{HTML}{BC8F8F}
\definecolor{royalblue}{HTML}{4169E1}
\definecolor{saddlebrown}{HTML}{8B4513}
\definecolor{salmon}{HTML}{FA8072}
\definecolor{sandybrown}{HTML}{F4A460}
\definecolor{seagreen}{HTML}{2E8B57}
\definecolor{seashell}{HTML}{FFF5EE}
\definecolor{sienna}{HTML}{A0522D}
\definecolor{silver}{HTML}{C0C0C0}
\definecolor{skyblue}{HTML}{87CEEB}
\definecolor{slateblue}{HTML}{6A5ACD}
\definecolor{slategray}{HTML}{708090}
\definecolor{slategrey}{HTML}{708090}
\definecolor{snow}{HTML}{FFFAFA}
\definecolor{springgreen}{HTML}{00FF7F}
\definecolor{steelblue}{HTML}{4682B4}
\definecolor{tan}{HTML}{D2B48C}
\definecolor{teal}{HTML}{008080}
\definecolor{thistle}{HTML}{D8BFD8}
\definecolor{tomato}{HTML}{FF6347}
\definecolor{turquoise}{HTML}{40E0D0}
\definecolor{violet}{HTML}{EE82EE}
\definecolor{wheat}{HTML}{F5DEB3}
\definecolor{white}{HTML}{FFFFFF}
\definecolor{whitesmoke}{HTML}{F5F5F5}
\definecolor{yellow}{HTML}{FFFF00}
\definecolor{yellowgreen}{HTML}{9ACD32}
\usepackage[most]{tcolorbox}

\usepackage{ifthen}
\provideboolean{admonitiontwoside}
\makeatletter%
\if@twoside%
\setboolean{admonitiontwoside}{true}
\else%
\setboolean{admonitiontwoside}{false}
\fi%
\makeatother%

\newenvironment{env-eda4deab-32af-4a92-829c-9092f376a0cb}
{
    \savenotes\tcolorbox[blanker,breakable,left=5pt,borderline west={2pt}{-4pt}{firebrick}]
}
{
    \endtcolorbox\spewnotes
}
                

\newenvironment{env-48e9bfc8-c3dd-4504-a5fa-3b34b9a27ccf}
{
    \savenotes\tcolorbox[blanker,breakable,left=5pt,borderline west={2pt}{-4pt}{blue}]
}
{
    \endtcolorbox\spewnotes
}
                

\newenvironment{env-69c6d983-4e25-44a5-9a13-662902d85e33}
{
    \savenotes\tcolorbox[blanker,breakable,left=5pt,borderline west={2pt}{-4pt}{green}]
}
{
    \endtcolorbox\spewnotes
}
                

\newenvironment{env-a57f3fff-c46d-4aeb-b36c-ada0d8066098}
{
    \savenotes\tcolorbox[blanker,breakable,left=5pt,borderline west={2pt}{-4pt}{aquamarine}]
}
{
    \endtcolorbox\spewnotes
}
                

\newenvironment{env-f3fdeea8-9a69-4242-ac7f-874202f0281b}
{
    \savenotes\tcolorbox[blanker,breakable,left=5pt,borderline west={2pt}{-4pt}{orange}]
}
{
    \endtcolorbox\spewnotes
}
                

\newenvironment{env-3a4a363c-8d41-4783-bfde-af85d2b9ad5f}
{
    \savenotes\tcolorbox[blanker,breakable,left=5pt,borderline west={2pt}{-4pt}{gold}]
}
{
    \endtcolorbox\spewnotes
}
                

\newenvironment{env-9193a5b6-f9d9-4a5c-9922-4e221c0d52be}
{
    \savenotes\tcolorbox[blanker,breakable,left=5pt,borderline west={2pt}{-4pt}{darkred}]
}
{
    \endtcolorbox\spewnotes
}
                

\newenvironment{env-b225a0d9-0c32-46a2-8b5d-c7fc92570012}
{
    \savenotes\tcolorbox[blanker,breakable,left=5pt,borderline west={2pt}{-4pt}{pink}]
}
{
    \endtcolorbox\spewnotes
}
                

\newenvironment{env-8073ffbe-6656-43d2-bf4d-97e604cb2369}
{
    \savenotes\tcolorbox[blanker,breakable,left=5pt,borderline west={2pt}{-4pt}{cyan}]
}
{
    \endtcolorbox\spewnotes
}
                

\newenvironment{env-8890de02-1ba1-42c2-9b89-cb105d97c320}
{
    \savenotes\tcolorbox[blanker,breakable,left=5pt,borderline west={2pt}{-4pt}{cyan}]
}
{
    \endtcolorbox\spewnotes
}
                

\newenvironment{env-c5a21b51-dc78-4e82-bd56-c0130c2151f9}
{
    \savenotes\tcolorbox[blanker,breakable,left=5pt,borderline west={2pt}{-4pt}{purple}]
}
{
    \endtcolorbox\spewnotes
}
                

\newenvironment{env-d439fdb3-4d6e-4b59-86b8-9f15e60f671d}
{
    \savenotes\tcolorbox[blanker,breakable,left=5pt,borderline west={2pt}{-4pt}{darksalmon}]
}
{
    \endtcolorbox\spewnotes
}
                

\newenvironment{env-b56569b0-a3e1-47fe-b37c-6cf3ba4ee6a2}
{
    \savenotes\tcolorbox[blanker,breakable,left=5pt,borderline west={2pt}{-4pt}{gray}]
}
{
    \endtcolorbox\spewnotes
}
                
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Acquisition Terms},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

\title{Acquisition Terms}
\author{}
\date{}

\begin{document}
\maketitle

{
\setcounter{tocdepth}{3}
\tableofcontents
}
\hypertarget{spa}{%
\section{SPA}\label{spa}}

\hypertarget{structure-of-agreement}{%
\subsection{Structure of Agreement}\label{structure-of-agreement}}

Order of the agreement:

\begin{itemize}
\tightlist
\item
  parties and date;
\item
  operative provisions;
\item
  schedules;
\item
  execution by the parties.
\end{itemize}

Think about if there are any language requirements (e.g., the French
insist on French). If there are multiple languages, say which prevails
in the event of a conflict.

\hypertarget{parties}{%
\subsection{Parties}\label{parties}}

Where there is more than one seller, the details of all the sellers will
usually be set out in a schedule to the agreement. Usually, all sellers
give warranties. If not, parties described as sellers, warrantors and
buyers. There may also be guarantors of the seller's/ buyer's
obligations.

\hypertarget{execution}{%
\subsubsection{Execution}\label{execution}}

When the parties execute the agreement, it will be dated, so creating a
binding contractual agreement to buy and sell either the shares of the
target company or the assets of the business.

Civil law countries do things differently, SPA is usually far shorter.
This is because contracts in civil code jurisdictions rely on the
underlying code and interpreting the provisions according to good faith.

\hypertarget{operative-provisions}{%
\subsection{Operative Provisions}\label{operative-provisions}}

The operative provisions set out the basis on which the buyer agrees to
purchase the shares or assets of the business, and on which the seller
agrees to sell. Precise details often in a schedule.

\hypertarget{definitions-and-interpretation}{%
\subsection{Definitions and
Interpretation}\label{definitions-and-interpretation}}

Start with these clauses. Makes the document cleaner.

The buyer will usually want it to be provided expressly that references
to statutory provisions are to be interpreted as including subsequent
amendments to those provisions. This is dangerous for the seller,
particularly in relation to its potential liability under the
warranties. Means the seller risks these liabilities increasing.

\hypertarget{conditions-precedent}{%
\subsection{Conditions Precedent}\label{conditions-precedent}}

Completion of acquisition usually simultaneous with exchange of
contracts. But sometimes there are CPs.

If the agreement is conditional, it is usual to include an obligation on
the parties to seek fulfilment of the condition, a longstop date by
which the condition must be fulfilled or waived, and a time frame within
which the acquisition must complete once the condition has been
fulfilled.

\hypertarget{agreement-to-purchase}{%
\subsection{Agreement to Purchase}\label{agreement-to-purchase}}

The agreement will also provide that the parties agree to sell/ purchase
the specified shares/ assets. Usually provides that buyer is not obliged
to buy any shares/ assets until it can purchase simultaneously.

\hypertarget{consideration}{%
\subsection{Consideration}\label{consideration}}

The sale and purchase agreement will stipulate the agreed price, the
form that it will take, and the timing of payment. For multiple sellers,
price payable to each set out in a schedule.

On an asset sale, the amount of the consideration which is attributable
to each separate asset will be specified in the agreement (the agreed
apportionment is usually set out in a schedule).

\hypertarget{completion-accounts}{%
\subsubsection{Completion Accounts}\label{completion-accounts}}

\hypertarget{reasons-and-form}{%
\paragraph{Reasons and Form}\label{reasons-and-form}}

Parties may have based their valuation of the target on dated
information. Buyer may want confirmation that the figures have not
altered significantly since the original accounts.

Parties may agree to adjust the purchase price to reflect the actual net
assets or earnings of the target on the completion date, as determined
by a set of agreed accounts drawn up after completion.

Such completion accounts may consist of:

\begin{itemize}
\tightlist
\item
  Full P\&L
\item
  Balance sheet
\item
  Statement of net assets.
\end{itemize}

\hypertarget{control}{%
\paragraph{Control}\label{control}}

The seller's accountant, who is likely to be the target company's
auditor up until completion, is often nominated to prepare the first
draft of the completion accounts since it will already be familiar with
the business.

But the buyer may prefer to do this itself (to gain familiarity with the
business).

\hypertarget{typical-provisions}{%
\paragraph{Typical Provisions}\label{typical-provisions}}

\begin{itemize}
\tightlist
\item
  Maximum amount of the purchase price, with a further adjustment
  payment/ repayment to be made once the completion accounts have been
  prepared.
\item
  Adjustment usually made in cash. Actual adjustment depends on what the
  parties agree.
\item
  Might be one-sided or two-sided, might apply a multiplier to any
  shortfall in earnings.
\item
  It is also important that the sale and purchase agreement contains
  clear provisions guarding against possible double recovery that could
  occur where, for example, a reduction in net assets at completion also
  constitutes evidence of breach of warranty. The buyer should not be
  able to recover under an action for breach of warranty, whilst also
  benefitting from paying a smaller purchase price as a result of the
  completion accounts.
\item
  Agreed mechanism for completing accounts, usually in a schedule.
\end{itemize}

\begin{env-c5a21b51-dc78-4e82-bd56-c0130c2151f9}

Typical provisions

\begin{enumerate}
\tightlist
\item
  Completion accounts drawn up within 30 business days. Accountant
  preparing accounts instructed to do so on a basis consistent with the
  latest set of audited accounts.
\item
  Buyer and accountant have the right to review completion accounts.
\item
  In the absence of agreement as to the accounts, a limited period
  during which the parties will attempt to resolve any disputes;
\item
  in the case of continued dispute, the matter to be referred to an
  independent firm of accountants acting as experts.
\end{enumerate}

\end{env-c5a21b51-dc78-4e82-bd56-c0130c2151f9}

\hypertarget{specific-accounting-policies}{%
\subsubsection{Specific Accounting
Policies}\label{specific-accounting-policies}}

Clearly set them out. Could be GAAP, UK GAAP or IFRS. Also set out
specifics to avoid any grey areas, e.g.,

\begin{itemize}
\tightlist
\item
  In relation to bad debts.
\item
  Valuation of fixed assets
\item
  Applicable depreciation policy.
\end{itemize}

Consider imposing restrictive covenants to avoid a party trying to
improve the results of accounts by making unnecessary changes within
their business.

The seller, for example, should be bound in the pre-completion period
preventing it from, say, overstocking the business.

\hypertarget{fixed-price}{%
\subsubsection{Fixed Price}\label{fixed-price}}

There is a recent trend towards fixed price, or so-called `locked box',
transactions where the parties agree a price for the target company (it
is most commonly used in share acquisitions) which is fixed before the
sale and purchase agreement is signed.

Risky for the buyer, whose only opportunity for a price adjustment after
completion is by bringing a valid warranty claim.

\begin{env-3a4a363c-8d41-4783-bfde-af85d2b9ad5f}

Definition

In a locked box transaction, the economic interest in the target passes
to the buyer at a pre-signing date (the `locked box date'), and the
buyer therefore has the benefit of the cash profits generated by the
business from that date. The seller, however, does not receive the sale
proceeds until completion, and to compensate the seller for this, the
buyer will typically pay interest on the purchase price for the period
between the locked box date and closing.

\end{env-3a4a363c-8d41-4783-bfde-af85d2b9ad5f}

The buyer is wise to include provisions in the sale and purchase
agreement to protect against `leakage' -- that is, the seller extracting
value from the target between the date of the agreed accounts and the
completion date. Could include payment of a dividend, unsanctioned
transfer of assets to target shareholders, or inappropriate payment of
expenses by the target.

Specify any permitted leakage in the SPA and details of whether this
will result in a reduction of price.

Considerations:

\begin{itemize}
\tightlist
\item
  Certainty of price at an early stage
\item
  Simplicity of process means lower costs
\item
  Seller has increased control over setting the price. Useful especially
  in an auction process to compare bids.
\item
  Seller must have a suitable balance sheet available
\item
  Seller must ensure interest charged between locked box date and
  closing is not too low.
\item
  Buyer must negotiate price adjustments earlier in the process with
  less knowledge of the target.
\item
  Buyer must ensure the SPA clearly defines who is seller/ regulated
  person for the purposes of identifying leakage.
\end{itemize}

\hypertarget{earn-outs}{%
\subsubsection{Earn Outs}\label{earn-outs}}

\begin{env-3a4a363c-8d41-4783-bfde-af85d2b9ad5f}

Definition

An arrangement whereby at least part of the consideration is determined
by reference to the future profitability of the target for a specified
period after completion (eg for the three accounting periods following
completion).

\end{env-3a4a363c-8d41-4783-bfde-af85d2b9ad5f}

This avoids a buyer paying too much if the target fails to perform, but
motivates sellers/ managers who are staying on post-acquisition.

Provisions usually provide for accounts to be prepared for relevant
periods by the buyer's accountant, with an adjustment to the purchase
price dependent on a certain level of profitability. Will be based on a
multiple of the profits achieved.

Parties will agree a timetable for review of accounts, agreement of
earn-out profit and payment, and how any dispute will be resolved.

Civil law: take care to draft such provisions carefully. French law
requires the price to be defined, else it is void. Clauses where a
party's rights are disproportionate to its obligations can also be held
void under French law.

\begin{env-48e9bfc8-c3dd-4504-a5fa-3b34b9a27ccf}

Note

Seller and buyer may be pulling in different directions post-acquisition
(maximising short term profit vs long-term gain). Parties may limit the
tension by defining the objectives of the company post-completion.

\end{env-48e9bfc8-c3dd-4504-a5fa-3b34b9a27ccf}

\hypertarget{payment-terms}{%
\subsubsection{Payment Terms}\label{payment-terms}}

Often a fixed amount is paid in cash on the completion date, but parties
free to agree other terms (e.g., deferred payment, payment through
issuing of shares).

\hypertarget{deferred-consideration}{%
\subsubsection{Deferred Consideration}\label{deferred-consideration}}

Seller will usually demand interest on the outstanding amount and will
insist on security for payment. Purchase price often subject to an
adjustment dependent on completion accounts. Buyer may insist on a
retention of part of the purchase price as security in case a warranty
claim arises in relation to the target. Usually only appropriate where
there is an identified but unquantified liability.

\hypertarget{tax-treatment}{%
\paragraph{Tax Treatment}\label{tax-treatment}}

\begin{env-9193a5b6-f9d9-4a5c-9922-4e221c0d52be}

General rule

Seller will be charged to tax as at the date of sale on the total amount
of the consideration, regardless of whether that consideration has
actually been received. If the consideration is certain, the fact that
some of it is deferred will not affect the seller's tax position.

\end{env-9193a5b6-f9d9-4a5c-9922-4e221c0d52be}

Where the consideration cannot be ascertained at completion because it
is dependent on future events, such as the profitability of the target
after the sale (as with an earn-out arrangement), the rule set out in
Marren (Inspector of Taxes) v Ingles {[}1980{]} 3 All ER 95 applies and
tax is charged on the disposal of two separate capital assets.

\begin{itemize}
\tightlist
\item
  First disposal is of the shares of the target company/ assets of the
  target business.

  \begin{itemize}
  \tightlist
  \item
    Seller treated as having received as consideration both the
    consideration actually received at completion + a "chose in action"
    (right to receive further payment if future profits attained).
  \item
    Chose in action given a value by HMRC.
  \end{itemize}
\item
  Where the earn-out is received, treated as a second disposal by the
  seller.

  \begin{itemize}
  \tightlist
  \item
    If the payment exceeds original valuation by HMRC, seller treated as
    having made a capital gain, and charged to CGT/ corporation tax.
  \item
    Loss is an allowable loss for seller's capital gains purposes.
  \end{itemize}
\end{itemize}

\begin{env-f3fdeea8-9a69-4242-ac7f-874202f0281b}

Warning

The principle applies only where the parties do not, as at Completion,
have the information required to determine the deferred consideration.
If consideration is deferred pending agreement of Completion Accounts,
the parties do have the information required to determine the the
target's financial condition as at Completion but have not yet processed
the information.

\end{env-f3fdeea8-9a69-4242-ac7f-874202f0281b}

Buyer's tax position for stamp duty purposes also dependent on whether
deferred consideration is ascertainable at completion.

\begin{enumerate}
\tightlist
\item
  Where the consideration is certain but deferred, stamp duty is payable
  within 30 days of the completion date -- no allowance for the deferral
  is made.
\item
  Where the amount of the deferred consideration has not yet been
  determined but is ascertainable at completion, for example because it
  is dependent on the preparation of completion accounts, the Stamp
  Office applies the `wait and see' principle (stamp duty will be
  charged once the consideration has been finally determined).
\end{enumerate}

\begin{env-a57f3fff-c46d-4aeb-b36c-ada0d8066098}

Action

The buyer should nonetheless submit the stock transfer forms for
stamping within 30 days of completion, and make a payment on account of
stamp duty, to avoid penalties and reduce interest charges (which run
from 30 days after the completion date even if the consideration has yet
to be ascertained).

\end{env-a57f3fff-c46d-4aeb-b36c-ada0d8066098}

\begin{enumerate}
\tightlist
\item
  Where the deferred consideration is unascertainable at completion, as
  with an earn-out arrangement, no stamp duty is payable on it unless
  the `contingency' principle applies. Under the contingency principle,
  if the consideration is stated to be subject to a minimum or a maximum
  amount, or is to be calculated by reference to a basic or `benchmark'
  figure, then stamp duty is chargeable on that minimum, maximum or
  benchmark sum. Where both a minimum and a maximum are specified, stamp
  duty is charged on the maximum amount.
\end{enumerate}

\hypertarget{security}{%
\subsubsection{Security}\label{security}}

Where part of the purchase price does remain outstanding on completion,
the seller may seek security for the outstanding sums:

\begin{enumerate}
\tightlist
\item
  by taking a \textbf{charge} over some or all of the assets transferred
  to the buyer;
\item
  by requiring a \textbf{guarantee} of the buyer's obligation to pay the
  balance of the price (e.g., from parent company/ directors);
\item
  by providing in the agreement that \textbf{title} to specified assets
  is to remain with the seller until the purchase price is paid in full;
\item
  by obliging the buyer to place a specified sum in a \textbf{joint
  deposit account} (or escrow account) on completion and defining the
  circumstances in which this (and accrued interest) can be released to
  the parties.
\end{enumerate}

\hypertarget{form}{%
\subsubsection{Form}\label{form}}

The buyer will usually be borrowing. Where the buyer is a company, all
or part of the consideration may be satisfied by the company issuing
equity securities (shares) or debt securities (loan notes) to the
seller. Avoids the need to increase borrowing/ make disposals of assets
to fund the acquisition.

Tax: seller on a share sale may be able to roll over any capital gain
made on the disposal of shares in the target, deferring the capital tax
which it would otherwise pay until disposal of the securities in the
acquiring company received in exchange.

\hypertarget{shares}{%
\paragraph{Shares}\label{shares}}

The seller will accept consideration in the form of shares only if the
shares are readily marketable, i.e., they can in the future be realised
for cash. Where shares are issued as consideration, the agreement should
specify how they rank with the other shares of the acquiring company and
what rights the seller will have to any dividend declared in relation to
a period in which completion falls.

The buyer may insist on a clause restricting the seller from disposing
of more than a specified percentage of shares within a specified period
(to avoid depressing the price by selling lots of shares at once).

Where the acquiring company wishes to issue shares as consideration, but
the seller only wants cash, it can be achieved through ``vendor
placing''. This is an arrangement whereby the acquiring company issues
shares to the seller, but arranges (through its financial advisers) for
the shares to be sold on immediately to institutional investors. The
acquiring company undertakes in the sale and purchase agreement that the
sale of the consideration shares will yield a specified sum.

\hypertarget{debt-securities}{%
\paragraph{Debt Securities}\label{debt-securities}}

The acquiring company may issue debt securities like loan notes to the
seller for the purchase price/ part of it. Usually, the seller is
entitled to demand repayment of all/ part at 6 monthly intervals.

It may suit the seller to receive staggered payments in this way with
interest on the balance outstanding, particularly as roll-over relief
from capital tax is available in the UK on the same basis as for shares
under a securities exchange.

The form of the loan note will usually be agreed between the parties and
set out in a schedule to the sale and purchase agreement. ). A loan note
may or may not be secured, and will usually be an acceptable form of
payment for the seller only if the creditworthiness of the buyer can be
assured.

\hypertarget{contractual-provisions}{%
\subsection{Contractual Provisions}\label{contractual-provisions}}

The remaining section of the operative provisions in the sale and
purchase agreement will set out the basis on which the seller will
provide any contractual reassurances, or warranties, about the target
business or company. The actual warranty usually appears in a schedule,
but the operative provisions set out the basis on which warranties are
given.

\begin{env-c5a21b51-dc78-4e82-bd56-c0130c2151f9}

Example

Typically, restrictive covenants limiting the Seller's business
activities post Completion are included to protect the goodwill of the
target company.

\end{env-c5a21b51-dc78-4e82-bd56-c0130c2151f9}

The agreement may also provide for specific indemnities in relation to
identified problems. On a share purchase, this includes tax indemnities,
where the seller agrees to indemnify the buyer in respect of any tax
liability arising in the target.

Tax indemnities/ covenants may be drafted as a separate document, but
more usual to incorporate them into the main body of the agreement.

\hypertarget{dispute-resolution-and-governing-law}{%
\subsection{Dispute Resolution and Governing
Law}\label{dispute-resolution-and-governing-law}}

Specify any ADR methods to be used and/ or provisions for litigation,
including a forum selection provision and governing law provisions.
Provides certainty and prevents future litigation of such issues.

Usual to also contain a notice clause specifying a mechanism for the
service of notices from one party to the other as a way of exercising
contractual rights/ demanding compliance with certain obligations. Pay
attention to time ones, business days and national holidays.

\hypertarget{schedules}{%
\subsection{Schedules}\label{schedules}}

The bulk of the SPA is usually attributable to schedules. Issues covered
will be determined by the terms of the transaction and whether it is an
asset or share acquisition. Include:

\begin{itemize}
\tightlist
\item
  Details of assets/ shareholdings to be transferred and prices
\item
  Details of price adjustment mechanism
\item
  Warranty statements about conditions of target
\item
  Limitations on liability in relation to warranty claims
\item
  Specific indemnities agreed between parties in relation to identified
  risks
\item
  Agreed forms of documentation governing associated transactions
\item
  Agreed restrictions on running the target business pending completion,
  if the agreement is exchanged conditionally (meaning a gap between
  exchange and completion).
\item
  Details of the completion procedure.
\end{itemize}

\hypertarget{execution-1}{%
\subsection{Execution}\label{execution-1}}

The sale and purchase agreement is purely a contract setting out the
terms on which the acquisition will be undertaken, so it can simply be
signed by the parties. However, to avoid any question about whether
appropriate consideration has been given in relation to specific
indemnities and restrictive covenants included in the agreement, the
buyer may prefer the agreement to be executed as a deed.

Prior to exchange and completion, the legal advisers will confirm the
parties' capability to enter into the agreement.

\hypertarget{risk---buyers-protection}{%
\section{Risk - Buyer's Protection}\label{risk---buyers-protection}}

\hypertarget{protections-implied-into-the-contract}{%
\subsection{Protections Implied into the
Contract}\label{protections-implied-into-the-contract}}

\hypertarget{shares-1}{%
\subsubsection{Shares}\label{shares-1}}

On a share acquisition under English law, neither common law nor statute
protects the buyer by implying terms into the contract. \emph{Caveat
emptor} applies.

In the absence of express provisions, an aggrieved buyer has very little
comeback on the seller unless a misrepresentation can be established.

\hypertarget{assets}{%
\subsubsection{Assets}\label{assets}}

The buyer of the assets of a business is not in such an exposed
position, as it does not assume automatically all the liabilities of the
business. The buyer may also get the benefit of some limited warranties
under the Sale of Goods Act 1979 for some of the assets being
transferred.

\hypertarget{civil-code-jurisdictions}{%
\subsubsection{Civil Code
Jurisdictions}\label{civil-code-jurisdictions}}

The buyer is seen as the party most at risk in the acquisition.
Statutory warranties are implied into the terms of the acquisition.
These outline the minimum protection to be provided to any buyer. Often
expressed as a guarantee by the seller that individual warranty
statements are true.

Civil code jurisdictions impose a duty of good faith between the parties
throughout the negotiations for the proposed acquisition. Scope varies
between countries. But even here, the buyer will usually seek additional
contractual warranties, though these can often be looser/ wider than
common law jurisdictions. Statutory warranties commonly included
explicitly.

\hypertarget{title-guarantee}{%
\subsection{Title Guarantee}\label{title-guarantee}}

The covenants are implied by the Law of Property (Miscellaneous
Provisions) Act 1994, and the alternative dispositions are sale with
full title guarantee and sale with limited title guarantee.

Full title guarantee implies:

\begin{enumerate}
\tightlist
\item
  the seller has the \textbf{right to dispose} of the property;
\item
  the seller will do all it \textbf{reasonably} can, at its own expense,
  to pass to the buyer the \textbf{title} it purports to give; and
\item
  the property is free from all \textbf{charges, encumbrances and third
  party rights}, other than those of which the seller is unaware and
  could not reasonably be expected to be aware.
\end{enumerate}

Limited title guarantee implies (1) and (2), and also that the seller
has neither created nor allowed to be created any subsisting charge or
encumbrance, and is not aware that anyone else has done so.

The type of guarantee to be given is to be negotiated. Implied covenants
often modified by express provisions relating to the nature and
perfection of title to the assets. The receiver or trustee selling
assets may be prepared to offer no more comfort than a limited title
guarantee.

\hypertarget{warranties}{%
\subsection{Warranties}\label{warranties}}

Much of the SPA is taken up with specific warranties by the seller in
favour of the buyer:

\begin{itemize}
\tightlist
\item
  Accounts
\item
  Subsisting contracts
\item
  Employees
\item
  Pensions
\item
  IP
\end{itemize}

The warranties are contractual statements about what is to be acquired.
On a share acquisition, the warranties will include statements about the
shares to be acquired.

As the warranties are terms of the contract, if they prove to be untrue,
the buyer will have a claim for damages against the warrantor(s) for
breach of contract.

The warranties are made about the state of the target as at a specified
point. Where exchange and completion are simultaneous, the warranties
will be statements given at the moment of completion.

Where there is a gap between exchange and completion, warranties are
made at exchange when the parties become contractually bound, and may
then be confirmed at completion.

\hypertarget{damages-recoverable}{%
\subsubsection{Damages Recoverable}\label{damages-recoverable}}

In Hadley v Baxendale (1854) 9 Ex 341, loss will be recoverable if it is
not too remote.

2 categories:

\begin{enumerate}
\tightlist
\item
  Loss which flows naturally from the breach
\item
  Loss which was fairly and reasonably in the contemplation of both
  parties.
\end{enumerate}

The aim of contractual damages is to put the buyer in the position it
would have been in if the contract had not been breached, subject to the
duty to mitigate.

\hypertarget{general-principles}{%
\subsubsection{General Principles}\label{general-principles}}

Application to breach of warranty:

The buyer's loss is the difference between the value of the shares if
the warranty had been true and their actual value (i.e., the value of
the shares with and without the breach). The basis for valuing the
shares will be relevant in assessing the loss.

\begin{env-c5a21b51-dc78-4e82-bd56-c0130c2151f9}

Example

Where there is a breach of warranty which results in the assets of the
target being less than expected, this may not have a direct effect on
the value of the shares, unless they were valued on an assets-related
basis.

\end{env-c5a21b51-dc78-4e82-bd56-c0130c2151f9}

\hypertarget{pre-estimate-of-damages}{%
\subsubsection{Pre-estimate of Damages}\label{pre-estimate-of-damages}}

The parties often specify in the agreement how the loss should be
quantified if certain warranties are broken. Common to include a clause
specifying the appropriate multiplier for any shortfall.

There is commonly a clause obliging the seller to pay for any deficiency
in assets or any undisclosed liability arising from a breach of
warranty. Avoids the need to prove an equivalent diminution in the value
of the shares.

Liquidated damages and penalties:

\begin{itemize}
\tightlist
\item
  Where the contractually agreed consequences of a breach are
  oppressive, the penalty rule applies (Cavendish Square Holdings BV v
  Talal El Makdessi {[}2015{]} UKSC 67)
\end{itemize}

\begin{env-b225a0d9-0c32-46a2-8b5d-c7fc92570012}

\emph{Makdessi} test

The true test for a penalty is therefore now `whether the impugned
provision is a secondary obligation which imposes a detriment on the
contract-breaker out of all proportion to any legitimate interest of the
innocent party in the enforcement of the primary obligation'.

\end{env-b225a0d9-0c32-46a2-8b5d-c7fc92570012}

In many civil code jurisdictions, warranties are likely to be structured
as guarantees, whereby the seller guarantees that the individual
warranties are correct. In the event that the object of the purchase
does not meet the standards set out in the warranties, the buyer will be
entitled to the remedies as set out in the sale and purchase agreement.

\hypertarget{representations}{%
\subsection{Representations}\label{representations}}

If any statements which have been made by or on behalf of the seller
prove to be untrue, the buyer may under common law have a claim for
misrepresentation.

\begin{env-48e9bfc8-c3dd-4504-a5fa-3b34b9a27ccf}

Note

Misrepresentation is a tort.

\end{env-48e9bfc8-c3dd-4504-a5fa-3b34b9a27ccf}

\hypertarget{general-principles-1}{%
\subsubsection{General Principles}\label{general-principles-1}}

A misrepresentation is a false statement of fact made by one party to
the contract to the other, which induces the other party to enter into
the contract. The usual remedies are rescission or damages.

\hypertarget{rescission}{%
\paragraph{Rescission}\label{rescission}}

Rescission is a remedy aimed at putting the parties back into the
pre-contract position.

Qualifications:

\begin{enumerate}
\tightlist
\item
  s 2(2) of the Misrepresentation Act 1967 enables the court to award
  damages instead of rescission for innocent or negligent
  misrepresentation. The court may exercise this discretion if it feels
  that rescission is too drastic a remedy in the circumstances.
\item
  Equitable remedy and therefore subject to equitable bars

  \begin{enumerate}
  \tightlist
  \item
    Impossible to restore parties to their pre-contract position
  \item
    Bona fide third party rights acquired
  \item
    Innocent party takes action affirming the agreement
  \item
    Undue delay seeking relief.
  \end{enumerate}
\end{enumerate}

\hypertarget{damages}{%
\paragraph{Damages}\label{damages}}

If the misrepresentation is innocent or negligent, s 2(2) of the
Misrepresentation Act 1967 provides that the court `may declare the
contract subsisting and award damages in lieu of rescission'. This is a
court discretion to award damages; no contractual right.

\begin{env-b56569b0-a3e1-47fe-b37c-6cf3ba4ee6a2}

Can the court still award damages in lieu of rescission if rescission
itself has been lost due to one of the equitable bars?

No: Salt v Stratstone Specialist Ltd {[}2015{]} EWCA Civ 745

\end{env-b56569b0-a3e1-47fe-b37c-6cf3ba4ee6a2}

If the misrepresentation is negligent, the innocent party is
\textbf{entitled} to damages under s 2(1). Under s 2(1), damages are
measured on the tortious basis, which aims to put the innocent party in
the position it would have been in if the tort had not been committed.

\hypertarget{representations-warranties}{%
\subsubsection{Representations \&
Warranties}\label{representations-warranties}}

In an SPA, any representations made by the seller which induce the buyer
to contract will invariably be included as warranties in the agreement.

It is rare for a claim for misrepresentation to be founded on a matter
which is not included in the contract. It is common for the agreement to
incorporate an acknowledgement by the buyer that it has not relied on
any representations which are not contained in the contract
(``non-reliance provision'') and an ``entire agreement'' clause.

If a clause attempts to exclude liability for misrepresentation (as is
common), the exclusion must be reasonable to be effective (s 3
Misrepresentation Act 1967).

\hypertarget{entire-agreement-clause}{%
\paragraph{Entire Agreement Clause}\label{entire-agreement-clause}}

In Thomas Witter Ltd v TBP Industries Ltd {[}1996{]} 2 All ER 573 it was
held that a contractual term purporting to confine the parties'
agreement to the terms of the contract will not, per se, exclude
liability for misrepresentation.

To exclude liability for misrepresentation, an express exclusion
provision in clear terms must be included in the contract (confirmed in
Axa Sun Life Services plc v Campbell Martin Ltd and others {[}2011{]}
EWCA Civ 133).

\hypertarget{non-reliance-provision}{%
\paragraph{Non-reliance Provision}\label{non-reliance-provision}}

Estops the buyer from claiming that it entered into an agreement in
reliance on pre-contractual statements made by the seller.

\begin{env-b56569b0-a3e1-47fe-b37c-6cf3ba4ee6a2}

Since non-reliance statements prevented liability for misrepresentation
rather than strictly excluding it, are they subject to the
reasonableness test in s 3 MA 1967?

Yes, Springwell Navigation Corp v JP Morgan Chase Bank {[}2010{]} EWCA
Civ 1221. The practical reality is excluding or restricting liability.

\end{env-b56569b0-a3e1-47fe-b37c-6cf3ba4ee6a2}

To make such statements ``reasonable'', it is common for the seller to
ensure that the buyer confirms its non-reliance only in relation to
non-fraudulent misrepresentation, with an express exclusion in relation
to fraudulent misrepresentation. This is despite a House of Lords House
of Lords decision in HIH Casualty and General Insurance Ltd v Chase
Manhattan Bank {[}2003{]} UKHL 6 suggesting this is not necessary. Held
that `in the absence of words which expressly refer to dishonesty, it
goes without saying that underlying the contractual arrangements of the
parties will be a common assumption that the persons involved will
behave honestly'.

\begin{env-a57f3fff-c46d-4aeb-b36c-ada0d8066098}

Action

It is nevertheless considered good practice that the Buyer expressly
agrees to waive its right to damages or recission only in relation to
non-fraudulent misrepresentation.

\end{env-a57f3fff-c46d-4aeb-b36c-ada0d8066098}

\hypertarget{right-of-rescission}{%
\paragraph{Right of Rescission}\label{right-of-rescission}}

Even when including an entire agreement clause and non-reliance clause,
many representations which the seller makes to the buyer will be
included as express warranties in the agreement. If any of these
warranties are broken, the buyer will be able to claim damages for
breach of contract.

\begin{env-48e9bfc8-c3dd-4504-a5fa-3b34b9a27ccf}

Note

s 1 Misrepresentation Act 1967: where a misrepresentation has become a
term of the contract, this will not affect the innocent party's right to
rescind (provided there are no equitable bars). Could be important if
the buyer becomes aware of the falsity of the representation after
exchange but before completion.

\end{env-48e9bfc8-c3dd-4504-a5fa-3b34b9a27ccf}

If there is a gap between exchange and purchase, the buyer will usually
negotiate a contractual right to withdraw from the contract if it
becomes aware of any breaches of warranty. Backed up by obliging the
seller to inform the buyer if it becomes aware of anything inconsistent
with the warranties.

Advantage: buyer can rescind regardless of whether the breach of
warranty was also a misrepresentation.

\hypertarget{damages-1}{%
\paragraph{Damages}\label{damages-1}}

The difference in the measure of damages for breach of contract and
misrepresentation may also be significant.

\begin{itemize}
\tightlist
\item
  The contractual measure allows for loss of bargain and, on a share
  purchase, involves assessing the difference between the actual value
  of the shares and their value if the warranty had been true.
\item
  Tortious damages aim to put the buyer in the position it would have
  been in if it had not entered the agreement ("restitutionary").
  Involves calculating the difference between the price paid for the
  shares and their market value.
\end{itemize}

Contractual measures often lead to higher damages (due to loss of
bargain) but not always.

\hypertarget{indemnities}{%
\subsection{Indemnities}\label{indemnities}}

A \textbf{warranty} is an undertaking by the seller that a particular
state of affairs exists.

\begin{env-3a4a363c-8d41-4783-bfde-af85d2b9ad5f}

Definition

An \textbf{indemnity} is a promise to reimburse the buyer in respect of
a designated type of liability which may arise in the future.

\end{env-3a4a363c-8d41-4783-bfde-af85d2b9ad5f}

On a share sale, the recipient of the indemnity receives an amount equal
to the actual liability (+ costs and expenses).

\begin{env-8890de02-1ba1-42c2-9b89-cb105d97c320}

Important

Another difference between a warranty and an indemnity is that, in the
case of the latter, there is no duty to mitigate the loss; such a duty
may, however, be incorporated in the contract.

\end{env-8890de02-1ba1-42c2-9b89-cb105d97c320}

\begin{itemize}
\tightlist
\item
  In the US, losses for breach of warranty are often qualified
\item
  Many civil code jurisdictions do not recognise the distinction between
  warranties and indemnities.
\end{itemize}

\hypertarget{taxation-indemnities}{%
\subsubsection{Taxation Indemnities}\label{taxation-indemnities}}

It is common for buyers of shares to seek protection from tax
liabilities in the form of both warranties and indemnities.

\begin{itemize}
\tightlist
\item
  The warranties usually deal with the target's compliance with tax and
  VAT requirements.
\item
  The indemnities will cover specific tax charges which may arise over
  and above those provided for in the accounts.
\end{itemize}

\begin{env-8073ffbe-6656-43d2-bf4d-97e604cb2369}

Tip

Dealing with most tax matters by way of indemnity has the advantage to
the buyer that there is no need to prove the link between the unexpected
tax liability and the value of the shares.

\end{env-8073ffbe-6656-43d2-bf4d-97e604cb2369}

\hypertarget{who-gives}{%
\subsection{Who Gives}\label{who-gives}}

Some sellers may try to limit their share of liability. Others may be
unwilling to accept any liability. The buyer will want all sellers to
give warranties and indemnities both jointly and severally.

\hypertarget{joint-and-several-liability}{%
\subsubsection{Joint and Several
Liability}\label{joint-and-several-liability}}

\begin{itemize}
\tightlist
\item
  Where there is more than one warrantor the buyer will invariably
  insist that they accept joint and several liability.
\item
  s 1 Civil Liability (Contribution) Act 1978 gives the party called
  upon to discharge the liability the right to recover a contribution
  from the other warrantors who are liable in respect of the same
  damage.

  \begin{itemize}
  \tightlist
  \item
    The court determines the amount of the contribution recoverable from
    the others on the basis of what it considers just and equitable
    having regard to the extent of their responsibility for the
    particular damage (s 2).
  \item
    Common for warrantors to agree how any liability should be borne
    between them to avoid uncertainty.

    \begin{itemize}
    \tightlist
    \item
      May be set out in a ``deed of contribution'' or incorporated into
      the SPA (expressed ``as between the warrantors'').
    \end{itemize}
  \end{itemize}
\end{itemize}

Where there are a large number of selling shareholders., the buyer may
not insist that they all give warranties and indemnities (minority
shareholders often excluded).

\hypertarget{trustees}{%
\subsubsection{Trustees}\label{trustees}}

Suppose some of the shares of a company are held in trust. Trustees will
wish to avoid personal liability and may refuse to give any warranties
or undertakings, other than that they have unencumbered title to the
shares.

\begin{env-a57f3fff-c46d-4aeb-b36c-ada0d8066098}

Possible solution

Limit the liability of trustees to the net value (after tax) of the
capital of the trust. The danger to the buyer of the capital being
distributed shortly after completion may be met by obliging the trustees
to require any beneficiary receiving capital to give appropriate
warranties to the buyer.

\end{env-a57f3fff-c46d-4aeb-b36c-ada0d8066098}

\hypertarget{substantial-shareholders}{%
\subsubsection{Substantial
Shareholders}\label{substantial-shareholders}}

Some powerful investors (PE funds) may avoid giving warranties and
indemnities on the basis that they have no involvement with the
management of the target company and cannot be expected to give promises
in relation to matters they know nothing about.

\begin{env-a57f3fff-c46d-4aeb-b36c-ada0d8066098}

Action

The role of warranties is to allocate risk between seller and buyer
{\(\rightarrow\)} the price negotiated is on the assumption that the
warranties are true. So could have a deal where the shareholders who
give a warranty get more cash than others.

\end{env-a57f3fff-c46d-4aeb-b36c-ada0d8066098}

\hypertarget{sellers-rights-against-management}{%
\subsubsection{Seller's Rights Against
Management}\label{sellers-rights-against-management}}

Suppose there is a breach of warranty. The sellers may try to shift
responsibility back onto the directors of the target. But these
directors may be employed by the buyer. So the buyer may ask the selling
shareholders to waive any right they may have to bring a claim against
the management/ the target.

\hypertarget{assigning-warranties-and-indemnities}{%
\subsubsection{Assigning Warranties and
Indemnities}\label{assigning-warranties-and-indemnities}}

\begin{env-9193a5b6-f9d9-4a5c-9922-4e221c0d52be}

General rule

The general position under English law is that, unless the contracting
parties agree otherwise, the benefit of an agreement can be assigned to
a third party without the consent of the other party to the contract.

\end{env-9193a5b6-f9d9-4a5c-9922-4e221c0d52be}

\begin{itemize}
\tightlist
\item
  The seller is therefore likely to insist on an express provision
  stating that no assignment is permitted/ only permitted on approval.
\item
  The buyer may try to negotiate with the seller an express right to
  assign the benefit of the warranties and indemnities.
\item
  Common for the seller to agree to the buyer assigning the benefit of
  warranties and indemnities to members of its own group.
\end{itemize}

\hypertarget{buyers-security-for-breach}{%
\subsection{Buyer's Security for
Breach}\label{buyers-security-for-breach}}

The buyer is in an exposed position:

\begin{itemize}
\tightlist
\item
  Buyer or target may be directly liable to 3rd parties
\item
  May be unable to recover the loss if the seller is in financial
  difficulties.
\item
  Common to ask for security from the seller.
\end{itemize}

\hypertarget{guarantees}{%
\subsubsection{Guarantees}\label{guarantees}}

Where the seller of a business or shares is a company which is part of a
group, the seller's parent company may be prepared to guarantee the
seller's obligations.

For a company owned by individuals, insist on guarantees from the
shareholders.

\hypertarget{retentions-from-purchase-price}{%
\subsubsection{Retentions From Purchase
Price}\label{retentions-from-purchase-price}}

May be used to protect the buyer's position:

\begin{itemize}
\tightlist
\item
  Payment of the retention into a joint account in the names of the
  seller's and buyer's solicitors;
\item
  The retention to be remitted to the seller on a certain date after
  completion, subject to any claims made before that date.
\item
  Where a claim is made, the balance of the retention to be paid to the
  seller.
\item
  The buyer to be paid the amount of the claim within a short period.
\item
  Payment of accrued interest on the fund to the parties.
\item
  An obligation on the buyer to pursue any claims against the seller
  which may delay the remittance of the fund promptly and diligently.
\item
  Buyer obliged to provide full details of any claim backed up by an
  opinion of counsel.
\end{itemize}

The retention will not affect the capital tax position of parties
(unless used to pay warranty claims).

\hypertarget{restrictions-on-the-seller}{%
\subsection{Restrictions on the
Seller}\label{restrictions-on-the-seller}}

The value of the buyer's investment could be reduced substantially by
the post-completion activities of the seller, whose detailed knowledge
of the target would enable the seller to cause it considerable harm.

\hypertarget{no-express-restraints}{%
\subsubsection{No Express Restraints}\label{no-express-restraints}}

Trego v Hunt {[}1896{]} AC 7 established that on the sale of the
goodwill of a business, the courts will refuse to imply a covenant by
the seller not to set up in competition.

The only implied undertakings of the seller are:

\begin{enumerate}
\tightlist
\item
  Not to use or disclose confidential information relating to the
  business.
\item
  Not to represent itself as a successor to the business, or carrying on
  the same business.
\item
  Not to solicit customers of the business.
\end{enumerate}

\hypertarget{express-restrictions}{%
\subsubsection{Express Restrictions}\label{express-restrictions}}

As the seller is being paid for goodwill, the buyer is justified in
seeking ways of preventing the seller from damaging it. But the buyer
must go no further than is reasonably necessary to protect the goodwill
of the target.

Common undertakings:

\begin{itemize}
\tightlist
\item
  Covenant not to be engaged or concerned in any competing business for
  a specified period after completion.
\item
  Covenant not to try to solicit/ entice away from the target customers
  or suppliers who have recently dealt with the target, for a specified
  period after completion.
\item
  Covenant not to try to solicit or entice away employees of the target
  for a specified period after completion.
\item
  Undertaking not to use or disclose any confidential information about
  the target or its customers.
\item
  Covenant not to use the name of the target.
\end{itemize}

\hypertarget{validity-at-common-law}{%
\subsubsection{Validity at Common Law}\label{validity-at-common-law}}

\begin{env-8073ffbe-6656-43d2-bf4d-97e604cb2369}

Tip

Under English law, all covenants in restraint of trade are prima facie
void at common law. However, the court will not strike down restrictive
covenants which it considers to be reasonable to protect a legitimate
interest of the buyer.

\end{env-8073ffbe-6656-43d2-bf4d-97e604cb2369}

The court will consider:

\begin{longtable}[]{@{}ll@{}}
\toprule()
Consideration & Details \\
\midrule()
\endhead
\textbf{Duration} of the restraint & 1-5 years after completion is
common. \\
\textbf{Geographical area} of restraint & If the covenant is
appropriately worded, it may not be necessary to stipulate a
geographical limit. If the business restricted is defined, specify a
reasonable area of restriction. \\
\textbf{Activities} restricted & Any attempt to extend the restrictions
to activities which are not carried on by the target itself will fail,
on the grounds that they are excessive in order to protect the goodwill
of the target. in Ronbar Enterprises Ltd v Green {[}1954{]} 2 All ER
266, the court considered a clause preventing the seller carrying on a
business `similar to' that of the target to be too wide. The restriction
should also be reasonable given the extent to which the party giving the
covenant has been actively involved in the business. Non-solicitation
clauses should not be so wide as to include those with no recent
dealings with the target. \\
\bottomrule()
\end{longtable}

\hypertarget{enforcement}{%
\subsubsection{Enforcement}\label{enforcement}}

\begin{env-b56569b0-a3e1-47fe-b37c-6cf3ba4ee6a2}

What if only some of the restrictions are reasonable?

If each restraint is contained in a separate clause or sub-clause and
expressed to be separate and independent, they remaind independently
enforceable. The court is sometimes prepared to "blue pencil" certain
parts of a clause, leaving the remaining parts enforceable (Tillman v
Egon Zehnder {[}2019{]} UKSC 32) though the court will not rewrite a
clause to make it reasonable.

\end{env-b56569b0-a3e1-47fe-b37c-6cf3ba4ee6a2}

\hypertarget{competition-act-1998}{%
\subsubsection{Competition Act 1998}\label{competition-act-1998}}

Unlikely that UK courts will deviate from the European regime. There
remains a general prohibition of anticompetitive agreements and a
prohibition of abuse of a dominant position. If an agreement is found to
breach s 2 or s 18 of the Competition Act 1998 , the agreement or
(subject to severance) the anticompetitive terms will be unenforceable.

\hypertarget{anticompetitive-agreements}{%
\paragraph{Anticompetitive
Agreements}\label{anticompetitive-agreements}}

\begin{env-69c6d983-4e25-44a5-9a13-662902d85e33}

s 2 Competition Act 1998

(1) Subject to section 3, agreements between undertakings, decisions by
associations of undertakings or concerted practices which---

\begin{itemize}
\tightlist
\item
  (a) may affect trade within the United Kingdom, and
\item
  (b) have as their object or effect the prevention, restriction or
  distortion of competition within the United Kingdom,
\end{itemize}

are prohibited unless they are exempt in accordance with the provisions
of this Part.

(2) Subsection (1) applies, in particular, to agreements, decisions or
practices which---

\begin{itemize}
\tightlist
\item
  (a) directly or indirectly fix purchase or selling prices or any other
  trading conditions;
\item
  (b) limit or control production, markets, technical development or
  investment;
\item
  (c) share markets or sources of supply;
\item
  (d) apply dissimilar conditions to equivalent transactions with other
  trading parties, thereby placing them at a competitive disadvantage;
\item
  (e) make the conclusion of contracts subject to acceptance by the
  other parties of supplementary obligations which, by their nature or
  according to commercial usage, have no connection with the subject of
  such contracts.
\end{itemize}

(3) Subsection (1) applies only if the agreement, decision or practice
is, or is intended to be, implemented in the United Kingdom.

(4) Any agreement or decision which is prohibited by subsection (1) is
void.

\end{env-69c6d983-4e25-44a5-9a13-662902d85e33}

N.B. any agreement that is exempt under EU law will also be exempt under
UK legislation.

\hypertarget{abuse-of-a-dominant-position}{%
\paragraph{Abuse of a Dominant
Position}\label{abuse-of-a-dominant-position}}

The wording of s 18 of the 1998 Act concerning abuse of a dominant
position is similar to that of Art 102 TFEU. There is a prohibition on
the activities of one or more undertakings which amount to the abuse of
a dominant position within the UK or a substantial part of it.

\begin{env-69c6d983-4e25-44a5-9a13-662902d85e33}

s 18 Competition Act 1998 - Abuse of dominant position.

(1) Subject to section 19, any conduct on the part of one or more
undertakings which amounts to the abuse of a dominant position in a
market is prohibited if it may affect trade within the United Kingdom.

(2) Conduct may, in particular, constitute such an abuse if it consists
in---

\begin{itemize}
\tightlist
\item
  (a) directly or indirectly imposing unfair purchase or selling prices
  or other unfair trading conditions;
\item
  (b) limiting production, markets or technical development to the
  prejudice of consumers;
\item
  (c) applying dissimilar conditions to equivalent transactions with
  other trading parties, thereby placing them at a competitive
  disadvantage;
\item
  (d) making the conclusion of contracts subject to acceptance by the
  other parties of supplementary obligations which, by their nature or
  according to commercial usage, have no connection with the subject of
  the contracts.
\end{itemize}

\end{env-69c6d983-4e25-44a5-9a13-662902d85e33}

\hypertarget{consequences}{%
\paragraph{Consequences}\label{consequences}}

\begin{itemize}
\tightlist
\item
  Restrictions in breach of the 1998 Act will be void.
\item
  The CMA is empowered to impose fines of up to 10\% of an undertaking's
  turnover.
\item
  The parties may be ordered to cease or modify the agreement.
\item
  CMA may conduct a dawn raid following a search warrant obtained from a
  High Court judge.
\end{itemize}

\hypertarget{eu-law}{%
\subsubsection{EU Law}\label{eu-law}}

Restrictive covenants in the sale and purchase agreement may breach Arts
101 and 102 TFEU (particularly Art 101) if trade between Member States
is affected. The European Commission's Notice of March 2005 confirms
that, on a transfer of goodwill, a non-compete period of up to two years
will be appropriate, whereas a period of three years is justified where
the buyer acquires know-how and goodwill.

\hypertarget{risk---sellers-limitations}{%
\section{Risk - Seller's Limitations}\label{risk---sellers-limitations}}

\hypertarget{limiting-scope}{%
\subsection{Limiting Scope}\label{limiting-scope}}

The first draft of the sale and purchase agreement prepared by the buyer
will often include extensive warranties and indemnities, with few
specific limitations on the seller's liability. In most cases, there is
considerable scope for negotiating the agreement.

Where warranties are given by reference to the `knowledge, information
or belief' of the seller, the buyer will usually insist on the seller
acknowledging that in giving the warranties it has made all due
enquiries and taken all reasonable steps to ensure their accuracy.

\hypertarget{limitations-on-claims}{%
\subsection{Limitations on Claims}\label{limitations-on-claims}}

The seller will also seek to restrict its liability in relation to the
warranties and indemnities generally by including a set of standard
limitation clauses in the agreement.

\begin{env-8890de02-1ba1-42c2-9b89-cb105d97c320}

Share sale

Limitations on the seller's liability for breach of warranty contained
in a share sale agreement are not controlled by UCTA 1977 since Sch 1,
para 1(e) precludes the application of its relevant provisions to any
contract so far as it relates to the `creation or transfer of securities
or of any right or interest in securities'.

\end{env-8890de02-1ba1-42c2-9b89-cb105d97c320}

But UCTA 1977 applies for asset sale agreements.

\begin{env-69c6d983-4e25-44a5-9a13-662902d85e33}

s 3 UCTA 1977 - Liability arising in contract

(1) This section applies as between contracting parties where one of
them deals on the other's written standard terms of business.

(2) As against that party, the other cannot by reference to any contract
term---

\begin{itemize}
\tightlist
\item
  (a) when himself in breach of contract, exclude or restrict any
  liability of his in respect of the breach; or
\item
  (b) claim to be entitled---

  \begin{itemize}
  \tightlist
  \item
    (i) to render a contractual performance substantially different from
    that which was reasonably expected of him, or
  \item
    (ii) in respect of the whole or any part of his contractual
    obligation, to render no performance at all,
  \end{itemize}
\end{itemize}

except in so far as (in any of the cases mentioned above in this
subsection) the contract term satisfies the requirement of
\textbf{reasonableness}.

\end{env-69c6d983-4e25-44a5-9a13-662902d85e33}

\begin{itemize}
\tightlist
\item
  Section 3 UCTA 1977 applies only where one party deals on the other's
  written standard terms of business
\item
  The reasonableness test will be relevant in connection with any
  exclusion of liability for misrepresentation.
\item
  Freely negotiated warranty limitations contained in an acquisition
  agreement are unlikely to be the subject of a successful challenge.
\end{itemize}

\hypertarget{limits-on-amount-of-claims}{%
\subsubsection{Limits on Amount of
Claims}\label{limits-on-amount-of-claims}}

\hypertarget{maximum-limits}{%
\paragraph{Maximum Limits}\label{maximum-limits}}

The seller will negotiate for a maximum liability under the warranties
and indemnities (commonly the price paid by the buyer for the target).

\begin{env-48e9bfc8-c3dd-4504-a5fa-3b34b9a27ccf}

Note

In some circumstances, damages claimed for warranty or indemnity
breaches can exceed the purchase price (e.g., where a target company has
been sold at an undervalue). But even here, a properly drafted maximum
liability clause may protect a seller and limit its liability to the
maximum liability it has contractually agreed with a buyer (116 Cardamon
Ltd v MacAlister {[}2019{]} EWHC 1200 (Comm)).

\end{env-48e9bfc8-c3dd-4504-a5fa-3b34b9a27ccf}

\hypertarget{minimum-limits}{%
\paragraph{Minimum Limits}\label{minimum-limits}}

The parties will often agree a minimum (aggregate) threshold for claims,
typically 0.5-1\% of the purchase price. The buyer should insist that
once the threshold has been breached, it is entitled to recover the full
value of the claim.

The seller may seek a \emph{de minimis} limit on individual claims. But
difficulties can arise, however, in respect of continuing breaches and
claims which are similar in nature or arise from the same default.

\hypertarget{time-limits}{%
\subsubsection{Time Limits}\label{time-limits}}

\begin{longtable}[]{@{}ll@{}}
\toprule()
Breach & Limitation period \\
\midrule()
\endhead
Breach of contract & 6 years from the date of breach (which for a
warranty is the date of the SPA). \\
Breach of deed & 12 years from the date of breach. \\
\bottomrule()
\end{longtable}

\hypertarget{tax-matters}{%
\paragraph{Tax Matters}\label{tax-matters}}

It is common to link the period in which the buyer can make claims
relating to tax matters to HMRC's time limit for making an assessment to
tax (four years; previously six years) from the end of the relevant tax
year, except to recover tax lost through fraudulent or negligent
conduct.

\hypertarget{non-tax-matters}{%
\paragraph{Non-tax Matters}\label{non-tax-matters}}

A much shorter period can be negotiated, generally linked to the target
company's audit (since the audit may ``flush out'' any breached of
warranty). A period allowing for two full audits to be carried out and
their results digested is commonly agreed. Consider a long-stop time
limit on bringing claims, along with a shorter deadline for bringing
claims without informing the party of the breach.

\hypertarget{insurance}{%
\subsubsection{Insurance}\label{insurance}}

The seller will invariably seek to prevent the buyer from bringing a
claim where the loss is covered by insurance taken out in relation to
the target.

\hypertarget{recovery-from-3rd-parties}{%
\subsubsection{Recovery From 3rd
Parties}\label{recovery-from-3rd-parties}}

The buyer may also be required to give credit for any sums received from
third parties in relation to the subject matter of the claim. Usually
expressed to be wide enough to cover tax reductions.

\hypertarget{understated-assets}{%
\subsubsection{Understated Assets}\label{understated-assets}}

The seller may want to include a provision that any claim will be
reduced by the amount by which the seller can show that the net assets
of the target were understated in the accounts drawn up before
completion. Buyer beware of accepting this.

\hypertarget{conduct-of-claims}{%
\subsubsection{Conduct of Claims}\label{conduct-of-claims}}

Suppose:

MERMAID1

In this situation, the seller may argue that it should have the right to
conduct the claim (if the buyer knows it can sue under a warranty, may
defend the claim less vigorously). The buyer is under a duty to mitigate
its loss, but a failure to mitigate is difficult to prove.

The buyer may have reservations about allowing the seller to have
control in case it damages the reputation and goodwill of the target.

Often a compromise is reached, e.g., consent of the other party required
before settling/ indemnity for legal costs.

\hypertarget{mitigation}{%
\paragraph{Mitigation}\label{mitigation}}

Where one party has suffered loss resulting from the other party's
breach of contract, the injured party should take 'reasonable steps' to
minimise the effect of the breach (British Westinghouse Electric and
Manufacturing Co v Underground Electric Rail Co {[}1912{]} AC 673).

There is no obligation for the claimant to mitigate their loss, but
losses attributable to a failure to do so are not legally recoverable.
So the innocent party cannot seek compensation by the party in default
for the loss which is due to the failure to behave reasonably after the
breach.

The question of which steps count as 'reasonable' is one of fact
(Pilkington v Wood {[}1953{]} CH 770). It was held that there was no
expectation that the claimant should embark on 'a complicated and
difficult piece of litigation' in order to minimise the effects of the
defendant's breach.

Payzu v Saunders {[}1919{]} 2 KB 581: reasonable steps to mitigate may
in some circumstances, including accepting the performance offered by
the defendant under a new contract even when performance amounts to a
breach of the original contract.

If the defendant's offer of performance remains the best substitute
performance (as it was in \emph{Payzu}) then it would seem unreasonable
not to go to that source.

Banco de Portugal v Waterlow \& Sons {[}1932{]} AC 452 establishes that
when considering whether the claimant has taken reasonable steps to
mitigate, the claimant's actions 'ought not to be weighed in nice scales
at the instance of the party whose breach of contract has occasioned the
difficulty'.

There is no duty to mitigate a claim for a payment of a debt. This
includes a claim for liquidated damages. This is because the amount is
payable as a contractual right rather than as damages.

\hypertarget{notification-of-claim}{%
\subsubsection{Notification of Claim}\label{notification-of-claim}}

A seller may attempt to (or may be able to) restrict its liability by
negotiating and demanding specific requirements that should be met
following a breach of warranty or indemnity claim. In Teoco UK Ltd v
Aircom Jersey 4 Ltd {[}2018{]} EWCA Civ 23,, claims for warranty
breaches were struck out for failing to comply with agreed notification
requirements.

\hypertarget{disclosure-letter}{%
\subsection{Disclosure Letter}\label{disclosure-letter}}

\hypertarget{purpose}{%
\subsubsection{Purpose}\label{purpose}}

The incentive for the seller in making disclosures is that they avoid it
being in breach of warranty in relation to those matters disclosed. The
disclosure letter will be subject to careful scrutiny. Advisable to make
full and early disclosure of potential issues.

\hypertarget{format}{%
\subsubsection{Format}\label{format}}

In the UK, the disclosures will usually be in the form of a letter from
the seller to the buyer. A bundle of documents (referred to as the
disclosure bundle) will be attached to the letter, comprising all the
documents referred to in the letter as providing information which
qualifies the warranties.

The parties' legal advisers will negotiate the disclosure letter or
schedule and bundle until it is in an agreed final form ready for
completion of the acquisition.

In the UK, on exchange of contracts, two copies of the agreed form of
disclosure letter and bundle will be available; one to be given to the
buyer and one to be retained by the seller. The letters and associated
bundles will be initialled by the parties as evidence that they are
identical.

\hypertarget{deemed-disclosures}{%
\subsubsection{Deemed Disclosures}\label{deemed-disclosures}}

\begin{env-3a4a363c-8d41-4783-bfde-af85d2b9ad5f}

General disclosures

The disclosure letter or schedule will state that a number of matters
are deemed to have been disclosed to the buyer. These deemed disclosures
are often referred to as `general disclosures'.

\end{env-3a4a363c-8d41-4783-bfde-af85d2b9ad5f}

These relate to publicly available information and occasionally
information made available to the buyer during due diligence. The effect
of deemed or general disclosure is to pass to the buyer the risk that
elements of the information will qualify the warranties. Usually
include:

\begin{enumerate}
\tightlist
\item
  Information held at Companies House
\item
  Information held at the Land Registry, Land Charges Department and any
  matters apparent in deeds of properties.
\item
  Matters which would be disclosed by physical inspection of each
  property (buyer should not accept unless a survey has been
  commissioned).
\item
  Matters in the public domain.
\item
  Matters disclosed or referred to in audited target company accounts.
\item
  Matters included or referred to in the accountants' report prepared on
  behalf of the buyer.
\item
  Matters disclosed or referred to in the replies to the preliminary
  enquiries and the documents enclosed within those replies.
\item
  Information and documentation passed to the buyer and its advisers in
  the pre-contract period.
\end{enumerate}

\hypertarget{specific-disclosure}{%
\subsubsection{Specific Disclosure}\label{specific-disclosure}}

Specific disclosures draw to the buyer's attention specific information
about the target which is inconsistent with one or more warranties given
by the seller. Disclosures usually refer to specific warranty
statements, but also include a catch-all that each disclosure is deemed
to be in respect of all the warranties.

\hypertarget{standard-of-disclosure}{%
\subsubsection{Standard of Disclosure}\label{standard-of-disclosure}}

It is in the interests of both parties for disclosures to be made
fairly.

\hypertarget{adequate-disclosure}{%
\paragraph{Adequate Disclosure}\label{adequate-disclosure}}

SPA will provide that the warranties are given subject to matters
disclosed in the disclosure letter. The buyer will often insist on a
warranty that all disclosures have been made ``fully, clearly and
accurately''.

In Levison v Farin {[}1978{]} 2 All ER 1149, the buyers were generally
aware of the rundown condition of the business, but this did not
preclude them from successfully claiming for breach of a warranty which
said `there will have been no material adverse change in the overall
value of the net assets of the Company'. Held that

\begin{quote}
`a clause in this form is primarily designed and intended to require a
party who wishes by disclosure to avoid a breach of warranty to give
specific notice for the purpose of the agreement, and protection by
disclosure will not normally be achieved by merely making known the
means of knowledge which may or do enable the other party to work out
certain facts and conclusions'.
\end{quote}

In Infiniteland Ltd v Artisan Contracting Ltd {[}2005{]} EWCA Civ 758 it
was held that the sufficiency of disclosure must be measured by
reference to the agreed contractual provisions and to the context of the
acquisition.

In MAN Nutzfahrzeuge AG v Freightliner Ltd {[}2005{]} EWHC 2347 the
judge considered (obiter) `the disclosure of all matters revealed by an
inspection of specified documents' would not extend so far as to include
inferences which might be drawn from the documents inspected.

\hypertarget{buyers-knowledge}{%
\paragraph{Buyer's Knowledge}\label{buyers-knowledge}}

The buyer will usually insist on limiting disclosure specifically to
matters revealed in the disclosure letter and bundle.

There are limits to the extent to which this will be successful. In
Eurocopy plc v Teesdale {[}1992{]} BCLC 1067 the SPA included an
agreement that the warranties were given subject to the disclosure
letter but no other information. The buyer tried to sue the seller for
failure to disclose certain material facts, which the seller claimed was
in the actual knowledge of the buyer following due diligence. The court
refused to strike out the defence on application by the buyer.

So it is always preferable for a buyer to acknowledge full disclosure of
a breach and seek an indemnity or reduction in the purchase price before
completion.

\hypertarget{actual-constructive-imputed-knowledge}{%
\paragraph{Actual/ constructive/ Imputed
Knowledge}\label{actual-constructive-imputed-knowledge}}

In Infiniteland Ltd v Artisan Contracting Ltd {[}2005{]} EWCA Civ 758,
the SPA contained a clause stating that the rights and remedies of the
purchaser in respect of any breach of warranty would not be affected by
any investigations made by it or on its behalf into the affairs, except
to the extent that such investigations gave it actual knowledge of the
relevant facts and circumstances.

\begin{itemize}
\tightlist
\item
  The Court confirmed that it would give effect to an express provision
  in the sale agreement prohibiting warranty claims of which the buyer
  had constructive knowledge.
\item
  On a 2:1 majority, it was held that `actual' and `imputed' knowledge
  are different. The knowledge of the professional adviser was not the
  buyer's knowledge, and the seller was therefore prevented from raising
  the buyer's knowledge as a defence to the warranty claim.
\end{itemize}

\begin{env-a57f3fff-c46d-4aeb-b36c-ada0d8066098}

Action

The parties need to stipulate in the agreement what type of knowledge
will or will not affect a buyer's rights and remedies for a claim for
breach of warranty.

\end{env-a57f3fff-c46d-4aeb-b36c-ada0d8066098}

\hypertarget{liability-for-information-from-third-parties}{%
\paragraph{Liability for Information From Third
Parties}\label{liability-for-information-from-third-parties}}

The impact of statements made by employees or agents of the seller or
the target company during negotiations or the due diligence process must
be considered. In MAN Nutzfahrzeuge AG v Freightliner Ltd {[}2007{]}
EWCA Civ 910, the seller was held to be vicariously liable for the
fraudulent pre-contractual statements of its agent, who had been put
forward as having authority to speak on its behalf.

\hypertarget{deliberate-non-disclosure}{%
\subsubsection{Deliberate
Non-disclosure}\label{deliberate-non-disclosure}}

Suppose the seller deliberately fails to disclose the existence of
particular facts to complete the sale. The buyer is induced to enter the
SPA by way of a deliberate concealment {\(\rightarrow\)} fraud.

In the context of a \textbf{share acquisition}, ss 89 and 90 of the
Financial Services Act 2012 make it a criminal offence for a person
knowingly or recklessly to make a statement, promise or forecast which
is misleading, false or deceptive. It is also an offence under the
section for a person with the requisite intent to conceal dishonestly
any material facts. A person guilty of an offence is liable to a fine,
or imprisonment for a maximum term of seven years for conviction on
indictment, or to both.

\hypertarget{insurance-1}{%
\subsection{Insurance}\label{insurance-1}}

Common to take out warranty and indemnity insurance (``W\&I policy'').
Premiums are typically between 0.80\% and 1.50\% of the insured limit of
liability in the London market.

The insurers will not agree cover until they have analysed the scope of
each warranty, the effect of the disclosures and the extent of the
general limitations on the seller's liability. Any policy is likely to
contain a number of general exclusions.

In addition to the set premium, it should be remembered that insurance
premium tax will be due -- in the UK, this represents an additional 12\%
tax on the value of the premium.

\hypertarget{completion}{%
\section{Completion}\label{completion}}

SPA is a contractual agreement setting out the acquisition terms. On
exchange, the parties undertake to perform the agreed steps to transfer
legal interest in the shares/ assets. Usually, exchange and completion
are simultaneous.

\begin{longtable}[]{@{}ll@{}}
\toprule()
Stage & Details \\
\midrule()
\endhead
Signing and exchange & The parties make a legal commitment to proceed
with the transaction by delivering their signed share purchase agreement
to the other party by way of exchange \\
Completion/ closing & The necessary formalities to conclude the
transaction are performed. For example, the seller delivers a duly
executed stock transfer form for the target shares and the buyer pays
the purchase price to the seller \\
\bottomrule()
\end{longtable}

It is usual for the SPA to set out the details of each party's
completion obligations.

\hypertarget{reasons-for-conditionality}{%
\subsection{Reasons for
Conditionality}\label{reasons-for-conditionality}}

Approvals, consents and clearances should ideally be obtained before
signing of the SPA, but can take time. SPA may be made conditional on:

\begin{enumerate}
\tightlist
\item
  HMRC clearance (e.g., on roll-over relief from capital tax on a
  securities exchange)
\item
  CMA notification that buyer will not be subject to a Phase 2
  investigation
\item
  Regulatory licence for the industry (e.g., environmental licence)
\item
  a substantial customer or landlord providing its consent to an
  assignment/ confirming it will not exercise a contractual right to
  terminate.
\item
  Corporate buyer shareholders passing a resolution for the issue of
  shares in consideration for the target.
\item
  Clearance under National Security and Investment Act 2021.
\end{enumerate}

Parties may not want to make the required applications (and make the
deal public) until after the contract is signed.

\hypertarget{risk-allocation-on-conditional-contracts}{%
\subsection{Risk Allocation on Conditional
Contracts}\label{risk-allocation-on-conditional-contracts}}

If completion is to be made conditional, the agreement must state:

\begin{itemize}
\tightlist
\item
  Each party's obligations
\item
  Date by which each condition must be met
\item
  Any right to terminate prior to completion.
\end{itemize}

The buyer will seek a contractual right to withdraw from the purchase if
a major problem occurs prior to completion.

\hypertarget{conditions-precedent-1}{%
\subsubsection{Conditions Precedent}\label{conditions-precedent-1}}

Typical CPs include:

\begin{enumerate}
\tightlist
\item
  An obligation on one party to take all reasonable steps to try to
  procure the satisfaction of the condition as early as possible

  \begin{itemize}
  \tightlist
  \item
    Implied in certain jurisdictions
  \end{itemize}
\item
  A long-stop date by which conditions must be satisfied. If not
  satisfied by this date, provision for the agreement to terminate
  automatically without liability attaching to either party.
\item
  If conditions are satisfied prior to the long-stop, provision for
  completion within a specific period.
\end{enumerate}

In France \& Germany, a condition precedent may be unenforceable if its
fulfilment depends solely on one party.

\hypertarget{management-restrictions}{%
\subsubsection{Management Restrictions}\label{management-restrictions}}

After exchange of a conditional SPA, the buyer is contractually
committed to proceeding with the acquisition. The buyer runs the risk
that the seller may neglect the business, knowing that the buyer must go
ahead with the purchase. The buyer will therefore require undertakings
from the seller about how the target company or business will be run in
the period between signing and completion.

The buyer runs the risk that the seller may neglect the business,
knowing that the buyer must go ahead with the purchase. May:

\begin{itemize}
\tightlist
\item
  Require undertakings from the seller about how the target company or
  business will be run in the period.
\item
  Require seller to allow it to attend board meetings
\item
  Require undertakings that the seller will not:

  \begin{itemize}
  \tightlist
  \item
    Lend or borrow money except for routine matters in the ordinary
    course of business
  \item
    Grant any mortgage/ charge/ debenture over assets
  \item
    Settle any claim/ dispute
  \item
    Acquire or agree to acquire any property
  \item
    Commit to capital expenditure
  \item
    Enter into hire purchase agreements
  \item
    Alter employment terms of employees/ directors
  \item
    Appoint any additional directors of the target.
  \end{itemize}
\end{itemize}

\hypertarget{repetition-of-warranties}{%
\subsubsection{Repetition of
Warranties}\label{repetition-of-warranties}}

\begin{env-48e9bfc8-c3dd-4504-a5fa-3b34b9a27ccf}

Note

A buyer often seeks to protect its position by requiring that the
representations and warranties made in the sale and purchase agreement
are repeated as at the date of completion.

\end{env-48e9bfc8-c3dd-4504-a5fa-3b34b9a27ccf}

A strong buyer may include this as an additional CP (giving the buyer
the right to withdraw from the purchase in the event of breach). Should
be strongly resisted by the seller/ limited to material breaches.

Consideration should also be given to the effect of any disclosure
against such repeated warranties. The buyer is at risk in relation to
disclosures made between exchange of contracts and completion because,
regardless of the breach disclosed, it is still bound to complete the
purchase.

\hypertarget{mac-clause}{%
\subsubsection{MAC Clause}\label{mac-clause}}

In addition to the repetition of warranties, a buyer in a strong
negotiating position may also seek to include a right to terminate the
agreement in the event of a material adverse change in the business,
assets or profits of the target.

Intended to protect the buyer against the general risk of adverse
events. Should be strongly resisted by the seller (general risk should
pass to the buyer on signing)/ limited to specified events.

Very standard in US transactions; unusual here. Generally interpreted
narrowly in a way favouring the seller.

\hypertarget{completing-acquisition}{%
\subsection{Completing Acquisition}\label{completing-acquisition}}

Much of the ``closing'' process is managed using an agreed checklist, or
completion agenda. This sets out the steps and documentation required,
and who is responsible for them. Regularly circulated and updated.
Detail depends on the transaction.

If there is to be a delay between signing the sale and purchase
agreement and completing the acquisition, the completion agenda will be
an agreed document referred to in the sale and purchase agreement or set
out in a schedule to it.

\hypertarget{preparation}{%
\subsubsection{Preparation}\label{preparation}}

Check who should attend and that they have been correctly authorised to
execute

\begin{env-48e9bfc8-c3dd-4504-a5fa-3b34b9a27ccf}

Power of attorney

If an individual is unable to attend the completion meeting in person,
they may execute a power of attorney giving another person the power to
sign the completion documents on their behalf. The terms of the power of
attorney will usually be agreed between the parties, and will often be
drafted in very broad terms to include a power to sign any additional
documentation that may be required in order to effect completion.

\end{env-48e9bfc8-c3dd-4504-a5fa-3b34b9a27ccf}

\hypertarget{seller-preparation}{%
\paragraph{Seller Preparation}\label{seller-preparation}}

\begin{itemize}
\tightlist
\item
  Board meeting of the seller company to approve the terms of the draft
  SPA and to appoint an authorised signatory to execute it.

  \begin{itemize}
  \tightlist
  \item
    A single director/ committee may be appointed to attend on behalf of
    the company.
  \end{itemize}
\item
  A certified copy of the minutes of this meeting must be given to the
  buyer on completion as evidence of the seller's representatives'
  authority to act (and sign documents) on its behalf.

  \begin{itemize}
  \tightlist
  \item
    Buyer can rely on this as evidence (as a 3rd party acting in good
    faith)
  \item
    Parties' lawyers will check articles of association \& securities
    documentation regardless.
  \end{itemize}
\item
  Lawyers check whether the sale is in accordance with directors'
  duties.
\item
  If sale is to a director or connected person, an ordinary resolution
  may be needed to approval of substantial property transaction (s 190)
\item
  On a share sale, the buyer will require that a board meeting of the
  target company is held at completion to approve registration of the
  transfer of shares.

  \begin{itemize}
  \tightlist
  \item
    Seller should ensure BM is quorate and there is a sufficient
    majority.
  \end{itemize}
\end{itemize}

\hypertarget{buyer-preparation}{%
\paragraph{Buyer Preparation}\label{buyer-preparation}}

\begin{itemize}
\tightlist
\item
  A corporate buyer must hold a meeting to approve the terms of the
  draft SPA and authorise its signing and completion.

  \begin{itemize}
  \tightlist
  \item
    A single director/ committee may be appointed to attend on behalf of
    the company.
  \end{itemize}
\item
  A certified copy of the minutes of this meeting must be given to the
  buyer on completion as evidence of the buyer's representatives'
  authority to act (and sign documents) on its behalf.
\end{itemize}

\begin{longtable}[]{@{}ll@{}}
\toprule()
Funding of acquisition & Preparations \\
\midrule()
\endhead
Equity & Shareholder resolutions may be required to authorise their
issue. \\
Debt & Buyer's lawyers must confirm that requirements necessary for the
provision of the finance have been complied with {\(\rightarrow\)} funds
will be available at completion. \\
\bottomrule()
\end{longtable}

\begin{itemize}
\tightlist
\item
  Repeat of some searches as a final check

  \begin{itemize}
  \tightlist
  \item
    Companies Registry search
  \item
    Bankruptcy search against individual sellers
  \item
    Trademark/ patent searches if these are important to the
    acquisition.
  \end{itemize}
\item
  Share acquisition: searches made against the target company itself.
\item
  On a share purchase, the buyer may also want to inspect the statutory
  books of the target company to check compliance with formalities.
\end{itemize}

\hypertarget{completion-of-asset-acquisition}{%
\subsubsection{Completion of Asset
Acquisition}\label{completion-of-asset-acquisition}}

UK asset purchase: all assets must be transferred individually. SPA
usually provides that transfers take place simultaneously.

\hypertarget{transfer-of-title}{%
\paragraph{Transfer of Title}\label{transfer-of-title}}

Formalities to transfer the legal interest in the assets of a business
to the buyer:

\begin{itemize}
\tightlist
\item
  conveyances, transfers or assignments of land and premises;
\item
  assignments of goodwill, certain intellectual property rights
  (including copyrights, patents and trade marks) and the benefit of
  contracts.
\end{itemize}

Stock and moveable assets are transferable on delivery.

\hypertarget{documents-to-hand-over}{%
\paragraph{Documents to Hand Over}\label{documents-to-hand-over}}

The seller will hand over:

\begin{itemize}
\tightlist
\item
  Where there is no delay between exchange and completion, the agreed
  form of disclosure letter or schedule
\item
  where the seller is a company, a certified copy of a board resolution
  authorising a representative to attend to completion arrangements and
  to sign the documentation.
\item
  Documents transferring title to assets
\item
  Deeds and documents of title to the assets
\item
  Duly executed releases of charges over the assets and confirmation of
  non-crystallisation of floating charges
\item
  Financial records, customer records, promo materials, PAYE and VAT
  records, employee records\ldots{}

  \begin{itemize}
  \tightlist
  \item
    Seller should retain a right to inspect these for a while after
    completion (in case needed)
  \end{itemize}
\item
  Originals, counterparts or certified copies as appropriate of licences
  to assign leasehold property, novation agreements and consents from
  third parties to assignment of contracts
\item
  A certified copy of a special resolution of the selling company
  resolving to change its name if it has been agreed that the buyer will
  acquire the right to use the seller's corporate name.
\end{itemize}

\hypertarget{completion-of-share-acquisition}{%
\subsubsection{Completion of Share
Acquisition}\label{completion-of-share-acquisition}}

Parties agree a completion timetable. Only the shares are transferred.

\begin{env-48e9bfc8-c3dd-4504-a5fa-3b34b9a27ccf}

Transferring a legal interest in shares

Under English law, the following documents are required to transfer the
legal interest in the shares:\\
( a) stock transfer form; and\\
( b) share certificate.

\end{env-48e9bfc8-c3dd-4504-a5fa-3b34b9a27ccf}

The buyer is not entitled to exercise its shareholders' rights until the
transfer is registered by the target. The buyer will therefore require
the seller to procure registration of the transferred shares, usually
(in a private company) by procuring a board meeting of the target
company to consider the transferred shares for registration. The
existing shareholders may also be required to waive any rights of
pre-emption on the transfer of shares that may exist under the
constitution of the company.

\hypertarget{completion-board-meeting}{%
\paragraph{Completion Board Meeting}\label{completion-board-meeting}}

Board meeting to implement any decisions the buyer requires to take
immediate control of the target. The minutes of this meeting must be
prepared in advance and circulated for approval.

The meeting usually deals with:

\begin{itemize}
\tightlist
\item
  Approval of the transfer of shares (subject to stamp duty)
\item
  Appointment of new directors and company secretary (consents to act,
  disclosure of interests, service contracts)
\item
  Acceptance of resignations of directors and company secretary.
\item
  Acceptance of resignation of company's auditors and appointment of new
  auditors.
\item
  Change in accounting reference date
\item
  Change in registered office
\item
  Alteration of existing bank mandates and completion of new finance
  arrangements.
\end{itemize}

\begin{env-48e9bfc8-c3dd-4504-a5fa-3b34b9a27ccf}

Note

A seller of shares is under a general duty not to obstruct or delay the
registration of the buyer as the holder of the shares
(\emph{\href{https://uk.westlaw.com/D-106-5776?originationContext=document\&transitionType=PLDocumentLink\&contextData=(sc.Default)\&ppcid=499185a9ecb940b1a3a505768c861e1b}{Hooper
v Herts {[}1906{]} 1 Ch 549}}). However, unless the parties expressly
agree otherwise, the seller is not obliged to procure that the target
company actually approves and registers the transfer
(\emph{\href{https://uk.westlaw.com/D-106-5778?originationContext=document\&transitionType=PLDocumentLink\&contextData=(sc.Default)\&ppcid=499185a9ecb940b1a3a505768c861e1b}{London
Founders Association Ltd v Clarke {[}1888{]} 20 QBD 576}}). As a result,
the buyer typically includes a provision in the SPA which requires the
seller to procure that the target company holds a board meeting at
completion for the purpose of approving the transfer of the target
shares to the buyer, subject only to the transfer being stamped.

\end{env-48e9bfc8-c3dd-4504-a5fa-3b34b9a27ccf}

Subsidiaries may also need board meetings to implements changes.

\hypertarget{completion-gm}{%
\paragraph{Completion GM}\label{completion-gm}}

Matters that may require shareholder approval:

\begin{itemize}
\tightlist
\item
  Consent to SPT (s 190)
\item
  Consent to directors' service contract {\(> 2\)} years
\item
  Consent to alter company articles.
\end{itemize}

Changes/ approvals to be effected by the shareholders will usually be
dealt with by written resolution at the completion board meeting of the
target company.

\begin{env-48e9bfc8-c3dd-4504-a5fa-3b34b9a27ccf}

Note

The buyer does not become a member of the target company until its name
is entered into the register of members (CA 2006, s 112) -- after
stamping of the stock transfer form. So the sellers must usually pass
the resolution.

\end{env-48e9bfc8-c3dd-4504-a5fa-3b34b9a27ccf}

\hypertarget{completion-documents}{%
\paragraph{Completion Documents}\label{completion-documents}}

The parties will agree a completion agenda. Usually, the buyer takes
responsibility for producing the final form SPA. Seller required to hand
over:

\begin{itemize}
\tightlist
\item
  where there is no delay between exchange and completion, the
  \textbf{agreed form of disclosure letter} or schedule;
\item
  if the seller is a company, a certified copy of a \textbf{board
  resolution authorising its representative} to attend to completion
  arrangements on behalf of the company and to sign the documentation;
\item
  \textbf{stock transfer form} duly executed and delivered by each
  seller together with relevant share certificates;

  \begin{itemize}
  \tightlist
  \item
    A proper instrument of transfer is required to transfer legal title
    to the target shares to the buyer (s 770 CA 2006).
  \item
    For fully paid shares, this is almost always the form prescribed by
    Sch 1 Stock Transfer Act 1963.
  \end{itemize}
\item
  minutes of a \textbf{completion board meeting} of the target to
  implement any specified board decisions if required;
\item
  any \textbf{written resolutions} required, duly executed by the
  sellers
\item
  \textbf{discharge of outstanding loans} made by the seller to the
  company/ vice versa

  \begin{itemize}
  \tightlist
  \item
    Deed of release of the charge
  \item
    Letter of non-crystallisation of a floatng charge
  \item
    Written consent of the charge holder to the sale of shares.
  \end{itemize}
\item
  statutory books of the target company (if required), title deeds to
  its properties, any certificates of title, documents of title for
  other assets, financial records, contracts, insurance policies etc.
\item
  \textbf{Officers' and auditor's letters of resignation} if required by
  buyer
\item
  \textbf{Deed of indemnity for lost share certificates} if any lost/
  destroyed.
\item
  \textbf{Power of attorney pending registration of the transfer} --
  allows the buyer to control the rights attached to the target shares
  pending registration of the transfer.
\item
  \textbf{Waiver of seller's claim}s -- waiver of any subsisting claims
  against the target, if required by the buyer.
\item
  \textbf{Loan note documents} if a loan note is used to fund the
  transaction
\item
  \textbf{Transitory services agreement} -- if the buyer requires the
  seller to continue to provide certain support/ services for a limited
  period following completion.
\item
  \textbf{New services agreement} for employees, if required by the
  buyer.
\end{itemize}

Buyer pays the purchase price on completion.

\hypertarget{virtual-completion}{%
\subsubsection{Virtual Completion}\label{virtual-completion}}

If completion is not face to face, take care to ensure legal validity.
Ensure that final versions of all documentation are sent to signatories
before completion occurs, so the signatories can subsequently give
authority for signed pages to be attached to a hard copy.

Historically, the non-availability of a signatory for an exchange or
completion meeting might have been addressed by arranging for the absent
signatory to sign separate signature pages in advance, and then
attaching the pre-signed pages to the relevant documents once they were
finalised. However, this practice has been curtailed following the High
Court decision in R (Mercury Tax Group Ltd and another) v HMRC
{[}2008{]} EWHC 2721 (Admin).

\begin{env-a57f3fff-c46d-4aeb-b36c-ada0d8066098}

Recommended options for virtual completion: JWP note

\begin{itemize}
\item
  \textbf{Option 1.} The signatory prints and signs the relevant
  signature page of the final version of the document. They then return
  a single email to their lawyers attaching both the entire final
  version of the document and a scanned copy of the signed signature
  page. This approach is suitable for all types of documents, including
  deeds.
\item
  \textbf{Option 2.} The signatory prints and signs the relevant
  signature page of the final version of the document. They then return
  to their lawyers a scanned copy of the signed signature page only and
  authorise them to attach the signed page to the final version of the
  document. This approach is \textbf{not} suitable for deeds or real
  estate contracts.
\item
  \textbf{Option 3.} The signatory pre-signs a signature page for the
  relevant document. They then returns this page to their lawyers to
  attach to the document once the final version is agreed. This approach
  is \textbf{not} suitable for deeds or real estate contracts.
\end{itemize}

\end{env-a57f3fff-c46d-4aeb-b36c-ada0d8066098}

When a party signs a document and sends this to their lawyer in advance
of exchange or completion, they should give or confirm in writing the
authority to deliver the relevant signed document, clearly stating any
conditions.

\hypertarget{authorising-execution}{%
\subsubsection{Authorising Execution}\label{authorising-execution}}

Where all required signatories cannot be physically present at exchange
or completion, rather than signing the transaction documents in advance,
the absent signatories could authorise another person to execute
documents on their behalf.

2 options:

\begin{enumerate}
\tightlist
\item
  Authorising the third party to sign as the party's agent.
\item
  Granting an ordinary power of attorney. This approach tends to be more
  common in practice and will be the only viable option if the
  transaction documents include a deed, given that authority to execute
  a deed must be given by deed (Steiglitz v Eggington {[}1815{]} Holt
  141).
\end{enumerate}

\begin{env-f3fdeea8-9a69-4242-ac7f-874202f0281b}

Warning

It is not usually possible for an absent director to appoint an attorney
to sign a document on their behalf in their capacity as a director of a
company.

\end{env-f3fdeea8-9a69-4242-ac7f-874202f0281b}

\hypertarget{payment-arrangements}{%
\subsubsection{Payment Arrangements}\label{payment-arrangements}}

The seller will expect the buyer to pay any cash element of the
transaction in cleared funds on the completion date. Check:

\begin{itemize}
\tightlist
\item
  The buyer has the requisite funds
\item
  The parties have agreed a completion funds flow dealing with
  logistics.
\item
  Agree timings in advance, or they can be a pain
\item
  Consider having the buyer's lawyer undertake that funds will be
  transferred on completion.
\item
  Consider setting up a special purpose escrow account in the parties'
  joint names and appointing a holding bank/ independent third agent as
  the escrow agent.
\end{itemize}

\hypertarget{using-sale-proceeds-to-discharge-security-interests}{%
\paragraph{Using Sale Proceeds to Discharge Security
Interests}\label{using-sale-proceeds-to-discharge-security-interests}}

Where a seller needs the sale proceeds to discharge a security interest
at completion, this can cause cash flow issues.

\begin{env-a57f3fff-c46d-4aeb-b36c-ada0d8066098}

Action

The lender's lawyers giving an undertaking to the buyer's lawyers to
forward the lender's deed of release to them once the lender has
received the monies following completion of the share sale may resolve
this timing issue.

\end{env-a57f3fff-c46d-4aeb-b36c-ada0d8066098}

\hypertarget{funding-acquisition}{%
\subsubsection{Funding Acquisition}\label{funding-acquisition}}

If a buyer is seeking to raise the requisite funds for the acquisition
through an issue of equity or by obtaining debt finance, this can cause
timing issues for completion:

\begin{env-48e9bfc8-c3dd-4504-a5fa-3b34b9a27ccf}

Catch-22

\begin{itemize}
\tightlist
\item
  The buyer may not want to raise funds unless acquisition is certain to
  complete.
\item
  Buyer unable to complete the transaction if it does not raise the
  necessary funds {\(\rightarrow\)} wants to avoid being subject to an
  obligation to complete in these circumstances.
\end{itemize}

\end{env-48e9bfc8-c3dd-4504-a5fa-3b34b9a27ccf}

\begin{itemize}
\tightlist
\item
  Acquisition funded by a placing

  \begin{itemize}
  \tightlist
  \item
    SPA often provides that completion is conditional on the allotment
    and listing of placing shares
  \item
    Placing agreement says placing conditional on SPA becoming
    unconditional in all respects other than placing conditions
  \end{itemize}
\item
  Acquisition funded by rights issue

  \begin{itemize}
  \tightlist
  \item
    Buyer risks raising too much cash if the acquisition is subject to a
    condition that cannot be satisfied before the rights issue is
    launched.
  \item
    Agree that if the acquisition does not complete, proceeds of the
    rights issue can be used for other purposes
  \item
    Or pay in 2 instalments("trombone rights issue")
  \end{itemize}
\end{itemize}

\hypertarget{post-completion}{%
\subsection{Post-completion}\label{post-completion}}

\hypertarget{stamp-duty-and-filing}{%
\subsubsection{Stamp Duty and Filing}\label{stamp-duty-and-filing}}

\begin{env-8890de02-1ba1-42c2-9b89-cb105d97c320}

Important

On a share purchase of a company registered in England and Wales, the
buyer's lawyers must ensure that the duly executed stock transfer forms
are delivered to the HMRC Stamp Office within \textbf{30 days} together
with the requisite transfer duty.

\end{env-8890de02-1ba1-42c2-9b89-cb105d97c320}

Usually done immediately by the buyer's lawyers. Stamp duty is payable
on the transfer of certificated shares unless the consideration for the
transfer is 1,000 or less, or another exemption applies.

The buyer's lawyers may also need to file the following at Companies
Registry:

\begin{itemize}
\tightlist
\item
  Forms notifying any change of registered office (s 87 CA 2006),
  directors and secretary, accounting reference period:

  \begin{itemize}
  \tightlist
  \item
    Change of registered office address
    (\emph{\href{https://uk.westlaw.com/7-422-1650?originationContext=document\&transitionType=DocumentItem\&contextData=(sc.Default)\&ppcid=499185a9ecb940b1a3a505768c861e1b}{Form
    AD01}}).
  \item
    Appointment of new director(s)
    (\emph{\href{https://uk.westlaw.com/9-422-1649?originationContext=document\&transitionType=DocumentItem\&contextData=(sc.Default)\&ppcid=499185a9ecb940b1a3a505768c861e1b}{Forms
    AP01}} and
    \emph{\href{https://uk.westlaw.com/2-422-2044?originationContext=document\&transitionType=DocumentItem\&contextData=(sc.Default)\&ppcid=499185a9ecb940b1a3a505768c861e1b}{AP02}}).
  \item
    Appointment of new company secretary
    (\emph{\href{https://uk.westlaw.com/1-422-1648?originationContext=document\&transitionType=DocumentItem\&contextData=(sc.Default)\&ppcid=499185a9ecb940b1a3a505768c861e1b}{Forms
    AP03}} and
    \emph{\href{https://uk.westlaw.com/4-422-2043?originationContext=document\&transitionType=DocumentItem\&contextData=(sc.Default)\&ppcid=499185a9ecb940b1a3a505768c861e1b}{AP04}}).
  \item
    Resignation of director(s)
    (\emph{\href{https://uk.westlaw.com/3-422-1647?originationContext=document\&transitionType=DocumentItem\&contextData=(sc.Default)\&ppcid=499185a9ecb940b1a3a505768c861e1b}{Form
    TM01}}).
  \item
    Resignation of company secretary
    (\emph{\href{https://uk.westlaw.com/6-422-2042?originationContext=document\&transitionType=DocumentItem\&contextData=(sc.Default)\&ppcid=499185a9ecb940b1a3a505768c861e1b}{Form
    TM02}}).
  \item
    Change of accounting reference date
    (\emph{\href{https://uk.westlaw.com/0-422-1291?originationContext=document\&transitionType=DocumentItem\&contextData=(sc.Default)\&ppcid=499185a9ecb940b1a3a505768c861e1b}{Form
    AA01}}).
  \end{itemize}
\item
  Copy of any special resolution
\item
  Copy of new articles of association.
\end{itemize}

Where shares are issued as consideration, the buyer's lawyers may also
need to file:

\begin{itemize}
\tightlist
\item
  Ordinary resolution triggering the directors' authority to allot
  shares under s 550
\item
  Ordinary resolution removing ceiling on authorised share capital (CA
  1985 company)
\item
  Ordinary resolution authorising directors to allot shares (s 551) -- a
  company other than a private company with only one class of shares in
  issue
\item
  Special resolution suspending any pre-emption rights
\item
  Return of allotments (s 555) together with a statement of capital.
\end{itemize}

Internal administration:

\begin{itemize}
\tightlist
\item
  Update register of members (section 128B, CA 2006) after stock
  transfer forms signed
\item
  Register of people with significant control (section 790X, CA 2006).
\item
  Register of directors (section 167A, CA 2006).
\item
  Register of directors' residential addresses (section 167A, CA 2006).
\item
  Register of secretaries (section 279A, CA 2006).
\item
  PSC register
\item
  Maintain internal record of board minutes and resolutions.
\end{itemize}

\hypertarget{outstanding-matters}{%
\subsubsection{Outstanding Matters}\label{outstanding-matters}}

There are often matters that may have been left outstanding on
completion of a transaction. Agree a timetable for these and consider
having indemnities in place in relation to any costs incurred. Take
practical steps to implement any changes required to incorporate the
target into the existing business.

\hypertarget{bibles}{%
\paragraph{Bibles}\label{bibles}}

Often compiled after completion, containing a complete set of documents
used in the acquisition.

\hypertarget{execution-of-deeds-and-documents}{%
\subsection{Execution of Deeds and
Documents}\label{execution-of-deeds-and-documents}}

English law makes a distinction between deeds and simple contracts.
Execution formalities vary according to the legal personality of the
executing party.

\begin{itemize}
\tightlist
\item
  The default position in English law favours flexibility.
\item
  Formalities are imposed where a transaction is considered sufficiently
  important.
\item
  A signatory to a contract is generally bound by its terms (in the
  absence of fraud or misrepresentation) irrespective of whether the
  party has in fact read the terms
  (\emph{\href{https://uk.practicallaw.thomsonreuters.com/D-016-1583?originationContext=document\&transitionType=PLDocumentLink\&contextData=(sc.Default)\&ppcid=0e69ecc7bd0b425c8c07be4517611bac}{L'Estrange
  v Graucob {[}1934{]} 2 KB 394}}).
\item
  English law mandates the use of a deed in certain transactions or
  circumstances. There is no requirement for the signature to be
  witnessed.
\item
  A deed is a written instrument which is executed with the necessary
  formality and by which an interest/ right or property passes or is
  confirmed.
\end{itemize}

Purposes of deed formalities:

\begin{itemize}
\tightlist
\item
  Evidential
\item
  Cautionary
\item
  Labelling
\end{itemize}

\hypertarget{effect-of-a-deed}{%
\subsubsection{Effect of a Deed}\label{effect-of-a-deed}}

\begin{itemize}
\tightlist
\item
  No requirement for consideration.

  \begin{itemize}
  \tightlist
  \item
    But note the courts will not normally order specific performance of
    a voluntary undertaking even if it made by deed (equity does not
    assist a volunteer).
  \end{itemize}
\item
  Longer limitation period

  \begin{itemize}
  \tightlist
  \item
    Generally 12 years (s 8 Limitation Act 1980), except for some rent/
    mortgage exceptions.
  \item
    For a contract, limitation period is 6 years from the date on which
    the cause of action accrued (s 5 Limitation Act 1980)
  \end{itemize}
\item
  Estoppel by deed

  \begin{itemize}
  \tightlist
  \item
    A party is estopped from disputing the validity or effectiveness of
    the grant.
  \end{itemize}
\end{itemize}

\hypertarget{written-contract}{%
\subsubsection{Written Contract}\label{written-contract}}

In a limited number of situations, the law requires an agreement to be
made or evidenced in writing and, in some instances, to be signed by the
parties in order to have legal effect.

\begin{itemize}
\tightlist
\item
  Contract for the sale of land

  \begin{itemize}
  \tightlist
  \item
    s 2 LP(MP)A applies; must be in writing, signed by or on behalf of
    each party, and incorporate all terms expressly agreed by the
    parties.
  \end{itemize}
\item
  The assignment of the benefit of a contract

  \begin{itemize}
  \tightlist
  \item
    An assignment that meets the requirements set out in
    \emph{\href{https://uk.practicallaw.thomsonreuters.com/2-505-5703?originationContext=document\&transitionType=PLDocumentLink\&contextData=(sc.Default)\&ppcid=0e69ecc7bd0b425c8c07be4517611bac}{section
    136}} of the LPA 1925 is a legal assignment
  \item
    These include that the assignment must be in writing and signed by
    the assignor.
  \end{itemize}
\item
  Transfer of legal title to certificated shares

  \begin{itemize}
  \tightlist
  \item
    s 770 CA 2006 requires a written transfer signed by or on behalf of
    the transferor to be lodged with the company before the new
    shareholder is registered as the owner of the legal title to the
    shares.
  \end{itemize}
\item
  Assignment of a number of IP rights
\item
  Equitable mortgage
\item
  Assent (legal interest in land)
\item
  Guarantee

  \begin{itemize}
  \tightlist
  \item
    Guarantee is unenforceable unless (s 4 Statute of Frauds) -

    \begin{itemize}
    \tightlist
    \item
      It is in writing (or there is a memorandum or note of the
      agreement).
    \item
      It is signed by the guarantor or at its direction.
    \end{itemize}
  \item
    Usually structured as deeds to avoid the problem of consideration.
  \end{itemize}
\item
  Declaration of trust over land (or any interest in it)
\item
  Financial collateral arrangement
\item
  Regulated credit agreement (Consumer Credit Act 1974)
\item
  Agreement to waive/ settle a claim against the employer in return for
  payment.
\end{itemize}

\hypertarget{deed}{%
\subsubsection{Deed}\label{deed}}

Statute demands the use of a deed for the following transactions:

\begin{itemize}
\tightlist
\item
  Transfers of land or interests in land.

  \begin{itemize}
  \tightlist
  \item
    Must be by deed or it will be void for the purpose of conveying/
    creating a legal estate (s 52 LPA 1925).
  \end{itemize}
\item
  Leases, except

  \begin{itemize}
  \tightlist
  \item
    Those taking effect in possession for a term not exceeding 3 years
    at best rent reasonably obtainable without taking a fine.
  \item
    Relevant social housing tenancies.
  \end{itemize}
\item
  Mortgages and charges.

  \begin{itemize}
  \tightlist
  \item
    A legal mortgage or charge by way of legal mortgage over land can
    only be created by deed (sections 52(2), 85(1) and 86(1), LPA 1925).
  \item
    Any mortgage or charge of land or other property (whether legal or
    equitable) must be by deed if the mortgagee or chargee is to have
    the statutory powers of sale and to appoint a receiver (section 101,
    LPA 1925).
  \end{itemize}
\item
  Sales by mortgagee.

  \begin{itemize}
  \tightlist
  \item
    Must be by deed if the sale is to overreach subsequent mortgages and
    charges (section 104, LPA 1925).
  \end{itemize}
\item
  Certain actions under the Trustee Act 1925 (TA 1925).

  \begin{itemize}
  \tightlist
  \item
    The power to replace a trustee under section 36 of the TA 1925 must
    be exercised in writing and is typically exercised by deed, because
    a deed allows automatic vesting of trust property under section
    40(1).
  \end{itemize}
\item
  Powers of attorney, including lasting powers of attorney.

  \begin{itemize}
  \tightlist
  \item
    An ordinary power of attorney must be executed as a deed by the
    donor of the power (section 1(1), Powers of Attorney Act 1971).
  \end{itemize}
\item
  Bills of sale.
\end{itemize}

Common law requires a deed for:

\begin{itemize}
\tightlist
\item
  Any agreement that is made without consideration.
\item
  A gift of tangible goods, without delivery of possession

  \begin{itemize}
  \tightlist
  \item
    (\emph{\href{https://uk.practicallaw.thomsonreuters.com/D-016-1484?originationContext=document\&transitionType=PLDocumentLink\&contextData=(sc.Default)\&ppcid=0e69ecc7bd0b425c8c07be4517611bac}{Irons
    v Smallpiece (1819) 2 B\&E Ald 551}} and
    \emph{\href{https://uk.practicallaw.thomsonreuters.com/D-016-1485?originationContext=document\&transitionType=PLDocumentLink\&contextData=(sc.Default)\&ppcid=0e69ecc7bd0b425c8c07be4517611bac}{Cochrane
    v Moore (1890) 25 QBD 57}}).
  \end{itemize}
\item
  An express release of certain rights..
\end{itemize}

\hypertarget{principles}{%
\subsubsection{Principles}\label{principles}}

\begin{env-8890de02-1ba1-42c2-9b89-cb105d97c320}

Important

\begin{itemize}
\tightlist
\item
  A contract contained in a deed can only be varied or discharged by
  another contract contained in a deed.
\item
  The release of a debt, liability or obligation must be effected by
  deed.
\end{itemize}

\end{env-8890de02-1ba1-42c2-9b89-cb105d97c320}

However, a variation or release can take effect in equity in the absence
of a deed provided that valuable consideration has been given
(\emph{\href{https://uk.practicallaw.thomsonreuters.com/D-009-7407?originationContext=document\&transitionType=PLDocumentLink\&contextData=(sc.Default)\&ppcid=0e69ecc7bd0b425c8c07be4517611bac}{Webb
v Hewitt (1857) 3 K and J 438}}).

The doctrine of promissory estoppel is also relevant in this context and
oral evidence is also permissible to show a variation of the terms of a
deed.

\hypertarget{formalities}{%
\subsubsection{Formalities}\label{formalities}}

There are four key requirements for a valid deed:

\begin{itemize}
\tightlist
\item
  \textbf{Writing.} A deed must be in writing. This is a common law
  requirement modified by section 1(1)(a) of the LP(MP)A 1989 which
  abolished any rule of law restricting the substances on which a deed
  may be written.
\item
  \textbf{Face value requirement.} It must be clear from the face of the
  instrument that it is intended to take effect as a deed
  (\emph{\href{https://uk.practicallaw.thomsonreuters.com/7-505-7361?originationContext=document\&transitionType=PLDocumentLink\&contextData=(sc.Default)\&ppcid=0e69ecc7bd0b425c8c07be4517611bac}{section
  1(2)(a)}, LP(MP)A 1989}). For further information, see
  \emph{\href{https://uk.practicallaw.thomsonreuters.com/Document/I2a05f3d71cb811e38578f7ccc38dcbee/View/FullText.html?navigationPath=Search\%2Fv1\%2Fresults\%2Fnavigation\%2Fi0ad6ad3f000001841ed7d74a8e6914e1\%3Fppcid\%3D6c2ae8a868b54456ac4cdc97bf99f24b\%26Nav\%3DKNOWHOW_UK\%26fragmentIdentifier\%3DI2a05f3d71cb811e38578f7ccc38dcbee\%26parentRank\%3D0\%26startIndex\%3D1\%26contextData\%3D\%2528sc.Search\%2529\%26transitionType\%3DSearchItem\&listSource=Search\&listPageSource=c156b6df63e9ac1dbe2bd9174c514051\&list=KNOWHOW_UK\&rank=2\&sessionScopeId=d84126a1818ed3fb6e1aadaf57885dfabcc9a95b48d8801a1ae90e19604d0dc0\&ppcid=6c2ae8a868b54456ac4cdc97bf99f24b\&originationContext=Search\%20Result\&transitionType=SearchItem\&contextData=(sc.Search)\&navId=DADDF17324157881A30EF74E4476B30F\&comp=pluk\#co_anchor_a243821}{The
  face value requirement}}.
\item
  \textbf{Execution.} The instrument must be validly executed as a deed
  by the person making it or, as the case may be, by one or more of the
  parties to it
  (\emph{\href{https://uk.practicallaw.thomsonreuters.com/7-505-7361?originationContext=document\&transitionType=PLDocumentLink\&contextData=(sc.Default)\&ppcid=0e69ecc7bd0b425c8c07be4517611bac}{section
  1(2)(b)}, LP(MP)A 1989}). For further information, see
  \emph{\href{https://uk.practicallaw.thomsonreuters.com/Document/I2a05f3d71cb811e38578f7ccc38dcbee/View/FullText.html?navigationPath=Search\%2Fv1\%2Fresults\%2Fnavigation\%2Fi0ad6ad3f000001841ed7d74a8e6914e1\%3Fppcid\%3D6c2ae8a868b54456ac4cdc97bf99f24b\%26Nav\%3DKNOWHOW_UK\%26fragmentIdentifier\%3DI2a05f3d71cb811e38578f7ccc38dcbee\%26parentRank\%3D0\%26startIndex\%3D1\%26contextData\%3D\%2528sc.Search\%2529\%26transitionType\%3DSearchItem\&listSource=Search\&listPageSource=c156b6df63e9ac1dbe2bd9174c514051\&list=KNOWHOW_UK\&rank=2\&sessionScopeId=d84126a1818ed3fb6e1aadaf57885dfabcc9a95b48d8801a1ae90e19604d0dc0\&ppcid=6c2ae8a868b54456ac4cdc97bf99f24b\&originationContext=Search\%20Result\&transitionType=SearchItem\&contextData=(sc.Search)\&navId=DADDF17324157881A30EF74E4476B30F\&comp=pluk\#co_anchor_a160961}{Execution
  of deeds}}.
\item
  \textbf{Delivery.} A deed must be delivered. Although this was
  historically a common law requirement, the rule has been modified by
  statute (for example,
  \emph{\href{https://uk.practicallaw.thomsonreuters.com/7-505-7361?originationContext=document\&transitionType=PLDocumentLink\&contextData=(sc.Default)\&ppcid=0e69ecc7bd0b425c8c07be4517611bac}{section
  1(1)(c)}, LP(MP)A 1989} and
  \emph{\href{https://uk.practicallaw.thomsonreuters.com/0-506-8297?originationContext=document\&transitionType=PLDocumentLink\&contextData=(sc.Default)\&ppcid=0e69ecc7bd0b425c8c07be4517611bac}{section
  46(2)}, CA 2006}). For further information, see
  \emph{\href{https://uk.practicallaw.thomsonreuters.com/Document/I2a05f3d71cb811e38578f7ccc38dcbee/View/FullText.html?navigationPath=Search\%2Fv1\%2Fresults\%2Fnavigation\%2Fi0ad6ad3f000001841ed7d74a8e6914e1\%3Fppcid\%3D6c2ae8a868b54456ac4cdc97bf99f24b\%26Nav\%3DKNOWHOW_UK\%26fragmentIdentifier\%3DI2a05f3d71cb811e38578f7ccc38dcbee\%26parentRank\%3D0\%26startIndex\%3D1\%26contextData\%3D\%2528sc.Search\%2529\%26transitionType\%3DSearchItem\&listSource=Search\&listPageSource=c156b6df63e9ac1dbe2bd9174c514051\&list=KNOWHOW_UK\&rank=2\&sessionScopeId=d84126a1818ed3fb6e1aadaf57885dfabcc9a95b48d8801a1ae90e19604d0dc0\&ppcid=6c2ae8a868b54456ac4cdc97bf99f24b\&originationContext=Search\%20Result\&transitionType=SearchItem\&contextData=(sc.Search)\&navId=DADDF17324157881A30EF74E4476B30F\&comp=pluk\#co_anchor_a980095}{Delivery
  of deeds}}.
\end{itemize}

\end{document}
