% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{article}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage[margin=1in]{geometry}
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\usepackage{xcolor}
\definecolor{aliceblue}{HTML}{F0F8FF}
\definecolor{antiquewhite}{HTML}{FAEBD7}
\definecolor{aqua}{HTML}{00FFFF}
\definecolor{aquamarine}{HTML}{7FFFD4}
\definecolor{azure}{HTML}{F0FFFF}
\definecolor{beige}{HTML}{F5F5DC}
\definecolor{bisque}{HTML}{FFE4C4}
\definecolor{black}{HTML}{000000}
\definecolor{blanchedalmond}{HTML}{FFEBCD}
\definecolor{blue}{HTML}{0000FF}
\definecolor{blueviolet}{HTML}{8A2BE2}
\definecolor{brown}{HTML}{A52A2A}
\definecolor{burlywood}{HTML}{DEB887}
\definecolor{cadetblue}{HTML}{5F9EA0}
\definecolor{chartreuse}{HTML}{7FFF00}
\definecolor{chocolate}{HTML}{D2691E}
\definecolor{coral}{HTML}{FF7F50}
\definecolor{cornflowerblue}{HTML}{6495ED}
\definecolor{cornsilk}{HTML}{FFF8DC}
\definecolor{crimson}{HTML}{DC143C}
\definecolor{cyan}{HTML}{00FFFF}
\definecolor{darkblue}{HTML}{00008B}
\definecolor{darkcyan}{HTML}{008B8B}
\definecolor{darkgoldenrod}{HTML}{B8860B}
\definecolor{darkgray}{HTML}{A9A9A9}
\definecolor{darkgreen}{HTML}{006400}
\definecolor{darkgrey}{HTML}{A9A9A9}
\definecolor{darkkhaki}{HTML}{BDB76B}
\definecolor{darkmagenta}{HTML}{8B008B}
\definecolor{darkolivegreen}{HTML}{556B2F}
\definecolor{darkorange}{HTML}{FF8C00}
\definecolor{darkorchid}{HTML}{9932CC}
\definecolor{darkred}{HTML}{8B0000}
\definecolor{darksalmon}{HTML}{E9967A}
\definecolor{darkseagreen}{HTML}{8FBC8F}
\definecolor{darkslateblue}{HTML}{483D8B}
\definecolor{darkslategray}{HTML}{2F4F4F}
\definecolor{darkslategrey}{HTML}{2F4F4F}
\definecolor{darkturquoise}{HTML}{00CED1}
\definecolor{darkviolet}{HTML}{9400D3}
\definecolor{deeppink}{HTML}{FF1493}
\definecolor{deepskyblue}{HTML}{00BFFF}
\definecolor{dimgray}{HTML}{696969}
\definecolor{dimgrey}{HTML}{696969}
\definecolor{dodgerblue}{HTML}{1E90FF}
\definecolor{firebrick}{HTML}{B22222}
\definecolor{floralwhite}{HTML}{FFFAF0}
\definecolor{forestgreen}{HTML}{228B22}
\definecolor{fuchsia}{HTML}{FF00FF}
\definecolor{gainsboro}{HTML}{DCDCDC}
\definecolor{ghostwhite}{HTML}{F8F8FF}
\definecolor{gold}{HTML}{FFD700}
\definecolor{goldenrod}{HTML}{DAA520}
\definecolor{gray}{HTML}{808080}
\definecolor{green}{HTML}{008000}
\definecolor{greenyellow}{HTML}{ADFF2F}
\definecolor{grey}{HTML}{808080}
\definecolor{honeydew}{HTML}{F0FFF0}
\definecolor{hotpink}{HTML}{FF69B4}
\definecolor{indianred}{HTML}{CD5C5C}
\definecolor{indigo}{HTML}{4B0082}
\definecolor{ivory}{HTML}{FFFFF0}
\definecolor{khaki}{HTML}{F0E68C}
\definecolor{lavender}{HTML}{E6E6FA}
\definecolor{lavenderblush}{HTML}{FFF0F5}
\definecolor{lawngreen}{HTML}{7CFC00}
\definecolor{lemonchiffon}{HTML}{FFFACD}
\definecolor{lightblue}{HTML}{ADD8E6}
\definecolor{lightcoral}{HTML}{F08080}
\definecolor{lightcyan}{HTML}{E0FFFF}
\definecolor{lightgoldenrodyellow}{HTML}{FAFAD2}
\definecolor{lightgray}{HTML}{D3D3D3}
\definecolor{lightgreen}{HTML}{90EE90}
\definecolor{lightgrey}{HTML}{D3D3D3}
\definecolor{lightpink}{HTML}{FFB6C1}
\definecolor{lightsalmon}{HTML}{FFA07A}
\definecolor{lightseagreen}{HTML}{20B2AA}
\definecolor{lightskyblue}{HTML}{87CEFA}
\definecolor{lightslategray}{HTML}{778899}
\definecolor{lightslategrey}{HTML}{778899}
\definecolor{lightsteelblue}{HTML}{B0C4DE}
\definecolor{lightyellow}{HTML}{FFFFE0}
\definecolor{lime}{HTML}{00FF00}
\definecolor{limegreen}{HTML}{32CD32}
\definecolor{linen}{HTML}{FAF0E6}
\definecolor{magenta}{HTML}{FF00FF}
\definecolor{maroon}{HTML}{800000}
\definecolor{mediumaquamarine}{HTML}{66CDAA}
\definecolor{mediumblue}{HTML}{0000CD}
\definecolor{mediumorchid}{HTML}{BA55D3}
\definecolor{mediumpurple}{HTML}{9370DB}
\definecolor{mediumseagreen}{HTML}{3CB371}
\definecolor{mediumslateblue}{HTML}{7B68EE}
\definecolor{mediumspringgreen}{HTML}{00FA9A}
\definecolor{mediumturquoise}{HTML}{48D1CC}
\definecolor{mediumvioletred}{HTML}{C71585}
\definecolor{midnightblue}{HTML}{191970}
\definecolor{mintcream}{HTML}{F5FFFA}
\definecolor{mistyrose}{HTML}{FFE4E1}
\definecolor{moccasin}{HTML}{FFE4B5}
\definecolor{navajowhite}{HTML}{FFDEAD}
\definecolor{navy}{HTML}{000080}
\definecolor{oldlace}{HTML}{FDF5E6}
\definecolor{olive}{HTML}{808000}
\definecolor{olivedrab}{HTML}{6B8E23}
\definecolor{orange}{HTML}{FFA500}
\definecolor{orangered}{HTML}{FF4500}
\definecolor{orchid}{HTML}{DA70D6}
\definecolor{palegoldenrod}{HTML}{EEE8AA}
\definecolor{palegreen}{HTML}{98FB98}
\definecolor{paleturquoise}{HTML}{AFEEEE}
\definecolor{palevioletred}{HTML}{DB7093}
\definecolor{papayawhip}{HTML}{FFEFD5}
\definecolor{peachpuff}{HTML}{FFDAB9}
\definecolor{peru}{HTML}{CD853F}
\definecolor{pink}{HTML}{FFC0CB}
\definecolor{plum}{HTML}{DDA0DD}
\definecolor{powderblue}{HTML}{B0E0E6}
\definecolor{purple}{HTML}{800080}
\definecolor{red}{HTML}{FF0000}
\definecolor{rosybrown}{HTML}{BC8F8F}
\definecolor{royalblue}{HTML}{4169E1}
\definecolor{saddlebrown}{HTML}{8B4513}
\definecolor{salmon}{HTML}{FA8072}
\definecolor{sandybrown}{HTML}{F4A460}
\definecolor{seagreen}{HTML}{2E8B57}
\definecolor{seashell}{HTML}{FFF5EE}
\definecolor{sienna}{HTML}{A0522D}
\definecolor{silver}{HTML}{C0C0C0}
\definecolor{skyblue}{HTML}{87CEEB}
\definecolor{slateblue}{HTML}{6A5ACD}
\definecolor{slategray}{HTML}{708090}
\definecolor{slategrey}{HTML}{708090}
\definecolor{snow}{HTML}{FFFAFA}
\definecolor{springgreen}{HTML}{00FF7F}
\definecolor{steelblue}{HTML}{4682B4}
\definecolor{tan}{HTML}{D2B48C}
\definecolor{teal}{HTML}{008080}
\definecolor{thistle}{HTML}{D8BFD8}
\definecolor{tomato}{HTML}{FF6347}
\definecolor{turquoise}{HTML}{40E0D0}
\definecolor{violet}{HTML}{EE82EE}
\definecolor{wheat}{HTML}{F5DEB3}
\definecolor{white}{HTML}{FFFFFF}
\definecolor{whitesmoke}{HTML}{F5F5F5}
\definecolor{yellow}{HTML}{FFFF00}
\definecolor{yellowgreen}{HTML}{9ACD32}
\usepackage[most]{tcolorbox}

\usepackage{ifthen}
\provideboolean{admonitiontwoside}
\makeatletter%
\if@twoside%
\setboolean{admonitiontwoside}{true}
\else%
\setboolean{admonitiontwoside}{false}
\fi%
\makeatother%
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Contract Legislation},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

\title{Contract Legislation}
\author{}
\date{}

\begin{document}
\maketitle

{
\setcounter{tocdepth}{3}
\tableofcontents
}
\hypertarget{sale-of-goods-act-1979}{%
\section{Sale of Goods Act 1979}\label{sale-of-goods-act-1979}}

\href{https://www.legislation.gov.uk/ukpga/1979/54/contents}{Sale of
Goods Act 1979}, as amended by the Sale and Supply of Goods Act 1994,
(SGA).

Implied terms set out do not apply to consumer contracts (e.g.~between a
trader and a consumer) defined in Consumer Rights Act 2015. Similar
terms but not the same.

Note that s 14(2) imposes strict liability.

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.4211}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.5789}}@{}}
\toprule()
\begin{minipage}[b]{\linewidth}\raggedright
Sections
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Explanation
\end{minipage} \\
\midrule()
\endhead
s 12 & Provides for the implication that the seller has the
\textbf{right to sell} the goods. Categorised as a condition of the
contract. \\
s 13 & Where there is a contract for the sale of goods by description,
there is an implied term that the goods will \textbf{correspond with the
description}. Categorised as a condition pursuant to s 13(1A). But
consider also s 15(A), breach so slight. So in reality more of a HK Fir
style test. \\
s 14(2) & Where seller sells goods, implied term that goods supplied are
of a \textbf{satisfactory quality}. Defined in terms of reasonable
person test, taking into account description of goods, price etc (s
14(2A)). Satisfactory means fitness for purposes, appearance and finish,
freedom from minor defects (if not examined/ highlighted in advance),
safety, durability (s 14(2B)). Breach is a breach of condition (s
14(6)). \\
s 14(3) & Goods should be \textbf{fit for purpose} - whether or not that
is a purpose for which such goods are commonly supplied, except where
buyer does not rely/ cannot reasonably rely on skill/judgement of
seller/credit-broker. Breach classified as a breach of a condition. \\
s 15 & For sale by sample: \textbf{bulk will correspond with the sample}
in quality, goods free from any defect not apparent on reasonable
inspection of sample. Breach is breach of condition. If breach is so
slight that unreasonable for buyer to reject the goods, breach should be
treated as breach of warranty (and burden of proof on seller) \\
s 55 & A seller's liability under above sections can be excluded and/or
restricted by agreement of parties, subject to Unfair Contract Terms Act
1977 (UCTA). But UCTA pretty toothless. \\
\bottomrule()
\end{longtable}

\hypertarget{implied-terms-as-to-title-s-12}{%
\subsection{Implied Terms as to Title -- S
12}\label{implied-terms-as-to-title-s-12}}

Section 12 provides for the implication that the seller has the right to
sell the goods.

The term implied by subsection 12(1) is categorised as a condition of
the contract. The distinction between conditions and warranties and the
relevance of that distinction are not detailed in this element.

\hypertarget{correspondence-with-description-s-13}{%
\subsection{Correspondence with Description -- S
13}\label{correspondence-with-description-s-13}}

Section 13(1) provides:

\begin{quote}
``Where there is a contract for the sale of goods by description, there
is an implied term that the goods will \textbf{correspond with the
description}''
\end{quote}

Where goods are described as having certain characteristics and
specifications then they must correspond with that description.

The term implied by subsection 13(1) is categorised as a condition under
s13(1A).

\hypertarget{satisfactory-quality-s142}{%
\subsection{Satisfactory Quality --
s14(2)}\label{satisfactory-quality-s142}}

Section 14(2) of the Sale of Goods Act 1979 (as amended) provides:

\begin{quote}
``Where the seller sells goods in the course of a business, there is an
implied term that the goods supplied under the contract are of
\textbf{satisfactory quality}''
\end{quote}

According to s14(2A), the goods are of satisfactory quality if:

\begin{quote}
``they meet the standard that a reasonable person would regard as
satisfactory, taking account of any description of the goods, the price
(if relevant) and all the other relevant circumstances''
\end{quote}

Section 14(2B) provides a checklist of what may be taken into account in
assessing whether the goods are of satisfactory quality. These are:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Fitness for all the purposes for which goods of the kind in question
  are commonly supplied;
\item
  Appearance and finish;
\item
  Freedom from minor defects;
\item
  Safety; and
\item
  Durability.
\end{enumerate}

If goods have a self-evident purpose, eg a hot-water bottle or a pair of
underpants, this self-evident purpose will clearly be covered by
s14(2B)(a) as a purpose for which the `goods of the kind in question are
commonly supplied'.

The exceptions to the implied term as to satisfactory quality are to be
found in s14(2C)(a) and (b). Thus, there is no condition as regards
defects specifically drawn to the buyer's attention before the contract
was made, or, if the buyer examines the goods before the contract is
made, as regards defects which that examination ought to have revealed.
Nor will a condition be implied where the seller is not selling in the
course of a business.

Breach of s14(2) is classified as breach of a condition by s14(6).

\hypertarget{fitness-for-a-particular-purpose-s143}{%
\subsection{Fitness for a Particular Purpose --
s14(3)}\label{fitness-for-a-particular-purpose-s143}}

In addition to satisfactory quality, the goods should also be fit for
purpose.

Section 14(3) states:

\begin{quote}
``Where the seller sells goods in the course of a business and the
buyer, expressly or by implication, makes known:

\begin{enumerate}
\def\labelenumi{\alph{enumi})}
\tightlist
\item
  to the seller, or\\
\item
  where the purchase price or part of it is payable by instalments and
  the goods
\end{enumerate}

were previously sold by a credit-broker to the seller, to that
credit-broker, any particular purpose for which the goods are being
bought, there is an implied term that the goods supplied under the
contract are \textbf{reasonably fit for that purpose}, whether or not
that is a purpose for which such goods are commonly supplied, except
where the circumstances show that the buyer does not rely, or that it is
unreasonable for him to rely, on the skill or judgment of the seller or
credit-broker.''
\end{quote}

The section above makes clear that no implied condition as to fitness
for a particular purpose arises where the circumstances show that the
buyer does not rely, or that it is unreasonable for him to rely, on the
seller's skill or judgment.

Breach of s14(3) is classified as breach of a condition by s14(6).

\hypertarget{sale-by-sample-s-15}{%
\subsection{Sale by Sample -- S 15}\label{sale-by-sample-s-15}}

Where a sale is by sample, two conditions are implied into the contract
by the Sale of Goods Act 1979 by virtue of s15(2):

\begin{quote}
\begin{enumerate}
\def\labelenumi{(\alph{enumi})}
\tightlist
\item
  that the bulk will correspond with the sample in quality;\\
\item
  that the goods will be free from any defect, making their quality
  unsatisfactory, which would not be apparent on reasonable examination
  of the sample.
\end{enumerate}
\end{quote}

Breach of s15(2) is classified as breach of a condition by s15(3), but
the remedies available are subject to s15A SGA 1979 (see later in this
element).

\hypertarget{modification-of-remedies-for-breach-of-condition---s15a}{%
\subsubsection{Modification of Remedies for Breach of Condition -
s15A}\label{modification-of-remedies-for-breach-of-condition---s15a}}

Although a breach of ss13, 14(2), 14(3) and 15 is breach of a condition,
as stated above, this is subject to s15A.

Section 15A provides that if the breach is so slight that it would be
unreasonable for the buyer to reject the goods and repudiate the
contract, the breach should be treated as a breach of warranty, which
will only entitle the buyer to claim damages.

The burden of proving that the breach is so slight and therefore
unreasonable for the buyer to reject the goods falls on the seller.

\textbf{Contracting out of the implied terms under the Sale of Goods Act
1979 (as amended)}

Section 55 indicates that a seller's liability under ss12, 13, 14 and 15
of the SGA 1979 can be excluded and/or restricted by agreement of the
parties, subject to the Unfair Contract Terms Act 1977 (UCTA).

UCTA does in fact significantly limit the ability to exclude / restrict
these terms.

Under s 6(1)(a) of UCTA, the implied undertaking as to title contained
in s 12 SGA 1979 cannot be excluded or restricted.

Under s6(1A) of UCTA, the implied undertakings as to description,
quality, fitness for purpose or sample contained in ss13-15 of the SGA
can be excluded / restricted subject to the requirement of
reasonableness. The detail of this is not addressed in this element.

\hypertarget{supply-of-goods-and-services-act-1982}{%
\section{Supply of Goods and Services Act
1982}\label{supply-of-goods-and-services-act-1982}}

Provides for the implication of terms in

\begin{itemize}
\tightlist
\item
  Certain contracts for transfer of property in goods
\item
  Contracts for hire of goods
\item
  Contracts for supply of services
\end{itemize}

Implied terms do not apply to consumer contracts defined in Consumer
Rights Act 2015.

\hypertarget{contracts-for-transfer-of-property-in-goods}{%
\subsection{Contracts for Transfer of Property in
Goods}\label{contracts-for-transfer-of-property-in-goods}}

A contract under which one person transfers or agrees to transfer to
another the property in goods, \textbf{other than} any of the following
(s 1):

\begin{itemize}
\tightlist
\item
  Contract of sale of goods
\item
  Hire purchase agreement
\item
  Contract under which property in goods is or is to be transferred in
  exchange for trading stamps
\item
  Transfer made by deed for which there is no consideration other than
  presumed consideration
\item
  Contract intended to operate by way of security
\end{itemize}

In any contract for transfer of goods, terms will be implied
corresponding to those implied in case of contracts for sale of goods
(title (s 2), description (s 3), quality and fitness of purpose (s 4),
transfer by sample where applicable (s 5)).

\hypertarget{contracts-for-hire-of-goods}{%
\subsection{Contracts for Hire of
Goods}\label{contracts-for-hire-of-goods}}

Means a contract under which one person bails or agrees to bail goods to
another by way of hire (s 6) other than the following:

\begin{itemize}
\tightlist
\item
  Hire purchase agreement, and
\item
  Contract under which goods bailed in exchange for trading stamps
\end{itemize}

In any contract for hire of goods, terms will be implied corresponding
to those implied in the case of contracts for the sale of goods,
including

\begin{itemize}
\tightlist
\item
  implied terms about the right to transfer possession (s 7)
\item
  implied terms where hire is by description (s 8)
\item
  implied terms about quality or fitness (s 9)
\item
  implied terms where hire is by sample (s 10)
\end{itemize}

\hypertarget{contracts-for-the-supply-of-services}{%
\subsection{Contracts for the Supply of
Services}\label{contracts-for-the-supply-of-services}}

SGSA provides implied terms:

\begin{itemize}
\tightlist
\item
  \textbf{Care and skill}: where supplier acting in course of business,
  implied term that the supplier will carry out service with reasonable
  care and skill (s 13)
\item
  \textbf{Time of performance}: where supplier acting in course of
  business and time for service not fixed by contract, implied term that
  supplier will carry out service within a reasonable time. What is a
  reasonable time is a question of fact. (s 14)
\item
  \textbf{Consideration}: where consideration for service not determined
  by contract, implied term that party contracting with supplier will
  pay a reasonable charge. What is a reasonable charge is a question of
  fact. (s 15)
\end{itemize}

\hypertarget{contracting-out-these-terms}{%
\subsection{Contracting Out These
Terms}\label{contracting-out-these-terms}}

\begin{itemize}
\tightlist
\item
  For contracts for transfer of goods or hire of goods, under s 11 the
  supplier may negative or vary these terms, subject to provisions of
  Unfair Contract Terms Act 1977.
\item
  For contracts for supply of services, under s 16, a party may contract
  out implied terms set out, subject to provisions of UCTA.
\item
  But UCTA pretty toothless
\end{itemize}

\hypertarget{classification-of-terms}{%
\subsection{Classification of Terms}\label{classification-of-terms}}

s 13 SGSA is an \textbf{innominate term}.

\hypertarget{unfair-contract-terms-act-1977}{%
\section{Unfair Contract Terms Act
1977}\label{unfair-contract-terms-act-1977}}

\hypertarget{purpose}{%
\subsection{Purpose}\label{purpose}}

From preamble of
\href{https://www.legislation.gov.uk/ukpga/1977/50/contents}{UCTA 1977}:

\begin{quote}
An Act to impose further limits on the extent to which \ldots{}
liability for breach of contract, or for negligence or other breach of
duty, can be avoided by means of contract terms and otherwise.
\end{quote}

Provisions ensure:

\begin{itemize}
\tightlist
\item
  Certain types of exemption clause have no effect
\item
  Other types of exemption clause are effective only so far as they
  satisfy the requirements of reasonableness
\end{itemize}

\hypertarget{scope}{%
\subsection{Scope}\label{scope}}

\hypertarget{consumer-contracts}{%
\subsubsection{Consumer Contracts}\label{consumer-contracts}}

Provisions do not apply to `consumer contracts'; these are instead
governed by {[}\protect\hyperlink{consumer-rights-act-2015}{Consumer
Rights Act 2015}{]}.

\hypertarget{business-liability}{%
\subsubsection{Business Liability}\label{business-liability}}

s 1(3):

\begin{quote}
In the case of both contract and tort, sections 2 to 7 apply (except
where the contrary is stated in section 6(4)) only to \textbf{business
liability}, that is liability for breach of obligations or duties
arising---
\end{quote}

\begin{quote}
\begin{itemize}
\item
  \begin{enumerate}
  \def\labelenumi{(\alph{enumi})}
  \tightlist
  \item
    from things done or to be done by a person in the course of a
    business (whether his own business or another's); or
  \end{enumerate}
\end{itemize}
\end{quote}

\begin{quote}
\begin{itemize}
\item
  \begin{enumerate}
  \def\labelenumi{(\alph{enumi})}
  \setcounter{enumi}{1}
  \tightlist
  \item
    from the occupation of premises used for business purposes of the
    occupier;
  \end{enumerate}
\end{itemize}
\end{quote}

The effect of this is:

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.5000}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.5000}}@{}}
\toprule()
\begin{minipage}[b]{\linewidth}\raggedright
Number of parties acting in the course of a business
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Statutory regime applicable
\end{minipage} \\
\midrule()
\endhead
Both parties & UCTA \\
One party & CRA 2015 \\
Neither party & Neither statutory regime applies \\
\bottomrule()
\end{longtable}

\hypertarget{unfair-clauses}{%
\subsubsection{Unfair Clauses}\label{unfair-clauses}}

Also note, UCTA \textbf{only applies to exemption clauses} (those which
limit or exclude liability, either directly or indirectly), not to all
`unfair' clauses.

\hypertarget{negligence-liability}{%
\subsection{Negligence Liability}\label{negligence-liability}}

s 1(1):

\begin{quote}
For the purposes of this Part of this Act, ``negligence'' means the
breach---
\end{quote}

\begin{quote}
\begin{itemize}
\item
  \begin{enumerate}
  \def\labelenumi{(\alph{enumi})}
  \tightlist
  \item
    of any obligation, arising from the express or implied terms of a
    contract, to take reasonable care or exercise reasonable skill in
    the performance of the contract;
  \end{enumerate}
\end{itemize}
\end{quote}

\begin{quote}
\begin{itemize}
\item
  \begin{enumerate}
  \def\labelenumi{(\alph{enumi})}
  \setcounter{enumi}{1}
  \tightlist
  \item
    of any common law duty to take reasonable care or exercise
    reasonable skill (but not any stricter duty);
  \end{enumerate}
\end{itemize}
\end{quote}

\begin{enumerate}
\def\labelenumi{(\alph{enumi})}
\tightlist
\item
  covers duties of care imposed by contract (e.g., a breach of s 13 of
  Supply of Goods and Services Act 1982, to carry out a service in the
  course of business with reasonable skill and care). (b) is the tort of
  negligence!
\end{enumerate}

\hypertarget{exempting-liability-for-losses-resulting-from-negligence}{%
\subsubsection{Exempting Liability for Losses Resulting From
Negligence}\label{exempting-liability-for-losses-resulting-from-negligence}}

s 2(1):

\begin{quote}
A person cannot by reference to any contract term or to a notice given
to persons generally or to particular persons exclude or restrict his
liability for \textbf{death or personal injury} resulting from
negligence.
\end{quote}

Accordingly, any attempt to exclude or restrict liability for death or
personal injury resulting from negligence will be void.

s 2(2):

\begin{quote}
In the case of \textbf{other loss or damage}, a person cannot so exclude
or restrict his liability for negligence, except in so far as the term
or notice satisfies the requirement of reasonableness.
\end{quote}

\hypertarget{exempting-liability-for-breach-of-statutory-implied-terms-about-quality-of-goods}{%
\subsubsection{Exempting Liability for Breach of Statutory Implied Terms
About Quality of
Goods}\label{exempting-liability-for-breach-of-statutory-implied-terms-about-quality-of-goods}}

s 6(1a):\\
Liability for breach of s 13-15 of
{[}\protect\hyperlink{sale-of-goods-act-1979}{Sale of Goods Act 1979}{]}
(conformity with description or sample, quality or fitness for a
particular purpose),

\begin{quote}
{[}\ldots{]} cannot be excluded or restricted by reference to a contract
term except in so far as the term satisfies the requirement of
\textbf{reasonableness}.
\end{quote}

\hypertarget{exempting-liability-arising-in-contract}{%
\subsubsection{Exempting Liability Arising in
Contract}\label{exempting-liability-arising-in-contract}}

s 3:

\begin{quote}
(1)This section applies as between contracting parties where one of them
deals \ldots{} on the other's \textbf{written standard terms} of
business.\\
(2)As against that party, the other cannot by reference to any contract
term - (a) when himself in breach of contract, exclude or restrict any
liability of his in respect of the breach; or - (b) claim to be
entitled--- - (i) to render a contractual performance substantially
different from that which was reasonably expected of him, or - (ii) in
respect of the whole or any part of his contractual obligation, to
render no performance at all,\\
except in so far as (in any of the cases mentioned above in this
subsection) the contract term satisfies the requirement of
\textbf{reasonableness}.
\end{quote}

`Written standard terms of business' is not defined, but the court uses
a common sense approach. In {[}{[}St Albans City Council v.
International Computers Ltd {[}1995{]} FSR 686{]}{]} held that even
where a party's general terms had been the subject of negotiation, they
were still dealing on standard terms for purposes of UCTA as terms
remained effectively untouched.

Recently: if exemption clauses are from one party's standard terms, then
even if the other party negotiates other clauses, UCTA will apply
({[}{[}Commercial Management (Investments) Ltd v Mitchell Design and
Construct Ltd {[}2016{]} EWHC 76 (TCC){]}{]}).

Key requirement here is for dealing on `written standard terms'. So,
exclusions/ limitations of liability for breach of contract in
individually negotiated business to business contracts are generally not
regulated by UCTA. That is, unless they are regulated by other areas of
UCTA, such as attempts to limit liability for death or personal injury
caused by negligence.

\hypertarget{reasonableness-test}{%
\subsection{Reasonableness Test}\label{reasonableness-test}}

Set out in s 11(1). To pass the test, the term:

\begin{quote}
{[}\ldots{]} shall have been a fair and reasonable one to be included
having regard to the circumstances which were, or ought reasonably to
have been, known to or in the contemplation of the parties \textbf{when
the contract was made}.
\end{quote}

``Fair and reasonable'' is a difficult test to apply. According to s
11(2), when applied to s 6-7, this should be considered alongside
matters specified in Schedule 2 of the Act. Courts have clarified that
Schedule 2 guidelines may be used more widely, at any time the UCTA
reasonableness test is being applied ({[}{[}Stewart Gill Ltd v Horatio
Myer \& Co.~Ltd {[}1992{]} 1 QB 600{]}{]}).

Guiding factors in Schedule 2:

\begin{itemize}
\item
  \begin{enumerate}
  \def\labelenumi{(\alph{enumi})}
  \tightlist
  \item
    the strength of the bargaining positions of the parties relative to
    each other, taking into account (among other things) alternative
    means by which the customer's requirements could have been met;
  \end{enumerate}
\item
  \begin{enumerate}
  \def\labelenumi{(\alph{enumi})}
  \setcounter{enumi}{1}
  \tightlist
  \item
    whether the customer received an inducement to agree to the term, or
    in accepting it had an opportunity of entering into a similar
    contract with other persons, but without having a similar term;
  \end{enumerate}
\item
  \begin{enumerate}
  \def\labelenumi{(\alph{enumi})}
  \setcounter{enumi}{2}
  \tightlist
  \item
    whether the customer knew or ought reasonably to have known of the
    existence and the extent of the term (having regard, among other
    things, to any custom of the trade and any previous course of
    dealing between the parties);
  \end{enumerate}
\item
  \begin{enumerate}
  \def\labelenumi{(\alph{enumi})}
  \setcounter{enumi}{3}
  \tightlist
  \item
    where the term excludes or restricts any relevant liability if some
    condition was not complied with, whether it was reasonable at the
    time of the contract to expect that compliance with that condition
    would be practicable;
  \end{enumerate}
\item
  \begin{enumerate}
  \def\labelenumi{(\alph{enumi})}
  \setcounter{enumi}{4}
  \tightlist
  \item
    whether the goods were manufactured, processed or adapted to the
    special order of the customer.
  \end{enumerate}
\end{itemize}

Reasonableness in case law: {[}{[}George Mitchell (Chester Hall) Ltd v
Finney Lock Seeds Ltd (1983) 2 AC 803{]}{]}. Lord Bridge emphasised that
the question of whether an exemption clause is reasonable was largely
down to judicial discretion. So

\begin{quote}
``the appellate court should treat the original decision {[}of the trial
judge{]} with the utmost respect and refrain from interference with it
unless satisfied that it proceeded on some erroneous principle or was
plainly and obviously wrong.''
\end{quote}

This suggests that there will be little precedent value in decisions
themselves.

\hypertarget{summary}{%
\subsection{Summary}\label{summary}}

For business-to-business agreements, under UCTA 1977:

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.3333}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.3333}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.3333}}@{}}
\toprule()
\begin{minipage}[b]{\linewidth}\raggedright
Type of liability
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Limitation
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Reference
\end{minipage} \\
\midrule()
\endhead
Death or personal injury resulting from negligence & Void &
\href{https://www.legislation.gov.uk/ukpga/1977/50/section/2}{s 2(1)} \\
Other loss resulting from negligence & Valid if reasonable &
\href{https://www.legislation.gov.uk/ukpga/1977/50/section/2}{s 2(2)} \\
Breach of statutory implied term about quality of goods & Valid if
reasonable &
\href{https://www.legislation.gov.uk/ukpga/1977/50/section/6}{6(1A)(a)} \\
Breach of contract if express term limiting liability is in a negotiated
agreement & Valid &
\href{https://www.legislation.gov.uk/ukpga/1977/50/section/3}{s 3} \\
Breach of contract if express term limiting liability is in a party's
`standard terms' & Valid if reasonable &
\href{https://www.legislation.gov.uk/ukpga/1977/50/section/3}{s 3} \\
\bottomrule()
\end{longtable}

\hypertarget{consumer-rights-act-2015}{%
\section{Consumer Rights Act 2015}\label{consumer-rights-act-2015}}

Three parts:

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.2135}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.7865}}@{}}
\toprule()
\begin{minipage}[b]{\linewidth}\raggedright
Part of Act
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Description
\end{minipage} \\
\midrule()
\endhead
{[}\protect\hyperlink{cra-2015-part-1}{CRA 2015 Part 1}{]} & Deals with
consumer contracts for goods, digital content, and services \\
{[}\protect\hyperlink{cra-2015-part-2}{CRA 2015 Part 2}{]} & Regulates
unfair terms (in consumer contracts) \\
CRA 2015 Part 3 & Miscellaneous \\
\bottomrule()
\end{longtable}

Applies to contracts between a {[}{[}Trader{]}{]} and a
{[}{[}Consumer{]}{]}.

s 2(4) says that where a trader claims that an individual was not acting
for purposes wholly or mainly outside the individual's trade, business,
craft or profession, then it is for the trader to prove it.

\begin{quote}
Contracts between a consumer and business entered into prior to 1/10/15
governed by Sales of Goods Act 1979, Supply of Goods and Services Act
1982 and Unfair Contract Terms Act 1977.
\end{quote}

s 48(2) states that contracts of employment are \textbf{not} covered by
the Act.

\hypertarget{cra-2015-part-1}{%
\section{CRA 2015 Part 1}\label{cra-2015-part-1}}

\hypertarget{implied-terms}{%
\subsection{Implied Terms}\label{implied-terms}}

\hypertarget{contracts-for-goods}{%
\subsubsection{Contracts for Goods}\label{contracts-for-goods}}

Rights:

\begin{itemize}
\tightlist
\item
  Goods should be of a \textbf{satisfactory quality} (s 9)
\item
  Goods should be \textbf{reasonably fit for any purpose} which the
  consumer makes known to the trader that the consumer intends to use
  the goods for (s 10)
\item
  Where goods are sold by description, the goods should match that
  \textbf{description} (s 11)
\end{itemize}

\hypertarget{contracts-for-digital-content}{%
\subsubsection{Contracts for Digital
Content}\label{contracts-for-digital-content}}

Digital content:

\begin{itemize}
\tightlist
\item
  Should be of \textbf{satisfactory quality} (s 34)

  \begin{itemize}
  \tightlist
  \item
    i.e., should meet the standard that a reasonable person would
    consider satisfactory (s 34(2))
  \item
    RP would consider description, price, all other relevant
    circumstances
  \item
    Such circumstances include `any public statement about specific
    characteristics of the digital content made by the trader, the
    producer, or any representative of the trader or producer' (s 34(5))
  \end{itemize}
\item
  Should be reasonably \textbf{fit for purpose} (s 35)
\item
  Should match the \textbf{description} of it given by trader to
  consumer (s 36)
\end{itemize}

\hypertarget{contracts-for-services}{%
\subsubsection{Contracts for Services}\label{contracts-for-services}}

Rights:

\begin{itemize}
\tightlist
\item
  Service performed with \textbf{reasonable care and skill} (s 49)
\item
  (where a price not agreed) a \textbf{reasonable price} is to be paid
  (s 51)
\item
  (where a time not fixed) service is to be provided in
  \textbf{reasonable time} (s 52)
\end{itemize}

s 50(1) provides that anything said or written to the consumer, on or on
behalf of trader, about trader or service is to be included as a term
where it is taken into account by the consumer either

\begin{itemize}
\tightlist
\item
  when deciding to enter into contract, or
\item
  when making any decision about the service after entering into the
  contract
\end{itemize}

\hypertarget{additional-remedies}{%
\subsection{Additional Remedies}\label{additional-remedies}}

CRA 2015 specifies remedies for consumers when terms breached. These sit
alongside remedies available to the consumer under general law.

\hypertarget{contracts-for-goods-1}{%
\subsubsection{Contracts for Goods}\label{contracts-for-goods-1}}

Where goods sold to consumer fail to meet requirements of s 9, s 10 or s
11, then goods required as non-conforming. Three remedies available to
consumer:

\begin{itemize}
\tightlist
\item
  Short term right to reject

  \begin{itemize}
  \tightlist
  \item
    Usually available for 30 days from time

    \begin{itemize}
    \tightlist
    \item
      That ownership has passed, and
    \item
      The goods have been delivered, and
    \item
      In cases where trader required to install goods/ take other setup
      steps, trader has notified consumer that required steps have been
      taken (s 22)
    \end{itemize}
  \end{itemize}
\item
  Right to repair or replacement

  \begin{itemize}
  \tightlist
  \item
    Available unless impossible or disproportionate (unreasonable cost
    to trader relative to consumer interests and other remedies) (s 23)
  \end{itemize}
\item
  Right to price reduction or final right to reject

  \begin{itemize}
  \tightlist
  \item
    Not entitled to both
  \item
    In either case, can only be exercised where

    \begin{itemize}
    \tightlist
    \item
      After one repair or replacement, goods do not conform to contract
    \item
      Consumer cannot request repair or replacement since impossible/
      disproportionate
    \item
      Consumer has required trader to repair or replace goods, but
      trader in breach of requirement to do so in reasonable time
      without significant inconvenience to consumer (s 24)
    \item
      General rule: where final right to reject exercised within 6
      months (clock running as with short term right to reject), should
      be full refund with no deduction for use (s 24(10)). Does not
      apply to motor vehicles and some other exceptions.
    \end{itemize}
  \end{itemize}
\end{itemize}

Consumers cannot treat the contract as at an end as a result of a breach
of a term implied by s 9, 10 or 11, save to the extent set out above. So
implied terms are neither conditions nor warranties; the extent to which
breach gives rise to the right to treat contract as at an end set out in
the Act.

\hypertarget{contracts-for-digital-goods}{%
\subsubsection{Contracts for Digital
Goods}\label{contracts-for-digital-goods}}

s 42: Where digital content is non-conforming in breach of terms implied
by s 34, s 35 or s 36: two remedial options available:

\begin{itemize}
\tightlist
\item
  Right to repair or replacement

  \begin{itemize}
  \tightlist
  \item
    Within reasonable time and without significant inconvenience (s
    43(2)(a)) (judged by nature and purpose of digital content (s
    43(5)))
  \item
    When not impossible/ disproportionate (s 43(3))
  \end{itemize}
\item
  Right to price reduction

  \begin{itemize}
  \tightlist
  \item
    Only exercisable where consumer cannot require repair or replacement
    (impossible or disproportionate), or where trader has failed to
    replace/repair within reasonable time and without significant
    inconvenience (s 44)
  \end{itemize}
\end{itemize}

s 45: Consumer has a right to receive a refund of all money paid for the
digital content. Refund must be given in 14 days, using same payment
method, without imposing any fee.

Digital content which does not conform to the contract at any time
within 6 months, beginning with the day on which supplied, must be taken
not to have conformed to the contract when it was supplied (s 42(9)).

Similar to with goods, consumers cannot treat the contract as at an end
as a result of a breach of a term.

s 46: Where

\begin{itemize}
\tightlist
\item
  Trader supplies digital content to consumer under contract
\item
  Digital content causes damage to device or other digital content
\item
  Device or digital content damaged belongs to consumer
\item
  Damage of a kind which would not have occurred if trader had exercised
  reasonable care and skill\\
  then consumer entitled to repair/ compensatory payment.
\end{itemize}

\hypertarget{contracts-for-services-1}{%
\subsubsection{Contracts for Services}\label{contracts-for-services-1}}

Where services non-conforming in breach of s 49 (reasonable care and
skill), two remedial options available:

\begin{itemize}
\tightlist
\item
  Right to require repeat performance

  \begin{itemize}
  \tightlist
  \item
    Reasonable time and without significant inconvenience (s 55(2)(a))
  \item
    Not if impossible (s 55(3))
  \end{itemize}
\item
  Right to price reduction (s 56(3))

  \begin{itemize}
  \tightlist
  \item
    When repeat performance impossible
  \item
    Or when trader has failed to provide repeat performance within
    reasonable time and without significant inconvenience to consumer
  \end{itemize}
\end{itemize}

Where the services are non-conforming in breach of the term implied by s
52 (performance within a reasonable time) the remedial option is the
right to a price reduction.

\hypertarget{cra-2015-part-2}{%
\section{CRA 2015 Part 2}\label{cra-2015-part-2}}

\hypertarget{scope-1}{%
\subsection{Scope}\label{scope-1}}

CRA regulates unfair terms in two areas:

\begin{itemize}
\tightlist
\item
  It regulates attempts to exclude/ limit liability for breach of
  contract (similar to
  {[}\protect\hyperlink{unfair-contract-terms-act-1977}{Unfair Contract
  Terms Act 1977}{]}, just for consumer contracts)
\item
  It regulates unfair terms more generally
\end{itemize}

There is no distinction here between contracts on standard terms and
those that have been individually negotiated.

\hypertarget{breach-of-contract-liability}{%
\subsection{Breach of Contract
Liability}\label{breach-of-contract-liability}}

Negligence defined in s 65(4) as breach of:

\begin{quote}
\begin{itemize}
\item
  \begin{enumerate}
  \def\labelenumi{(\alph{enumi})}
  \tightlist
  \item
    Any obligation to take reasonable care or exercise reasonable skill
    in the performance of a contract where the obligation arises from an
    express or implied term of the contract,
  \end{enumerate}
\item
  \begin{enumerate}
  \def\labelenumi{(\alph{enumi})}
  \setcounter{enumi}{1}
  \tightlist
  \item
    A common law duty to take reasonable care or exercise reasonable
    skill,
  \end{enumerate}
\item
  \begin{enumerate}
  \def\labelenumi{(\alph{enumi})}
  \setcounter{enumi}{2}
  \tightlist
  \item
    The common duty of care imposed by the Occupiers' Liability Act 1957
  \end{enumerate}
\end{itemize}
\end{quote}

\begin{enumerate}
\def\labelenumi{(\alph{enumi})}
\setcounter{enumi}{1}
\tightlist
\item
  is the tort of negligence, (a) is duties of care imposed by contract.
  Recall, s 49 {[}\protect\hyperlink{cra-2015-part-1}{CRA 2015 Part
  1}{]} implies a term into a contract for the supply of a service that
  the supplier will carry out the service with reasonable skill and
  care. So, breach of such a term would constitute negligence for
  purposes of CRA.
\end{enumerate}

\hypertarget{death-or-personal-injury-resulting-from-negligence}{%
\subsubsection{Death or Personal Injury Resulting From
Negligence}\label{death-or-personal-injury-resulting-from-negligence}}

s 65(1):

\begin{quote}
A trader cannot be a term of a consumer contract or by a consumer notice
exclude or restrict liability for death or personal injury resulting
from negligence.
\end{quote}

So any attempts to exclude liability like this will not be binding on
the consumer.

\hypertarget{implied-terms-about-goods}{%
\subsubsection{Implied Terms About
Goods}\label{implied-terms-about-goods}}

s 31 (1):

\begin{quote}
A term of a contract to supply goods is not binding on the consumer to
the extent that it would exclude or restrict the trader's liability
arising under any of these provisions--- - (a) section 9 (goods to be of
satisfactory quality); - (b) section 10 (goods to be fit for particular
purpose); - (c) section 11 (goods to be as described);
\end{quote}

\hypertarget{implied-terms-about-digital-content}{%
\subsubsection{Implied Terms About Digital
Content}\label{implied-terms-about-digital-content}}

s 47(1):

\begin{quote}
A term of a contract to supply digital content is not binding on the
consumer to the extent that it would exclude or restrict the trader's
liability arising under any of these provisions--- - (a) section 34
(digital content to be of satisfactory quality), - (b) section 35
(digital content to be fit for particular purpose), - (c) section 36
(digital content to be as described),
\end{quote}

\hypertarget{implied-terms-about-services}{%
\subsubsection{Implied Terms About
Services}\label{implied-terms-about-services}}

s 57(1):

\begin{quote}
A term of a contract to supply services is not binding on the consumer
to the extent that it would exclude the trader's liability arising under
section 49 (service to be performed with reasonable care and skill).
\end{quote}

This section is complicated but:

\begin{itemize}
\tightlist
\item
  Any attempt to \textbf{entirely exclude} section 49 will \textbf{not}
  be binding
\item
  Any attempt to \textbf{restrict/ limit liability} under section 49
  will \textbf{not be binding to the extent that it would prevent
  consumer from recovering price paid}. So liability cannot be limited
  to less than price paid.
\end{itemize}

(interpret these in the logical way)

\hypertarget{regulation-of-other-unfair-terms}{%
\subsection{Regulation of Other Unfair
Terms}\label{regulation-of-other-unfair-terms}}

s 62(1):

\begin{quote}
An unfair term of a consumer contract is not binding on the consumer.
\end{quote}

Note this applies to \textbf{any} term. But this is limited by s 64(1):

\begin{quote}
A term of a consumer contract may not be assessed for fairness under
section 62 to the extent that--- - (a) it specifies the main subject
matter of the contract, or - (b) the assessment is of the
appropriateness of the price payable under the contract by comparison
with the goods, digital content or services supplied under it.
\end{quote}

The above holds provided that the terms are \textbf{transparent} and
\textbf{prominent} (s 64).

Unfairness if defined in s 62(4):

\begin{quote}
A term is unfair if, contrary to the requirement of good faith, it
causes a significant imbalance in the parties' rights and obligations
under the contract to the detriment of the consumer.
\end{quote}

This is judged at the time the contract is entered into.

\href{https://www.legislation.gov.uk/ukpga/2015/15/schedule/2/enacted}{Part
1 Schedule 2} contains a list of terms which `may be regarded as
unfair'. Interesting inclusion:

\begin{quote}
A term which has the object or effect of inappropriately excluding or
limiting the legal rights of the consumer in relation to the trader or
another party in the event of total or partial non-performance or
inadequate performance by the trader of any of the contractual
obligations.
\end{quote}

\hypertarget{anti-avoidance-provisions}{%
\subsection{Anti-avoidance Provisions}\label{anti-avoidance-provisions}}

CRA restricts the parties' contractual freedom. Generally it tries to
cover all the loopholes parties may try to use to get around this too,
e.g.,

\begin{itemize}
\tightlist
\item
  Using a secondary separate contract containing terms pertaining to the
  restriction of liabilities (prohibited under s 72)
\item
  Making the liability or its enforcement subject to certain onerous
  conditions (generally prohibited under s 31(2))
\end{itemize}

\hypertarget{summary-1}{%
\subsection{Summary}\label{summary-1}}

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.3333}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.3333}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.3333}}@{}}
\toprule()
\begin{minipage}[b]{\linewidth}\raggedright
Type of term
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Impact of CRA
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Reference
\end{minipage} \\
\midrule()
\endhead
Limitation of liability in relation to death or personal injury
resulting from negligence & Not binding on consumer &
\href{https://www.legislation.gov.uk/ukpga/2015/15/section/65/enacted}{s
65(1)} \\
Limitation of liability in relation to breach of statutory implied terms
under ss 9-11 about goods & Not binding on consumer &
\href{https://www.legislation.gov.uk/ukpga/2015/15/section/31/enacted}{s
31(1)} \\
Limitation of liability in relation to breach of statutory implied terms
under ss 34-36 about digital content & Not binding on consumer &
\href{https://www.legislation.gov.uk/ukpga/2015/15/section/47/enacted}{s
47(1)} \\
Limitation of liability in relation to breach of statutory implied terms
under s 49 about services & Not binding if preventing recovery of price
paid &
\href{https://www.legislation.gov.uk/ukpga/2015/15/section/57/enacted}{s
57} \\
Transparent and prominent terms specifying main subject of content/
price & Cannot be assessed for fairness &
\href{https://www.legislation.gov.uk/ukpga/2015/15/section/64/enacted}{s
64} \\
Any other term & Not binding if contrary to the requirement of good
faith it causes a significant imbalance in parties' rights and
obligations under the contract, to the detriment of the consumer &
\href{https://www.legislation.gov.uk/ukpga/2015/15/section/62/enacted}{s
62} \\
\bottomrule()
\end{longtable}

\end{document}
