% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{article}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage[margin=1in]{geometry}
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\usepackage{xcolor}
\definecolor{aliceblue}{HTML}{F0F8FF}
\definecolor{antiquewhite}{HTML}{FAEBD7}
\definecolor{aqua}{HTML}{00FFFF}
\definecolor{aquamarine}{HTML}{7FFFD4}
\definecolor{azure}{HTML}{F0FFFF}
\definecolor{beige}{HTML}{F5F5DC}
\definecolor{bisque}{HTML}{FFE4C4}
\definecolor{black}{HTML}{000000}
\definecolor{blanchedalmond}{HTML}{FFEBCD}
\definecolor{blue}{HTML}{0000FF}
\definecolor{blueviolet}{HTML}{8A2BE2}
\definecolor{brown}{HTML}{A52A2A}
\definecolor{burlywood}{HTML}{DEB887}
\definecolor{cadetblue}{HTML}{5F9EA0}
\definecolor{chartreuse}{HTML}{7FFF00}
\definecolor{chocolate}{HTML}{D2691E}
\definecolor{coral}{HTML}{FF7F50}
\definecolor{cornflowerblue}{HTML}{6495ED}
\definecolor{cornsilk}{HTML}{FFF8DC}
\definecolor{crimson}{HTML}{DC143C}
\definecolor{cyan}{HTML}{00FFFF}
\definecolor{darkblue}{HTML}{00008B}
\definecolor{darkcyan}{HTML}{008B8B}
\definecolor{darkgoldenrod}{HTML}{B8860B}
\definecolor{darkgray}{HTML}{A9A9A9}
\definecolor{darkgreen}{HTML}{006400}
\definecolor{darkgrey}{HTML}{A9A9A9}
\definecolor{darkkhaki}{HTML}{BDB76B}
\definecolor{darkmagenta}{HTML}{8B008B}
\definecolor{darkolivegreen}{HTML}{556B2F}
\definecolor{darkorange}{HTML}{FF8C00}
\definecolor{darkorchid}{HTML}{9932CC}
\definecolor{darkred}{HTML}{8B0000}
\definecolor{darksalmon}{HTML}{E9967A}
\definecolor{darkseagreen}{HTML}{8FBC8F}
\definecolor{darkslateblue}{HTML}{483D8B}
\definecolor{darkslategray}{HTML}{2F4F4F}
\definecolor{darkslategrey}{HTML}{2F4F4F}
\definecolor{darkturquoise}{HTML}{00CED1}
\definecolor{darkviolet}{HTML}{9400D3}
\definecolor{deeppink}{HTML}{FF1493}
\definecolor{deepskyblue}{HTML}{00BFFF}
\definecolor{dimgray}{HTML}{696969}
\definecolor{dimgrey}{HTML}{696969}
\definecolor{dodgerblue}{HTML}{1E90FF}
\definecolor{firebrick}{HTML}{B22222}
\definecolor{floralwhite}{HTML}{FFFAF0}
\definecolor{forestgreen}{HTML}{228B22}
\definecolor{fuchsia}{HTML}{FF00FF}
\definecolor{gainsboro}{HTML}{DCDCDC}
\definecolor{ghostwhite}{HTML}{F8F8FF}
\definecolor{gold}{HTML}{FFD700}
\definecolor{goldenrod}{HTML}{DAA520}
\definecolor{gray}{HTML}{808080}
\definecolor{green}{HTML}{008000}
\definecolor{greenyellow}{HTML}{ADFF2F}
\definecolor{grey}{HTML}{808080}
\definecolor{honeydew}{HTML}{F0FFF0}
\definecolor{hotpink}{HTML}{FF69B4}
\definecolor{indianred}{HTML}{CD5C5C}
\definecolor{indigo}{HTML}{4B0082}
\definecolor{ivory}{HTML}{FFFFF0}
\definecolor{khaki}{HTML}{F0E68C}
\definecolor{lavender}{HTML}{E6E6FA}
\definecolor{lavenderblush}{HTML}{FFF0F5}
\definecolor{lawngreen}{HTML}{7CFC00}
\definecolor{lemonchiffon}{HTML}{FFFACD}
\definecolor{lightblue}{HTML}{ADD8E6}
\definecolor{lightcoral}{HTML}{F08080}
\definecolor{lightcyan}{HTML}{E0FFFF}
\definecolor{lightgoldenrodyellow}{HTML}{FAFAD2}
\definecolor{lightgray}{HTML}{D3D3D3}
\definecolor{lightgreen}{HTML}{90EE90}
\definecolor{lightgrey}{HTML}{D3D3D3}
\definecolor{lightpink}{HTML}{FFB6C1}
\definecolor{lightsalmon}{HTML}{FFA07A}
\definecolor{lightseagreen}{HTML}{20B2AA}
\definecolor{lightskyblue}{HTML}{87CEFA}
\definecolor{lightslategray}{HTML}{778899}
\definecolor{lightslategrey}{HTML}{778899}
\definecolor{lightsteelblue}{HTML}{B0C4DE}
\definecolor{lightyellow}{HTML}{FFFFE0}
\definecolor{lime}{HTML}{00FF00}
\definecolor{limegreen}{HTML}{32CD32}
\definecolor{linen}{HTML}{FAF0E6}
\definecolor{magenta}{HTML}{FF00FF}
\definecolor{maroon}{HTML}{800000}
\definecolor{mediumaquamarine}{HTML}{66CDAA}
\definecolor{mediumblue}{HTML}{0000CD}
\definecolor{mediumorchid}{HTML}{BA55D3}
\definecolor{mediumpurple}{HTML}{9370DB}
\definecolor{mediumseagreen}{HTML}{3CB371}
\definecolor{mediumslateblue}{HTML}{7B68EE}
\definecolor{mediumspringgreen}{HTML}{00FA9A}
\definecolor{mediumturquoise}{HTML}{48D1CC}
\definecolor{mediumvioletred}{HTML}{C71585}
\definecolor{midnightblue}{HTML}{191970}
\definecolor{mintcream}{HTML}{F5FFFA}
\definecolor{mistyrose}{HTML}{FFE4E1}
\definecolor{moccasin}{HTML}{FFE4B5}
\definecolor{navajowhite}{HTML}{FFDEAD}
\definecolor{navy}{HTML}{000080}
\definecolor{oldlace}{HTML}{FDF5E6}
\definecolor{olive}{HTML}{808000}
\definecolor{olivedrab}{HTML}{6B8E23}
\definecolor{orange}{HTML}{FFA500}
\definecolor{orangered}{HTML}{FF4500}
\definecolor{orchid}{HTML}{DA70D6}
\definecolor{palegoldenrod}{HTML}{EEE8AA}
\definecolor{palegreen}{HTML}{98FB98}
\definecolor{paleturquoise}{HTML}{AFEEEE}
\definecolor{palevioletred}{HTML}{DB7093}
\definecolor{papayawhip}{HTML}{FFEFD5}
\definecolor{peachpuff}{HTML}{FFDAB9}
\definecolor{peru}{HTML}{CD853F}
\definecolor{pink}{HTML}{FFC0CB}
\definecolor{plum}{HTML}{DDA0DD}
\definecolor{powderblue}{HTML}{B0E0E6}
\definecolor{purple}{HTML}{800080}
\definecolor{red}{HTML}{FF0000}
\definecolor{rosybrown}{HTML}{BC8F8F}
\definecolor{royalblue}{HTML}{4169E1}
\definecolor{saddlebrown}{HTML}{8B4513}
\definecolor{salmon}{HTML}{FA8072}
\definecolor{sandybrown}{HTML}{F4A460}
\definecolor{seagreen}{HTML}{2E8B57}
\definecolor{seashell}{HTML}{FFF5EE}
\definecolor{sienna}{HTML}{A0522D}
\definecolor{silver}{HTML}{C0C0C0}
\definecolor{skyblue}{HTML}{87CEEB}
\definecolor{slateblue}{HTML}{6A5ACD}
\definecolor{slategray}{HTML}{708090}
\definecolor{slategrey}{HTML}{708090}
\definecolor{snow}{HTML}{FFFAFA}
\definecolor{springgreen}{HTML}{00FF7F}
\definecolor{steelblue}{HTML}{4682B4}
\definecolor{tan}{HTML}{D2B48C}
\definecolor{teal}{HTML}{008080}
\definecolor{thistle}{HTML}{D8BFD8}
\definecolor{tomato}{HTML}{FF6347}
\definecolor{turquoise}{HTML}{40E0D0}
\definecolor{violet}{HTML}{EE82EE}
\definecolor{wheat}{HTML}{F5DEB3}
\definecolor{white}{HTML}{FFFFFF}
\definecolor{whitesmoke}{HTML}{F5F5F5}
\definecolor{yellow}{HTML}{FFFF00}
\definecolor{yellowgreen}{HTML}{9ACD32}
\usepackage[most]{tcolorbox}

\usepackage{ifthen}
\provideboolean{admonitiontwoside}
\makeatletter%
\if@twoside%
\setboolean{admonitiontwoside}{true}
\else%
\setboolean{admonitiontwoside}{false}
\fi%
\makeatother%

\newenvironment{env-f9227045-5e0a-4928-96e3-3aedd444fec4}
{
    \savenotes\tcolorbox[blanker,breakable,left=5pt,borderline west={2pt}{-4pt}{firebrick}]
}
{
    \endtcolorbox\spewnotes
}
                

\newenvironment{env-486fc4cd-eb32-4215-aaa2-fc77a67eebc5}
{
    \savenotes\tcolorbox[blanker,breakable,left=5pt,borderline west={2pt}{-4pt}{blue}]
}
{
    \endtcolorbox\spewnotes
}
                

\newenvironment{env-99ba4262-5297-4318-9547-86307715f49a}
{
    \savenotes\tcolorbox[blanker,breakable,left=5pt,borderline west={2pt}{-4pt}{green}]
}
{
    \endtcolorbox\spewnotes
}
                

\newenvironment{env-0174d31b-3f44-410b-9ae6-7a7b66b1e278}
{
    \savenotes\tcolorbox[blanker,breakable,left=5pt,borderline west={2pt}{-4pt}{aquamarine}]
}
{
    \endtcolorbox\spewnotes
}
                

\newenvironment{env-31409176-8a91-46bb-85fd-48f2d6a04234}
{
    \savenotes\tcolorbox[blanker,breakable,left=5pt,borderline west={2pt}{-4pt}{orange}]
}
{
    \endtcolorbox\spewnotes
}
                

\newenvironment{env-40bb7fc6-f900-4e05-843f-ce2c0cce552d}
{
    \savenotes\tcolorbox[blanker,breakable,left=5pt,borderline west={2pt}{-4pt}{blue}]
}
{
    \endtcolorbox\spewnotes
}
                

\newenvironment{env-8872a5c8-2356-4b38-87ff-b8a967f137b3}
{
    \savenotes\tcolorbox[blanker,breakable,left=5pt,borderline west={2pt}{-4pt}{yellow}]
}
{
    \endtcolorbox\spewnotes
}
                

\newenvironment{env-828b13a4-7762-4c89-995d-f5fea7d7cc53}
{
    \savenotes\tcolorbox[blanker,breakable,left=5pt,borderline west={2pt}{-4pt}{darkred}]
}
{
    \endtcolorbox\spewnotes
}
                

\newenvironment{env-f70373ea-d099-4f9a-bab1-dd1716394de8}
{
    \savenotes\tcolorbox[blanker,breakable,left=5pt,borderline west={2pt}{-4pt}{pink}]
}
{
    \endtcolorbox\spewnotes
}
                

\newenvironment{env-0c49cb39-e036-4df4-8974-f6a48144dfd4}
{
    \savenotes\tcolorbox[blanker,breakable,left=5pt,borderline west={2pt}{-4pt}{cyan}]
}
{
    \endtcolorbox\spewnotes
}
                

\newenvironment{env-be598745-44b1-4815-8875-ed2abeb511dd}
{
    \savenotes\tcolorbox[blanker,breakable,left=5pt,borderline west={2pt}{-4pt}{cyan}]
}
{
    \endtcolorbox\spewnotes
}
                

\newenvironment{env-91b82eae-616f-475e-aa24-1cbca07fe183}
{
    \savenotes\tcolorbox[blanker,breakable,left=5pt,borderline west={2pt}{-4pt}{purple}]
}
{
    \endtcolorbox\spewnotes
}
                

\newenvironment{env-ed99ec69-8464-4a28-8f3a-7402fdec6a62}
{
    \savenotes\tcolorbox[blanker,breakable,left=5pt,borderline west={2pt}{-4pt}{darksalmon}]
}
{
    \endtcolorbox\spewnotes
}
                

\newenvironment{env-a9aa42a3-f5ef-44bc-9cb3-81fc5caedba7}
{
    \savenotes\tcolorbox[blanker,breakable,left=5pt,borderline west={2pt}{-4pt}{gray}]
}
{
    \endtcolorbox\spewnotes
}
                
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Secured Lending},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

\title{Secured Lending}
\author{}
\date{}

\begin{document}
\maketitle

{
\setcounter{tocdepth}{3}
\tableofcontents
}
\hypertarget{common-securities}{%
\section{Common Securities}\label{common-securities}}

\hypertarget{categorising-security}{%
\subsection{Categorising Security}\label{categorising-security}}

Different types of loan security have different effects:

\begin{itemize}
\tightlist
\item
  Security which gives the bank rights over an asset
\item
  Security under which ownership in the secured asset is transferred to
  the bank
\item
  Security under which a bank has actual (or "constructive") possession
  of the secured asset.
\end{itemize}

\begin{env-8872a5c8-2356-4b38-87ff-b8a967f137b3}

Ownership

Having (legal or equitable) title to an asset.

\end{env-8872a5c8-2356-4b38-87ff-b8a967f137b3}

\begin{env-8872a5c8-2356-4b38-87ff-b8a967f137b3}

Possession

Having physical possession of an asset.

\end{env-8872a5c8-2356-4b38-87ff-b8a967f137b3}

\hypertarget{giving-rights-over-assets}{%
\subsection{Giving Rights Over Assets}\label{giving-rights-over-assets}}

Charges may be `fixed' or `floating' and are always `equitable' (other
than the `charge by deed expressed to be by way of legal mortgage' (LPA
1925, s 87) which is arguably not a true `charge' anyway).

\hypertarget{fixed-charges}{%
\subsection{Fixed Charges}\label{fixed-charges}}

\begin{itemize}
\tightlist
\item
  The most certain way to ensure a secured asset is available to meet
  unpaid monies is to take possession of it (a ``pledge'').

  \begin{itemize}
  \tightlist
  \item
    Prevents the borrower disposing of the asset or diminishing its
    value.
  \item
    Impractical or impossible for a bank to take possession of certain
    assets (e.g., machinery)
  \end{itemize}
\item
  A fixed charge (``specific equitable charge'') gives the bank rights
  over an asset which:

  \begin{itemize}
  \tightlist
  \item
    prohibit the borrower from disposing of the asset without
    permission;
  \item
    attempt to maintain the asset's value whilst it is in the borrower's
    hands; and
  \item
    allow the bank recourse to the asset in the event of the borrower's
    default under the loan.
  \end{itemize}
\end{itemize}

This allows the borrower to use the charged assets unless and until any
``enforcement event'' occurs.

On an `enforcement event', under its contractual rights in the charge
document, the bank may require the borrower to sell the asset, or the
bank may take possession of, and sell, the asset, or appoint a receiver
to do the same (this carries less risk for the bank since the receiver
acts as the borrower's agent).

\begin{env-0c49cb39-e036-4df4-8974-f6a48144dfd4}

Tip

If the borrower sells an asset subject to a fixed charge, the rights
created by the charge usually remain with the asset.

\end{env-0c49cb39-e036-4df4-8974-f6a48144dfd4}

\hypertarget{floating-charges}{%
\subsection{Floating Charges}\label{floating-charges}}

A fixed charge is still unsuitable to secure assets which a borrower
might want to deal with quickly at an unspecified future date (e.g.,
stock). Most companies are unable to grant fixed charges for the
majority of their assets. With a floating charge, whatever assets in a
particular class that a borrower owns at any point in time are subject
to the floating charge.

\begin{itemize}
\tightlist
\item
  A floating charge secures a group of assets, which may fluctuate from
  time to time.
\item
  The assets secured by a floating charge are identified generically,
  for example a borrower's `trading stock' or its `undertaking and
  assets'.
\item
  Unlike a fixed charge, a floating charge specifically allows the
  borrower to deal with the charged assets in the \textbf{ordinary
  course of business} (construed widely).
\item
  Possible to create more than one floating charge over the same group
  of assets.
\end{itemize}

\begin{env-8872a5c8-2356-4b38-87ff-b8a967f137b3}

Definition

In Re Yorkshire Woolcombers Association {[}1904{]} AC 355, a floating
charge was defined as

\begin{itemize}
\tightlist
\item
  A class of assets, present and future; and
\item
  Which in the ordinary course of the company's business changes from
  time to time; and
\item
  It is contemplated that until the holders of the charge take steps to
  enforce it, the company may carry on business in the ordinary way as
  far as concerns the assets charged.
\end{itemize}

\end{env-8872a5c8-2356-4b38-87ff-b8a967f137b3}

Subsequent cases have emphasised the third characteristic identified by
Romer LJ. Thus, if a charge allows a borrower to deal with the charged
assets without the bank's consent, it almost certainly will not be a
fixed charge. Similarly, if a borrower is not free to deal with secured
assets without the bank's consent, the security will be fixed even if it
is described as `floating' in the security document (Russell Cooke Trust
Company Ltd v Elliott {[}2007{]} EWHC 1443 (Ch)).

\hypertarget{crystallisation}{%
\subsubsection{Crystallisation}\label{crystallisation}}

When a floating charge crystallises, it ceases to float over all the
assets in a class and instead fixes onto the assets in the class charged
at the time of crystallisation. The lender then has control of these
assets, and the borrower is unable to deal with them, as if the assets
were subject to a fixed charge.

However, the assets subject to crystallisation of the floating charge
are \textbf{not} treated as fixed charge assets for distribution
purposes on a winding up.

If the company receives more assets of the same class after
crystallisation, these are automatically subject to the crystallisation
charge (NW Robbie and Co v Whitney Warehouse Co Ltd {[}1963{]} 1 WLR
1324, CA). The triggering events for crystallisation are set out in
contract.

\begin{env-486fc4cd-eb32-4215-aaa2-fc77a67eebc5}

Partial crystallisation

Most charges provide for crystallisation to be simultaneously effective
over all the assets secured by the charge. There is, however, no reason
why a floating charge should not be partially crystallised.

\end{env-486fc4cd-eb32-4215-aaa2-fc77a67eebc5}

\begin{itemize}
\tightlist
\item
  Advantages

  \begin{itemize}
  \tightlist
  \item
    Allows the company to deal with the secured assets on a day-to-day
    basis.
  \item
    Allows the company to maximise the amount they are able to borrow.
  \item
    Charge may be granted over the whole of the company's business.
  \end{itemize}
\item
  Disadvantages

  \begin{itemize}
  \tightlist
  \item
    On crystallisation, a floating charge ranks behind any fixed
    security created before it crystallised and is subject to any
    3rd-party rights, including legal or equitable rights, which accrue
    prior to crystallisation.
  \item
    Prior to crystallisation, the company could sell existing stock and
    not purchase new stock to replace it, if there is just a floating
    charge over stock.
  \item
    Preferential creditors also take priority over the holder of a
    floating charge (e.g., employees, pension payments (Bloom and others
    v Pensions Regulator (Nortel, Re) {[}2010{]} EWHC 3010 (Ch)),
    ring-fenced amount for the unsecured creditors).
  \item
    Under s 245 IA 1986, in certain circumstances a floating charge is
    automatically void and a liquidator/ administrator can set it aside
    (basically if created at the 'relevant time' where it was obvious
    the company would fail).

    \begin{itemize}
    \tightlist
    \item
      In Re Comet Group Limited, the judge described its purpose as ``to
      prevent a company that is on its last legs from creating a
      floating charge to secure past debts or to secure moneys which do
      not go to swell its assets and become available for creditors''.
    \end{itemize}
  \item
    Crystallising a floating charge will create an effective
    stranglehold on a borrower: without the use of its current assets it
    cannot usually make any money.
  \end{itemize}
\end{itemize}

Crystallisation occurs:

\begin{enumerate}
\tightlist
\item
  In common law
\item
  Specified event---as defined in the loan agreement. This usually
  occurs where a borrower defaults on the loan repayments or interest
  payments, or where another lender enforces their security.
\end{enumerate}

\hypertarget{under-common-law}{%
\paragraph{Under Common Law}\label{under-common-law}}

Events causing a floating charge to crystallise from case law:

\begin{itemize}
\tightlist
\item
  The borrower's liquidation

  \begin{itemize}
  \tightlist
  \item
    All types of winding-up, including a voluntary winding-up
  \item
    But not just a petition for winding up being presented
  \end{itemize}
\item
  Appointment of a receiver/ administrative receiver by the bank

  \begin{itemize}
  \tightlist
  \item
    If the purpose of the administrative order is to prevent the
    borrower from continuing to run the business as a going concern
  \item
    Mere crystallisation of another floating charge will not cause
    crystallisation under common law.
  \end{itemize}
\item
  Borrower ceasing to carry on business
\item
  Intervention by the bank.
\end{itemize}

\hypertarget{under-the-charge-document}{%
\paragraph{Under the Charge Document}\label{under-the-charge-document}}

Circumstances precipitating crystallisation commonly agreed
contractually between the borrower and bank in the charge document:

\begin{itemize}
\tightlist
\item
  An immediate right to crystallise the floating charge upon giving
  notice to the borrower.

  \begin{itemize}
  \tightlist
  \item
    Borrower may add ``if the bank reasonably believes the secured
    assets are in jeopardy''
  \end{itemize}
\item
  {\(\exists\)} specified events which crystallise the charge without
  any action by the bank - "automatic crystallisation"

  \begin{itemize}
  \tightlist
  \item
    e.g., appointment of an administrator.
  \end{itemize}
\end{itemize}

\hypertarget{post-crystallisation}{%
\paragraph{Post-crystallisation}\label{post-crystallisation}}

\begin{env-be598745-44b1-4815-8875-ed2abeb511dd}

Important

A floating charge creates an immediate security interest but does not
attach to any individual asset until crystallisation.

\end{env-be598745-44b1-4815-8875-ed2abeb511dd}

Crystallisation does not in itself create a new charge to be registered
under the CA 2006 (Re Manurewa Transport Ltd {[}1971{]} NZLR 909).

\hypertarget{weaknesses-of-floating-charges}{%
\paragraph{Weaknesses of Floating
Charges}\label{weaknesses-of-floating-charges}}

\begin{itemize}
\tightlist
\item
  Later fixed charges

  \begin{itemize}
  \tightlist
  \item
    Vulnerable to later fixed charges (unless there is a prohibition on
    creating further security)
  \end{itemize}
\item
  Administrator's costs

  \begin{itemize}
  \tightlist
  \item
    Floating charges rank behind administrator costs and expenses.
  \item
    All liquidation expenses (other than litigation expenses) have
    automatic priority over floating charge holders, and are recoverable
    from the assets of the company
  \end{itemize}
\item
  Preferential creditors

  \begin{itemize}
  \tightlist
  \item
    Any proceeds realised from assets secured by a floating charge must
    be used to pay any `preferential debts' of the company before
    meeting claims under the floating charge (IA 1986, s 40).

    \begin{itemize}
    \tightlist
    \item
      `Preferential debts' are defined as including (s 386 \& Sch 6 IA
      1986) outstanding employees' wages and pension scheme
      contributions.
    \item
      Prescribed part fund for unsecured creditors; 50\% of the first
      £10,000 of the prescribed part and 20\% thereafter, subject to a
      maximum amount of £800,000 if created on or after 6 April 2020
      (£600,000 if created before) (see IA 1986, s 176A).
    \end{itemize}
  \end{itemize}
\item
  Avoidance of floating charges.

  \begin{itemize}
  \tightlist
  \item
    Liquidator/ administrator can apply to court to have a floating
    charge set aside (s 245);

    \begin{itemize}
    \tightlist
    \item
      If granted to a connected person within 2 years of the onset of
      insolvency, or
    \item
      If granted to an unconnected person within 12 months of the onset
      of insolvency.
    \end{itemize}
  \item
    Not invalid in either case to the extent that it is in return for
    valuable consideration given on or after the creation of the charge.
  \end{itemize}
\end{itemize}

\hypertarget{transfer-of-ownership}{%
\subsection{Transfer of Ownership}\label{transfer-of-ownership}}

\hypertarget{mortgages}{%
\subsubsection{Mortgages}\label{mortgages}}

The security provider retains possession of the asset, but transfers
ownership to the creditor. Security taken by transferring ownership
(i.e., legal or equitable title) to the bank, along with a right to sell
on default and an obligation to re-transfer title on satisfaction of the
debt. Highest form of security.

Subject to the security provider's right to require the creditor to
transfer the asset back to it when the debt is repaid. Right is known is
the 'equity of redemption'. The equity of redemption is held in reserve
and only exercised if the borrowed money is not repaid.

A type of mortgage (known as ``charge by way of legal mortgage'' or
``charge by deed expressed to be by way of legal mortgage''--s 87 LPA
1925) is usually taken over land. Ownership will remain vested in the
security provider in this case.

\begin{env-0c49cb39-e036-4df4-8974-f6a48144dfd4}

Tip

A mortgage can therefore be thought of as an `enhanced' charge, since it
gives rights over an asset which are backed with a transfer of legal or
equitable title.\\
{\[\text{Mortgage} = \text{Right\ of\ appropriation} + \text{Transfer\ of\ ownership}\]}

\end{env-0c49cb39-e036-4df4-8974-f6a48144dfd4}

Subject to contrary intention (and statutory restrictions in the case of
land), a (legal) mortgage entitles the bank to take immediate possession
of the mortgaged asset. But taking possession is not required to perfect
the security.

\hypertarget{mortgage-over-land}{%
\paragraph{Mortgage Over Land}\label{mortgage-over-land}}

The only way to create legal security over freehold or leasehold land is
by way of `a charge by deed expressed to be by way of legal mortgage'
(LPA 1925, s 87). The net result is to provide the bank with rights
equivalent to granting it a 3,000-year lease (or, for leasehold, a lease
for one day less than the borrower's lease).

\hypertarget{equitable-mortgages}{%
\paragraph{Equitable Mortgages}\label{equitable-mortgages}}

If a mortgage does not transfer legal title, it will be an equitable
mortgage. These can be created by:

\begin{itemize}
\tightlist
\item
  Written agreement by a borrower that the property is secured by
  equitable mortgage
\item
  Agreement to provide a legal mortgage if the borrower intended the
  agreement to have immediate legal effect.
\item
  Mortgage over an equitable interest

  \begin{itemize}
  \tightlist
  \item
    Where the borrower holds an equitable interest in the land (i.e.,
    they are not a legal owner, e.g., a beneficiary in a trust of land),
    any mortgage of that interest will be equitable in nature.
  \item
    Such a mortgage can be created very informally. In accordance with
    \href{https://www.legislation.gov.uk/ukpga/Geo5/15-16/20/section/53}{s
    53(1)(c) LPA 1925}, such a mortgage need only be in writing and
    signed by the grantor in order to be validly created.
  \end{itemize}
\item
  Purported legal mortgage that does not comply with the necessary
  formalities.

  \begin{itemize}
  \tightlist
  \item
    A mortgage over registered land which is not granted by a valid deed
    or that is not completed by registration will not take effect as a
    legal mortgage (it will be defective).
  \item
    However, it may be regarded as an equitable mortgage if it complies
    with \href{https://www.legislation.gov.uk/ukpga/1989/34/section/2}{s
    2 LP(MP)A 1989}. Equity will recognise it as a `contract to grant a
    legal mortgage' providing it is in writing, contains all the agreed
    terms and is signed by both the mortgagor and mortgagee.
  \end{itemize}
\end{itemize}

There is little practical difference between an equitable mortgage and
equitable fixed charge.

\hypertarget{disadvantages-of-equitable-mortgages}{%
\subparagraph{Disadvantages of Equitable
Mortgages}\label{disadvantages-of-equitable-mortgages}}

\begin{env-31409176-8a91-46bb-85fd-48f2d6a04234}

Warning

An equitable mortgage can be ignored by a bona fide purchaser (or
`security taker') for value of the legal title to an asset without
notice (actual or constructive) of the equitable security.

\end{env-31409176-8a91-46bb-85fd-48f2d6a04234}

But this disadvantage is largely overcome if the equitable security is
registered since this provides notice of the security to anyone
searching the register.

An equitable mortgage may not benefit from the `automatic' powers vested
in legal mortgagees under LPA 1925, s 101, which include the power of
sale (although these powers should be specifically given in the security
document).

\hypertarget{giving-physical-possession}{%
\subsection{Giving Physical
Possession}\label{giving-physical-possession}}

\hypertarget{pledge}{%
\subsubsection{Pledge}\label{pledge}}

\begin{env-8872a5c8-2356-4b38-87ff-b8a967f137b3}

Definition

A pledge is a way of creating security by the actual (or constructive)
delivery of a tangible movable asset to a creditor, to be held until the
performance of an obligation (eg, repaying a debt). The bank will take
possession while the borrower retains ownership.

\end{env-8872a5c8-2356-4b38-87ff-b8a967f137b3}

There is a right to sell the asset to settle the debt owed, upon the
giving of reasonable notice.

\hypertarget{pledgee-rights}{%
\paragraph{Pledgee Rights}\label{pledgee-rights}}

The security is the possession of an asset. A pledged asset must be
``delivered'' by the borrower. This can be through physical delivery or
``constructive delivery'' i.e., vesting control (e.g., keys to a safe
deposit box).

\begin{env-486fc4cd-eb32-4215-aaa2-fc77a67eebc5}

Delivery of documents

Assets may be pledged by delivery of documents, if these documents
effectively provide control. So holding the share certificate of a
registered share does not give sufficient control, even if accompanied
by a blank stock transfer form.

\end{env-486fc4cd-eb32-4215-aaa2-fc77a67eebc5}

The pledge ends when the pledged asset is returned to the borrower. The
most important right implied by a pledge is the right to sell pledged
assets to meet a defaulted debt.

\hypertarget{problems}{%
\paragraph{Problems}\label{problems}}

A bank will probably want to take a pledge only over valuable assets -
how can these be kept secure? The bank will probably be liable as bailee
while it holds the pledged assets, and this may entail insurance costs.

No legal requirement to document a pledge, but documentation is a good
idea (``letter of pledge'' / ``memorandum of deposit'').

\hypertarget{pledge-able-assets}{%
\paragraph{Pledge-able Assets}\label{pledge-able-assets}}

\begin{itemize}
\tightlist
\item
  Asset must be capable of actual/ constructive delivery.
\item
  Common to pledge bearer instruments (e.g., bills of exchange, bearer
  bonds).
\end{itemize}

\hypertarget{lien}{%
\subsubsection{Lien}\label{lien}}

The right to physical possession of the debtor's goods until the debt is
paid. But no right to sell these assets to settle the debt owed. Liens
arise automatically under English law in certain types of commercial
relationship.

3 types:

\begin{enumerate}
\tightlist
\item
  Common law/ particular liens

  \begin{itemize}
  \tightlist
  \item
    Allows retention of the goods under which the debt arises (e.g., a
    dry-cleaner retaining cleaned suits until paid).
  \end{itemize}
\item
  Statutory/ general liens

  \begin{itemize}
  \tightlist
  \item
    Statutory liens allow certain persons to retain any goods in their
    possession against any obligation owed to them by the owner of those
    goods.
  \item
    Arise by operation of law without any agreement
  \item
    e.g., stockbrokers, solicitors, innkeepers benefit.
  \end{itemize}
\item
  Equitable liens

  \begin{itemize}
  \tightlist
  \item
    Lien does not require possession of the property to which it
    attaches (usually land) and will give a right of sale through
    application to the court.
  \item
    Also possible for a lien to be created by a contract.
  \end{itemize}
\end{enumerate}

Common law and statutory liens are both types of legal lien. Legal liens
will usually require actual possession of the property which has been
rightfully acquired. Generally no right to sell the property, with the
exception of a banker's lien.

\hypertarget{assignment}{%
\subsection{Assignment}\label{assignment}}

An assignment may be used to create a mortgage over certain assets.

The most common assets secured by assignment are a borrower's rights
against a third party (`choses in action'), for example debts, and
rights under various contracts.

An assignment by way of security can be legal or equitable:

\begin{itemize}
\tightlist
\item
  Legal if s 136 LPA 1925 complied with (``statutory assignment'') -
  notice in writing.

  \begin{itemize}
  \tightlist
  \item
    Written notice~alone will not necessarily be sufficient to make an
    assignment legal as opposed to equitable. For example, an assignment
    of part will always be equitable even if notice is given.
  \end{itemize}
\item
  Equitable ``assignment by way of charge''.
\end{itemize}

Assignment `by way of charge' is expressly excluded from s 136 and so
assigning rights by way of security will constitute a legal assignment
only if it is an absolute assignment with a proviso for re-assignment on
satisfaction of the secured obligation.

\begin{env-99ba4262-5297-4318-9547-86307715f49a}

136 - Legal assignments of things in action.

(1) Any absolute assignment by writing under the hand of the assignor
(not purporting to be by way of charge only) of any debt or other legal
thing in action, of which \textbf{express notice in writing} has been
given to the debtor, trustee or other person from whom the assignor
would have been entitled to claim such debt or thing in action, is
effectual in law (subject to equities having priority over the right of
the assignee) to pass and transfer from the date of such notice---

\begin{itemize}
\tightlist
\item
  (a) the legal right to such debt or thing in action;
\item
  (b) all legal and other remedies for the same; and
\item
  (c) the power to give a good discharge for the same without the
  concurrence of the assignor:
\end{itemize}

Provided that, if the debtor, trustee or other person liable in respect
of such debt or thing in action has notice---

\begin{itemize}
\tightlist
\item
  (a) that the assignment is disputed by the assignor or any person
  claiming under him; or
\item
  (b) of any other opposing or conflicting claims to such debt or thing
  in action;he may, if he thinks fit, either call upon the persons
  making claim thereto to interplead concerning the same, or pay the
  debt or other thing in action into court under the provisions of the
  M1Trustee Act, 1925.
\end{itemize}

\end{env-99ba4262-5297-4318-9547-86307715f49a}

Assignments by way of security will always include a provision (either
explicit or implied) for reassignment on satisfaction of the debt.
Priority between assignments depends on notice being given to the third
party.

\begin{env-31409176-8a91-46bb-85fd-48f2d6a04234}

Warning

It is not possible to take a legal assignment over future choses in
action. So an assignment can never cover e.g., future debts.

\end{env-31409176-8a91-46bb-85fd-48f2d6a04234}

\hypertarget{terminology}{%
\subsection{Terminology}\label{terminology}}

\begin{itemize}
\tightlist
\item
  Debenture

  \begin{itemize}
  \tightlist
  \item
    At common law, any document which a company issues to acknowledge,
    or create, a debt.
  \item
    In banking, the document under which a borrower creates the security
    for a secured loan, usually incorporating fixed and floating charges
    along with other security interests.
  \end{itemize}
\item
  Charge

  \begin{itemize}
  \tightlist
  \item
    Appropriation of assets to the satisfaction of a debt (creating an
    equitable security interest)
  \item
    Generic label for most securities
  \end{itemize}
\item
  Security

  \begin{itemize}
  \tightlist
  \item
    Refers to the provision of rights against a person's assets which
    are intended to enhance the probability of recovering a debt claim
    against them.
  \item
    In banking, refers to the instrument (piece of paper) on which the
    borrower acknowledges the debt etc.
  \end{itemize}
\item
  Statutory definitions

  \begin{itemize}
  \tightlist
  \item
    In the LPA 1925, `mortgage' is defined to include `any charge or
    lien on any property for securing money or money's worth'.
  \item
    Under the IA 1986, `security' means `any mortgage, charge, lien or
    other security' (s 248(b)).
  \item
    In the CA 2006, the term `charge' includes a `mortgage'.
  \end{itemize}
\end{itemize}

\hypertarget{key-features}{%
\subsection{Key Features}\label{key-features}}

\begin{longtable}[]{@{}ll@{}}
\toprule()
Type of Security & Key feature \\
\midrule()
\endhead
Charge & The beneficiary takes rights over the asset. \\
Pledge & The beneficiary takes possession over the asset. \\
Mortgage & The beneficiary takes title to the asset or, alternatively
(in the case of land), rights equivalent to those enjoyed by a title
holder. \\
Common law lien & The beneficiary retains possession of the asset. \\
\bottomrule()
\end{longtable}

\hypertarget{quasi-security}{%
\section{Quasi-Security}\label{quasi-security}}

There are various ways a bank might choose to enhance its prospects of
repayment, which are not strictly security interests because they are
not ``rights in rem''.

\hypertarget{guarantees-and-comfort-letters}{%
\subsection{Guarantees and Comfort
Letters}\label{guarantees-and-comfort-letters}}

\begin{env-8872a5c8-2356-4b38-87ff-b8a967f137b3}

Obligors

The borrower, together with any third parties willing to support the
borrower's repayment obligations.

\end{env-8872a5c8-2356-4b38-87ff-b8a967f137b3}

Guarantors are also known as sureties.

\hypertarget{guarantees}{%
\subsubsection{Guarantees}\label{guarantees}}

\begin{itemize}
\tightlist
\item
  A guarantee is a form of undertaking by one party to answer for
  another party's liabilities, usually on default.
\item
  A guarantor will frequently be a company in the same group as the
  borrower.
\item
  The company which has given the guarantee may be required to give
  security over its own assets.
\item
  A parent company usually also guarantees punctual performance of the
  borrower's obligations.
\end{itemize}

\hypertarget{guarantees-vs-indemnities}{%
\subsubsection{Guarantees Vs
Indemnities}\label{guarantees-vs-indemnities}}

A guarantee creates a contractual `secondary' obligation, which relies
on there being a valid primary obligation between two parties other than
the guarantor. The guarantor's obligation is contingent on the primary
obligation. It will therefore never be greater than that of the obligor
under the underlying agreement.

An indemnity creates an obligation to indemnify a party for a specified
loss which it may incur. Unlike a guarantee, an indemnity creates a
`stand alone' (or primary) obligation which is independent of the
liability or default of another party. It is independent of, and not
contingent on, the obligations of the primary obligor. This means that,
if the underlying transaction is set aside for any reason, the indemnity
will remain valid.

It is possible for both a guarantee and an indemnity to be given at the
same time. For the lender, it is essential that a guarantee of a loan is
supported by an indemnity if the lender is to be properly protected.

\begin{env-0174d31b-3f44-410b-9ae6-7a7b66b1e278}

Action

When including both an indemnity and a guarantee in the same document,
it is very important to spell out the terms of the obligations, as the
onus will be on the lender to prove that the document is both a primary
and secondary obligation. The title of the document will not determine
whether it is a guarantee or an indemnity, or both, and neither will the
practice of referring to a guarantor and indemnifier simply as a
guarantor, even though this is widespread.

\end{env-0174d31b-3f44-410b-9ae6-7a7b66b1e278}

Assume that A has borrowed money from B, and so A owes a contractual
obligation to repay that money to B.

\begin{quote}
A guarantee is a promise by a party to ensure that another party carries
out its obligations, or a promise to fulfil those obligations itself if
the other party does not do so.
\end{quote}

\begin{quote}
An indemnity is a promise to reimburse someone in the event that they
suffer a stated loss.
\end{quote}

Suppose C indemnifies B against non-repayment of the loan to A. Then C's
obligation is a primary obligation, so C's liability might ultimately be
even more than A's was. Provides greater protection for B: it can look
to either A or C for payment.

\begin{longtable}[]{@{}ll@{}}
\toprule()
Guarantee & Indemnity \\
\midrule()
\endhead
Must be in writing or evidenced in writing and signed by the guarantor
or its agent (s 4 Statute of Frauds) & No formal requirements for a
valid indemnity \\
Constitutes a secondary liability & Constitutes a primary liability \\
Secondary nature of the liability under a guarantee means if the
underlying transaction is set aside for any reason/ underlying obligor
discharged from liability, the guarantors obligations under the
guarantee will also be unenforceable or discharged. & Primary nature of
the liability under an indemnity means if the underlying transaction is
set aside for any reason or the underlying obligor is discharged from
liability, the indemnity will remain valid. \\
Guarantor only liable to the beneficiary to the same extent as the
obligor is liable to the beneficiary & Indemnifier liable to the
beneficiary independently of the obligor, and its liability may exceed
the obligor's liability. \\
Variation of guaranteed obligations may discharge the guarantee. &
Variation of underlying obligations will not generally discharge the
indemnity. \\
Beneficiary can claim against the guarantor after the principle obligor
defaults in its obligations. & The beneficiary can claim against the
indemnifier irrespective of a default by the principal obligor. \\
\bottomrule()
\end{longtable}

\hypertarget{indemnity-advantages}{%
\subsubsection{Indemnity Advantages}\label{indemnity-advantages}}

\begin{itemize}
\tightlist
\item
  Primary obligation is more robust than a secondary obligation since it
  is not dependent on the underlying obligation.
\item
  Guarantee requires formalities (in writing and signed by the guarantor
  or a person authorised by it)
\item
  Guarantees vulnerable if any changes made to the underlying
  obligation.

  \begin{itemize}
  \tightlist
  \item
    \emph{\href{https://uk.westlaw.com/Document/IC2F31920E42711DA8FC2A0F0355337E9/View/FullText.html}{Holme
    v Brunskill (1878) 3 QBD 495}} established the rule that any
    amendments of the primary underlying contract, after the giving of
    the guarantee, will discharge the guarantor's liability under the
    guarantee unless either:

    \begin{itemize}
    \tightlist
    \item
      The guarantor consents to the variation.
    \item
      The variation is patently insubstantial or incapable of adversely
      affecting the guarantor.
    \end{itemize}
  \item
    Above rule does not apply to indemnities
    (\emph{\href{https://uk.westlaw.com/D-102-7705?originationContext=document\&transitionType=PLDocumentLink\&contextData=\%28sc.Default\%29\&comp=pluk}{GPP
    Big Field LLP v Solar EPC Solutions SL {[}2018{]} EWHC 2866
    (Comm)}})
  \end{itemize}
\end{itemize}

\hypertarget{practical-implications}{%
\subsubsection{Practical Implications}\label{practical-implications}}

\begin{itemize}
\tightlist
\item
  Substance not form is important
\item
  If A's obligations cease, then if C has provided a guarantee, C's
  obligation under the guarantee will also cease, because it is
  dependent on A's obligation
\item
  If C provides an indemnity, this would remain in place.
\item
  If there is a change to the contract between A and B after the
  guarantee/ indemnity is given, then a guarantee would almost always be
  discharged, but an indemnity would remain in force.
\item
  To execute a guarantee, it must be in writing and signed by the
  guarantor. Not necessary for an indemnity.
\end{itemize}

\hypertarget{legal-implications}{%
\subsubsection{Legal Implications}\label{legal-implications}}

\begin{enumerate}
\tightlist
\item
  Since the guarantee creates a `secondary' obligation, the guarantor
  will not be liable unless the borrower's debt becomes due.

  \begin{enumerate}
  \tightlist
  \item
    Most banks will therefore require that the `guarantee' is `on
    demand' and payable as if a `principal debt'
  \item
    So once the borrower defaults, the bank can immediately demand
    performance from the guarantor (McGuinness v Norwich and
    Peterborough Building Society {[}2010{]} EWHC 2989 (Ch)).
  \end{enumerate}
\item
  If the contract between bank and borrower (the primary contract) is
  void (or if the borrower's obligation is discharged), the guarantor's
  obligation will fall away.

  \begin{enumerate}
  \tightlist
  \item
    A bank will therefore usually include a provision that the guarantee
    survives failure of the underlying obligation
  \item
    Also an indemnity within the guarantee document in order to create a
    primary obligation on the `guarantor' if the guarantee fails.
  \end{enumerate}
\item
  If the contract between bank and borrower is varied in any material
  way without the guarantor's consent, the guarantor is released from
  the guarantee (even if the variation was not prejudicial to the
  guarantor).

  \begin{enumerate}
  \tightlist
  \item
    Similarly, certain acts which might be prejudicial to the guarantor
    (e.g., if the bank fails to register security for the debt) will
    void the guarantee.
  \item
    Triodos Bank v Ashley Charles Dobbs {[}2005{]} EWCA Civ 630: the
    court decided that in respect of a guarantee purporting to allow
    ``any amendments, variation, waiver or release from an obligation'',
    any amendment or variation must be `within the general purview of
    the original guarantee'.
  \item
    In Maxted, Lorimer v Investec Bank PLC {[}2017{]} EWHC 1997 (Ch),
    the High Court held that amendments to a loan (including an increase
    in amount and extension of the term) were within the purview of
    guarantees given by the borrowing company's directors.
  \item
    The judge in Maxted, Lorimer v Investec Bank PLC {[}2017{]} EWHC
    1997 (Ch) confirmed that if facility amendments were beyond the
    purview of the guarantees, this could be cured by the guarantors
    consenting to them (so the bank will ask guarantors to do this).
  \end{enumerate}
\end{enumerate}

Other rights -- usually waived or postponed:

\begin{enumerate}
\tightlist
\item
  Once a guaranteed debt has crystallised, the guarantor can make
  payment and claim indemnity from the obligor without waiting for a
  demand for payment of the debt. This is the case even if the guarantee
  requires payment only on demand
  (\emph{\href{https://uk.westlaw.com/D-000-5038?originationContext=document\&transitionType=PLDocumentLink\&contextData=(sc.Default)\&ppcid=9fc18fbc2d79468f9bd5272f7e22e9db}{Thomas
  v Nottingham Incorporated Football Club Ltd {[}1972{]} Ch 596}}).
\item
  A guarantor may use most rights of set-off or counterclaim which the
  borrower has against the bank to reduce its liability under guarantee.

  \begin{enumerate}
  \tightlist
  \item
    Where the underlying obligor's payment obligations are satisfied by
    way of set-off against the creditor's corresponding liabilities, the
    guarantor is entitled to that right of set-off and will therefore be
    discharged from its obligations
    (\emph{\href{https://uk.westlaw.com/D-000-4199?originationContext=document\&transitionType=PLDocumentLink\&contextData=(sc.Default)\&ppcid=9fc18fbc2d79468f9bd5272f7e22e9db}{BOC
    Group plc v Centeon LLC {[}1999{]} 1 All ER (Comm) 53}}).
  \end{enumerate}
\item
  Once paid out under the guarantee, a guarantor will be entitled to
  reimbursement from the borrower.

  \begin{enumerate}
  \tightlist
  \item
    There is an implied contract between the guarantor and the obligor
    that the obligor will provide such an indemnity
    (\emph{\href{https://uk.westlaw.com/D-000-5037?originationContext=document\&transitionType=PLDocumentLink\&contextData=(sc.Default)\&ppcid=31b1b10b11a346b28acbfa5d800fec19}{Re
    a Debtor (No.627 of 1936) {[}1937{]} Ch 156}}).
  \end{enumerate}
\item
  Guarantor entitled to any security the bank holds for the borrower's
  debt, under the doctrine of \textbf{subrogation} (equitable remedy to
  prevent unjust enrichment).

  \begin{enumerate}
  \tightlist
  \item
    Subrogation arises only once a guarantor has paid the full amount
    for which it is liable under the guarantee.
  \item
    Subrogation allows the guarantor to assume all the rights the lender
    has in relation to the borrower under the underlying contract.
  \item
    Once the guarantor has repaid the beneficiary, it is entitled to
    step into the shoes of the beneficiary and take the benefit of any
    rights of set-off and any security that the beneficiary has taken
    from the underlying obligor
    (\emph{\href{https://uk.westlaw.com/D-000-5042?originationContext=document\&transitionType=PLDocumentLink\&contextData=(sc.Default)\&ppcid=9fc18fbc2d79468f9bd5272f7e22e9db}{Craythorne
    v Swinburne (1807) 14 Ves 160}}).
  \end{enumerate}
\item
  If there is more than one guarantor of the debt, any guarantor that
  pays out will have a right to claim back a share of the payment from
  co-guarantors (``right of contribution'').
\end{enumerate}

The guarantee document should postpone the rights until such time as the
bank has been repaid in full and no longer requires the security. It
will also want any obligation of contribution to be waived in respect of
any subsidiary it wants to sell.

\hypertarget{drafting-guarantee-indemnity}{%
\subsubsection{Drafting guarantee/
Indemnity}\label{drafting-guarantee-indemnity}}

\begin{itemize}
\tightlist
\item
  If there is any uncertainty, the courts will presume that the
  interpretation that is less onerous for the giver of the document is
  correct.
\item
  When distinguishing between indemnities and guarantees, the courts
  have traditionally looked at the extent to which the promisor is
  connected with the underlying transaction.
\item
  A promisor unconnected with the underlying transaction, except by
  means of a promise to pay, has been held to be a guarantor, while a
  promise made by someone who will derive some benefit from the
  transaction was an indemnity
  (\emph{\href{https://uk.westlaw.com/D-000-4146?originationContext=document\&transitionType=PLDocumentLink\&contextData=(sc.Default)\&ppcid=31b1b10b11a346b28acbfa5d800fec19}{Couturier
  v Hastie (1852) 8 Exch 40}}).
\item
  The courts also consider the extent to which the promise can be
  construed as being the main purpose of the transaction (a guarantee)
  or incidental to the main purpose of the transaction (an indemnity)
  (\emph{\href{https://uk.westlaw.com/D-000-5044?originationContext=document\&transitionType=PLDocumentLink\&contextData=(sc.Default)\&ppcid=31b1b10b11a346b28acbfa5d800fec19}{Harburg
  India Rubber Comb Co v Martin {[}1902{]} 1 KB 778}}).
\item
  Identifying a guarantee involves making a distinction between the
  promisor having a mere motive for making the promise and having a
  ``real interest'' in (that is, deriving actual benefit from) the
  underlying transaction
  (\emph{\href{https://uk.westlaw.com/D-000-2980?originationContext=document\&transitionType=PLDocumentLink\&contextData=(sc.Default)\&ppcid=31b1b10b11a346b28acbfa5d800fec19}{Pitts
  v Jones {[}2007{]} EWCA Civ 1301}}).
\end{itemize}

\begin{env-8872a5c8-2356-4b38-87ff-b8a967f137b3}

Principle of co-extensiveness

The key feature of a guarantee that a guarantor is only liable to the
beneficiary to the same extent as the underlying obligor.

\end{env-8872a5c8-2356-4b38-87ff-b8a967f137b3}

\hypertarget{principle-debtor}{%
\paragraph{Principle Debtor}\label{principle-debtor}}

A so-called principal debtor clause states, broadly, that a guarantor
will be liable to the lender as if it were the principal obligor. This
allows the lender to treat the guarantor as a debtor in certain
circumstances.

The generally accepted view is that a principal debtor clause is
unlikely to convert a guarantee into an indemnity
(\emph{\href{https://uk.westlaw.com/D-000-1341?originationContext=document\&transitionType=PLDocumentLink\&contextData=(sc.Default)\&ppcid=31b1b10b11a346b28acbfa5d800fec19}{Carey
Value Added, S.L. v Group Urvasco, S.A. {[}2010{]} EWHC 1905}}). If the
beneficiary wants an indemnity, it should expressly include one.

\hypertarget{general-considerations}{%
\subsubsection{General Considerations}\label{general-considerations}}

\begin{itemize}
\tightlist
\item
  Governing law of the guarantee should if possible mirror that of the
  other loan documents.
\item
  If the contract is legally a guarantee, it must be evidenced in
  writing before it can be enforced (Statute of Frauds 1677, s 4).
\item
  As a contract, a guarantee is enforceable only if it is given for
  \textbf{consideration} (potentially void for past consideration if it
  is pre-dated by the obligation). So create as deed.

  \begin{itemize}
  \tightlist
  \item
    If a guarantee is executed as a deed, the guarantor will be bound
    from the date on which the deed is delivered.
  \end{itemize}
\item
  A guarantee should be expressed as `continuing' to avoid the rule in
  `Clayton's case' and to help it to survive variations to the primary
  obligation.

  \begin{itemize}
  \tightlist
  \item
    Generally, receipts from the borrower into an account will discharge
    the liabilities in the order in which they were incurred, according
    to the rule in
    \emph{\href{https://uk.westlaw.com/D-000-4111?originationContext=document\&transitionType=PLDocumentLink\&contextData=\%28sc.Default\%29\&comp=pluk}{Devaynes
    v Noble, Clayton's case (1816) 1 Mer 529}}.
  \item
    In the context of continuing guarantees, it is not necessary under
    English law to state whether the guarantor can revoke its
    obligations under the guarantee.
  \end{itemize}
\item
  The guarantee should provide that the bank's evidence of the
  borrower's liability is conclusive.
\item
  As with any other contract, a company must have the requisite power to
  enter into a guarantee, and its officers must be properly authorised.

  \begin{itemize}
  \tightlist
  \item
    Consider \textbf{s 172} duty to promote the success of the company.
  \item
    The higher the risk that a company is assuming by entering into a
    contract, the greater the commercial benefit (sometimes called
    ``corporate benefit'') obtained from the contract needs to be.

    \begin{itemize}
    \tightlist
    \item
      It is assumed that, if a parent company guarantees or indemnifies
      the obligations of a subsidiary (a ``downstream'' guarantee),
      there is sufficient commercial benefit because of the increased
      profitability of the subsidiary, higher dividends and increased
      value of the shares.
    \item
      If a subsidiary guarantees or indemnifies the obligations of a
      parent company (an ``upstream'' guarantee), this is more
      difficult.

      \begin{itemize}
      \tightlist
      \item
        Subsidiaries often rely on parent companies for treasury
        services, which can justify an upstream guarantee.
      \item
        The guarantee or indemnity provided by each subsidiary helps
        reduce the funding costs for the group as a whole.
      \end{itemize}
    \end{itemize}
  \end{itemize}
\item
  A guarantee is capable of being `financial assistance' under the
  provisions of CA 2006, ss 678--680.

  \begin{itemize}
  \tightlist
  \item
    It is unlawful for a public company whose shares are being, or have
    been, acquired (or for any of that company's subsidiaries) to give
    financial assistance directly or indirectly for the purpose of that
    acquisition unless certain exceptions apply.
  \item
    The term ``financial assistance'' includes the provision of
    guarantees, security or indemnities and any form of assistance
    whereby the net assets of the assisting company are reduced to a
    material extent or where the assisting company has no net assets.
  \item
    Penalties include the guarantee/ indemnity being void, and directors
    being liable for a fine and {\(\leq 2\)} years imprisonment.
  \end{itemize}
\end{itemize}

\begin{env-f70373ea-d099-4f9a-bab1-dd1716394de8}

Rule in Clayton's case

Technically, the rule applying to withdrawals from a current bank
account is the rule in Clayton's Case (1816) 1 Mer 529. Grant MR stated:
`it is the sum first paid in that is first drawn out.' The rule is
commonly described as `the first in, first out' rule.

\end{env-f70373ea-d099-4f9a-bab1-dd1716394de8}

\begin{env-91b82eae-616f-475e-aa24-1cbca07fe183}

Example

T misapplies £1,000 of trust fund A and pays it into a current bank
account (`the account'). T misapplies £1,000 of trust fund B and pays it
into the account. T withdraws £1,000 from the account and uses it to
purchase shares. T misapplies £1,000 of trust fund C and pays it into
the account. T withdraws £1,000 from the account and dissipates it.
£1,000 is still credited to the account.

\end{env-91b82eae-616f-475e-aa24-1cbca07fe183}

\begin{env-be598745-44b1-4815-8875-ed2abeb511dd}

Illegality

Azimut-Bennetti v Healay -- if the underlying obligation was void for
public policy (e.g. as a penalty), then the guarantee will also be void
for reasons of public policy and there is nothing that the guarantee can
say to change it.

\end{env-be598745-44b1-4815-8875-ed2abeb511dd}

\hypertarget{insolvency-act-1986}{%
\subsubsection{Insolvency Act 1986}\label{insolvency-act-1986}}

A guarantee is arguably capable of being a transaction at an undervalue
(because, unlike a borrower granting security, it usually increases the
guarantor's liabilities) and, in limited circumstances (e.g., where the
guarantor had previously only given third party security, with no
guarantee), a preference, and so being set aside under IA 1986, ss 238
and 239 respectively.

\hypertarget{counter-indemnity}{%
\subsubsection{Counter-indemnity}\label{counter-indemnity}}

There is no case law that says that payment under an indemnity will give
rise to the same right to be counter-indemnified as applies in the case
of a guarantee, but it seems likely that it does
(\emph{\href{https://uk.westlaw.com/D-009-8767?originationContext=document\&transitionType=PLDocumentLink\&contextData=(sc.Default)\&ppcid=9fc18fbc2d79468f9bd5272f7e22e9db}{Berghoff
Trading Ltd v Swinbrook Developments Ltd {[}2009{]} EWCA Civ 413}}).

However, as an indemnity is not co-extensive with the underlying
obligations, the indemnifier may be liable to pay more than the
underlying obligor. There would therefore be a question about whether
the counter-indemnity would be limited to the amount payable by the
underlying obligor. Consequently, it is not unusual, when entering into
a combined guarantee and indemnity, to insist that a counter-indemnity
agreement between the guarantor and the underlying obligor is also
entered into.

\hypertarget{comfort-letters}{%
\subsection{Comfort Letters}\label{comfort-letters}}

Suppose a parent company/ other potential obligor refuses to provide a
guarantee. The bank may demand a comfort letter (``support letter''/
``letter of intent'') from the obligor. This could be a written
acknowledgement that the subsidiary is undertaking the obligations
created by the loan, to a statement of intention to maintain an interest
in, and support for, a subsidiary, and to ensure it is capable of
fulfilling those obligations ("keep-well agreement").

Not normally intended to be legally binding. Phrasing usually
deliberately vague enough not to be enforceable under contract law.
Consider offer and acceptance, ICLR and consideration.

In the case of Re Atlantic Computers plc (in administration); National
Australia Bank Ltd v Soden {[}1995{]} BCC 696, it was held that the
common intention of the parties to a comfort letter must be ascertained
by reading it as a whole. So include wording like `an expression of
present intention by way of comfort only'.

\hypertarget{negative-pledge}{%
\subsection{Negative Pledge}\label{negative-pledge}}

Arguably a restriction on how the borrower may use its assets.
Technically not a security interest (although it might provide a claim
in tort for damages against a third party which induces a borrower to
breach the prohibition, and there is an argument that the third party
holds any such security on trust for the negative pledge holder).

\hypertarget{retention-of-title}{%
\subsection{Retention of Title}\label{retention-of-title}}

A retention of title clause allows a seller to retain legal title to
goods which have been supplied to a buyer until the buyer settles its
account. If payment is not forthcoming, the seller can recover the goods
from the buyer.

This may undermine a bank's security, so it is common for the bank to
require representations and/or warranties to the effect that the
borrower is not party to any agreements which contain a retention of
title clause.

\hypertarget{set-off}{%
\subsection{Set-off}\label{set-off}}

A right to set a debt owed by a creditor to a debtor against the
debtor's debt, and so reduce or extinguish that debt.

\begin{env-0c49cb39-e036-4df4-8974-f6a48144dfd4}

Tip

Basically the right to take money from the borrower's current account if
it defaults on its loan.

\end{env-0c49cb39-e036-4df4-8974-f6a48144dfd4}

3 ways in which a bank is likely to have a right of set-off against the
borrower:

\begin{longtable}[]{@{}ll@{}}
\toprule()
Means & Details \\
\midrule()
\endhead
Equitable right of set-off & A bank may set off a liquidated debt owed
to it by a borrower against a liquidated debt owed by it to that
borrower. \\
Contractual set-off & Most security documents supporting a loan will
enhance the equitable right of set-off by allowing the bank to set off
unliquidated (contingent) claims. \\
Set-off on liquidation & r 490 Insolvency Rules 1986 provides for
mandatory set-off. Basically, any `mutual' credits, debts or dealings
can be set off so long as the bank did not know of the petition for
insolvency against the borrower when it gave credit. Set off of
contingent claims also allowed. \\
\bottomrule()
\end{longtable}

\begin{env-486fc4cd-eb32-4215-aaa2-fc77a67eebc5}

Note

The right of set-off is particularly important if a debtor is in danger
of becoming insolvent (but no petition has been lodged), since it allows
access to the debtor's unsecured assets (ie, cash) without risk of being
a transaction at an undervalue or a preference under IA 1986, ss 238 and
239.

\end{env-486fc4cd-eb32-4215-aaa2-fc77a67eebc5}

\hypertarget{bankers-right-of-combination}{%
\subsubsection{Banker's Right of
Combination}\label{bankers-right-of-combination}}

There is an implied right under common law which is peculiar to banks
and which allows them, subject to contrary agreement, to combine two or
more accounts of a borrower. So an overdrawn account and an account in
credit can be combined.

\hypertarget{why-take-security}{%
\subsection{Why Take Security}\label{why-take-security}}

Aside from increasing the likelihood of getting repaid, there are other
reasons for taking security:

\hypertarget{appointing-administrator-receiver}{%
\subsubsection{Appointing administrator/
Receiver}\label{appointing-administrator-receiver}}

If a bank wants to enforce security it will invariably do so through a
third party, and usually when the borrower is, or is about to become,
insolvent.

\hypertarget{administration-pre-ea-2002}{%
\paragraph{Administration pre-EA
2002}\label{administration-pre-ea-2002}}

The administration procedure was introduced by the IA 1986. For an
administrator, the first, and key, purpose was to ensure `the survival
of the company, and the whole or any part of its undertaking, as a going
concern'. The administrator was given power to effect a moratorium on
the rights of all secured and unsecured creditors, to help further this
purpose.

Banks didn't like this, so successfully lobbied for the right to block
an administration order if they held security (which included a floating
charge) over all, or substantially all, of a company's assets. This was
achieved by appointing an administrative receiver under IA 1986 (s 9).

\hypertarget{administrative-receiver-pre-ea-2002}{%
\paragraph{Administrative Receiver pre-EA
2002}\label{administrative-receiver-pre-ea-2002}}

\begin{env-99ba4262-5297-4318-9547-86307715f49a}

s 29(2) IA 1986

``Administrative receiver'' means---

\begin{itemize}
\tightlist
\item
  (a) a receiver or manager of the whole (or substantially the whole) of
  a company's property appointed by or on behalf of the holders of any
  debentures of the company secured by a charge which, as created, was a
  floating charge, or by such a charge and one or more other securities;
  or
\item
  (b) a person who would be such a receiver or manager but for the
  appointment of some other person as the receiver of part of the
  company's property.
\end{itemize}

\end{env-99ba4262-5297-4318-9547-86307715f49a}

The term `debenture' `includes \ldots{} bonds and any other securities
of a company \ldots'. So long as a bank held security (which included a
floating charge) over `all or substantially all' of a borrower's assets,
it could appoint an administrative receiver, with wide-ranging powers to
manage the business and sell assets to repay the bank debt and,
crucially, the power to block an administration order and moratorium.

\hypertarget{administration-post-ea-2002}{%
\paragraph{Administration post-EA
2002}\label{administration-post-ea-2002}}

New purposes of administration (cascading downwards):

\begin{enumerate}
\tightlist
\item
  attempt to rescue the company as a going concern -- if this is not
  reasonably practical, or would not produce the best result for the
  creditors as a whole;
\item
  attempt to achieve a better result for creditors than if the company
  were wound up; and (only if the administrator thinks the first two
  purposes are not reasonably practical)
\item
  realise the company's property to distribute to secured or
  preferential creditors, so long as the administrator does not
  unnecessarily harm the interests of the creditors as a whole.
\end{enumerate}

These apply regardless of who appoints the administrator. Administrator
can be appointed:

\begin{enumerate}
\tightlist
\item
  By court order
\item
  By Qualifying Floating Charge holder. Floating charge document must
  state that EA 2002 applies to it/ purport to allow the holder to
  appoint an administrator or administrative receiver.
\end{enumerate}

\hypertarget{administrative-receivership-post-ea-2002}{%
\paragraph{Administrative Receivership post-EA
2002}\label{administrative-receivership-post-ea-2002}}

In addition to any floating charges created before 15 September 2003,
there are eight `exceptions' under which an administrative receiver may
still be appointed:

\begin{enumerate}
\tightlist
\item
  `capital market' exception,
\item
  `public--private partnership' exception,
\item
  `utility projects' exception,
\item
  `urban regeneration projects' exception,
\item
  `project finance' exception,
\item
  `financial markets' exception,
\item
  `protected railways and other companies' exception and
\item
  `registered social landlord' exception.
\end{enumerate}

Applies only to specialist financing.

\begin{env-a9aa42a3-f5ef-44bc-9cb3-81fc5caedba7}

Summary

Where the bank holds a `qualifying floating charge' created after 15
September 2003, but does not fall within one of the eight statutory
exceptions, it will be permitted to appoint an administrator only to
realise security on insolvency. The bank's administrator must consider
the two prior purposes of administration before realising assets on the
bank's behalf.

\end{env-a9aa42a3-f5ef-44bc-9cb3-81fc5caedba7}

\hypertarget{peripheral-benefits}{%
\subsubsection{Peripheral Benefits}\label{peripheral-benefits}}

\begin{itemize}
\tightlist
\item
  Taking fixed security will help to prevent a borrower disposing of
  assets and ensure they remain in the business to create value,
  enhancing disposal undertakings.
\item
  It may deter other banks from lending to the borrower, enhancing the
  negative pledge.
\item
  It may also dissuade unsecured creditors from taking action to wind up
  the borrower.
\item
  Personal security from directors of a small company may act as an
  incentive for them to work to make the business prosper.
\end{itemize}

\hypertarget{costs}{%
\subsubsection{Costs}\label{costs}}

Secured loans will carry less risk for a bank than unsecured loans, and
compliance with regulatory capital requirements will make some types of
secured loan `cheaper' than unsecured loans.

\hypertarget{banks-perspective}{%
\subsection{Bank's Perspective}\label{banks-perspective}}

\begin{itemize}
\tightlist
\item
  Taking security is expensive and time-consuming. Investment grade
  loans are unsecured; highly leveraged loans almost always secured.
\item
  A bank's solicitor should check that the borrower and any other party
  providing security has capacity to do so (there may be restrictions in
  the articles).
\item
  Check there is no existing negative pledge.
\end{itemize}

\hypertarget{borrowers-perspective}{%
\subsection{Borrower's Perspective}\label{borrowers-perspective}}

\begin{itemize}
\tightlist
\item
  Usually liable for all reasonable costs of providing security,
  including solicitors' fees and registration costs.
\item
  Two concessions frequently negotiated:

  \begin{enumerate}
  \tightlist
  \item
    In acquisition financings, only share security is provided at
    closing (when the loan is available). Borrower given a period of
    time (up to 120 days) to provide remaining asset security.
  \item
    Common to agree security principles early in negotiation --
    pre-agreed parameters for security, costs and any obligations on the
    borrower.
  \end{enumerate}
\end{itemize}

\hypertarget{taking-security}{%
\section{Taking Security}\label{taking-security}}

3 steps to creating a valid security interest:

\begin{enumerate}
\tightlist
\item
  Creation -- agreement to create a particular security interest between
  debtor and creditor.
\item
  Attachment -- ensuring the security interest binds on specific assets
\item
  Perfection -- ensuring validity against third parties and priority.
\end{enumerate}

Perfection and priority are here examined.

\hypertarget{registration}{%
\subsection{Registration}\label{registration}}

The safest way to secure an asset is to take possession of it (take a
``pledge''). But most loan security is non-possessory. To stop a secured
asset being sold on/ re-secured, there is a statutory requirement for
the registration of non-possessory security interests.

\hypertarget{ca-regime}{%
\subsubsection{CA Regime}\label{ca-regime}}

The requirement for companies to register certain security interests at
Companies House is long-standing. The primary aim is a mechanism to give
third parties notice of the security and thereby make it valid against
their claims.

The registration regime changed on 06/04/13. It used to be compulsory to
register most charges, now it is voluntary. But if the charge is not
registered in time, there are consequences (e.g., may become void).

\hypertarget{what-to-register}{%
\paragraph{What to Register}\label{what-to-register}}

The registration requirement applies to `charges' created by companies
and LLPs. The term `charge' is defined to include mortgages (s 859A(7))
and security assignments. By implication, registration is not required
for pledges or liens for which actual or constructive possession of the
secured asset by the security holder operates as notice to third
parties.

Guarantees don't need registering, mortgages and fixed/ floating charges
usually do. For other stuff (loans, pledges), solicitors will usually
register out of an abundance of caution.

All charges (remember this includes mortgages) may be registered, with
the following three (minor) exceptions (s 859A(6)):

\begin{enumerate}
\tightlist
\item
  a charge in favour of a landlord on a cash deposit given as security
  in connection with the lease of land;
\item
  a charge created by a member of Lloyd's to secure its obligations in
  connection with its underwriting business;
\item
  a charge excluded by any other Act.
\end{enumerate}

\begin{env-99ba4262-5297-4318-9547-86307715f49a}

s 859A CA 2006 - Charges created by a company

(1) Subject to subsection (6), this section applies where a company
creates a charge.

(2) The registrar must register the charge if, before the end of the
period allowed for delivery, the company or any person interested in the
charge delivers to the registrar for registration a section 859D
statement of particulars.

(3) Where the charge is created or evidenced by an instrument, the
registrar is required to register it only if a certified copy of the
instrument is delivered to the registrar with the statement of
particulars.

(4) ``The period allowed for delivery'' is 21 days beginning with the
day after the date of creation of the charge (see section 859E), unless
an order allowing an extended period is made under section 859F(3).

(5) Where an order is made under section 859F(3) a copy of the order
must be delivered to the registrar with the statement of particulars.

(6) This section does not apply to---

\begin{itemize}
\tightlist
\item
  (a) a charge in favour of a landlord on a cash deposit given as a
  security in connection with the lease of land;
\item
  (b) a charge created by a member of Lloyd's (within the meaning of the
  Lloyd's Act 1982 F2) to secure its obligations in connection with its
  underwriting business at Lloyd's;
\item
  (c) a charge excluded from the application of this section by or under
  any other Act.
\end{itemize}

(7) In this Part---

``cash'' includes foreign currency,

``charge'' includes---

\begin{itemize}
\tightlist
\item
  (a) a mortgage;
\item
  (b) a standard security, assignation in security, and any other right
  in security constituted under the law of Scotland, including any
  heritable security, but not including a pledge, and
\end{itemize}

``company'' means a UK-registered company.

\end{env-99ba4262-5297-4318-9547-86307715f49a}

\begin{env-31409176-8a91-46bb-85fd-48f2d6a04234}

Warning

Charges over shares~are, strictly speaking,~exempt from registration
under the Companies Act 2006 because of the~effect of the Financial
Collateral Arrangements (No. 2) Regulations 2003 (as amended). However,
in practice, security over shares is~usually registered, even when
registration is not strictly required, because often the share security
document covers more than just the shares, thereby giving rise to the
concern that payments on the shares (e.g. dividends) may be book debts
and, therefore, the security may be registrable.

\end{env-31409176-8a91-46bb-85fd-48f2d6a04234}

\hypertarget{who-registers}{%
\paragraph{Who Registers}\label{who-registers}}

Either the company or ``any person interested in the charge'' (s
859A(2)). This includes the charge-holder as well as the legal advisors
to the company or charge-holder. In practice, the charge-holder register
the charge.

\hypertarget{requirements}{%
\paragraph{Requirements}\label{requirements}}

\begin{env-0174d31b-3f44-410b-9ae6-7a7b66b1e278}

Action

Once agreement has been entered into, the company/ any person interested
in the charge should deliver (s 859A(2)):

\begin{enumerate}
\tightlist
\item
  A s 859D statement of particulars to Companies House (s 859A(2))

  \begin{itemize}
  \tightlist
  \item
    This will usually be Form MR01 for a company (or MR08 where there is
    no charge instrument) or the equivalent forms LLMR01 and LLMR08 for
    a limited liability partnership.
  \end{itemize}
\item
  A certified copy of the instrument creating the charge (s 859A(3))
\item
  The requisite fee
\end{enumerate}

\end{env-0174d31b-3f44-410b-9ae6-7a7b66b1e278}

s 859G allows certain details to be obscured in the certified copy of
the charge instrument (known as `redaction'), such as

\begin{itemize}
\tightlist
\item
  Personal information relating to an individual
\item
  Bank account number of a person/ individual
\item
  Signature.
\end{itemize}

The paragraph certifying it as a true copy of the original should state
`save for material redacted pursuant to s 859G'.

\hypertarget{time-limit-for-registration}{%
\paragraph{Time Limit for
Registration}\label{time-limit-for-registration}}

The registration documents must be delivered \textbf{before the end of
21 days beginning with the day after the date of creation of the charge}
(s 859A(4)).

If there is a defect with the submission, it must be corrected within
the 21-day time limit for submission. whilst there is provision at s
859F for the company (or a person interested) to apply to court for an
extension to the 21-day period, there is no guarantee it will be
granted.

\hypertarget{late-inaccurate-delivery}{%
\paragraph{Late/ Inaccurate Delivery}\label{late-inaccurate-delivery}}

Under \href{https://www.legislation.gov.uk/ukpga/2006/46/section/859F}{s
859F CA 2006}, the court has the power to extend the period for
registration where the grounds of s 859F(2) are met:

\begin{quote}
(a) that the failure to deliver those documents-\\
(i) was accidental or due to inadvertence or to some other sufficient
cause, or\\
(ii) is not of a nature to prejudice the position of creditors or
shareholders of the company, or\\
(b) that on other grounds it is just and equitable to grant relief.
\end{quote}

The court will allow the register to be rectified provided this does not
prejudice any other charges created between the date of creation of the
unregistered charge and the date of eventual registration (Barclays Bank
plc v Stuart Landon Ltd {[}2001{]} 2 BCLC 316).

If an application is successful, the charge will have priority only from
the date of actual registration, so may lose priority due to the delay
if other charges have been registered in the meantime.

The court has also refused to allow a charge to be registered late,
where the time elapsed is too long (Victoria Housing Estates Ltd v
Ashpurton Estates Ltd {[}1982{]}).

The court has power:

\begin{itemize}
\tightlist
\item
  Under s 859M to allow rectification of a statement/ notice delivered
  to the Registrar for inaccurate details.
\item
  Under s 859N if the charging document was defective/ wrong document
  was sent, can be replaced.
\end{itemize}

\hypertarget{effect-of-valid-registration}{%
\paragraph{Effect of Valid
Registration}\label{effect-of-valid-registration}}

A validly registered charge is valid against a liquidator, administrator
and any creditor of the company that created the charge. But it will not
prevent the terms of a charge being challenged, challenge as a
preference, transaction at an undervalue or similar.

The form MR01 also has a `tick box' for noting whether the terms of the
charge include a negative pledge. The intention is to put anyone who
searches the register on notice of the negative pledge.

\hypertarget{effect-of-invalid-registration}{%
\paragraph{Effect of Invalid
Registration}\label{effect-of-invalid-registration}}

Failure to register a charge within the 21-day period means (s 859H):

\begin{enumerate}
\tightlist
\item
  Security is void against a liquidator, administrator and any creditor
\item
  The amount secured by the void security becomes immediately payable.
\end{enumerate}

The charge remains valid against the company creating it unless and
until that company goes into liquidation or administration. However, the
money secured by the charge becomes payable immediately (CA 2006, s
859H(4)).

\hypertarget{documents}{%
\paragraph{Documents}\label{documents}}

The company must keep the following documents available at its
registered office (s 859P) or SAIL (s 859Q(2)):

\begin{enumerate}
\tightlist
\item
  a copy of every instrument creating a charge capable of registration
  under the CA 2006;
\item
  a copy of every instrument that amends or varies a charge capable of
  registration under the CA 2006;
\item
  a copy of any documents containing particulars required for
  registration if those particulars are not in the charge instrument;
\item
  a copy of any translation required as part of the registration (see CA
  2006, s 1105).
\end{enumerate}

The documents may be certified copies (other than the translation) and
be available without charge to a creditor or shareholder, and to any
other person on payment of a prescribed fee.

\hypertarget{security-release}{%
\paragraph{Security Release}\label{security-release}}

Some security documents may purport to allow a bank to retain the
security for a period even after repayment as `insurance' against the
repayment being clawed back under the IA 1986, or may deem the debt and
security to continue after repayment if the security agent considers
clawback a possibility. The efficacy of such provision is questionable.

When the loan is repaid by the company to the lender, a person with an
interest in the registration of the charge (such as a director of the
company) may, but is not obliged to, complete, sign and send
\textbf{Form MR04} to the Registrar of Companies at Companies House.
This is usually done to ensure the company's file is up-to-date.

Security might also be released to allow for disposal of a particular
asset. Release is usually achieved by a deed of release drawn up by the
borrower's solicitor. If the lender decides to release the company's
property from the charge or allows the company to sell the asset covered
by the charge, a person with an interest in the registration of the
charge (such as a director of the company) must complete, sign and send
\textbf{Form MR05} to the Registrar of Companies at Companies House.

\hypertarget{property-already-subject-to-charge}{%
\paragraph{Property Already Subject to
Charge}\label{property-already-subject-to-charge}}

The current regime allows a company (or person interested) to register a
charge which already exists on property it acquires (assuming the charge
would have been capable of registration if the company had created it).
There is no time limit, and non-registration is not an offence.

\hypertarget{overseas-companies}{%
\paragraph{Overseas Companies}\label{overseas-companies}}

Registration of charges applies to all charges created by UK companies/
LLPs, even if the secured assets are overseas.

\hypertarget{asset-specific-registration-requirements}{%
\subsubsection{Asset-specific Registration
Requirements}\label{asset-specific-registration-requirements}}

\begin{longtable}[]{@{}ll@{}}
\toprule()
Asset & Registration \\
\midrule()
\endhead
Land & Registered land registered at the District Land Registry. \\
Ships & Registered at ship's port of registration (Merchant Shipping Act
1894) \\
Aircraft & Aircraft Mortgage Register \\
IP & Various statutory provisions. \\
\bottomrule()
\end{longtable}

\hypertarget{security-created-by-individuals}{%
\subsubsection{Security Created by
Individuals}\label{security-created-by-individuals}}

The Bills of Sale Acts 1878 and 1882 impose registration requirements on
security created by individuals.

\begin{env-486fc4cd-eb32-4215-aaa2-fc77a67eebc5}

Note

The Acts require most written, non-possessory charges given by an
individual or a partnership over tangible property to be registered in
prescribed form at the Supreme Court as a bill of sale.

\end{env-486fc4cd-eb32-4215-aaa2-fc77a67eebc5}

IA 1986, s 344 requires most forms of assignment of a trader's book
debts to be registered as a bill of sale if they are to be valid against
a trustee in bankruptcy.

\hypertarget{priority-of-security}{%
\subsection{Priority of Security}\label{priority-of-security}}

\hypertarget{priority-between-charges-and-mortgages}{%
\subsubsection{Priority Between Charges and
Mortgages}\label{priority-between-charges-and-mortgages}}

\begin{itemize}
\tightlist
\item
  Fixed charges and mortgages rank in order of creation (if properly
  registered).
\item
  At common law, a legal mortgage takes priority over an existing
  equitable security interest only if the legal interest was taken bona
  fide, for value and without notice of the prior equitable security.
\item
  Registration gives notice to anyone who has, or ought to have,
  searched the register, which will usually protect a prior registered
  equitable security.
\end{itemize}

\hypertarget{fixed-and-floating-charges}{%
\subsubsection{Fixed and Floating
Charges}\label{fixed-and-floating-charges}}

\begin{itemize}
\tightlist
\item
  Fixed charge will rank in priority to a floating charge
\item
  Note the floating charge could contain a negative pledge
  {\(\rightarrow\)} then if the later fixed charge holder has actual/
  constructive notice of the prohibition through registration, the
  floating charge would rank first.
\item
  Priority between floating charges governed by their time of creation.
\item
  The date of the charge from which priority can be calculated is the
  date of creation, not the date of registration

  \begin{itemize}
  \tightlist
  \item
    A search made on the day of registering a charge might not show a
    charge created within 21 days which could take priority.
  \end{itemize}
\end{itemize}

\hypertarget{security-over-land}{%
\subsubsection{Security Over Land}\label{security-over-land}}

\begin{itemize}
\tightlist
\item
  Under the registered land system, the bank proposing to take a legal
  security must complete a ``priority search'' at the Land Registry.
\item
  This shows any existing encumbrances over the registered land, as well
  as actual/ proposed dealings with land.
\item
  If no other encumbrances/ dealings/ priority searches are revealed,
  the bank has 30 working days within which to register the new legal
  mortgage and take priority.
\item
  Registration then ensures priority even over an earlier dated
  unregistered legal mortgage without the benefit of a priority search.
\end{itemize}

\hypertarget{dearle-v-hall}{%
\subsubsection{\texorpdfstring{\emph{Dearle V
Hall}}{Dearle V Hall}}\label{dearle-v-hall}}

Dearle v Hall (1828) 3 Russ 1 establishes a rule of priority applicable
to all assignments or charges of equitable interests and any interest in
debts.

\begin{env-0174d31b-3f44-410b-9ae6-7a7b66b1e278}

Action

To ensure priority, notice of the charge or assignment must be given to
the third party which owes the obligation.

\end{env-0174d31b-3f44-410b-9ae6-7a7b66b1e278}

The same rule applies to charges. The rule will not apply, however,
where a subsequent assignee (or chargee) knows of the earlier assignment
when taking its assignment. Thus, registration of the security at the
Companies Registry will usually defeat the rule.

\hypertarget{contractual-subordination}{%
\subsection{Contractual Subordination}\label{contractual-subordination}}

\hypertarget{contractual-priority-arrangements}{%
\subsubsection{Contractual Priority
Arrangements}\label{contractual-priority-arrangements}}

Possible for creditors to decide between themselves the order in which
they will benefit from claims against a defaulting debtor. Priority
arrangements frequently used in large financings with layers of debt.

Document creating rights known as an ``intercreditor agreement''/
``subordination agreement''/ ``deed of priorities''.

\hypertarget{typical-arrangement}{%
\subsubsection{Typical Arrangement}\label{typical-arrangement}}

Participants restricted so that:

\begin{itemize}
\tightlist
\item
  Participating lenders divided into an agreed order of priority
  (tranches) for repayment of their credit.

  \begin{itemize}
  \tightlist
  \item
    Senior creditors
  \item
    Mezzanine
  \item
    Loan notes
  \item
    High yield.
  \end{itemize}
\item
  No participating bank may enforce security or otherwise proceed
  against the borrower without the agreement of senior banks/ majority
  consent
\item
  When security is enforced:

  \begin{itemize}
  \tightlist
  \item
    Subordinated creditors contractually agree not to claim against the
    borrower until the senior creditors are repaid (``contingent debt
    subordination'').
  \item
    Subordinated creditors agree to hold any money they receive on trust
    for senior creditors (known as a `turnover trust').

    \begin{itemize}
    \tightlist
    \item
      More robust than contingent debt subordination.
    \end{itemize}
  \end{itemize}
\item
  The pooled money will then be distributed in accordance with the
  agreement.
\end{itemize}

\begin{env-a9aa42a3-f5ef-44bc-9cb3-81fc5caedba7}

Does a turnover trust create a proprietory interest over the debt which
requires registration as a charge under CA 2006?

SSSL Realisations Limited {[}2004{]} EWHC (Ch) 1760 held that the
turnover trust in that case did not create a charge: it did create a
proprietory right, but since it was limited to the debt owed there was
no charge over the `junior' creditor's property.

Some practitioners register anyway just to be sure.

\end{env-a9aa42a3-f5ef-44bc-9cb3-81fc5caedba7}

There is also \textbf{structural subordination}: ordinary shareholders
automatically subordinated to debt-holding creditors because debts must
be repaid before capital can be realised.

Subordination between shareholders through preference shares.

\hypertarget{when-is-contractual-subordination-used}{%
\subsubsection{When is Contractual Subordination
Used}\label{when-is-contractual-subordination-used}}

\begin{itemize}
\tightlist
\item
  {\(\uparrow\)} risk {\(\;\Longrightarrow\;\)} {\(\uparrow\)} margin.
  All banks want a diversified spread of risk.
\item
  Contractual subordination might also be used where an unsecured bank
  requires any debts to a borrower's directors or parent company to be
  subordinated to its own.
\item
  Existing bank may agree to subordination to a new bank to secure new
  funding to help the business.
\end{itemize}

\hypertarget{tacking}{%
\subsection{Tacking}\label{tacking}}

\begin{env-8872a5c8-2356-4b38-87ff-b8a967f137b3}

Definition

Tacking is the ability to secure new advances under existing security.

\end{env-8872a5c8-2356-4b38-87ff-b8a967f137b3}

Depends on the asset:

\hypertarget{registered-land}{%
\subsubsection{Registered Land}\label{registered-land}}

LRA 2002, s 49.: a mortgage/ charge holder can tack if:

\begin{enumerate}
\tightlist
\item
  it has not received notice of the subsequent security; or
\item
  the new advance is under an obligation already on the register when
  the subsequent charge was created; or
\item
  the maximum amount of the first charge was agreed and on the register
  when the subsequent charge was created.
\end{enumerate}

\hypertarget{personalty}{%
\subsubsection{Personalty}\label{personalty}}

Mortgaged/charged personalty is subject to the provisions of LPA 1925, s
94. Any actual or constructive notice of a subsequent mortgage/charge
will end a prior mortgagee's/ chargee's right to tack unless:

\begin{itemize}
\tightlist
\item
  an arrangement has been made to that effect with the subsequent
  mortgagees; or
\item
  where the mortgage imposes an obligation on the first mortgagee to
  make further advances.
\end{itemize}

\hypertarget{ruling-off-accounts}{%
\subsection{Ruling Off Accounts}\label{ruling-off-accounts}}

When a borrower makes a payment to a bank, the borrower may instruct the
bank as to which account it should be appropriated or which loan should
be repaid.

In the absence of any specific agreement, the rule in Clayton's Case
(1816) 1 Mer 529 (Devaynes v Noble (1816) 1 Mer 572) applies. Money paid
to the bank will discharge the borrower's debt in the chronological
order in which the debt was incurred. But this might disadvantage the
bank:

\begin{enumerate}
\tightlist
\item
  May repay a secured loan rather than an unsecured loan, leaving the
  bank in a more precarious position than if the unsecured loan was
  repaid first
\item
  New security may be more vulnerable to be set aside than older
  ``hardened'' security.
\item
  In certain circumstances, new advances would rank behind a second loan
  to another lender. So a repaid and redrawn RCF may lose priority to a
  new loan.
\end{enumerate}

\begin{itemize}
\tightlist
\item
  To avoid (1) and (2), the bank may insist on choosing where payments
  are appropriated (if not received in full).
\item
  (3) dealt with by allowing a bank to ``rule off'' an account if it
  learns the borrower has taken a second secured loan.

  \begin{itemize}
  \tightlist
  \item
    Original secured loan is preserved ("ruled-off") and not treated as
    repaid by any amounts received from the borrower.
  \item
    Security should be specified as securing all amounts due from time
    to time from the borrower to the bank.
  \item
    Insert negative pledge clauses.
  \end{itemize}
\end{itemize}

\hypertarget{clayton-guarantees}{%
\subsubsection{Clayton \& Guarantees}\label{clayton-guarantees}}

Subject to contrary agreement, once a guaranteed liability has been
fixed in value, any payment by the borrower to the bank will be deemed
to reduce the guaranteed debt.

\begin{env-0174d31b-3f44-410b-9ae6-7a7b66b1e278}

Action

Can be avoided by specifying the guarantee to be `continuing', or
alternatively an `all monies guarantee'.

\end{env-0174d31b-3f44-410b-9ae6-7a7b66b1e278}

\hypertarget{fixed-or-floating}{%
\subsection{Fixed or Floating}\label{fixed-or-floating}}

Floating charge advantages (for the bank):

\begin{itemize}
\tightlist
\item
  Allows a borrower to deal freely with the charged assets in the
  ordinary course of business
\item
  May allow the bank to appoint an administrative receiver

  \begin{itemize}
  \tightlist
  \item
    After changes made by the EA 2002, this is relevant only to
    `specialised' financial structures.
  \end{itemize}
\item
  The holder of a `qualifying floating charge' (as defined by IA 1986,
  Sch B1, para 14) can appoint an administrator using an out-of-court
  procedure.
\end{itemize}

\begin{env-8872a5c8-2356-4b38-87ff-b8a967f137b3}

``Lightweight''/ ``Phantom'' floating charge

Charge taken primarily for the purpose of allowing the bank to appoint
an administrative receiver.

\end{env-8872a5c8-2356-4b38-87ff-b8a967f137b3}

Floating charge disadvantages:

\begin{itemize}
\tightlist
\item
  On insolvency, a floating charge ranks behind any fixed security/
  preferential creditors.
\item
  Floating charge may be vulnerable to later fixed charges, even if
  containing a registered negative pledge.
\item
  Can be set aside under wide circumstances
\item
  Some of the proceeds available allocated to the prescribed part fund
\item
  Ranks behind liquidator's expenses
\item
  Borrower free to deal with assets until crystallisation.
\end{itemize}

See also Common Securities.

\hypertarget{hardening-periods}{%
\subsection{Hardening Periods}\label{hardening-periods}}

\begin{env-8872a5c8-2356-4b38-87ff-b8a967f137b3}

Hardening period

A period during which the security is vulnerable to provision in IA 1986
for setting aside certain transactions.

\end{env-8872a5c8-2356-4b38-87ff-b8a967f137b3}

Possible grounds for setting aside security:

\begin{itemize}
\tightlist
\item
  Transaction at an undervalue (s 238 IA 1986)
\item
  Preferences (s 239 IA 1986)
\item
  Floating charges (s 245 IA 1986)
\end{itemize}

See Business Law and Practice/Insolvency/Voidable Transactions.

\hypertarget{capacity}{%
\subsection{Capacity}\label{capacity}}

Before lending, check the capacity of the borrower to provide security.

The powers and limitations on trustees will be found in the trust deed.

\begin{env-99ba4262-5297-4318-9547-86307715f49a}

s 16 TA 1925 - Power to raise money by sale, mortgage, \&c.

(1) Where trustees are authorised by the instrument, if any, creating
the trust or by law to pay or apply capital money subject to the trust
for any purpose or in any manner, they shall have and shall be deemed
always to have had power to raise the money required by sale,
conversion, calling in, or mortgage of all or any part of the trust
property for the time being in possession.

(2) This section applies notwithstanding anything to the contrary
contained in the instrument, if any, creating the trust, but does not
apply to trustees of property held for charitable purposes, or to
trustees of a settlement for the purposes of the Settled Land Act, 1925,
not being also the statutory owners.

\end{env-99ba4262-5297-4318-9547-86307715f49a}

So trustees are granted the power to grant security provided that
borrowing is authorised and for the purpose of the trust, regardless of
what the trust deed says.

\hypertarget{corporate-benefit}{%
\subsection{Corporate Benefit}\label{corporate-benefit}}

A bank often requires additional security from a borrower's parent,
subsidiary, or sister company. The company providing the guarantee/
indemnity/ security must prove it derives sufficient ``corporate
benefit'' from granting the security. Consider s 172 duty to promote the
success of the company.

A subsidiary securing its parent (e.g., an `up-stream' guarantee) could
argue that the support from its parent (e.g., financial, marketing,
product development) provides sufficient corporate benefit. This
argument is more tenuous if the company is a small fish in a large group
pond, but can be strengthened by a down-stream support letter.

If a company believes there is insufficient corporate benefit to provide
security, there is authority to suggest that it may still be given,
provided that:

\begin{enumerate}
\tightlist
\item
  the company's shareholders unanimously direct the security to be
  given; and
\item
  the company is not insolvent either at the time of giving the
  security, or immediately following the giving of the security.
\end{enumerate}

\hypertarget{financial-assistance}{%
\subsection{Financial Assistance}\label{financial-assistance}}

\hypertarget{definition}{%
\subsubsection{Definition}\label{definition}}

Refers to a company providing financial assistance for the purchase of
its own shares. This includes financial assistance given by way of a
gift, loan, guarantee, security or indemnity
(\href{https://www.legislation.gov.uk/ukpga/2006/46/section/677}{s 677
CA 2006}).

Financial assistance was defined in British and Commonwealth Holding PLC
v Barclays Bank plc {[}1996{]} 1 WLR 1 as

\begin{quote}
``The section requires that there should be assistance or help for the
purpose of acquiring the shares and that that assistance should be
financial''.
\end{quote}

Financial assistance was originally made unlawful to counter a practice
whereby individuals were able to purchase cash rich companies by
arranging for the target company to advance the cash to finance the
acquisition. In effect, the company paid for its own acquisition.

\hypertarget{prohibitions}{%
\subsubsection{Prohibitions}\label{prohibitions}}

CA 2006 \textbf{removed} the prohibition in relation to private
companies, so now prohibition of financial assistance \textbf{only}
applies in general to public companies.

\begin{itemize}
\tightlist
\item
  Both public companies and their private limited subsidiaries are
  prohibited from providing financial assistance for the purchase of
  shares in the public company
  (\href{https://www.legislation.gov.uk/ukpga/2006/46/section/678}{s 678
  CA 2006}).
\item
  Public companies are also prohibited from providing financial
  assistance for the purchase of shares in their private limited
  \textbf{holding} companies
  (\href{https://www.legislation.gov.uk/ukpga/2006/46/section/679}{s 679
  CA 2006}).
\end{itemize}

The consequences of unlawful financial assistance are that the
transaction will be held void and the company, and any officer in
default, will be liable to a fine/ up to two years in prison.

\begin{env-be598745-44b1-4815-8875-ed2abeb511dd}

Important

The rules on financial assistance are relevant in the context of share
purchases and do not apply where the buyer is acquiring the target
company's assets.

\end{env-be598745-44b1-4815-8875-ed2abeb511dd}

\hypertarget{purpose}{%
\paragraph{Purpose}\label{purpose}}

Financial assistance will only be prohibited by
\emph{\href{https://uk.westlaw.com/5-505-7258?originationContext=document\&transitionType=PLDocumentLink\&contextData=(sc.Default)\&ppcid=52c063520c5f48d8a198d13504dc80c1}{section
678}} or
\emph{\href{https://uk.westlaw.com/6-506-2056?originationContext=document\&transitionType=PLDocumentLink\&contextData=(sc.Default)\&ppcid=52c063520c5f48d8a198d13504dc80c1}{679}}
if it is given ``for the purpose of'' the acquisition of shares or to
reduce or discharge previous liability in relation to an acquisition.
The company must have \textbf{intended} to facilitate the acquisition by
giving the assistance.

\hypertarget{time}{%
\paragraph{Time}\label{time}}

The prohibition on financial assistance in
\emph{\href{https://uk.westlaw.com/5-505-7258?originationContext=document\&transitionType=PLDocumentLink\&contextData=(sc.Default)\&ppcid=52c063520c5f48d8a198d13504dc80c1}{section
678}} and
\emph{\href{https://uk.westlaw.com/6-506-2056?originationContext=document\&transitionType=PLDocumentLink\&contextData=(sc.Default)\&ppcid=52c063520c5f48d8a198d13504dc80c1}{section
679}} is not limited in time.

\hypertarget{re-registration}{%
\paragraph{Re-registration}\label{re-registration}}

Public companies may be able to re-register as private to give financial
assistance for the acquisition of their own shares. Where a public
company has re-registered as a private company after the shares have
been acquired and it is a private company at the time the
post-acquisition assistance is given, the
\emph{\href{https://uk.westlaw.com/5-505-7258?originationContext=document\&transitionType=PLDocumentLink\&contextData=(sc.Default)\&ppcid=52c063520c5f48d8a198d13504dc80c1}{section
678}} prohibition will not apply.

\hypertarget{exceptions}{%
\subsubsection{Exceptions}\label{exceptions}}

There are some exceptions, but these are narrowly construed.

\hypertarget{unconditional-exceptions}{%
\paragraph{Unconditional Exceptions}\label{unconditional-exceptions}}

\href{https://www.legislation.gov.uk/ukpga/2006/46/section/681}{s 681 CA
2006} contains a wide list of unconditional exceptions, mainly relating
to financial assistance being offered for procedures which are
authorised in other sections (e.g., redemption of shares, reduction of
capital).

\hypertarget{conditional-exceptions}{%
\paragraph{Conditional Exceptions}\label{conditional-exceptions}}

\href{https://www.legislation.gov.uk/ukpga/2006/46/section/682}{s 682 CA
2006} lists a number of ``conditional'' exceptions, which apply only if
the company has net assets and either:

\begin{enumerate}
\tightlist
\item
  Those assets are not reduced by the giving of financial assistance, or
\item
  To the extent that those assets are reduced, the assistance is
  provided out of distributable profits.
\end{enumerate}

\hypertarget{principal-purpose-and-incidental-part-of-a-larger-purpose}{%
\paragraph{Principal Purpose and Incidental Part of a Larger
Purpose}\label{principal-purpose-and-incidental-part-of-a-larger-purpose}}

One example of a conditional exception is financial assistance by a
company for the purposes of an employee share scheme, provided this is
made in good faith in the interests of the company or its holding
company (s 682(2)(b)).

There are exceptions set out in s 678(2) and (3) and s 679 (2) and (3)
which are designed to ensure that the prohibition in s 678(1) and s 679
(1) does not prohibit genuine commercial transactions that are in the
interests of the company.

These provide that financial assistance is not prohibited:

\begin{itemize}
\tightlist
\item
  If the principal purpose of the assistance is not to give it for the
  purpose of an acquisition of shares, or where the assistance is
  incidental to some other larger purpose of the company; and
\item
  In either case, where the financial assistance is given in good faith
  in the interests of the company.

  \begin{itemize}
  \tightlist
  \item
    To satisfy the good faith test, the officers procuring the company
    to give the financial assistance must act in the genuine belief that
    the assistance is being given in the company's interests.
  \end{itemize}
\end{itemize}

These exceptions can be difficult to rely on (Brady v Brady {[}1989{]}
AC 755) since they are extremely narrowly construed by the courts.

\begin{env-99ba4262-5297-4318-9547-86307715f49a}

s 678 - Assistance for acquisition of shares in public company

(1) Where a person is acquiring or proposing to acquire shares in a
public company, it is not lawful for that company, or a company that is
a subsidiary of that company, to give financial assistance directly or
indirectly for the purpose of the acquisition \textbf{before or at the
same time} as the acquisition takes place.

(2) Subsection (1) does not prohibit a company from giving financial
assistance for the acquisition of shares in it or its holding company
if---

\begin{itemize}
\tightlist
\item
  (a) the company's principal purpose in giving the assistance is not to
  give it for the purpose of any such acquisition, or
\item
  (b) the giving of the assistance for that purpose is only an
  incidental part of some larger purpose of the company,
\end{itemize}

and the assistance is given in good faith in the interests of the
company.

(3) Where---

\begin{itemize}
\tightlist
\item
  (a) a person has acquired shares in a company, and
\item
  (b) a liability \textbf{has been incurred} (by that or another person)
  for the purpose of the acquisition,
\end{itemize}

it is not lawful for that company, or a company that is a subsidiary of
that company, to give financial assistance directly or indirectly for
the purpose of \textbf{reducing or discharging the liability} if,
\textbf{at the time the assistance is given}, the company in which the
shares were acquired is a public company.

(4) Subsection (3) does not prohibit a company from giving financial
assistance if---

\begin{itemize}
\tightlist
\item
  (a) the company's principal purpose in giving the assistance is not to
  reduce or discharge any liability incurred by a person for the purpose
  of the acquisition of shares in the company or its holding company, or
\item
  (b) the reduction or discharge of any such liability is only an
  incidental part of some larger purpose of the company,
\end{itemize}

and the assistance is given in good faith in the interests of the
company.

(5) This section has effect subject to sections 681 and 682
(unconditional and conditional exceptions to prohibition).

\end{env-99ba4262-5297-4318-9547-86307715f49a}

\begin{env-99ba4262-5297-4318-9547-86307715f49a}

s 679 - Assistance by public company for acquisition of shares in its
private holding company

(1) Where a person is acquiring or proposing to acquire shares in a
private company, it is not lawful for a public company that is a
subsidiary of that company to give financial assistance directly or
indirectly for the purpose of the acquisition \textbf{before or at the
same time} as the acquisition takes place.

(2) Subsection (1) does not prohibit a company from giving financial
assistance for the acquisition of shares in its holding company if---

\begin{itemize}
\tightlist
\item
  (a) the company's principal purpose in giving the assistance is not to
  give it for the purpose of any such acquisition, or
\item
  (b) the giving of the assistance for that purpose is only an
  incidental part of some larger purpose of the company,
\end{itemize}

and the assistance is given in good faith in the interests of the
company.

(3) Where---

\begin{itemize}
\tightlist
\item
  (a) a person has acquired shares in a private company, and
\item
  (b) a liability \textbf{has been incurred} (by that or another person)
  for the purpose of the acquisition,
\end{itemize}

it is not lawful for a public company that is a subsidiary of that
company to give financial assistance directly or indirectly for the
purpose of \textbf{reducing or discharging the liability}.

(4) Subsection (3) does not prohibit a company from giving financial
assistance if---

\begin{itemize}
\tightlist
\item
  (a) the company's principal purpose in giving the assistance is not to
  reduce or discharge any liability incurred by a person for the purpose
  of the acquisition of shares in its holding company, or
\item
  (b) the reduction or discharge of any such liability is only an
  incidental part of some larger purpose of the company,
\end{itemize}

and the assistance is given in good faith in the interests of the
company.

(5) This section has effect subject to sections 681 and 682
(unconditional and conditional exceptions to prohibition).

\end{env-99ba4262-5297-4318-9547-86307715f49a}

\hypertarget{types}{%
\subsubsection{Types}\label{types}}

\emph{\href{https://uk.westlaw.com/8-505-7474?originationContext=document\&transitionType=PLDocumentLink\&contextData=(sc.Default)\&ppcid=52c063520c5f48d8a198d13504dc80c1}{section
677(1)}} sets out a number of different heads of financial assistance:

\begin{itemize}
\tightlist
\item
  Financial assistance given by way of gift.

  \begin{itemize}
  \tightlist
  \item
    If there is some consideration, however inadequate, moving to the
    company, it will not be a gift. The transaction might still be
    caught by
    \emph{\href{https://uk.westlaw.com/8-505-7474?originationContext=document\&transitionType=PLDocumentLink\&contextData=(sc.Default)\&ppcid=52c063520c5f48d8a198d13504dc80c1}{section
    677(1)(d)}} (material reduction in net assets).
  \end{itemize}
\item
  Financial assistance given by way of guarantee, security or indemnity
  (other than an indemnity in respect of the indemnifier's own neglect
  or default), or by way of release or waiver.
\item
  Financial assistance given by way of loan or any other agreement under
  which any of the obligations of the person giving the assistance are
  to be fulfilled at a time when in accordance with the agreement any
  obligation of another party to the agreement remains unfulfilled.
\item
  Any other financial assistance given by a company where the net assets
  of the company are reduced to a material extent by the giving of the
  financial assistance or the company has no net assets.

  \begin{itemize}
  \tightlist
  \item
    Assets and liabilities here should be given their actual values and
    not their book values.
  \item
    For ``material extent'', it is commonly accepted that less than 1\%
    would fall outside the prohibition but 1\% and over would amount to
    a material reduction.

    \begin{itemize}
    \tightlist
    \item
      This is the only \emph{de minimis} rule in s 677.
    \end{itemize}
  \end{itemize}
\end{itemize}

\hypertarget{flowcharts}{%
\subsubsection{Flowcharts}\label{flowcharts}}

\hypertarget{s-678}{%
\paragraph{S 678}\label{s-678}}

MERMAID1

\hypertarget{s-679}{%
\paragraph{S 679}\label{s-679}}

MERMAID2

\hypertarget{share-purchases}{%
\subsubsection{Share Purchases}\label{share-purchases}}

There are several situations which may raise financial assistance
problems in the context of a share purchase (if transactions carried out
for the purpose of assisting the acquisition of shares in the target --
which is a public company/ private holding company of a public company
giving financial assistance):

\begin{enumerate}
\tightlist
\item
  The target public company grants security over its assets in respect
  of a loan taken out by the buyer of shares.
\item
  The target public company repays its existing indebtedness under a
  loan agreement to its private holding company (the seller) at the time
  of the sale of the company.

  \begin{enumerate}
  \tightlist
  \item
    If the loan from the seller is repayable on demand, then this
    repayment will not be financial assistance.
  \item
    Else, acceleration of payment could amount to financial assistance.
  \end{enumerate}
\item
  Target public company sells assets to the buyer at the time of the
  sale of the company.

  \begin{enumerate}
  \tightlist
  \item
    If the company sells the assets at less than their market value (for
    example, at book value), this could be financial assistance insofar
    as there is a material reduction in the company's net assets
    (section 677(1)(d)(i)).
  \item
    If the target company leaves any of the consideration on the sale
    outstanding, this may amount to financial assistance by way of a
    loan ``or any other agreement under which any of the obligations of
    the person giving the assistance are to be fulfilled at a time when
    in accordance with the agreement any obligation of another party to
    the agreement remains unfulfilled'' (section 677(1)(c)(i)).
  \end{enumerate}
\item
  Using loan finance, the buyer acquires the target public company
  (which has a wholly owned subsidiary). The subsidiary is subsequently
  transferred to the buyer and the subsidiary grants security for the
  loan.

  \begin{enumerate}
  \tightlist
  \item
    The subsidiary of the target company is not giving financial
    assistance here if it is no longer a subsidiary of the target
    company.
  \item
    But the transfer of the subsidiary could constitute financial
    assistance if it was not made at market value.
  \end{enumerate}
\item
  Using loan finance, the buyer acquires shares in the target public
  company. The target subsequently acquires (or incorporates) a
  subsidiary which grants security for the loan.

  \begin{enumerate}
  \tightlist
  \item
    Although the subsidiary was not a subsidiary at the time of the
    acquisition, it is one at the time of the assistance (the security)
    is given and this will be caught by
    \emph{\href{https://uk.westlaw.com/5-505-7258?originationContext=document\&transitionType=PLDocumentLink\&contextData=(sc.Default)\&ppcid=fc9f4d9e83af4b4c8d47bf5847a3205c}{section
    678(3)}}.
  \end{enumerate}
\item
  A company borrows money, all or some of which it uses to subscribe for
  shares in a public company subsidiary. That subsidiary uses the money
  to acquire the majority of shares in the target public company. The
  subsidiary and target give guarantees and security for the bank
  borrowing.

  \begin{enumerate}
  \tightlist
  \item
    The guarantees and security granted by the target will constitute
    financial assistance for the acquisition of its shares.
  \item
    It will also constitute financial assistance for the acquisition of
    shares in its parent company (namely the subscription of shares in
    the subsidiary).
  \end{enumerate}
\item
  Using loan finance, the buyer acquires shares in the target public
  company. Subsequently, the buyer refinances this borrowing (that is,
  it borrows from a new lender and repays the original lender). This new
  borrowing is secured by the target company.

  \begin{enumerate}
  \tightlist
  \item
    The grant of security by the target company will constitute
    financial assistance within
    \emph{\href{https://uk.westlaw.com/5-505-7258?originationContext=document\&transitionType=PLDocumentLink\&contextData=(sc.Default)\&ppcid=fc9f4d9e83af4b4c8d47bf5847a3205c}{section
    678(3)}} because its purpose is to assist in the discharge of a
    liability.
  \item
    However, if this new financing is subsequently itself refinanced (a
    second-round refinancing) security granted by the target company for
    that new lending will not be caught because that new lending is not
    discharging acquisition debt (but to discharge borrowings which
    themselves discharged acquisition debt).
  \end{enumerate}
\item
  Using loan finance, the buyer acquires shares in the target public
  company. Subsequently, the target company makes (otherwise lawful)
  dividend payments and loans to the buyer which the buyer uses to
  service or repay the loan.

  \begin{enumerate}
  \tightlist
  \item
    So far as the dividends are concerned, these fall within the
    unconditional exception in
    \emph{\href{https://uk.westlaw.com/8-506-2055?originationContext=document\&transitionType=PLDocumentLink\&contextData=(sc.Default)\&ppcid=fc9f4d9e83af4b4c8d47bf5847a3205c}{section
    681(2)(a)}}.
  \item
    However, the loans will be prohibited by
    \emph{\href{https://uk.westlaw.com/5-505-7258?originationContext=document\&transitionType=PLDocumentLink\&contextData=(sc.Default)\&ppcid=fc9f4d9e83af4b4c8d47bf5847a3205c}{section
    678(3)}} and
    \emph{\href{https://uk.westlaw.com/8-505-7474?originationContext=document\&transitionType=PLDocumentLink\&contextData=(sc.Default)\&ppcid=fc9f4d9e83af4b4c8d47bf5847a3205c}{section
    677(1)(c)}}.
  \end{enumerate}
\end{enumerate}

\hypertarget{break-fees}{%
\subsubsection{Break Fees}\label{break-fees}}

Where the company whose shares are potentially being acquired pays the
break fee (or agrees to do so) the break fee could amount to unlawful
financial assistance. Draft such provisions carefully.

\hypertarget{sanctions}{%
\subsubsection{Sanctions}\label{sanctions}}

\hypertarget{effect-of-transaction}{%
\paragraph{Effect of Transaction}\label{effect-of-transaction}}

The transaction itself is illegal and unenforceable, as is any security
or guarantee given in contravention of the prohibition. If a provision
for illegal financial assistance is contained in a share purchase
agreement, then it may be possible to sever the offending term.

\hypertarget{directors-liability}{%
\paragraph{Directors' Liability}\label{directors-liability}}

If the provision of financial assistance is not in accordance with the
directors' duty to act in the way they consider, in good faith, would be
most likely to promote the success of the company for the benefit of its
members as a whole
(\emph{\href{https://uk.westlaw.com/5-505-5377?originationContext=document\&transitionType=PLDocumentLink\&contextData=(sc.Default)\&ppcid=fc9f4d9e83af4b4c8d47bf5847a3205c}{section
172}}), and the breach has not been approved or ratified by an
appropriate shareholder resolution the directors will be in breach of
their duty and may have to account for any losses suffered by the
company.

A director may also be liable for disqualification as a result of a
company's involvement in a prohibited transaction
(\emph{\href{https://uk.westlaw.com/5-506-8501?originationContext=document\&transitionType=PLDocumentLink\&contextData=(sc.Default)\&ppcid=fc9f4d9e83af4b4c8d47bf5847a3205c}{section
1},
\href{https://uk.westlaw.com/4-505-7796?originationContext=document\&transitionType=PLDocumentLink\&contextData=(sc.Default)\&ppcid=fc9f4d9e83af4b4c8d47bf5847a3205c}{Company
Directors Disqualification Act 1986}}).

\begin{env-486fc4cd-eb32-4215-aaa2-fc77a67eebc5}

Note

Misfeasance can include illegal financial assistance (s 212)

\end{env-486fc4cd-eb32-4215-aaa2-fc77a67eebc5}

\hypertarget{recipient-liability}{%
\paragraph{Recipient Liability}\label{recipient-liability}}

If a person receives money from a company as a result of illegal
financial assistance and knew or ought to have known of the
circumstances rendering the assistance illegal, then a constructive
trust may arise in favour of the company
(\emph{\href{https://uk.westlaw.com/D-016-7350?originationContext=document\&transitionType=PLDocumentLink\&contextData=(sc.Default)\&ppcid=fc9f4d9e83af4b4c8d47bf5847a3205c}{Belmont
Finance Corp Ltd v Williams Furniture Ltd (No 2) {[}1980{]} 1 All ER
393}}). The third party may be required to account for any money or
property received in breach of
\emph{\href{https://uk.westlaw.com/5-505-7258?originationContext=document\&transitionType=PLDocumentLink\&contextData=(sc.Default)\&ppcid=fc9f4d9e83af4b4c8d47bf5847a3205c}{section
678}} or
\emph{\href{https://uk.westlaw.com/6-506-2056?originationContext=document\&transitionType=PLDocumentLink\&contextData=(sc.Default)\&ppcid=fc9f4d9e83af4b4c8d47bf5847a3205c}{679}}
as a constructive trustee.

\hypertarget{restrictions-on-relevant-shares}{%
\paragraph{Restrictions on Relevant
Shares}\label{restrictions-on-relevant-shares}}

A person who has been given illegal financial assistance by a public
company to acquire that company's shares and who holds those shares as a
trustee for the company must not exercise any voting rights attached to
the shares
(\emph{\href{https://uk.westlaw.com/5-506-0543?originationContext=document\&transitionType=PLDocumentLink\&contextData=(sc.Default)\&ppcid=fc9f4d9e83af4b4c8d47bf5847a3205c}{section
662(1)(e)}}).

\hypertarget{criminal-sanctions}{%
\paragraph{Criminal Sanctions}\label{criminal-sanctions}}

The only sanction set out in the
\emph{\href{https://uk.westlaw.com/3-503-8567?originationContext=document\&transitionType=PLDocumentLink\&contextData=(sc.Default)\&ppcid=fc9f4d9e83af4b4c8d47bf5847a3205c}{CA
2006}} provides that contravention of
\emph{\href{https://uk.westlaw.com/5-505-7258?originationContext=document\&transitionType=PLDocumentLink\&contextData=(sc.Default)\&ppcid=fc9f4d9e83af4b4c8d47bf5847a3205c}{section
678(1) or (3)}} or
\emph{\href{https://uk.westlaw.com/6-506-2056?originationContext=document\&transitionType=PLDocumentLink\&contextData=(sc.Default)\&ppcid=fc9f4d9e83af4b4c8d47bf5847a3205c}{section
679(1) or (3)}} renders the company liable to a fine and every officer
in default liable to a fine or up to two years' imprisonment or both
(\emph{\href{https://uk.westlaw.com/4-506-2057?originationContext=document\&transitionType=PLDocumentLink\&contextData=(sc.Default)\&ppcid=fc9f4d9e83af4b4c8d47bf5847a3205c}{section
680}}).

``Officers'' include directors, managers and secretaries and any person
who is to be treated as an officer of the company for the purposes of
the provision in question
(\emph{\href{https://uk.westlaw.com/2-506-8531?originationContext=document\&transitionType=PLDocumentLink\&contextData=(sc.Default)\&ppcid=fc9f4d9e83af4b4c8d47bf5847a3205c}{section
1121(2)}}). Officers who may be liable are those who authorise or
permit, participate in, or fail to take all reasonable steps to prevent
the contravention (\emph{section 1121(3)}).

\hypertarget{applicable-law}{%
\subsection{Applicable Law}\label{applicable-law}}

\hypertarget{security-interests}{%
\subsubsection{Security Interests}\label{security-interests}}

The legal system which governs a security interest will dictate how the
security will be perfected (eg, notice, registration, etc) and how it
may be enforced.

\begin{longtable}[]{@{}ll@{}}
\toprule()
Type of asset & Legal system \\
\midrule()
\endhead
Immovable assets & Security governed by the legal system of the place
where the assets are physically situated (known as the `\emph{lex
situs}'). \\
Tangible movable assets & Generally, the \emph{lex situs} of the asset
when the security is created will determine the validity of its creation
and perfection. Generally, the \emph{lex situs} of the asset when the
security is created will determine the validity of its creation and
perfection. \\
Intangible movable assets & Security over intangible assets (eg, debts)
is usually subject to the governing law of the contract creating the
asset, but the law of the place of performance of the asset (eg, where
the debt is to be paid) is also relevant. \\
\bottomrule()
\end{longtable}

\hypertarget{insolvency}{%
\subsubsection{Insolvency}\label{insolvency}}

From 31 May 2002, the EU Insolvency Regulation (EUIR) came into force in
all EU Member States except Denmark. The EUIR provides rules that
dictate which system of insolvency law applies in a `cross-border'
insolvency situation.

\begin{itemize}
\tightlist
\item
  Under the EUIR, the main insolvency proceedings will be opened in the
  jurisdiction of the entity's Centre of Main Interest (COMI).
\item
  There is a rebuttable presumption that the COMI is in the jurisdiction
  of an entity's registered office (Article 3(1)), although factors such
  as location of employees, head office, board meetings and where the
  company does most of its business can influence the COMI.
\item
  2 main circumstances in which assets may be subject to insolvency
  proceedings in jurisdictions other than the COMI.

  \begin{enumerate}
  \tightlist
  \item
    Assets outside the COMI which are subject to `rights in rem' are
    basically unaffected by the COMI proceedings.
  \item
    Supplemental proceedings may be brought in a jurisdiction where a
    company has an `establishment'.
  \end{enumerate}
\item
  The EUIR applies to companies, partnerships and individuals with their
  COMI in an EU Member State (other than Denmark), but not to most
  regulated entities (eg, banks, insurance companies) which have
  separate regimes.
\item
  The UK now has a slightly different regime but the EUIR largely
  operates unchanged.
\end{itemize}

\hypertarget{uncitral-model-law}{%
\subsection{UNCITRAL Model Law}\label{uncitral-model-law}}

Cross-Border Insolvency Regulations 2006 implements the main provisions
of the `Model Law' on Cross-Border Insolvency devised by UNCITRAL (the
United Nations Commission on International Trade Law). EUIR takes
precedence where there is a clash.

\hypertarget{to-consider}{%
\subsection{To Consider}\label{to-consider}}

When advising on security issues, also consider:

\begin{itemize}
\tightlist
\item
  Prior security

  \begin{itemize}
  \tightlist
  \item
    Check for negative pledges
  \item
    May need a subordination agreement.
  \end{itemize}
\item
  Notification requirements

  \begin{itemize}
  \tightlist
  \item
    May need to notify third parties like licensing authorities and
    landlords.
  \end{itemize}
\item
  Overseas registration.

  \begin{itemize}
  \tightlist
  \item
    Security interests over overseas properties may need registering in
    the jurisdiction in which situated.
  \end{itemize}
\item
  Parallel debt structures

  \begin{itemize}
  \tightlist
  \item
    Security for a syndicated loan will be held by a security trustee
    (usually one of the lenders): legal title to the security is held on
    trust and the syndicate banks are the beneficiaries.
  \item
    This allows syndicate members to change without danger of the
    security being treated as renewed (losing priority/ hardening
    periods).
  \item
    Trusts aren't a thing in civil law jurisdictions, so a "parallel
    debt" structure may be used instead.

    \begin{itemize}
    \tightlist
    \item
      A security agent holds the security on behalf of the syndicate.
    \item
      But most jurisdictions only allow a party to hold security to the
      extent it is owed a debt.
    \item
      Parallel debt structure creates a debt owed to the security agent
      in the same amount as the debt owed to the bank syndicate.
    \item
      Normally, if the borrower repays the syndicate, any amount it
      repays is deemed to reduce the syndicate debt by the same ("in
      parallel").
    \item
      On a default, however, the security agent becomes entitled to
      demand repayment of the debt owed to it and enforce the security
      if necessary. For every amount the security agent recovers, the
      syndicated debt will be deemed to reduce by the same.
    \item
      Not as good as a trust; the syndicate takes a risk on the security
      agent becoming insolvent.
    \end{itemize}
  \end{itemize}
\end{itemize}

\hypertarget{debentures}{%
\section{Debentures}\label{debentures}}

The document creating multiple security interests (e.g. charges,
mortgages, assignments) for a secured loan is commonly known as a
`debenture'.

\hypertarget{structure}{%
\subsection{Structure}\label{structure}}

\begin{itemize}
\tightlist
\item
  Most banks have their own short standard form security agreement.
\item
  More sophisticated secured loans require a bespoke debenture drafted
  by the bank's solicitor.
\item
  The broad structure of most debentures is similar to other commercial
  documents, ie: date, title and parties; preamble; definitions;
  operative provisions; schedules; and execution.
\end{itemize}

Operative provisions:

\hypertarget{covenant-to-pay}{%
\subsubsection{Covenant to Pay}\label{covenant-to-pay}}

\begin{itemize}
\tightlist
\item
  Acknowledges the debt obligation of the borrower to the bank, and may
  refer to a specific debt obligation/ to all the debts between the bank
  and borrower (``all monies'' clause).
\item
  Purposes:

  \begin{enumerate}
  \tightlist
  \item
    Brings the document within the common law definition of debenture --
    `a document which either creates a debt or acknowledges it' (Levy v
    Abercorris Slate and Slab Co (1887) 37 Ch D 260). Satisfies one of
    the conditions to appointing an administrative receiver under s
    29(2) IA 1986.
  \item
    In a syndicated loan, the covenant to pay is addressed to the
    security trustee (``Security Agent''). Allows the security trustee
    to demand repayment and enforce security on behalf of the syndicate
    members.
  \end{enumerate}
\item
  ``On demand'' intended to prevent Limitation Act 1939 from acting
  against the bank until it makes a demand.

  \begin{itemize}
  \tightlist
  \item
    It does not alter the payment obligation in a facility agreement
    which it secures.
  \item
    Some debentures refer to monies owing or incurred `on any account
    whatsoever', but this wording may be inadvisable because there is a
    danger that it would not include certain types of obligation (Re
    Quest Cae Limited {[}1985{]} BCLC 266)
  \end{itemize}
\end{itemize}

\hypertarget{charging-clause}{%
\subsubsection{Charging Clause}\label{charging-clause}}

Specifies which of the borrower's assets are charged to repay the debts
acknowledged in the covenant to pay, as well as how each type of asset
is secured.

\hypertarget{land}{%
\paragraph{Land}\label{land}}

\begin{itemize}
\tightlist
\item
  A debenture will usually secure leasehold and freehold land by way of
  a `charge by way of legal mortgage'.
\item
  Details of registered land to be secured included in a schedule to the
  debenture so the security can be registered at the Land Registry.
\item
  Check the terms of any leases to ensure they do not prohibit granting
  of security interest.
\item
  Future acquired land cannot be secured by legal mortgage, and is
  therefore secured by a separate equitable mortgage or charge over the
  property as soon as it is acquired.

  \begin{itemize}
  \tightlist
  \item
    The borrower may then be required subsequently to `upgrade' this to
    a legal mortgage.
  \end{itemize}
\end{itemize}

\hypertarget{fixed-plant-and-machinery}{%
\paragraph{Fixed Plant and Machinery}\label{fixed-plant-and-machinery}}

\begin{itemize}
\tightlist
\item
  A mortgage of land will automatically secure any assets which are, or
  become, permanently fixed to the secured property (known as
  `fixtures').
\item
  May be other assets which are not fixtures but are fixed assets.
\end{itemize}

\hypertarget{cash-deposits}{%
\paragraph{Cash Deposits}\label{cash-deposits}}

A bank may want to take security over a cash deposit made with it (e.g.,
if it has provided a performance guarantee).

\hypertarget{mortgage-charge}{%
\subparagraph{Mortgage/ Charge}\label{mortgage-charge}}

The obvious way to secure the account is for the depositor to create a
mortgage or charge over it in favour of the bank. But in Re Charge Card
Services {[}1987{]} Ch 150,

\begin{quote}
{[}a debt{]} \ldots{} cannot be made the subject of a legal or equitable
mortgage in favour of the debtor, since this requires a conveyance or
assignment by way of security, and this operates as a conditional
release;
\end{quote}

and

\begin{quote}
A charge in favour of a debtor of his own indebtedness to the chargor is
conceptually impossible.
\end{quote}

\begin{env-486fc4cd-eb32-4215-aaa2-fc77a67eebc5}

Logic

\begin{itemize}
\tightlist
\item
  An account is a debt owed by a bank to a depositor.
\item
  A debt is a right to sue the debtor for the amount owed.
\item
  If this was charged/ mortgaged in favour of the debt, this would give
  the debtor the right to sue itself.
\item
  So this would cancel/ reduce the debt.
\end{itemize}

\end{env-486fc4cd-eb32-4215-aaa2-fc77a67eebc5}

In Morris v Agrichemicals Ltd {[}1997{]} 3 WLR 909 (`BCCI No 8'), the
House of Lords, led by Lord Hoffmann, declared that a \textbf{charge} in
favour of a person over the benefit of a deposit made with that person,
by another, is not a conceptual impossibility. But an assignment
(mortgage) of the account to the bank remains ineffective.

\begin{env-a9aa42a3-f5ef-44bc-9cb3-81fc5caedba7}

Question

How shold a depositor provide a bank with security ("cash collateral")
over a deposit with that bank?

\end{env-a9aa42a3-f5ef-44bc-9cb3-81fc5caedba7}

The means of enforcing a cash collateral arrangement should be by
set-off -- contractual set-off prior to winding-up, and mandatory
set-off after winding-up. Any security granted over the account is
intended to ensure that the money is in the account when the bank needs
to set off. So the bank should require:

\begin{enumerate}
\tightlist
\item
  An express contractual right of set-off;
\item
  a fixed charge over the account (assuming the depositor will not be
  allowed to withdraw any money without the bank's prior consent);
\item
  a prohibition on the depositor assigning the account (to deal with the
  fact that a bona fide purchaser, for value, without notice could
  defeat the fixed charge); and
\item
  if the depositor's obligation is contingent, a `flawed asset'
  arrangement (that is, the deposit only becomes repayable at the same
  time that the depositor's debt becomes due and unpaid). This protects
  the account from a liquidator until the bank can exercise its rights
  of set-off.
\end{enumerate}

\begin{env-486fc4cd-eb32-4215-aaa2-fc77a67eebc5}

Note

\begin{itemize}
\tightlist
\item
  An assignment (mortgage) will usually be possible if security over an
  account is given to a security trustee or an agent bank on behalf of a
  syndicate
\item
  None of these issues will apply if a borrower gives security over an
  account held with a third party bank.
\end{itemize}

\end{env-486fc4cd-eb32-4215-aaa2-fc77a67eebc5}

\hypertarget{goodwill}{%
\paragraph{Goodwill}\label{goodwill}}

Any goodwill held by a borrower is probably automatically secured as an
inseparable part of the borrower's secured `undertaking'. But usually
the debenture explicitly secures the borrower's goodwill by fixed
charge.

\hypertarget{book-debts}{%
\paragraph{Book Debts}\label{book-debts}}

Book debt is an unpaid invoice, i.e., a sum owed to the company in
respect of goods or services supplied by it.

Book debts (or ``receivables'') are a fluctuating asset, and may be a
significant asset of the company.

In early cases, the courts took the approach that a charge granted over
book debts was a fixed charge, but this approach has now been overruled.

\begin{longtable}[]{@{}ll@{}}
\toprule()
Case & Authority \\
\midrule()
\endhead
Siebe Gorman and Co Ltd v Barclays Bank Ltd {[}1979{]} 2 Lloyd's Rep 142
& The court held that a charge over book debts was a fixed charge
because of the degree of control of the bank which could stop the
company making withdrawals, even when the account was in credit. Since
overruled. \\
Re Brightlife Ltd {[}1987{]} Ch 200 ChD & The court in this case
clarified that a company's bank balance is not a book debt and therefore
cannot be subject to a fixed charge. \\
Re Keenan Bros Ltd {[}1986{]} BCLC 242 & A fixed charge was created by
the means of a requirement that the funds collected by the company were
to be paid into a blocked account. The prior written consent of the bank
was required before any funds were allowed to be withdrawn from this
account. Here the court said that the charge was a fixed charge due to
the fact that the account was blocked. \\
Re Brumark Investments Ltd {[}2001{]} & The company's freedom to collect
and use the proceeds of the book debts as it wished without requiring
the lender's consent was inconsistent with the nature of a fixed charge.
This was a Privy Council case, which means that it is persuasive but not
binding. \\
\bottomrule()
\end{longtable}

Most cases suggest that \textbf{book debts have two elements, the debt
and the proceeds of the debt}, and that a bank must show control over
both elements to achieve fixed security over either. Taking control over
the debt element is usually uncontroversial: the bank must prevent the
borrower doing anything other than collecting in the debt. Controlling
the proceeds is more difficult, because the borrower usually needs
access to them to run its business.

In the early 1990s there was a run of judgments which moved against the
previous cases and was much more in favour of the banks. The leading
case was Re New Bullas Trading {[}1994{]} 1 BCLC 485, where the Court of
Appeal blessed a debenture allowing fixed security over the unrealised
debt and floating security over the proceeds. But this was subject to
lots of criticism.

Eventually, the position was resolved in National Westminster Bank plc v
Spectrum Plus Limited and others {[}2005{]} UKHL 41. It is now very
difficult to demonstrate the requisite level of control over book debts
for a fixed charge. It would require at least a blocked account.

\begin{env-be598745-44b1-4815-8875-ed2abeb511dd}

Charge over book debts

From the case of National Westminster Bank plc v Spectrum Plus Limited
and others {[}2005{]} UKHL 41 it is only possible to have a fixed charge
over book debts if they are paid into a blocked account which gives the
lender the degree of control required.

\textbf{Book debts can only be secured by a fixed charge where the
charge holder has control over both the debts and the proceeds once they
were paid.}

\end{env-be598745-44b1-4815-8875-ed2abeb511dd}

The vexed question for practitioners, still unanswered by the courts, is
how much control (if any) may be given to the borrower without fixed
security being lost.

The court will look at the substance of an arrangement, not the title.

\hypertarget{shares}{%
\paragraph{Shares}\label{shares}}

Shares may be in

\begin{itemize}
\tightlist
\item
  Registered form (in which a register of legal owners is maintained by
  the company),
\item
  Bearer form (where legal ownership is generally determined by
  possession of the share certificate) or
\item
  `Dematerialised' (where they are held and traded electronically).
\end{itemize}

Registered shares usually secured by either:

\begin{enumerate}
\tightlist
\item
  Creating a legal mortgage by transferring the shares into the bank's
  name and issuing a new share certificate to that effect.

  \begin{itemize}
  \tightlist
  \item
    Rare -- lots of admin and potential liabilities for the bank.
  \item
    Enviroco Ltd v Farstad Supply A/S {[}2009{]} EWCA Civ 1399: a
    company ceases to be a subsidiary of its holding company if the
    holding company gives a legal mortgage over the subsidiary's shares.
  \item
    The debenture will allow the bank to sell the shares on default of
    the borrower and provide for re-transfer when the loan is repaid.
  \end{itemize}
\item
  Creating an equitable mortgage (or charge) by taking custody of the
  share certificate together with a stock transfer form executed by the
  borrower, but with the transferee's name and other details left blank.

  \begin{itemize}
  \tightlist
  \item
    The security document should also contain a power of attorney
    allowing the bank to complete the stock transfer form and submit it,
    with the share certificates, to register a new owner.
  \item
    The bank does not have the `inconvenience' of ownership
    pre-enforcement (e.g., voting, passing dividend back to the
    borrower) and does not run the risk of the company whose shares are
    secured becoming a subsidiary of the bank.
  \end{itemize}
\end{enumerate}

The bank's solicitor should ensure:

\begin{itemize}
\tightlist
\item
  that the shares are fully paid up,
\item
  that (for (1)) any stamp duty payable on the transfer form has been
  paid, and
\item
  that there are no restrictions on transfer in the articles (e.g.,
  pre-emption rights).
\end{itemize}

In the case of bearer shares, security is usually achieved simply by
taking delivery of the share certificate and a memorandum of deposit
(i.e., a pledge).

\hypertarget{stock-in-trade-and-sundry-assets}{%
\paragraph{Stock in Trade and Sundry
Assets}\label{stock-in-trade-and-sundry-assets}}

A borrower's stock, sometimes known as `inventory', is by definition an
asset the borrower needs to sell and so will typically be secured by a
floating charge.

\hypertarget{contracts}{%
\paragraph{Contracts}\label{contracts}}

Contractual rights are choses in action and so may be secured by way of
assignment (legal or equitable) or a charge. Which of these security
devices is used will depend on a number of factors, including whether:

\begin{itemize}
\tightlist
\item
  the contract is specified as `non-transferable';
\item
  the borrower needs to alter the contract from time to time;
\item
  it is preferable for the borrower to enforce the contact rights;
\item
  all or only part of the contract is to be secured;
\item
  the counterparty can be notified; and
\item
  the contract has specific features (e.g. an insurance contract).
\end{itemize}

\hypertarget{further-assurance-clause}{%
\subsubsection{Further Assurance
Clause}\label{further-assurance-clause}}

\begin{env-8872a5c8-2356-4b38-87ff-b8a967f137b3}

Further assurance

An undertaking by the borrower to do anything necessary in order to
perfect the security.

\end{env-8872a5c8-2356-4b38-87ff-b8a967f137b3}

May include execution of documents, conversion of an equitable to legal
mortgage, deposit of documents of title or provision of additional
security.

\hypertarget{undertakings}{%
\subsubsection{Undertakings}\label{undertakings}}

Most debentures contain undertakings. The most common are:

\begin{itemize}
\tightlist
\item
  A negative pledge

  \begin{itemize}
  \tightlist
  \item
    Often included even if the facility agreement contains a negative
    pledge to help ensure registration at Companies House.
  \end{itemize}
\item
  Undertaking to execute, at the bank's request, a legal mortgage in
  place of an equitable one;
\item
  an undertaking to provide the bank with any documents of title it
  might request;
\item
  an undertaking to repair and maintain land, and insure all assets,
  together with a power of attorney allowing the bank to do these things
  (and recover any expenses it incurs) if the borrower does not.
\end{itemize}

\hypertarget{crystallisation-clause}{%
\subsubsection{Crystallisation Clause}\label{crystallisation-clause}}

If it creates a floating charge, the debenture will specify when the
floating charge crystallises. Usually:

\begin{enumerate}
\tightlist
\item
  At any time, the bank notifies the borrower in writing (though the
  borrower may resist this and insist on default/ perceived risk)
\item
  Automatically on specified events (e.g., any demand for repayment, on
  a petition for winding-up etc.)
\end{enumerate}

\hypertarget{enforcement}{%
\subsubsection{Enforcement}\label{enforcement}}

The bank will have a right (subject to contrary agreement) to immediate
possession of any assets over which it has a legal mortgage, but not
those over which it has an equitable mortgage/ charge.

The debenture should therefore give the bank

\begin{itemize}
\tightlist
\item
  an express contractual right to take possession of the charged assets
  on specified `enforcement events' (i.e., events of default).
\item
  enhanced powers to sell, lease and appropriate secured assets on
  enforcement.
\item
  the right, on an enforcement event, to appoint a receiver of the
  secured property, an administrator and, if permitted, an
  administrative receiver. Such appointees will be extended any enhanced
  rights given to the bank.
\end{itemize}

\hypertarget{collection-of-debts}{%
\subsubsection{Collection of Debts}\label{collection-of-debts}}

A debenture will usually require the borrower to collect in any debts
and claims of the business `promptly', and to pay them into specified
accounts.

\hypertarget{power-of-attorney}{%
\subsubsection{Power of Attorney}\label{power-of-attorney}}

A debenture will invariably provide the bank and any receiver it
appoints with a power of attorney.

\begin{itemize}
\tightlist
\item
  This will allow the bank to carry out any of the borrower's
  obligations under the debenture which the borrower fails to honour.
\item
  Also assists in enforcement of security.
\item
  The instrument creating a power of attorney must be executed as a deed
  and, provided it is given to secure a proprietary interest of the
  donee and also expressed to be `irrevocable', it will survive the
  borrower's liquidation (Powers of Attorney Act 1971, s 4).
\end{itemize}

\hypertarget{redemption-of-security}{%
\subsubsection{Redemption of Security}\label{redemption-of-security}}

The debenture must provide for the release of the security once the
secured obligations are satisfied.

The debenture may allow the bank, if it believes a preference (or
similar) claim is possible, to deem the debt and related security to
`continue' even after repayment. Unproven and probably not workable.

There are no strict formalities for the discharge of security, other
than for registered land.

\hypertarget{administrative-provisions}{%
\subsubsection{Administrative
Provisions}\label{administrative-provisions}}

A debenture will contain a number of boilerplate clauses concerning
administrative matters, such as notice periods, interpretation, service
of documents, and costs and expenses.

\hypertarget{execution}{%
\subsection{Execution}\label{execution}}

Debentures should be made by deed:

\begin{enumerate}
\tightlist
\item
  it avoids any doubt as to sufficiency of consideration;
\item
  it is necessary because debentures invariably create a power of
  attorney (see the Power of Attorney Act 1971, s 1);
\item
  it is necessary if a debenture conveys or creates a legal estate in
  land (LPA 1925, s 52); and
\item
  it is necessary if the bank is to take powers under LPA 1925, s 101
  (powers of mortgagees to sell mortgaged property/ enter into contracts
  etc.).
\end{enumerate}

A debenture should also be expressed as being made by the borrower `in
favour of the bank' (rather than `between' borrower and bank) to ensure
that it can be enforced by a party other than the bank (e.g., a receiver
or attorney).

\hypertarget{legal-opinions}{%
\section{Legal Opinions}\label{legal-opinions}}

First introduced in US, now usual for UK banks making substantial loans
or capital market investments.

\hypertarget{fundamentals}{%
\subsection{Fundamentals}\label{fundamentals}}

\hypertarget{purpose-1}{%
\subsubsection{Purpose}\label{purpose-1}}

Relate only to matters of law. In banking, a legal opinion will usually
confirm that all the documents associated with the loan are legally
valid and enforceable and highlight areas of risk.

\hypertarget{who}{%
\subsubsection{Who}\label{who}}

Typically provided by the bank's solicitor, addressed to the bank.
Generally, only an addressee may sue the solicitor if the opinion proves
wrong, although the solicitor will owe a general responsibility to
anyone they know has relied on their advice. If the transaction involves
a syndicate (or underwriters), the opinion may be addressed to the agent
bank on their behalf, or be addressed to each of them individually.

\hypertarget{when}{%
\subsubsection{When}\label{when}}

Almost always required for:

\begin{longtable}[]{@{}ll@{}}
\toprule()
Scenario & Details \\
\midrule()
\endhead
Large loans & Opinion by the bank's lawyers that the documents are
legal, valid and binding. \\
Secured lending & Statement confirming the priority and enforceability
of the security and to outline any risks. \\
Overseas jurisdictions & Legal opinion required by a local lawyer. \\
\bottomrule()
\end{longtable}

\begin{itemize}
\tightlist
\item
  Even if the loan is a repeat of an earlier one (a re-financing), the
  bank may still require a new opinion (to update effect).
\item
  A bank will usually make the delivery of a satisfactory legal opinion
  a condition precedent to the loan being drawn down.

  \begin{itemize}
  \tightlist
  \item
    A borrower ideally should insist that the bank and the opining
    solicitors agree the form of the opinion in advance of executing the
    facility agreement.
  \end{itemize}
\end{itemize}

\hypertarget{form}{%
\subsubsection{Form}\label{form}}

Written as a letter on the solicitor's headed notepaper, addressed to
the bank and signed in the firm's name.

Structure:

\begin{enumerate}
\tightlist
\item
  Preamble;
\item
  assumptions;
\item
  opinion; and
\item
  qualifications.
\end{enumerate}

\hypertarget{preamble}{%
\subsection{Preamble}\label{preamble}}

\begin{itemize}
\tightlist
\item
  Outlines form of loan transaction
\item
  Lists all the documents in respect of which the opinion is given
\item
  Lists specific documents which the solicitor has reviewed in order to
  give the opinion.
\item
  Usually delivered after the documents to which it refers have been
  executed -- if not be clear that it refers to a particular draft.
\item
  Record of any searches on which the opinion relies and the date on
  which they were done.
\end{itemize}

\hypertarget{assumptions}{%
\subsection{Assumptions}\label{assumptions}}

\begin{itemize}
\tightlist
\item
  All signatures appearing on reviewed documents are genuine
\item
  Documents submitted to the solicitors as originals are authentic and
  complete
\item
  All copies complete and accurate.
\item
  Any resolutions properly passed at quorate meetings duly held
\item
  Documents accurate
\item
  Draft documents executed unchanged
\item
  Searches remain unchanged
\item
  No insolvencies since the opinion
\item
  No foreign law affects the conclusion
\item
  That the bank had no notice of any prohibition or restriction on
  creating the security, or notice of any other security in the same
  assets;
\item
  that immediately after providing any guarantee (or other security),
  each obligor was fully solvent, and that the provision of the
  guarantee (or security) was in good faith, for the benefit of the
  obligor and did not constitute financial assistance.
\end{itemize}

\hypertarget{opinion}{%
\subsection{Opinion}\label{opinion}}

\begin{itemize}
\tightlist
\item
  Actual opinion appears in the form of numbered statements.
\item
  Short compared to the overall length of the letter.
\item
  The exact form of legal opinion will be negotiated between bank and
  solicitor.
\end{itemize}

Commonly includes:

\begin{itemize}
\tightlist
\item
  That borrower was properly incorporated and not in liquidation.
\item
  Borrower has requisite power to enter into and perform the facility
  agreement, and has taken action necessary to authorise execution.
\item
  Execution, delivery and performance will not contravene English law or
  regulation, and will not violate memorandum/ articles.
\item
  Payments made by the borrower not subject to UK taxes and no taxes
  payable on execution/ registration of facility agreement.
\item
  All necessary registration, filings, licences and consents have been
  obtained in order to validate the loan document.
\end{itemize}

\hypertarget{qualifications}{%
\subsection{Qualifications}\label{qualifications}}

Commonly:

\begin{itemize}
\tightlist
\item
  The validity, performance and enforcement of the facility agreement
  may be affected by insolvency proceedings.
\item
  Any declaration in the opinion that the facility agreement is
  enforceable must be read on the basis that equitable remedies are at
  the discretion of the court
\item
  A provision purporting to create a fixed charge may be construed as
  creating a floating charge
\item
  The enforcement of some security rights is controlled by law.
\item
  Any claims under the loan may be affected by lapse of time.
\item
  The opinion can only take account of the law of England and Wales.
\item
  A court may refuse to entertain proceedings if a claim has been, or is
  being, brought elsewhere.
\item
  Certain specific provisions may not be upheld by a court, most notably
  default interest, provisions for severability, and provisions
  requiring a borrower to pay the costs of litigation on enforcement.
\end{itemize}

\hypertarget{foreign-counsel-legal-opinions}{%
\subsection{Foreign Counsel Legal
Opinions}\label{foreign-counsel-legal-opinions}}

If the facility agreement is governed by English law but a borrower (or
guarantor) is incorporated outside of England and Wales, the banks will
usually require a legal opinion from counsel in that jurisdiction.

This should cover:

\begin{itemize}
\tightlist
\item
  Borrower's status (duly incorporated etc.)
\item
  Borrower has power and authority to execute transaction and perform
  obligations
\item
  No legal conflict (with law or constitution of the borrower).
\item
  Legal, valid, binding and enforceable
\item
  Filings, registrations and consent -- confirmation that no further
  action is needed (e.g., executing in front of a notary, which is
  common under civil law).
\item
  Withholding tax -- lenders sensitive to withholding tax being imposed.
\item
  Domicile -- confirmation that the lenders will not become domiciled in
  the local jurisdiction by virtue of the transaction documentation.
\item
  Choice of law -- confirmation that it will be recognised.
\item
  Lender's claim to rank \emph{pari passu} (at least)
\item
  Jurisdiction and agent for service of process.

  \begin{itemize}
  \tightlist
  \item
    Confirmation that the borrower's submission to jurisdiction is valid
    and cannot be revoked.
  \item
    Confirmation that the appointment of an agent in the jurisdiction of
    the governing law (an `agent for service of process') is valid and
    cannot be revoked.
  \end{itemize}
\item
  Validity of security and that it covers future as well current
  advances.
\end{itemize}

Other issues:

\begin{itemize}
\tightlist
\item
  No limitation on paying interest
\item
  No exchange control
\item
  Borrower solvent
\item
  Borrower cannot claim immunity to proceedings.
\item
  Whether there are any restrictions on providing guarantees and
  security.
\item
  Whether there are issues with financial assistances.
\end{itemize}

\hypertarget{reference-in-facility-agreement}{%
\subsection{Reference in Facility
Agreement}\label{reference-in-facility-agreement}}

\begin{env-31409176-8a91-46bb-85fd-48f2d6a04234}

Warning

The legal opinion is likely to be~referred to in the ``Representations''
section of the facility agreement and is likely to~be listed as a
``Condition Precedent'' in the facility agreement.

\end{env-31409176-8a91-46bb-85fd-48f2d6a04234}

\end{document}
