% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{article}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage[margin=1in]{geometry}
\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\newenvironment{Shaded}{}{}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{1.00,0.00,0.00}{\textbf{#1}}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textbf{\textit{#1}}}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.49,0.56,0.16}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.25,0.63,0.44}{#1}}
\newcommand{\BuiltInTok}[1]{#1}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{#1}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textit{#1}}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textbf{\textit{#1}}}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.53,0.00,0.00}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.00,0.44,0.13}{\textbf{#1}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.56,0.13,0.00}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.25,0.63,0.44}{#1}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.73,0.13,0.13}{\textit{#1}}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{1.00,0.00,0.00}{\textbf{#1}}}
\newcommand{\ExtensionTok}[1]{#1}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.25,0.63,0.44}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.02,0.16,0.49}{#1}}
\newcommand{\ImportTok}[1]{#1}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textbf{\textit{#1}}}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.00,0.44,0.13}{\textbf{#1}}}
\newcommand{\NormalTok}[1]{#1}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.40,0.40,0.40}{#1}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.00,0.44,0.13}{#1}}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.74,0.48,0.00}{#1}}
\newcommand{\RegionMarkerTok}[1]{#1}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{#1}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.73,0.40,0.53}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.10,0.09,0.49}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{#1}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textbf{\textit{#1}}}}
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\usepackage{xcolor}
\definecolor{aliceblue}{HTML}{F0F8FF}
\definecolor{antiquewhite}{HTML}{FAEBD7}
\definecolor{aqua}{HTML}{00FFFF}
\definecolor{aquamarine}{HTML}{7FFFD4}
\definecolor{azure}{HTML}{F0FFFF}
\definecolor{beige}{HTML}{F5F5DC}
\definecolor{bisque}{HTML}{FFE4C4}
\definecolor{black}{HTML}{000000}
\definecolor{blanchedalmond}{HTML}{FFEBCD}
\definecolor{blue}{HTML}{0000FF}
\definecolor{blueviolet}{HTML}{8A2BE2}
\definecolor{brown}{HTML}{A52A2A}
\definecolor{burlywood}{HTML}{DEB887}
\definecolor{cadetblue}{HTML}{5F9EA0}
\definecolor{chartreuse}{HTML}{7FFF00}
\definecolor{chocolate}{HTML}{D2691E}
\definecolor{coral}{HTML}{FF7F50}
\definecolor{cornflowerblue}{HTML}{6495ED}
\definecolor{cornsilk}{HTML}{FFF8DC}
\definecolor{crimson}{HTML}{DC143C}
\definecolor{cyan}{HTML}{00FFFF}
\definecolor{darkblue}{HTML}{00008B}
\definecolor{darkcyan}{HTML}{008B8B}
\definecolor{darkgoldenrod}{HTML}{B8860B}
\definecolor{darkgray}{HTML}{A9A9A9}
\definecolor{darkgreen}{HTML}{006400}
\definecolor{darkgrey}{HTML}{A9A9A9}
\definecolor{darkkhaki}{HTML}{BDB76B}
\definecolor{darkmagenta}{HTML}{8B008B}
\definecolor{darkolivegreen}{HTML}{556B2F}
\definecolor{darkorange}{HTML}{FF8C00}
\definecolor{darkorchid}{HTML}{9932CC}
\definecolor{darkred}{HTML}{8B0000}
\definecolor{darksalmon}{HTML}{E9967A}
\definecolor{darkseagreen}{HTML}{8FBC8F}
\definecolor{darkslateblue}{HTML}{483D8B}
\definecolor{darkslategray}{HTML}{2F4F4F}
\definecolor{darkslategrey}{HTML}{2F4F4F}
\definecolor{darkturquoise}{HTML}{00CED1}
\definecolor{darkviolet}{HTML}{9400D3}
\definecolor{deeppink}{HTML}{FF1493}
\definecolor{deepskyblue}{HTML}{00BFFF}
\definecolor{dimgray}{HTML}{696969}
\definecolor{dimgrey}{HTML}{696969}
\definecolor{dodgerblue}{HTML}{1E90FF}
\definecolor{firebrick}{HTML}{B22222}
\definecolor{floralwhite}{HTML}{FFFAF0}
\definecolor{forestgreen}{HTML}{228B22}
\definecolor{fuchsia}{HTML}{FF00FF}
\definecolor{gainsboro}{HTML}{DCDCDC}
\definecolor{ghostwhite}{HTML}{F8F8FF}
\definecolor{gold}{HTML}{FFD700}
\definecolor{goldenrod}{HTML}{DAA520}
\definecolor{gray}{HTML}{808080}
\definecolor{green}{HTML}{008000}
\definecolor{greenyellow}{HTML}{ADFF2F}
\definecolor{grey}{HTML}{808080}
\definecolor{honeydew}{HTML}{F0FFF0}
\definecolor{hotpink}{HTML}{FF69B4}
\definecolor{indianred}{HTML}{CD5C5C}
\definecolor{indigo}{HTML}{4B0082}
\definecolor{ivory}{HTML}{FFFFF0}
\definecolor{khaki}{HTML}{F0E68C}
\definecolor{lavender}{HTML}{E6E6FA}
\definecolor{lavenderblush}{HTML}{FFF0F5}
\definecolor{lawngreen}{HTML}{7CFC00}
\definecolor{lemonchiffon}{HTML}{FFFACD}
\definecolor{lightblue}{HTML}{ADD8E6}
\definecolor{lightcoral}{HTML}{F08080}
\definecolor{lightcyan}{HTML}{E0FFFF}
\definecolor{lightgoldenrodyellow}{HTML}{FAFAD2}
\definecolor{lightgray}{HTML}{D3D3D3}
\definecolor{lightgreen}{HTML}{90EE90}
\definecolor{lightgrey}{HTML}{D3D3D3}
\definecolor{lightpink}{HTML}{FFB6C1}
\definecolor{lightsalmon}{HTML}{FFA07A}
\definecolor{lightseagreen}{HTML}{20B2AA}
\definecolor{lightskyblue}{HTML}{87CEFA}
\definecolor{lightslategray}{HTML}{778899}
\definecolor{lightslategrey}{HTML}{778899}
\definecolor{lightsteelblue}{HTML}{B0C4DE}
\definecolor{lightyellow}{HTML}{FFFFE0}
\definecolor{lime}{HTML}{00FF00}
\definecolor{limegreen}{HTML}{32CD32}
\definecolor{linen}{HTML}{FAF0E6}
\definecolor{magenta}{HTML}{FF00FF}
\definecolor{maroon}{HTML}{800000}
\definecolor{mediumaquamarine}{HTML}{66CDAA}
\definecolor{mediumblue}{HTML}{0000CD}
\definecolor{mediumorchid}{HTML}{BA55D3}
\definecolor{mediumpurple}{HTML}{9370DB}
\definecolor{mediumseagreen}{HTML}{3CB371}
\definecolor{mediumslateblue}{HTML}{7B68EE}
\definecolor{mediumspringgreen}{HTML}{00FA9A}
\definecolor{mediumturquoise}{HTML}{48D1CC}
\definecolor{mediumvioletred}{HTML}{C71585}
\definecolor{midnightblue}{HTML}{191970}
\definecolor{mintcream}{HTML}{F5FFFA}
\definecolor{mistyrose}{HTML}{FFE4E1}
\definecolor{moccasin}{HTML}{FFE4B5}
\definecolor{navajowhite}{HTML}{FFDEAD}
\definecolor{navy}{HTML}{000080}
\definecolor{oldlace}{HTML}{FDF5E6}
\definecolor{olive}{HTML}{808000}
\definecolor{olivedrab}{HTML}{6B8E23}
\definecolor{orange}{HTML}{FFA500}
\definecolor{orangered}{HTML}{FF4500}
\definecolor{orchid}{HTML}{DA70D6}
\definecolor{palegoldenrod}{HTML}{EEE8AA}
\definecolor{palegreen}{HTML}{98FB98}
\definecolor{paleturquoise}{HTML}{AFEEEE}
\definecolor{palevioletred}{HTML}{DB7093}
\definecolor{papayawhip}{HTML}{FFEFD5}
\definecolor{peachpuff}{HTML}{FFDAB9}
\definecolor{peru}{HTML}{CD853F}
\definecolor{pink}{HTML}{FFC0CB}
\definecolor{plum}{HTML}{DDA0DD}
\definecolor{powderblue}{HTML}{B0E0E6}
\definecolor{purple}{HTML}{800080}
\definecolor{red}{HTML}{FF0000}
\definecolor{rosybrown}{HTML}{BC8F8F}
\definecolor{royalblue}{HTML}{4169E1}
\definecolor{saddlebrown}{HTML}{8B4513}
\definecolor{salmon}{HTML}{FA8072}
\definecolor{sandybrown}{HTML}{F4A460}
\definecolor{seagreen}{HTML}{2E8B57}
\definecolor{seashell}{HTML}{FFF5EE}
\definecolor{sienna}{HTML}{A0522D}
\definecolor{silver}{HTML}{C0C0C0}
\definecolor{skyblue}{HTML}{87CEEB}
\definecolor{slateblue}{HTML}{6A5ACD}
\definecolor{slategray}{HTML}{708090}
\definecolor{slategrey}{HTML}{708090}
\definecolor{snow}{HTML}{FFFAFA}
\definecolor{springgreen}{HTML}{00FF7F}
\definecolor{steelblue}{HTML}{4682B4}
\definecolor{tan}{HTML}{D2B48C}
\definecolor{teal}{HTML}{008080}
\definecolor{thistle}{HTML}{D8BFD8}
\definecolor{tomato}{HTML}{FF6347}
\definecolor{turquoise}{HTML}{40E0D0}
\definecolor{violet}{HTML}{EE82EE}
\definecolor{wheat}{HTML}{F5DEB3}
\definecolor{white}{HTML}{FFFFFF}
\definecolor{whitesmoke}{HTML}{F5F5F5}
\definecolor{yellow}{HTML}{FFFF00}
\definecolor{yellowgreen}{HTML}{9ACD32}
\usepackage[most]{tcolorbox}

\usepackage{ifthen}
\provideboolean{admonitiontwoside}
\makeatletter%
\if@twoside%
\setboolean{admonitiontwoside}{true}
\else%
\setboolean{admonitiontwoside}{false}
\fi%
\makeatother%
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

\author{}
\date{}

\begin{document}

{
\setcounter{tocdepth}{3}
\tableofcontents
}
\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{min\_depth: 1}
\end{Highlighting}
\end{Shaded}

\hypertarget{the-legal-profession}{%
\section{The Legal Profession}\label{the-legal-profession}}

\hypertarget{professional-conduct}{%
\subsection{Professional Conduct}\label{professional-conduct}}

\begin{itemize}
\tightlist
\item
  Being a professional, the public must be able to trust a solicitor.
\item
  Must be trusted to act in best interests of the client
\item
  Must comply with general regulatory requirements such as Bribery Act
  2010. Mostly the onus on firms is to have adequate procedures in place
  to prevent bad stuff happening.
\item
  SRA produced additional rules and regulations

  \begin{itemize}
  \tightlist
  \item
    Breaching these has severe consequences, including being struck off
  \end{itemize}
\end{itemize}

\hypertarget{legal-services-board}{%
\subsection{Legal Services Board}\label{legal-services-board}}

\begin{itemize}
\tightlist
\item
  Created by Legal Services Act 2007
\item
  Supervises the `approved regulators' of the legal profession,
  including the SRA (and the Bar Council etc.)
\end{itemize}

\hypertarget{sra}{%
\subsection{SRA}\label{sra}}

Established by The Law Society in Jan 2007. Purpose is to protect the
public by ensuring that solicitors meet high standards. Governed by a
board of 11 members, 5 of which are solicitors. HQ in Birmingham. Budget
funded by practising certificate fee.

\hypertarget{law-society}{%
\subsection{Law Society}\label{law-society}}

Representative body for solicitors in England and Wales. Founded in
1825. Lobbies the government about changes in the law, lobbies for
better working conditions for solicitors etc.

\begin{itemize}
\tightlist
\item
  Membership is optional.
\item
  Governed by a Council
\item
  There are 54 local law societies around the country
\end{itemize}

\hypertarget{aspects-of-practice}{%
\subsection{Aspects of Practice}\label{aspects-of-practice}}

Different ways in which law can be practised::

\begin{itemize}
\tightlist
\item
  Sole practice

  \begin{itemize}
  \tightlist
  \item
    Becoming less common (Westlaw and stuff is hella expensive)
  \item
    Practicality: going on holidays without having to hire a locum.
  \item
    Authorised in the same way as other types of firm
  \end{itemize}
\item
  As a recognised body

  \begin{itemize}
  \tightlist
  \item
    Legal services body (75\% of managers legally qualified \&
    non-solicitor managers approved by SRA) can apply for authorisation
    by the SRA as a recognised body.
  \item
    Recognised forms include:

    \begin{itemize}
    \tightlist
    \item
      Partnerships

      \begin{itemize}
      \tightlist
      \item
        Classic law firm structure: partners and then everyone else.
        Some large firms have salaried and equity partners.
      \end{itemize}
    \item
      Limited Liability Partnerships

      \begin{itemize}
      \tightlist
      \item
        Instead of partners, LLPs have members, who share in the
        profits.
      \end{itemize}
    \item
      Companies

      \begin{itemize}
      \tightlist
      \item
        Incorporate under Companies Act 2006
      \end{itemize}
    \end{itemize}
  \item
    Once authorised, a body is entitled to carry on all `reserved legal
    activities' (except notary services). This is defined in s 12 Legal
    Services Act 2007.
  \item
    The recognised body must have a Compliance Officer for Legal
    Practice (COLP) and Compliance Officer for Finance and
    Administration (COFA) and whom the SRA has approved.
  \end{itemize}
\item
  As a Licensed Body

  \begin{itemize}
  \tightlist
  \item
    The Legal Services Act 2007 also introduced alternative business
    structures, allowing non-lawyers to offer legal services alongside
    non-legal services. Also allowed non-lawyers to own and invest in
    firms providing legal services.
  \item
    Eligibility: must be one manager authorised by the SRA or another
    approved regular. Must also be a licensable body (confusing
    definition).
  \item
    Once authorised, a licensed body is entitled to carry out the same
    range of activities as a recognised body.
  \item
    Similarly as recognised bodies, must have a Head of Finance and
    Administration and Head of Legal Practice.
  \end{itemize}
\end{itemize}

\hypertarget{new-changes}{%
\subsubsection{New Changes}\label{new-changes}}

New SRA Standards and Regulations 2019: solicitors allowed to deliver
non-reserved legal activities to the public from within a business not
regulated by a legal services regulator. So e.g., a big four accountancy
firm could offer family/ employment law services without going through
the full authorisation process, but would not be able to hold client
money.

Further, individual solicitors can now work and offer legal services on
a freelance basis. Will not have to maintain the same level of client
protection usually required. Does not need to be authorised if:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Practice consists entirely of non-reserved activities, or
\item
  Any reserved legal activities are provided through an authorised body/
  authorised non-SRA firm.
\end{enumerate}

\hypertarget{employed-solicitors}{%
\subsubsection{Employed Solicitors}\label{employed-solicitors}}

This is a solicitor who is employed by a non-legal services employer
(i.e., in-house counsel). Subject to same rules as lawyers in private
practice.

\hypertarget{registered-european-lawyers-and-registered-foreign-lawyers}{%
\subsubsection{Registered European Lawyers and Registered Foreign
Lawyers}\label{registered-european-lawyers-and-registered-foreign-lawyers}}

Foreign lawyers who want to become managers owners of law firms
authorised by the SRA must become RFLs. RELs are no longer a thing after
Brexit, apart from the Swiss until 2024. These classifications allow
foreign lawyers to provide all unreserved legal services, and some
reserved legal services.

\hypertarget{law-centres}{%
\subsubsection{Law Centres}\label{law-centres}}

There are 43 of them nationwide. Provide free legal advice to
disadvantaged individuals within communities. Must take out appropriate
indemnity insurance.

\hypertarget{legal-advice-centres}{%
\subsubsection{Legal Advice Centres}\label{legal-advice-centres}}

No casework is undertaken, but legal advice provided. ULaw and BPP run
their own.

\hypertarget{duty-solicitor-schemes}{%
\subsubsection{Duty Solicitor Schemes}\label{duty-solicitor-schemes}}

Police station duty solicitor scheme aims to provide advice and
assistance to people at police stations who would otherwise not be
legally represented The court duty solicitor scheme aims to provide
emergency legal representation at magistrates' courts on a rota basis.
Fees for both cases are paid by the Legal Aid Agency to contracted
solicitors.

\hypertarget{regulating-the-profession}{%
\section{Regulating the Profession}\label{regulating-the-profession}}

\hypertarget{history}{%
\subsection{History}\label{history}}

\begin{itemize}
\tightlist
\item
  The SRA was created in anticipation of the Legal Services Act 2007.
  The Legal Services Board supervises the SRA's regulation of the
  profession.
\item
  Since 2006, alternative business structures have enabled lawyers and
  non-lawyers to share the management and control of businesses
  providing legal services.

  \begin{itemize}
  \tightlist
  \item
    The SRA undertook a complete review of the regulation framework in
    anticipation of this.
  \item
    Resulted in the SRA Code of Conduct 2011, replacing the Solicitors'
    Code of Conduct 2007
  \item
    This was much shorter and less prescriptive -- included more general
    principles
  \end{itemize}
\item
  2014: SRA began a programme of reform → new simplified Standards and
  Regulations: the two Codes of Conduct.

  \begin{itemize}
  \tightlist
  \item
    Much shorter than the 2011 Code
  \item
    New codes and regulations came into effect 25/11/19.
  \end{itemize}
\end{itemize}

\hypertarget{sra-standards-and-regulations}{%
\subsection{SRA Standards and
Regulations}\label{sra-standards-and-regulations}}

\begin{itemize}
\tightlist
\item
  Apply to everyone regulated by the SRA
\item
  Consist of Principles, the two Codes, SRA Accounts Rules and some
  others.
\item
  7 Principles which are mandatory and apply to all those regulated by
  the SRA
\item
  SRA Code of Conduct for Solicitors, Registered European Lawyers (RELs)
  and Registered Foreign Lawyers (RFLs) applies to individuals
  authorised by the SRA to provide legal services.
\item
  Code of Conduct for Firms applies to all bodies authorised by the SRA
  to provide legal services. Indicates that a serious failure can result
  in the SRA taking action against the body itself as an entity, rather
  than just individual solicitors.
\item
  SRA Overseas and Cross-border Practice Rules Part A (Overseas Rules)
  apply to regulated individuals practising overseas and authorised
  bodies with an overseas practice.
\item
  Part B (Cross-border Practice Rules applies to European cross-border
  practice
\end{itemize}

\hypertarget{complaints}{%
\subsection{Complaints}\label{complaints}}

Solicitor must establish and maintain/ participate in a procedure for
handling complaints (para 8.2 Code of Conduct for Solicitors). Must
ensure clients are informed in writing about

\begin{itemize}
\tightlist
\item
  Their right to complain about solicitor's services and charges
\item
  How complaints may be made and to whom
\item
  Any right to make a complaint to the Legal Ombudsman.
\end{itemize}

Information about complaints must be published on a firm's website (SRA
Transparency Rules).

If a client complains and it is not resolved in 8 weeks, solicitor must
inform client in writing about

\begin{itemize}
\tightlist
\item
  any right to complain to Legal Ombudsman and details of how to do this
\item
  If complaints procedure has been exhausted:

  \begin{itemize}
  \tightlist
  \item
    That the solicitor cannot settle the complaint
  \item
    Name and website of an alternative dispute resolution approved body
  \item
    Whether you agree to use a scheme operated by that body.
  \end{itemize}
\end{itemize}

The Chartered Trading Standards Institute (CTSI) has approved a number
of ADR entities.

\hypertarget{the-legal-ombudsman}{%
\subsection{The Legal Ombudsman}\label{the-legal-ombudsman}}

\hypertarget{details}{%
\subsubsection{Details}\label{details}}

\begin{itemize}
\tightlist
\item
  Legal Ombudsman for England and Wales (`LeO') became operational on
  6/10/10.
\item
  Set up by the Office for Legal Complaints under the Legal Services Act
  2007.
\item
  Deals with complaints against all types of lawyers.
\item
  Replaced Legal Complaints Service
\item
  Only certain types of client can complain to LeO, including
  individuals, small business, charities, clubs and associations and
  personal representatives/ beneficiaries of an estate.
\item
  Complain must relate to the act/ omission of a solicitor, or to
  services they offered/ provided/ refused to provide.
\end{itemize}

\hypertarget{when-used}{%
\subsubsection{When Used}\label{when-used}}

Can use LeO when:

\begin{itemize}
\tightlist
\item
  Complaint has not been resolved within 8 weeks of being made to
  solicitor
\item
  LeO considers there are exceptional reasons to consider the complaint
  sooner
\item
  LeO considers that in-house resolution is not possible
\end{itemize}

\hypertarget{time-limits}{%
\subsubsection{Time Limits}\label{time-limits}}

Should be brought no later than:

\begin{itemize}
\tightlist
\item
  6 years from act/ omission; or
\item
  3 years from when the complainant should reasonably have known there
  was cause for complaint.
\end{itemize}

But LeO has discretion to extend these limits to the extent it considers
fair. A hearing will only be held where LeO considers that the complaint
cannot be determined fairly without one.

\hypertarget{resolution}{%
\subsubsection{Resolution}\label{resolution}}

A `determination' will then be made and sent to the parties and SRA.
Complainant can accept or reject this. May direct the lawyer to:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Apologise
\item
  Pay compensation (+ interest)
\item
  Ensure putting right of any error or omission
\item
  Take specified action in the interests of the complainant
\item
  Limit solicitor's fees.
\end{enumerate}

Limit of £50,000 on total value that may be awarded in respect of (3)
and (4) - though pay awards for the others are not limited.

\hypertarget{possible-outcomes}{%
\subsubsection{Possible Outcomes}\label{possible-outcomes}}

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.2083}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.7917}}@{}}
\toprule()
\begin{minipage}[b]{\linewidth}\raggedright
Complainant's response
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Outcome
\end{minipage} \\
\midrule()
\endhead
Accepts the determination & Determination is binding on the parties and
final. Can be enforced in court by the complainant. \\
Rejects the determination & Parties free to pursue other legal
remedies. \\
\bottomrule()
\end{longtable}

The LeO can refer a particular legal question to court if unsure. If
professional misconduct is alleged in a complaint, LeO will inform the
SRA.

\hypertarget{breach-of-professional-conduct}{%
\subsubsection{Breach of Professional
Conduct}\label{breach-of-professional-conduct}}

\begin{itemize}
\tightlist
\item
  Professional misconduct primarily concerns breaches of the SRA's Codes
  of Conduct.
\item
  Complaints primarily concerning a breach of professional conduct will
  be dealt with by the SRA.
\item
  Role of the SRA is to protect the public. If the SRA is informed of
  alleged misconduct, it will investigate. The individual/ firm will be
  given notice and invited to respond with written recommendations.
\item
  Statutory power to carry out intervention into a practice.
\end{itemize}

\hypertarget{enforcement-strategy}{%
\paragraph{Enforcement Strategy}\label{enforcement-strategy}}

The SRA Enforcement Strategy sets out how the SRA will use its
enforcement powers and what it expects of those it regulates. Approach
to enforcement:

\begin{itemize}
\tightlist
\item
  Focus on the most serious breaches
\item
  Take account of aggravating or mitigating factors
\item
  Emphasise constructive engagement to support firms and individuals to
  achieve compliance.
\end{itemize}

The position of the firm and the individuals inside it will be
considered when deciding who to enforce against.

\hypertarget{outcomes}{%
\paragraph{Outcomes}\label{outcomes}}

The SRA may decide to do the following:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Take no further action with or without issuing advice or a warning
  about future conduct
\item
  Impose a financial penalty or written rebuke

  \begin{enumerate}
  \def\labelenumii{\arabic{enumii}.}
  \tightlist
  \item
    Max penalty £2,000 for solicitors and their employees, £250 million
    for licensed bodies, £50 million for managers and employees of
    licensed bodies).
  \item
    Since 2008, SRA has made regulatory decisions public.
  \end{enumerate}
\item
  Control how the firm/ regulated person practises

  \begin{enumerate}
  \def\labelenumii{\arabic{enumii}.}
  \tightlist
  \item
    Can impose conditions or suspend a solicitor's practising
    certificate
  \end{enumerate}
\item
  Refer the matter to the Solicitors Disciplinary Tribunal

  \begin{enumerate}
  \def\labelenumii{\arabic{enumii}.}
  \tightlist
  \item
    Tribunal is independent from the SRA
  \end{enumerate}
\item
  Disqualify an individual from actin as a HOLP, HOFA, manager or
  employee of a licenced body.
\end{enumerate}

\hypertarget{complaints-about-professional-services-vs-negligence}{%
\paragraph{Complaints About Professional Services Vs
Negligence}\label{complaints-about-professional-services-vs-negligence}}

Solicitor owes a duty of care in tort to their client. Solicitor breach
+ loss suffered = solicitor can be sued by the client for negligence.

For LeO to get involved, the client does not need to have suffered any
loss -- it is enough that services were not of a reasonable quality.

SRA and LeO have no power to adjudicate issues of negligence in a legal
sense, but where there is overlap they can act.

\hypertarget{solicitors-disciplinary-tribunal}{%
\paragraph{Solicitors Disciplinary
Tribunal}\label{solicitors-disciplinary-tribunal}}

\hypertarget{summary}{%
\subparagraph{Summary}\label{summary}}

Hears and determines applications relating to allegations of unbefitting
conduct/ breaches of requirements of professional conduct by solicitors
and others, including ABSs. Most applications to the Disciplinary
Tribunal are made on behalf of the SRA, though people can also make an
application to the SDT directly without going through the SRA.

\hypertarget{structure}{%
\subparagraph{Structure}\label{structure}}

\begin{itemize}
\tightlist
\item
  Established by s 46 Solicitors Act 1974.
\item
  Members appointed by the Master of the Rolls
\item
  Currently 45 members - about 2/3 solicitors
\item
  Sits in divisions of 3 members
\item
  Civil standard of proof.
\end{itemize}

\hypertarget{procedure}{%
\subparagraph{Procedure}\label{procedure}}

\begin{itemize}
\tightlist
\item
  Disciplinary Tribunal has the power to make rules governing its
  procedure and practice.
\item
  Maintains a panel of solicitors in private practice
\item
  No investigative powers
\end{itemize}

\hypertarget{application-to-the-tribunal}{%
\subparagraph{Application to the
Tribunal}\label{application-to-the-tribunal}}

\begin{itemize}
\tightlist
\item
  Must be made in the prescribed form, supported by evidence
\item
  Hearing date will be fixed where the Tribunal is satisfied that there
  is a case to answer
\item
  Each party can be represented by a solicitor or counsel.
\item
  Evidence given under oath
\item
  Can call witnesses
\item
  Decisions called `Judgments' or `Orders'
\item
  Order takes effect once filed with The Law Society
\item
  Where an application is made directly to the Tribunal, they can refer
  it to the SRA for investigation.
\end{itemize}

\hypertarget{powers}{%
\subparagraph{Powers}\label{powers}}

Can make orders including:

\begin{itemize}
\tightlist
\item
  Striking a solicitor off the roll
\item
  Suspending a solicitor from practice or imposing restrictions
\item
  Imposing an unlimited fine
\item
  Reprimanding solicitor
\item
  Requiring payment of costs/ contribution towards costs
\item
  Restoring a solicitor to the roll
\end{itemize}

\hypertarget{appeals}{%
\subparagraph{Appeals}\label{appeals}}

Disciplinary Tribunal orders can be appealed at the Administrative
Court.

\hypertarget{other-sra-powers}{%
\paragraph{Other SRA Powers}\label{other-sra-powers}}

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.0576}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.9424}}@{}}
\toprule()
\begin{minipage}[b]{\linewidth}\raggedright
Power
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Details
\end{minipage} \\
\midrule()
\endhead
Intervention & Aim is to protect client interests and money. Practical
effect is that a practice is closed and practice monies vest in the SRA.
Agent is appointed by the SRA to carry out the intervention. May also
occur when a sole practitioner dies/ is unable to continue practising
suddenly. \\
Delivery of files & To investigate complains, SRA has the power (s 44B
Solicitors Act 1974) to serve notice on a solicitor/ recognised body
requiring the delivery of a file/ documents in the possession of the
solicitor to the SRA. Code of Conduct obliges solicitor/ firm to respond
promptly. \\
\bottomrule()
\end{longtable}

\hypertarget{powers-of-the-court}{%
\paragraph{Powers of the Court}\label{powers-of-the-court}}

A solicitor is an officer of the court. So the court has jurisdiction to
discipline a solicitor in respects of any matter before the court.

\hypertarget{negligence}{%
\paragraph{Negligence}\label{negligence}}

A solicitor may be sued by their client in the tort of negligence, or
for breach of contract. Firms must carry compulsory indemnity insurance
against such actions.

\hypertarget{steps-to-take}{%
\subparagraph{Steps to Take}\label{steps-to-take}}

Code of Conduct states that solicitors and firms must be `honest and
open' with clients if things go wrong, and if a client suffers loss or
harm as a result, matters are to be put right and a full and prompt
explanation given.

If requested by the SRA, solicitor or firm must investigate whether
anyone may have a claim against them and report back to SRA.

Generally, if a client says they wish to sue a solicitor, there will
likely be a conflict of interest. Law Society guidance states solicitors
should respond to complaints even if they believe the allegations
concern negligence and a referral to the firm's insurance provider will
be necessary.

\hypertarget{sra-compensation-fund}{%
\paragraph{SRA Compensation Fund}\label{sra-compensation-fund}}

Maintained by the SRA under ss 36 and 36A of the Solicitors Act 1974.
Governed by the SRA Compensation Fund Rules and exists as a
discretionary fund of last resort (e.g., if money has been stolen or
otherwise not accounted for as the result of an act or omission of those
regulated by the SRA). Funded by mandatory contributions by all
solicitors and SRA authorised bodies. Applications can be made where an
individual/ organisation has suffered loss due to a defaulting
practitioner's dishonesty/ failure to account for money they received.

The fund does not pay out more than £2 million/ claim (unless there are
exceptional circumstances). For multiple related claims, the threshold
is £5 million.

Where payment is made from the Fund, the SRA will seek to be subrogated
to the rights of the applicant and can take proceedings against the
defaulting solicitor to recover the amount paid.

Codes of Conduct impose an obligation on solicitors and firms to ensure
clients understand the regulatory protections available to them.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{(1) The SRA is em powered by the Solicitors Act 1974 to publish rules and requirements  }
\NormalTok{for the regulation of conduct of solicitors, recog nised/licensed bodies, m embers of recognised/licensed bodies, and em ployees of recognis ed/licens ed bodies/sole  }
\NormalTok{practices.  }
\NormalTok{(2) These rules and requirements are enforced by the SRA and the Leg al Ombudsm an.  }
\NormalTok{(3) The SRA deals with breaches of the requirements of professional conduct, whereas the Legal Ombudsm an primarily deals with com plaints about services provided by legal practitioners, including solicitors.}
\NormalTok{(4) The Legal Ombudsm an deals with com plaints relating to an act/om iss ion by a legal practitioner and which relates to services provided by the practitioner. The client does not need to suffer any loss as a result of the service provide d in order to raise the issue.  }
\NormalTok{(5) The SRA may impose disciplinary sanctions for a breach of the requirem ents of  }
\NormalTok{professional conduct, or in serious cas es may refer the m atter to the Solicitors  }
\NormalTok{Disciplinary Tribunal.  }
\NormalTok{(6) The Legal Ombudsman has num erous sanctions at its disposal. These include  }
\NormalTok{directing the solicitor to limit their fees, or directing the solicitor to com pensate the client up to £50,000. The SRA/Legal Ombudsman has no power to pay  }
\NormalTok{compensation to the client itself.  }
\NormalTok{(7) The Solicitors Disciplinary Tribunal m ay fine a solicitor, suspend the solicitor or, as an ultim ate sanction, strike the solicitor of f the roll.  }
\NormalTok{(8) A solicitor will be negligent if they breach a duty of care to their client and  }
\NormalTok{foreseeable loss results as a consequence of that breach. Where the solicitor has  }
\NormalTok{been negligent, they may be sued by the client for damages.  }
\NormalTok{(9) The SRA Compensation Fund is maintained by the SRA. Payment m ay be made from  }
\NormalTok{the Fund when a client has suffered loss as a result of a defaulting practitioner’s act or om iss ion or to cover claims against uninsured firms.}
\end{Highlighting}
\end{Shaded}

\hypertarget{sra-principles}{%
\section{SRA Principles}\label{sra-principles}}

\hypertarget{introduction}{%
\subsection{Introduction}\label{introduction}}

The profession is regulated by the Government, the FCA and the SRA. The
SRA Principles are all-pervasive and underpin all the SRA's regulatory
arrangements.

The Principles are mandatory and aply to all individuals and bodies
regulated by the SRA. They apply outside of a solicitor's practice as
well as within it.

\hypertarget{justice-and-the-rule-of-law}{%
\subsection{1. Justice and the Rule of
Law}\label{justice-and-the-rule-of-law}}

Principle 1 provides that a soolicitor must act in a way that upholds
the constitutional rule of law and proper administration of justice.

\hypertarget{public-trust}{%
\subsection{2. Public Trust}\label{public-trust}}

A solicitor must act in a way that upholds public trust and confidence
in the solicitors' profession and in legal services. A solicitor may
breach Principle 2 through behaviour outside their solicitors' practice
(e.g., getting a criminal conviction).

\hypertarget{independence}{%
\subsection{3. Independence}\label{independence}}

A solicitor must act with independence. This includes indepedence from
the client - need to be mindful of their wider obligations to the
profession as a whole.

\hypertarget{honesty}{%
\subsection{4. Honesty}\label{honesty}}

Where a solicitor is convicted of an offence involving dishonesty,
likely that they will be struck off the roll of solicitors. The standard
of honesty required for solicitors is that they may be `trusted to the
ends of the earth' ({[}{[}Bolton v Law Society {[}1993{]} EWCA Civ
32{]}{]}).

\hypertarget{integrity}{%
\subsection{5. Integrity}\label{integrity}}

A solicitor must act with integrity; they are in a position of trust.
The concept of integrity is wider than just dishonesty - refers to the
higher standards which society expects of a professional.

\hypertarget{equality-diversity-and-inclusion}{%
\subsection{6. Equality, Diversity and
Inclusion}\label{equality-diversity-and-inclusion}}

Solicitors and firms must act in a way that encourages equality,
diversity and inclusion.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{{-} Making reasonable adjustments for disabled clients}
\NormalTok{{-} Treating employees fairly}
\NormalTok{{-} Having processes in place to deal with bullying, harassment and discrimination.}
\end{Highlighting}
\end{Shaded}

\hypertarget{best-interests-of-the-client}{%
\subsection{7. Best Interests of the
Client}\label{best-interests-of-the-client}}

A solicitor must act in the best interests of each client. Derives from
the common law and the fiduciary relationship.

This includes ensuring clients are provided with a proper standard of
service in terms of client care, competence and standard of work.

\hypertarget{confidentiality}{%
\section{Confidentiality}\label{confidentiality}}

\hypertarget{introduction-1}{%
\subsection{Introduction}\label{introduction-1}}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{title: Para 6.3 Codes of Conduct}
\NormalTok{You [must] keep the affairs of current and former clients confidential unless disclosure is required or permitted by law or the client consents. }
\end{Highlighting}
\end{Shaded}

\begin{itemize}
\tightlist
\item
  Obligation of confidentiality extends beyond obviously confidential
  information given to a solicitor by their client and includes all
  information about a client or matter, regardless of the source of
  information.
\item
  Firms must have good systems in place -- Code of Conduct for Firms
\item
  Duty continues after the retainer has been terminated.
\item
  Duty continues after the death of the client.
\item
  Any breach of the duty of confidentiality will be a breach of
  professional conduct.
\item
  Practical steps should be taken to safeguard the identity of their
  clients.
\item
  Beware in particular about mergers and solicitors moving firms.
\end{itemize}

\hypertarget{acceptable-disclosure}{%
\subsection{Acceptable Disclosure}\label{acceptable-disclosure}}

\begin{itemize}
\tightlist
\item
  Disclosure is permitted with the client's consent -- take steps to
  ensure the client knows the exact purpose.
\item
  Can also disclose pursuant to a statutory requirement in certain
  circumstances.
\item
  Can disclose under a court order or police warrant.
\end{itemize}

In the following situations, disclosure is still technically a breach
but is a mitigating factor in the context of disciplinary action:

\begin{itemize}
\tightlist
\item
  Client indicating their intention to commit suicide or serious
  self-harm

  \begin{itemize}
  \tightlist
  \item
    First ask for consent to disclose to third party
  \item
    May decide to disclose anyway
  \end{itemize}
\item
  Preventing harm to children or vulnerable adults

  \begin{itemize}
  \tightlist
  \item
    No requirement in law to disclose this information, but personal
    conscience
  \end{itemize}
\item
  Preventing the commission of a criminal offence

  \begin{itemize}
  \tightlist
  \item
    If a solicitor is being used by a client to perpetrate a crime, duty
    of confidentiality does not arise (no confidence under common law)
  \end{itemize}
\end{itemize}

\hypertarget{duty-of-disclosure}{%
\subsection{Duty of Disclosure}\label{duty-of-disclosure}}

Where a solicitor is acting for a client, they must make the client
aware of all information material to the matter of which the solicitor
has knowledge (para 6.4, Code of Conduct). This is a personal duty.

\hypertarget{limitations}{%
\subsubsection{Limitations}\label{limitations}}

There are specific exceptions:

\begin{itemize}
\tightlist
\item
  Disclosure of information which is prohibited by legal restrictions in
  the interests of national security/ the prevention of crime
\item
  Client gives informed consent (evidenced in writing) to the
  information not being disclosed to them
\item
  Solicitor has reason to believe that serious physical or mental injury
  will be caused to the client or another if the information is
  disclosed to the client
\item
  The information is contained in a privileged document that the
  solicitor has knowledge of only because mistakenly disclosed.
\end{itemize}

\hypertarget{confidential-material-received-by-mistake}{%
\subsubsection{Confidential Material Received by
Mistake}\label{confidential-material-received-by-mistake}}

Occasionally, mistakes may be made, and confidential information
accidentally sent to the wrong party. The receiving solicitor must
immediately return the papers to the original solicitor without reading
them ({[}{[}Ablitt v Mills \& Reeve (1995){]}{]}).

\hypertarget{placing-confidential-information-at-risk}{%
\subsection{Placing Confidential Information at
Risk}\label{placing-confidential-information-at-risk}}

A solicitor or firm must not act in certain circumstances, where the
client has an interest \textbf{adverse} to the interest of another
current or former client for whom confidential information is held,
unless either of the exceptions set out in the Codes is met (para 6.5 of
both codes).

\hypertarget{material-information}{%
\subsubsection{Material Information}\label{material-information}}

For the purpose of s 6.5, the confidential information must be
`material'. This is information which is:

\begin{itemize}
\tightlist
\item
  Relevant to the client's matter; and
\item
  Of more than inconsequential interest to the client
\end{itemize}

\hypertarget{adverse-interest}{%
\subsubsection{Adverse Interest}\label{adverse-interest}}

Adversity relates to the relationship between two respective clients. A
relationship is adverse if the clients are or are likely to become
opposing parties in a matter. This includes both litigation and
negotiation.

\hypertarget{when-it-is-possible-to-act}{%
\subsubsection{When it is Possible to
Act}\label{when-it-is-possible-to-act}}

When the same individual holds confidential information for one party
which is relevant to another client, the solicitor will be unable to act
for both clients unless the client to whom the duty of confidentiality
is owed consents to the information being disclosed or the other client
gives informed consent (evidenced in writing) to the information not
being disclosed to them. This is regardless of any adverse interests
between parties -- since the duties of disclosure and confidentiality
are \textbf{personal} to the solicitor.

There are some exceptions:

\begin{itemize}
\tightlist
\item
  Effective measures have been taken which result in there being no real
  risk of disclosure of the confidential information.
\item
  A current or former client whose information is held has given
  informed consent (evidenced in writing), including to any measures
  taken to protect their personal information.
\end{itemize}

More likely that these exceptions will apply in practice wrt a former
client.

\hypertarget{effective-measures-to-safeguard-information}{%
\subsubsection{Effective Measures to Safeguard
Information}\label{effective-measures-to-safeguard-information}}

Effective measures include information barriers (previously, Chinese
walls). This could be achieved through separate teams handling the
matters, separate servers and printers, information being encrypted and
password protected etc.

They are not appropriate for small firms.

\hypertarget{informed-consent}{%
\subsubsection{Informed Consent}\label{informed-consent}}

The client must consent after having understood and considered the risks
and rewards involved in the situation. It is often not possible to
disclose sufficient information about the identity and business of the
other client without breaching the client's confidentiality; something
to bear in mind. Generally, informed consent is only of use to
sophisticated clients.

\hypertarget{professional-embarrassment}{%
\subsubsection{Professional
Embarrassment}\label{professional-embarrassment}}

A solicitor may still be obliged to refuse to act on the grounds of
professional embarrassment. A firm should decline to act where the
information which it cannot disclose would cause severe embarrassment to
the firm if it came out.

The firm should also consider the Principles.

\hypertarget{confidentiality-and-privilege}{%
\subsubsection{Confidentiality and
Privilege}\label{confidentiality-and-privilege}}

Confidentiality is distinct from the issue of legal professional
privilege.

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.2137}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.7863}}@{}}
\toprule()
\begin{minipage}[b]{\linewidth}\raggedright
Principle
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Description
\end{minipage} \\
\midrule()
\endhead
Confidentiality & Prevents a solicitor from disclosing any information
relating to a client without the client's consent. \\
Legal professional privilege & Allows a solicitor to withhold specific
information which they would otherwise be required to disclose (e.g., in
court proceedings) \\
\bottomrule()
\end{longtable}

Legal professional privilege applies to information between a solicitor
and client, whether written or oral, direct or indirect.

Limitations are placed to prevent it being used as a cloak.

\hypertarget{legal-advice-privilege}{%
\paragraph{Legal Advice Privilege}\label{legal-advice-privilege}}

One type of legal professional privilege is legal advice privilege,
which applies only to information passed between the client and a
solicitor acting in the capacity of solicitor. Documents sent to or from
an independent third party are not covered by this privilege; `copying
in' a solicitor will not mean the information can be withheld. If the
communication is made for the purpose of committing a fraud or crime, it
will not attract privilege ({[}{[}R v Cox and Railton (1884){]}{]}).

\hypertarget{litigation-privilege}{%
\paragraph{Litigation Privilege}\label{litigation-privilege}}

May be claimed in respect of documents created for the sole/ dominant
purpose of litigation which has commenced/ is contemplated. This extends
to communications between a solicitor and third parties.

Anything said by a solicitor whilst speaking in court as an advocate is
privileged, so a solicitor cannot be sued for defamation in such
circumstances.

{[}{[}Three Rivers DC v Bank of England {[}2004{]} UKHL 48{]}{]}
examines the limits of legal professional privilege. Privilege continues
beyond the death of a client ({[}{[}Bullivant v AG for Victoria
{[}1901{]}{]}{]}).

\hypertarget{conflicts-of-interest}{%
\section{Conflicts of Interest}\label{conflicts-of-interest}}

Solicitors must act in the best interests of their clients (Principle
7). If the interests of two clients (or prospective clients) conflict,
it will be practically impossible to act in the beset interests of the
two clients simultaneously.

\hypertarget{types-of-conflict}{%
\subsection{Types of Conflict}\label{types-of-conflict}}

Two broad situations:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Conflict between two or more clients (`conflict of interest')
\item
  Conflict between client's interest and solicitor's interest (`own
  interest conflict')
\end{enumerate}

A solicitor can \textbf{never} act where there is a conflict between
themselves and a current client.

\hypertarget{conflict-between-clients}{%
\subsection{Conflict Between Clients}\label{conflict-between-clients}}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{title: Conflict of interest}
\NormalTok{A situation where the solicitor\textquotesingle{}s or firm\textquotesingle{}s separate duties to act in the best interests of two or more clients conflict (SRA glossary).}
\end{Highlighting}
\end{Shaded}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{title: Para 6.2, Code of Conduct}

\NormalTok{A solicitor or firm must not act in relation to a matter or particular aspect of it if there is a conflict of interest or a significant risk of such a conflict unless:}
\NormalTok{a) The clients have a substantially common interest in relation to the matter or aspect of it; or}
\NormalTok{b) The clients are competing for the same objective.}
\end{Highlighting}
\end{Shaded}

\hypertarget{former-clients}{%
\subsubsection{Former Clients}\label{former-clients}}

Where a client's retainer has been terminated, the firm no longer owes a
duty to act in that client's best interests, so a conflict cannot arise
in such a situation. Even in the absence of a conflict, a firm may be
prevented from acting if it holds confidential information for the
former client which would be material to the work done for the new
client.

\hypertarget{matter-or-aspect-of-it}{%
\subsubsection{Matter or Aspect of it}\label{matter-or-aspect-of-it}}

Where the only conflict between the parties is their wider business
interests, this will not create any conflict of interest issues.

\hypertarget{significant-risk-of-a-conflict-of-interest}{%
\subsubsection{Significant Risk of a Conflict of
Interest}\label{significant-risk-of-a-conflict-of-interest}}

An actual conflict is not required, just a significant risk. This can
arise where there is an inequality of bargaining power between clients.

\hypertarget{exceptions}{%
\subsection{Exceptions}\label{exceptions}}

The exceptions were introduced by the Law Society following a period of
lobbying by part of the profession to liberalise the conflict of
interest rules.

\hypertarget{substantially-common-interest-exception-para-6.2a}{%
\subsubsection{Substantially Common Interest Exception (para
6.2(a))}\label{substantially-common-interest-exception-para-6.2a}}

Where there is a conflict or a significant risk of one, a solicitor or
firm can still act where the clients have a substantially common
interest in relation to the matter.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{title: Substantially common interest}
\NormalTok{A situation where there is a clear common purpose between the clients and a strong consensus as to how it is to be achieved. }
\end{Highlighting}
\end{Shaded}

e.g., this exception may be used where a solicitor is instructed by a
group of people who want to set up a company.

\hypertarget{competing-for-the-same-interest-commercial-exception}{%
\subsubsection{Competing for the Same Interest (`commercial')
Exception}\label{competing-for-the-same-interest-commercial-exception}}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{title: Competing for the same interest}
\NormalTok{Two or more clients competing for an objective which, if attained by one client, will make the objective unattainable to the other client/ clients. }
\end{Highlighting}
\end{Shaded}

Here, an objective is an asset, contract or business opportunity which
clients are seeking to acquire or recover through liquidation or by
means of an auction or tender process, but not a public takeover.

Generally, this exception applies to corporate clients. It enables one
firm to act for two companies bidding to take over a third.

\hypertarget{conditions-for-acting-under-exceptions}{%
\subsubsection{Conditions for Acting Under
Exceptions}\label{conditions-for-acting-under-exceptions}}

Under para 6.2:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  All the clients must have given informed consent evidenced in writing
\item
  Where appropriate, effective safeguards put in place
\item
  The solicitor or firm is satisfied it is reasonable for them to act
  for all the clients.
\end{enumerate}

Informed consent means the client must appreciate the issues and risks
involved. Effective safeguards will be needed to protect confidential
information. It must be reasonable for the solicitor to act, considering
whether any clients are prejudiced, particularly if a client is
vulnerable.

\hypertarget{conflict-of-interest-between-existing-clients}{%
\subsection{Conflict of Interest Between Existing
Clients}\label{conflict-of-interest-between-existing-clients}}

There may be two clients who originally posed no conflict, but a
conflict arises during the retainer. The solicitor may continue to act
for one of the clients, but must ensure that their duty of
confidentiality is not put at risk.

\hypertarget{own-interest-conflict}{%
\subsubsection{Own Interest Conflict}\label{own-interest-conflict}}

If there is an own-interest conflict, the solicitor may not act (para
6.1). This can arise in cases such as:

\begin{itemize}
\tightlist
\item
  A client wanting to gift a significant amount to a solicitor
\item
  Solicitor lending money to client
\item
  Personal relationship with the client, impacting the solicitor's
  ability to advise them independently and impartially.
\end{itemize}

A solicitor may not make a secret profit while acting for the client.

\hypertarget{systems-for-identifying-conflict}{%
\subsubsection{Systems for Identifying
Conflict}\label{systems-for-identifying-conflict}}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{title: Code of Conduct for Firms para 2.1(a)}
\NormalTok{Firms must have effective systems and controls in place to ensure they comply with all the SRA\textquotesingle{}s regulatory requirements. }
\end{Highlighting}
\end{Shaded}

To identify `own interest' conflicts, the firm should assess all
relevant circumstances, including whether the solicitor's ability as an
individual, or that of anyone within the firm, to act in the best
interests of the client is impaired by:

\begin{itemize}
\tightlist
\item
  Financial interests
\item
  Personal relationships
\item
  Appointment to public office
\item
  Commercial relationships
\item
  Employment.
\end{itemize}

A conflict of interest check should be done before taking on any new
work.

\hypertarget{professional-embarrassment-1}{%
\subsubsection{Professional
Embarrassment}\label{professional-embarrassment-1}}

Even if no conflict, may still decline to act due to professional
embarrassment. A solicitor should consider whether accepting
instructions to act for a client would breach one of the Principles.

A solicitor can also take into account their own commercial
considerations.

\hypertarget{limited-retainer}{%
\subsubsection{Limited Retainer}\label{limited-retainer}}

A solicitor could be retained to act only in relation to those areas
where no conflict exists, with each client seeking independent advice on
the conflicting areas.

\hypertarget{conveyancing}{%
\subsubsection{Conveyancing}\label{conveyancing}}

Subject to additional requirements.

\hypertarget{undertakings}{%
\section{Undertakings}\label{undertakings}}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{title: Undertaking}
\NormalTok{A statement given orally or in writingk, whether or not it includes the word "undertake" or "undertaking" to someone who reasonably places reliance on it, that you or a third party will do something or cause something to be done, or refrain from doing something. (SRA Glossary)}
\end{Highlighting}
\end{Shaded}

A solicitor must perform all undertakings given by them within an agreed
timescale, or if no timescale is agreed, within a reasonable amount of
time (para 1.3 Codes of Conduct).

Even a promise to give an undertaking will usually be interpreted as an
undertaking and will be binding on the solicitor concerned. Once an
undertaking has been relied upon, it can be withdrawn only be agreement.
A solicitor is not obliged to give or accept an undertaking.

\hypertarget{why-necessary}{%
\subsection{Why Necessary}\label{why-necessary}}

Often given by solicitors to smooth the path of a transaction, and avoid
otherwise problematic areas. e.g., avoids the need for advance deposits
in some cases, since other solicitors will rely on the undertaking as
just as good.

\hypertarget{breach-of-an-undertaking}{%
\subsection{Breach of an Undertaking}\label{breach-of-an-undertaking}}

The terms of an undertaking are personally binding on the individual
solicitor concerned. So if they fail to comply, they can be sued
personally by the recipient. They will also breach the Code of Conduct
and will be disciplined by the SRA. The firm may also be disciplined for
a lack of supervision.

As there is an obligation on firms to maintain records to demonstrate
compliance, and on solicitors to be able to justify their actions, firms
and solicitors must maintain a system for recording undertakings that
have been given and discharged.

\hypertarget{oral-and-written-undertakings}{%
\subsection{Oral and Written
Undertakings}\label{oral-and-written-undertakings}}

An undertaking is binding even if oral, though it is advisable to give
them in writing. If given orally, it should be confirmed as soon as
possible.

\hypertarget{terms-of-the-undertaking}{%
\subsection{Terms of the Undertaking}\label{terms-of-the-undertaking}}

A undertaking is not a contract (no consideration). But arguably just as
important.

\hypertarget{wording}{%
\subsubsection{Wording}\label{wording}}

A previous Code of Conduct stated that any ambiguity in the wording of
an undertaking would be interpreted in favour of its recipient. So take
care drafting undertakings.

\hypertarget{acts-outside-the-solicitors-control}{%
\subsubsection{Acts Outside the Solicitor's
Control}\label{acts-outside-the-solicitors-control}}

Provided an undertaking which is wholly reliant on the action/ inaction
of a third party should be avoided given the binding nature of such an
undertaking. Could alternatively provide an undertaking to use their
reasonable endeavours to ensure something happens. But again, don't give
this where there is likely to be dispute over what constitutes
reasonable endeavours.

\hypertarget{undertaking-on-behalf-of-a-client}{%
\subsubsection{Undertaking `on Behalf' of a
Client}\label{undertaking-on-behalf-of-a-client}}

An undertaking may be drafted as being made `on behalf of' the client.
This will not prevent the undertaking being enforceable against the
solicitor personally. It is possible to exclude personal liability, but
in order to do so the solicitor must clearly and expressly disclaim
personal liability.

\hypertarget{timescale}{%
\subsubsection{Timescale}\label{timescale}}

Either agree a timescale or a `reasonable' timescale will be imposed.

\hypertarget{costs}{%
\subsubsection{Costs}\label{costs}}

Where an undertaking is given in respect of the payment of costs of
another party, the term `costs' will be implied to mean proper costs
unless as specific amount is agreed.

\hypertarget{clients-authority}{%
\subsection{Client's Authority}\label{clients-authority}}

A solicitor should ensure they have clear and express authority from
their client before giving any undertakings. When such authority has
been received, it may be withdrawn by the client any time until the
solicitor has acted (even if expressed to be irrevocable).

\hypertarget{change-of-circumstances}{%
\subsubsection{Change of Circumstances}\label{change-of-circumstances}}

An undertaking will remain binding if the circumstances change so that
it is impossible to fulfil it, either wholly or partially. The recipient
may agree to its variation or discharge. Good practice to always inform
the recipient of any changes to circumstances.

\hypertarget{enforcement}{%
\subsubsection{Enforcement}\label{enforcement}}

\hypertarget{the-courts}{%
\paragraph{The Courts}\label{the-courts}}

The court is able to enforce an undertaking against a solicitor as an
officer of the court. Firms are obliged to carry indemnity insurance to
cover such claims. But the value of such claims may fall within the
excess of these policies, leading to personal liability for the
solicitor concerned.

A breach of an undertaking is likely to be a breach of professional
conduct. SRA will generally not investigate a breach of an undertaking
given to the court itself, unless the court reported the matter to the
SRA.

An undertaking given by an incorporated law firm (limited company or
LLP) is not capable of being enforced against the firm under the court's
supervisory jurisdiction over solicitors as officers of the court, since
such a firm has a separate legal existence. So the recipient would have
to make a claim under contract, negligence or breach of trust. The firm
can still face action for professional misconduct.

\hypertarget{sra-and-solicitors-disciplinary-tribunal}{%
\paragraph{SRA and Solicitors Disciplinary
Tribunal}\label{sra-and-solicitors-disciplinary-tribunal}}

The Legal Ombudsman, SRA, and Solicitors Disciplinary Tribunal have no
power to enforce the performance of an undertaking, but may investigate
the solicitor's conduct and impose disciplinary sanctions.

\hypertarget{liability-of-others}{%
\paragraph{Liability of Others}\label{liability-of-others}}

In a firm, if an undertaking is given by an assistant solicitor and not
honoured, there will be a breach of professional conduct by the
solicitor and also the partners of the firm. So many firms have strict
procedures about undertakings.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{Be super careful. If you make an undertaking to pay £100k from the proceeds of sale of a house, and the house sells for 90k, you are personally liable for the 10k. So draft the undertaking to expressly limit liability to the money received through the sale. }
\end{Highlighting}
\end{Shaded}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{Why not just have the client undertake to do something directly? And sign an undertaking/ contract to that effect?}
\end{Highlighting}
\end{Shaded}

\hypertarget{money-laundering}{%
\section{Money Laundering}\label{money-laundering}}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{title: Money laundering}
\NormalTok{The process by which criminals seek to alter or launder their proceeds of crime so that it appears that these funds come from a legitimate source. }
\end{Highlighting}
\end{Shaded}

Solicitors are targets for criminals in their efforts to launder the
proceeds of crime.

\hypertarget{money-laundering-regulations-2017}{%
\subsection{Money Laundering Regulations
2017}\label{money-laundering-regulations-2017}}

The Money Laundering, Terrorist Financing and Transfer of Funds
(Information on the Payer) Regulations 2017 (SI 2017/692) came into
force 26/06/17 and repealed and replaced the Money Laundering
Regulations 2007. Since been amended by the Money Laundering and
Terrorist Financing (Amendment) Regulations 2019 and the Money
Laundering and Terrorist Financing (Amendment) (EU Exit) Regulations
2020.

The Regulations apply to persons acting in the course of businesses
carried out in the UK, including independent legal professions (reg 8).
The vast majority of solicitors' firms will be subject to the
Regulations. Failure to comply is a criminal offence.

There is a requirement for the `beneficial owners, officers or managers'
of the firm and sole practitioners to apply to the SRA for approval
under the Regulations, which will be granted unless they have been
convicted of a relevant offence. These offences include any which
involve deception/ dishonesty.

A firm is required to take appropriate steps to identify and assess the
risk of money laundering and terrorist financing to which the business
is subject.

Services provided by law firms most likely to be targeted by money
launders:

\begin{itemize}
\tightlist
\item
  Trust and company formation
\item
  Conveyancing and client account services
\end{itemize}

So firms need to ensure they have appropriate measures in place to
address these risks.

\hypertarget{nominated-officers-and-reporting-procedures}{%
\subsubsection{Nominated Officers and Reporting
Procedures}\label{nominated-officers-and-reporting-procedures}}

\begin{itemize}
\tightlist
\item
  A firm is required to establish and maintain written policies,
  controls and procedures to mitigate and manage effectively the money
  laundering and terrorist financing risks identified. This is mainly
  about systems and risk assessment.
\item
  A firm must have a `nominated officer' (MLRO) to receive reports
  within the firm of any instances of suspected money laundering and to
  liaise with the National Crime Agency (reg 21(3)).
\item
  Firm must appoint an individual from senior management to be
  responsible for compliance with the Regulations (reg 21(1)(a)).
\item
  Relevant employees (those in compliance/ identification of money
  laundering) must be screened prior to and during the course of their
  employment to assess their skills, knowledge, conduct and integrity.
  (reg 21(1)(b)).
\item
  Establishing an independent audit function to examine and monitor
  firm's policies adopted to comply with the Regulations (reg 21(1)(c)).
\item
  Controls which enable the firm to `respond fully and rapidly' to
  enquiries from law enforcement as to whether it maintains a business
  relationship with any person, and the nature of their relationship
  (reg 21(8)).
\end{itemize}

\hypertarget{client-due-diligence}{%
\subsection{Client Due Diligence}\label{client-due-diligence}}

Firms carrying out relevant business are obliged to obtain verification
of the identity of each of their clients (`customer due diligence').
This includes the following circumstances (reg 27):

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  If the client and solicitor agree to form a business relationship
\item
  Carrying out the occasional transaction that amounts to a `transfer of
  funds' exceeding EUR 1000
\item
  Carrying out an occasional transaction amounting to EUR 15,000 or
  more, whether in one transaction or several linked transactions
\item
  Where the solicitor suspects money laundering or terrorist financing
\item
  Where the solicitor doubts the veracity or adequacy of documents or
  information supplied to verify client identity.
\end{enumerate}

Verification must take place before a business relationship is
established/ before a transaction is carried out (reg 30). Proof of
registration must be collected before establishing a business
relationship with a company. There are certain exceptions, e.g., if
there is little risk of money laundering.

\hypertarget{simplified-due-diligence}{%
\subsubsection{Simplified Due
Diligence}\label{simplified-due-diligence}}

Simplified due diligence is permitted where a firm determines that the
business relationship or transaction presents a low risk of money
laundering or terrorist financing. If the transaction and client are
eligible for simplified due diligence, the solicitor will not
necessarily need to obtain information on the beneficial owners.

\hypertarget{standard-due-diligence}{%
\subsubsection{Standard Due Diligence}\label{standard-due-diligence}}

Will apply to most clients. The solicitor is obliged to verify the
identity of the client on the basis of documents/ information from a
reliable source which is independent of the person whose identity is
being verified.

\begin{itemize}
\tightlist
\item
  For companies/ trusts, take reasonable measures to understand
  ownership and control structures.
\item
  For natural persons, need passports/ driving licences.
\item
  For partnerships, obtain information on the constituent individuals
  (other than where partnerships are well-known reputable organisations
  with long histories)
\item
  For companies, need to know:

  \begin{itemize}
  \tightlist
  \item
    Name
  \item
    Company number
  \item
    Registered office address and principal place of business
  \item
    Proof of registration
  \item
    Reasonable efforts to obtain constitution, names of board of
    directors

    \begin{itemize}
    \tightlist
    \item
      This will be helped by the obligation on a UK body corporate to
      provide certain information on request when it forms a business
      relationship with a firm.
    \end{itemize}
  \end{itemize}
\end{itemize}

The definition of a beneficial owner varies depending on the nature of
the client.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{title: Beneficial owner of a company}
\NormalTok{For companies, reg 5(1) defines a beneficial owner as:}
\NormalTok{{-} Any individual who exercises ultimate control over the management of the body corporate}
\NormalTok{{-} Any individual who ultimately owns or controls more than 25\% of the shares/ voting rights in the body corporate. }
\NormalTok{{-} An individual who controls the body corporate}
\end{Highlighting}
\end{Shaded}

This does \textbf{not} apply to a company listed on a regulated market,
but does apply to LLPs.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{title: Beneficial owner of a partnership}
\NormalTok{Reg 5(3) defines the beneficial owner of a partnership as any individual who:}
\NormalTok{{-} Is ultimately entitled to or controls $\textgreater{}25\textbackslash{}\%$ of capital/ profits/ voting rights of the partnership}
\NormalTok{{-} Otherwise exercises control over the management of the partnership.}
\end{Highlighting}
\end{Shaded}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{title: Beneficial owner of a trust}
\NormalTok{Reg 6(1) says this includes:}
\NormalTok{{-} The settlor}
\NormalTok{{-} The trustees}
\NormalTok{{-} The beneficiary}
\NormalTok{{-} The class of persons in whose main interest the trust is set up}
\NormalTok{{-} Any individual who has control over the trust. }
\end{Highlighting}
\end{Shaded}

This is more thorough than for companies, since the trust does not have
legal personality itself. Beneficial owners are those individuals who
hold equivalent or similar positions to those set out in reg 6(1).

\hypertarget{enhanced-due-diligence}{%
\subsubsection{Enhanced Due Diligence}\label{enhanced-due-diligence}}

Enhanced due diligence must be carried out where (reg 33):

\begin{itemize}
\tightlist
\item
  High risk of money laundering/ terrorist financing, identified by the
  firm/ information provided by the SRA and Law Society
\item
  Client/ other party to transaction is established in a high-risk third
  party
\item
  Client has provided false or stolen identification documentation
\item
  Client is a politically exposed person (PEP), or a family member or
  close associate of a PEP
\item
  Transaction is complex or unusually large
\item
  Any other situation where there is a higher risk of money laundering
  or terrorist financing.
\end{itemize}

A solicitor must take measures to examine the background and purpose of
a transaction and consider whether it is appropriate to obtain further
independent verification of the client or beneficial owner's identity or
more detail on the ownership and control structure of the client.

A PEP is an individual who is entrusted with prominent public functions.
They have been a focus for the Financial Action Task Force and other
bodies because of concerns about using their political positions to
corruptly enrich themselves.

Where a solicitor is dealing with a PEP, additional obligations are
placed on the solicitor, such as having the approval of senior
management to act for the client, taking adequate measures to establish
the source of wealth and enhanced ongoing monitoring.

A solicitor is obliged to undertake ongoing monitoring of business
relationships to ensure transactions are consistent with the solicitor's
knowledge of the client (reg 28(11)).

\hypertarget{training}{%
\subsection{Training}\label{training}}

Firms are obliged to train employees and agents used about money
laundering (reg 24). Employees should be made aware of the law relating
to money laundering, terrorist financing and data protection.

Where no training is given, may provide the employee with a defence
under the Proceeds of Crime Act 2002.

\hypertarget{record-keeping}{%
\subsubsection{Record Keeping}\label{record-keeping}}

Keep detailed records to satisfy due diligence requirements. Records
must be kept for at least 5 years from when the business relationship
ends.

\hypertarget{proceeds-of-crime-act-2002}{%
\subsubsection{Proceeds of Crime Act
2002}\label{proceeds-of-crime-act-2002}}

This Act made it an offence to be involved in money laundering. There
are 3 offences that most concern solicitors.

\hypertarget{arrangements}{%
\subsubsection{Arrangements}\label{arrangements}}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{title: Proceeds of Crime Act 2002, s 328}

\NormalTok{(1) A person commits an offence if he enters into or becomes concerned in an arrangement which he knows or suspects facilitates (by whatever means) the acquisition, retention, use or control of criminal property by or on behalf of another person.}

\NormalTok{(2) But a person does not commit such an offence if—}

\NormalTok{{-} (a) he makes an authorised disclosure under section 338 and (if the disclosure is made before he does the act mentioned in subsection (1)) he has the appropriate consent;}

\NormalTok{{-} (b) he intended to make such a disclosure but had a reasonable excuse for not doing so;}

\NormalTok{{-} (c) the act he does is done in carrying out a function he has relating to the enforcement of any provision of this Act or of any other enactment relating to criminal conduct or benefit from criminal conduct.}

\NormalTok{(3) Nor does a person commit an offence under subsection (1) if—}

\NormalTok{{-} (a) he knows, or believes on reasonable grounds, that the relevant criminal conduct occurred in a particular country or territory outside the United Kingdom, and}

\NormalTok{{-} (b) the relevant criminal conduct—}

\NormalTok{    {-} (i) was not, at the time it occurred, unlawful under the criminal law then applying in that country or territory, and}
    
\NormalTok{    {-} (ii) is not of a description prescribed by an order made by the Secretary of State.}

\NormalTok{(4) In subsection (3) \_“the relevant criminal conduct”\_ is the criminal conduct by reference to which the property concerned is criminal property.}

\NormalTok{(5) A deposit{-}taking body, electronic money institution or payment institution]3 that does an act mentioned in subsection (1) does not commit an offence under that subsection if—}

\NormalTok{{-} (a) it does the act in operating an account maintained with it, and}

\NormalTok{{-} (b) the arrangement facilitates the acquisition, retention, use or control of criminal property of a value that is less than the threshold amount determined under section 339A for the act.}
\end{Highlighting}
\end{Shaded}

This is drafted very widely. There is no requirement that the funds pass
through the hands of the solicitor for them to be criminally liable. The
maximum penalty is 14 years' imprisonment.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{title: Criminal property}
\NormalTok{A person\textquotesingle{}s direct or indirect benefit from criminal conduct (s 340)}
\end{Highlighting}
\end{Shaded}

Criminal conduct is defined very widely to include basically all
offences, including conduct internationally which is criminalised in the
UK.

A solicitor does not commit an offence under s 328 if they make an
authorised disclosure to the firm's nominated officer as soon as is
practically possible. If the nominated officer decides to make a report
to the NCA, they are unable to consent to the transaction proceeding
until:

\begin{itemize}
\tightlist
\item
  The NCA consents
\item
  The NCA does not respond after 7 working days (starting the day after
  disclosure)
\item
  Where consent is refused, and 31 days have since passed.
\end{itemize}

A solicitor may seek to make an authorised disclosure whilst the
prohibited act is ongoing. To avoid breaching s 328, they must satisfy s
338(2A):

\begin{itemize}
\tightlist
\item
  Disclosure made whilst the prohibited act is ongoing
\item
  When the alleged offender began to do the act, they did not know/
  suspect criminal assets
\item
  Disclosure made as soon as is practicable.
\end{itemize}

A solicitor may also seek to make an authorised disclosure after the
prohibited act is completed, but must have a good reason for their
failure to disclose prior to their completing the act (s 338(3)).

\hypertarget{exception-litigation}{%
\paragraph{Exception: Litigation}\label{exception-litigation}}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{In [[Bowman v Fels [2005] EWCA Civ 226]], the Court of Appeal held that taking steps in litigation and the resolution of issues in a litigious context were excluded from the scope of s 328. }
\end{Highlighting}
\end{Shaded}

But the solicitor would not be able to take advantage of this exclusion
if the litigation was a sham created for the purpose of money
laundering.

\hypertarget{exception-overseas-crime}{%
\paragraph{Exception: Overseas Crime}\label{exception-overseas-crime}}

A solicitor will not commit an offence where the criminal conduct
occurred outside the UK and was not unlawful in the country where it
happened. But this defence is not available where the offence would
carry a sentence of \(\geq 12\) months in the UK.

\hypertarget{acquisition-use-or-possession}{%
\paragraph{Acquisition, Use or
Possession}\label{acquisition-use-or-possession}}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{title: Proceeds of Crime Act 2002, s 329}

\NormalTok{(1) A person commits an offence if he—}

\NormalTok{{-} (a) acquires criminal property;}

\NormalTok{{-} (b) uses criminal property;}

\NormalTok{{-} (c) has possession of criminal property.}

\NormalTok{(2) But a person does not commit such an offence if—}

\NormalTok{{-} (a) he makes an authorised disclosure under section 338 and (if the disclosure is made before he does the act mentioned in subsection (1)) he has the appropriate consent;}

\NormalTok{{-} (b) he intended to make such a disclosure but had a reasonable excuse for not doing so;}

\NormalTok{{-} (c) he acquired or used or had possession of the property for adequate consideration;}

\NormalTok{{-} (d) the act he does is done in carrying out a function he has relating to the enforcement of any provision of this Act or of any other enactment relating to criminal conduct or benefit from criminal conduct.}

\NormalTok{(2A) Nor does a person commit an offence under subsection (1) if—}

\NormalTok{{-} (a) he knows, or believes on reasonable grounds, that the relevant criminal conduct occurred in a particular country or territory outside the United Kingdom, and}

\NormalTok{{-} (b) the relevant criminal conduct—}

\NormalTok{    {-} (i) was not, at the time it occurred, unlawful under the criminal law then applying in that country or territory, and}
    
\NormalTok{    {-} (ii) is not of a description prescribed by an order made by the Secretary of State.}

\NormalTok{(2B) In subsection (2A) \_“the relevant criminal conduct”\_ is the criminal conduct by reference to which the property concerned is criminal property.}

\NormalTok{(2C) A deposit{-}taking body [, electronic money institution or payment institution]3 that does an act mentioned in subsection (1) does not commit an offence under that subsection if—}

\NormalTok{{-} (a) it does the act in operating an account maintained with it, and}

\NormalTok{{-} (b) the value of the criminal property concerned is less than the threshold amount determined under section 339A for the act.}

\NormalTok{(3) For the purposes of this section—}

\NormalTok{{-} (a) a person acquires property for inadequate consideration if the value of the consideration is significantly less than the value of the property;}

\NormalTok{{-} (b) a person uses or has possession of property for inadequate consideration if the value of the consideration is significantly less than the value of the use or possession;}

\NormalTok{{-} (c) the provision by a person of goods or services which he knows or suspects may help another to carry out criminal conduct is not consideration.}
\end{Highlighting}
\end{Shaded}

\hypertarget{adequate-consideration-defence}{%
\subparagraph{Adequate Consideration
Defence}\label{adequate-consideration-defence}}

An issue may arise where a solicitor received money for their costs for
work carried out for a client charged with a criminal offence, and there
is the possibility that the money in question is criminal property.

The defence applies where professional advisers receive money for costs,
but the fees charged must be reasonable. Returning the balance of money
in an account to a client may be an offence under s 328 if the solicitor
knows/ suspects the money is criminal property. In this case, an
authorised disclosure must be made.

\hypertarget{failure-to-disclose}{%
\paragraph{Failure to Disclose}\label{failure-to-disclose}}

A person commits an offence under s 330 Proceeds of Crime Act 2002 if:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  They know, suspect, or have reasonable grounds to know or suspect,
  that a person is engaged in money laundering; and
\item
  The information comes to them in the course of a business in the
  regulated sector; and
\item
  The information may assist in identifying the money launderer or the
  location of any laundered property; and
\item
  They do not make an authorised disclosure as soon as is practicable.
\end{enumerate}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{title: Proceeds of Crime Act 2002, s 330}

\NormalTok{(1) A person commits an offence if [the conditions in subsections (2) to (4) are satisfied.}

\NormalTok{(2) The first condition is that he—}

\NormalTok{{-} (a) knows or suspects, or}

\NormalTok{{-} (b) has reasonable grounds for knowing or suspecting,}

\NormalTok{that another person is engaged in money laundering.}

\NormalTok{(3) The second condition is that the information or other matter—}

\NormalTok{{-} (a) on which his knowledge or suspicion is based, or}

\NormalTok{{-} (b) which gives reasonable grounds for such knowledge or suspicion,}

\NormalTok{came to him in the course of a business in the regulated sector.}

\NormalTok{(3A) The third condition is—}

\NormalTok{{-} (a) that he can identify the other person mentioned in subsection (2) or the whereabouts of any of the laundered property, or}

\NormalTok{{-} (b) that he believes, or it is reasonable to expect him to believe, that the information or other matter mentioned in subsection (3) will or may assist in identifying that other person or the whereabouts of any of the laundered property.}

\NormalTok{(4) The fourth condition is that he does not make the required disclosure to—}

\NormalTok{{-} (a) a nominated officer, or}

\NormalTok{{-} (b) a person authorised for the purposes of this Part by [the Director General of the National Crime Agency,}

\NormalTok{as soon as is practicable after the information or other matter mentioned in subsection (3) comes to him.}
\end{Highlighting}
\end{Shaded}

Breaking this down:

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.1739}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.8261}}@{}}
\toprule()
\begin{minipage}[b]{\linewidth}\raggedright
Term
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Explanation
\end{minipage} \\
\midrule()
\endhead
Objective test & A solicitor will commit this offence even where they
genuinely did not know or suspect a person was engaged in money
laundering. So turning a blind eye is no defence. \\
Regulated sector & Most solicitors' firms fall within this category. \\
The information & The information obtained must be of some use to the
authorities. So the solicitor must identify/ believe the information
will assist authorities. Again, an objective test. \\
Authorised disclosure & A solicitor will be provided with a defence to s
330 if they make an authorised disclosure to the firm's nominated
officer as soon as practically possible. \\
Reasonable excuse for non-disclosure & Reasonable excuse is a defence \\
Training & If not trained, employee has a defence against s 330
offence \\
\bottomrule()
\end{longtable}

\hypertarget{confidentiality-and-privilege-1}{%
\subparagraph{Confidentiality and
Privilege}\label{confidentiality-and-privilege-1}}

A solicitor is under a duty at common law to keep confidential certain
information given to them by a client. This is referred to as legal
professional privilege (LPP). Communications will be protected from
disclosure if they fall within `advice privilege' or `litigation
privilege'. But LLP cannot be relied upon where communication takes
place for committing an offence.

A solicitor is not obliged under s 330 to disclose any information that
comes to them as a professional legal adviser in privileged
circumstances, or in relation to legal proceedings, whether contemplated
or actual. Does not apply where the information is communicated for the
intention of furthering a criminal purpose.

\hypertarget{overseas-defence}{%
\subparagraph{Overseas Defence}\label{overseas-defence}}

If the solicitor believes the money laundering is taking place outside
the UK, and it is not unlawful in that country, they do not have to
disclose anything.

\hypertarget{penalty}{%
\subparagraph{Penalty}\label{penalty}}

Maximum 5 years' imprisonment for s 330 offence.

\hypertarget{tipping-off}{%
\paragraph{Tipping Off}\label{tipping-off}}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{title: Disclosing a suspicious activity report (s 333A(1))}
\NormalTok{It is an offence to disclose to any person that a suspicious activity report (on money laundering) has been made if that disclosure is likely to prejudice any investigation that follows such a report. }
\end{Highlighting}
\end{Shaded}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{title: Disclosing an investigation (s 333A(3))}
\NormalTok{More general offence which is committed where a disclosure is made to any person that an investigation into money laundering is being carried out or contemplated, and that disclosure is likely to prejudice the money laundering investigation.}
\end{Highlighting}
\end{Shaded}

\hypertarget{defences}{%
\subparagraph{Defences}\label{defences}}

There are a number of defences to tipping off:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  the person who made the disclosure did not know or suspect that the
  disclosure would prejudice an investigation into money laundering (s
  333D(3) and (4)).
\item
  Disclosure is made to the client for the purposes of dissuading the
  client from engaging in the alleged money laundering (s 333D(2))
\end{enumerate}

\hypertarget{penalties}{%
\subparagraph{Penalties}\label{penalties}}

Carry a maximum penalty of an unlimited fine and/ or max. 2 year prison
sentence.

\hypertarget{regulated-sector}{%
\subparagraph{Regulated Sector}\label{regulated-sector}}

Above offences only apply to the regulated sector, though similar
provisions are in effect for work outside the regulated sector under s
342.

\hypertarget{confidentiality-1}{%
\paragraph{Confidentiality}\label{confidentiality-1}}

A solicitor is under a duty under para 6.3 Codes of Conduct to keep
confidential the affairs of clients (and former clients) unless
disclosure is required or permitted by law, or the client consents.

When making s 328 or s 330 disclosures, the legislation expressly
provides that such a disclosure will not breach this duty and is
permitted by law.

\hypertarget{criminal-finances-act-2017}{%
\paragraph{Criminal Finances Act
2017}\label{criminal-finances-act-2017}}

From 30/09/17, Criminal Finances Act 2017 introduced the corporate
offence of failure to prevent the criminal facilitation of tax evasion.
It applies to law firms as `relevant bodies'.

This makes a firm liable for failing to prevent tax evasion by its
employees/ other associated persons. There is strict liability for the
offence: no knowledge or intention is required on the part of the firm/
senior management.

The only defence available is that the firm had reasonable prevention
procedures in place (or it was reasonable not to have such measures).
The penalty is unlimited fines and confiscation of assets. A firm-wide
assessment of risk and internal procedures will be required.

Serious and extensive restrictions are imposed on those dealing with
people or entities on the sanctions list. This covers all firms and
imposes a requirement to inform the Office of Financial Sanctions
Implementation (OFSI) if it is known or reasonably suspected that a
person is a designated person has committed offences under financial
sanctions/ asset freezing regimes. If the firm still wants to act for
the person, needs a licence from OFSI.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{title: Summary}
\NormalTok{1. Money laundering is the process whereby the proceeds of crime are changed so that they appear to come from a legitimate source.  }
\NormalTok{2. The Government has introduced legislation to disrupt this process.  }
\NormalTok{3. Solicitors who undertake relevant business must comply with the Money Laundering, Terrorist Financing and Transfer of Funds (Information on the Payer) Regulations 2017 (SI 2017/692).  }
\NormalTok{4. Under the 2017 Regulations, firms must appoint a nominated officer, who will receive internal reports concerning money laundering and must consider whether to report the matter to the NCA.  }
\NormalTok{5. A person commits an offence under s 328 of the Proceeds of Crim e Act 2002 where they become concerned in an arrangement involving money laundering.  }
\NormalTok{6. Section 328 does not apply to steps taken in litigation.  }
\NormalTok{7. A person commits an offence under s 329 of the Proceeds of Crime Act 2002 if they acquire, use or have possession of criminal property  }
\NormalTok{8. A solicitor within the regulated sector is obliged to disclose any suspected money laundering activity to their nominated officer under s 330 of the Proceeds of Crime Act 2002.  }
\NormalTok{9. It is an offence to disclose to any person that a disclosure has been made where this may prejudice an investigation.}
\end{Highlighting}
\end{Shaded}

\hypertarget{financial-services}{%
\section{Financial Services}\label{financial-services}}

Solicitors often have to do financial services work and will have to
comply with FCA and SRA regulations.

Work could arise:

\begin{itemize}
\tightlist
\item
  In conveyancing (e.g., helping a client finding a mortgage and support
  package)
\item
  In probate (when executors sell off assets)
\item
  In litigation (helping a successful client to invest damages)
\item
  etc.
\end{itemize}

A solicitor would be in breach of their duty to act in the client's best
interests (Principle 7) if they did not have sufficient expertise in the
area concerned. For many activities, also require authorisation.

\hypertarget{source-material}{%
\subsection{Source Material}\label{source-material}}

\begin{itemize}
\tightlist
\item
  Financial Services and Markets Act 2000 (FSMA 2000) came into force
  30/11/2001.
\item
  Financial Services Act 2012 extensively amends FSMA 2000.

  \begin{itemize}
  \tightlist
  \item
    Currently partly in force
  \item
    Need to study it alongside the Orders of Council made by the
    Treasury - lots of the details only implemented through secondary
    legislation.
  \end{itemize}
\end{itemize}

\hypertarget{financial-services-regulatory-structure}{%
\subsection{Financial Services Regulatory
Structure}\label{financial-services-regulatory-structure}}

FSMA 2000 established the FSA, to regulate financial services. Following
the 2008 crisis, the Government abolished the FSA and established three
bodies:

\begin{itemize}
\tightlist
\item
  Financial Conduct Authority (FCA)
\item
  Prudential Regulation Authority (PRA)
\item
  Financial Policy Committee (FPC).
\end{itemize}

The FCA and PRA have taken over the majority of the FSA's functions,
whilst the FPC is responsible for monitoring the stability of the whole
UK financial system.

\hypertarget{fca}{%
\subsubsection{FCA}\label{fca}}

\begin{itemize}
\tightlist
\item
  Responsibility for the conduct of business regulation of all firms
\item
  Responsibility for prudential regulation of firms (their safety and
  soundness) not regulated by the PRA.
\end{itemize}

FCA objectives are:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Securing an appropriate degree of protection for consumers
\item
  Protecting and enhancing the integrity of the UK financial system
\item
  Promoting effective competition in the interests of consumers in the
  market.
\end{enumerate}

The FCA can require firms to withdraw or amend misleading financial
promotions with immediate effect and can block/ stop a service or
product. Also regulates consumer credit and second charge mortgages.

\hypertarget{pra}{%
\subsubsection{PRA}\label{pra}}

Subsidiary of the Bank of England responsible for the authorisation,
prudential regulation and general supervision of firms which manage
significant financial risks. Such firms are known as PRA-authorised or
dual-regulated firms.

\hypertarget{general-framework}{%
\subsubsection{General Framework}\label{general-framework}}

Businesses carrying out regulated activities need to be authorised by
the appropriate regulator. If certain conditions are satisfied, a firm
carrying out regulated activity incidentally in the course of carrying
out professional work does not have to obtain authority from the FCA
provided it is regulated and supervised by a particular Designated
Professional Body (`DPB'). The SRA is a DPB.

\hypertarget{general-restrictions}{%
\subsection{General Restrictions}\label{general-restrictions}}

Restrictions:

\begin{itemize}
\tightlist
\item
  s 19 FSMA: No person may carry on a regulated activity in the UK
  unless authorised or exempt.

  \begin{itemize}
  \tightlist
  \item
    Can result in agreements being rendered unenforceable.
  \end{itemize}
\item
  s 21 FSMA: making an unauthorised financial promotion is prohibited.
\item
  Any breaches of the above are criminal offences
\end{itemize}

\hypertarget{regulated-activity}{%
\subsection{Regulated Activity}\label{regulated-activity}}

{[}{[}Regulated activity flowchart.png{]}{]}

\hypertarget{four-tests}{%
\subsubsection{Four Tests}\label{four-tests}}

s 22 FSMA 2000 sets out the definition of a regulated activity, which
has been substantially amended by FSA 2012.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{title: Is an activity regulated?}
\NormalTok{Tests:}
\NormalTok{1. Are you in business?}
\NormalTok{2. Is there a specified activity?}
\NormalTok{3. Is there a specified investment, or does the specified actifvitiy relate to information about a person\textquotesingle{}s financial standing or administering a benchmark?}
\end{Highlighting}
\end{Shaded}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{RAO 2001 is shorthand for Financial Services and Markets Act 2000 (Regulated Activities) Order 2001, SI 2001/544}
\end{Highlighting}
\end{Shaded}

\begin{itemize}
\tightlist
\item
  Particular activities and investments are specified by the Treasure in
  RAO 2001, s 22.
\item
  Information about a person's financial standing includes credit
  references and stuff
\item
  `Benchmark' is defined in s 22(6A) FSMA 2000.
\end{itemize}

\hypertarget{specified-investments}{%
\subsection{Specified Investments}\label{specified-investments}}

These include:

\begin{itemize}
\tightlist
\item
  Company stocks and shares
\item
  Debentures, loan stock and bonds
\item
  Government securities
\item
  Unit trusts and open-ended investment companies (OEICs) - like a unit
  trust but company rather than trust
\item
  Insurance contracts
\item
  Certain mortgage contracts called regulated mortgage contracts (those
  over land of which \(\geq 40\%\) will be a dwelling)
\item
  Home reversion/ home purchase plans
\item
  Deposits. Note sums received by solicitors acting in the course of
  business are exempt.
\end{itemize}

\hypertarget{specified-investment-activities}{%
\subsubsection{Specified Investment
Activities}\label{specified-investment-activities}}

Include:

\begin{itemize}
\tightlist
\item
  Dealing as agent

  \begin{itemize}
  \tightlist
  \item
    Buying, selling, subscribing for or underwriting investments.
  \end{itemize}
\item
  Arranging

  \begin{itemize}
  \tightlist
  \item
    Solicitors have many clients whose transactions involve investments.
  \item
    Solicitor may be involved as the contact between client and broker/
    life company.
  \item
    Merely introducing a client to an FCA-authorised third-party does
    not constitute arranging, except where the transaction relates to an
    insurance contract.
  \end{itemize}
\item
  Managing

  \begin{itemize}
  \tightlist
  \item
    Requires active participation beyond merely holding investments.
  \end{itemize}
\item
  Safeguarding

  \begin{itemize}
  \tightlist
  \item
    Safeguarding and administering investments belonging to a client.
  \end{itemize}
\item
  Advising

  \begin{itemize}
  \tightlist
  \item
    Giving advice to a person in their capacity as an investor on the
    merits of their buying/ selling/ subscribing for/ underwriting an
    investment.
  \item
    Does not include generic investment
  \item
    You could refer your client to a person authorised by the FCA to
    provide advice.
  \end{itemize}
\item
  Lending money on regulated mortgage
\end{itemize}

\hypertarget{exclusions}{%
\subsubsection{Exclusions}\label{exclusions}}

Exclusions relevant to solicitors include:

\begin{itemize}
\tightlist
\item
  Introducing
\item
  Using an ATP (authorised third party)

  \begin{itemize}
  \tightlist
  \item
    If the transaction is entered into with or through an ATP, on the
    advice given to the client by the ATP, the effect of RAO 2001 is
    excluded.
  \item
    The exclusion does not apply to insurance contracts
  \item
    Does not apply if you receive a commission
  \end{itemize}
\item
  Acting for an execution-only client

  \begin{itemize}
  \tightlist
  \item
    The client, in their capacity as an investor, is not seeking or has
    not sought advice from the solicitor as to the merits of the
    client's entering into the transaction.
  \end{itemize}
\item
  Acting as a trustee or personal representative

  \begin{itemize}
  \tightlist
  \item
    Applies to arranging, managing, safeguarding or advising fellow
    trustees and/or beneficiaries.
  \item
    Has limitations: available to a solicitor acting as trustee or PR,
    but not a solicitor acting \textbf{for} a trustee/ PR.
  \item
    Does not apply to insurance contracts.
  \end{itemize}
\item
  The `professional/ necessary' exclusion

  \begin{itemize}
  \tightlist
  \item
    Applies to advising, arranging, safeguarding and dealing as agents.
  \item
    Exclusion if the activity is performed in the course of carrying on
    any profession or business, may be reasonably regarded as a
    necessary part of other services provided in the course of that
    profession or business.
  \item
    Necessary means they can't get from A to B without solicitor doing
    this. Strictly necessary.
  \end{itemize}
\item
  The `takeover' exclusion

  \begin{itemize}
  \tightlist
  \item
    Exclusion applies to arranging, advising and dealing as agent.
    Applies to a transaction to acquire or dispose of shares in a body
    corporate or for a transaction entered into for the purposes of such
    an acquisition or disposal, if:

    \begin{itemize}
    \tightlist
    \item
      Shares consist of 50\% or more of voting rights in the body
      corporate, and
    \item
      The acquisition or disposal is between parties each of whom is a
      body corporate, a partnership, a single individual or a group of
      connected individuals.
    \end{itemize}
  \item
    Extremely valuable exclusion for the corporate department where a
    client is seeking to take over or sell its interest in a company,
    whether public or private.
  \end{itemize}
\end{itemize}

\hypertarget{summary-1}{%
\paragraph{Summary}\label{summary-1}}

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.1698}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.8302}}@{}}
\toprule()
\begin{minipage}[b]{\linewidth}\raggedright
Specified activity
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Applicable exclusions
\end{minipage} \\
\midrule()
\endhead
Dealing as agent & ATP, execution-only, professional/necessary,
takeover \\
Arranging & Introducing, ATP, execution-only, professional/necessary,
acting as trustee/PR, takeover \\
Advising & Professional/necessary, acting as trustee/PR, takeover \\
Managing & Acting as trustee/PR \\
Safeguarding & Professional/necessary, acting as trustee/PR \\
\bottomrule()
\end{longtable}

\hypertarget{s-327-exemption-for-professional-firms}{%
\paragraph{S 327 Exemption for Professional
Firms}\label{s-327-exemption-for-professional-firms}}

Under s 326 FSMA 2000, the Treasury can designate a professional body
(DPB). It has designated the Law Society - with this role now undertaken
by the SRA.

Under s 327, the general prohibition in s 19 FSMA will not apply to
regulated activity carried on by a firm of solicitors if the following
are met:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  The firm must not receive from a person other than the client any
  pecuniary or other advantage out of the activity for which it does not
  account to its client;

  \begin{itemize}
  \tightlist
  \item
    Mirrored by para 4.1 Code for Solicitors and para 5.1 Code for Firms
  \end{itemize}
\item
  The manner of providing `any service in the course of carrying on the
  activities must be incidental to the provision' by the firm of
  professional services, i.e., services regulated by the SRA.

  \begin{itemize}
  \tightlist
  \item
    There are two tests for `incidental'; a specific and a general one.
  \item
    Specific test

    \begin{itemize}
    \tightlist
    \item
      Refers to the particular client concerned: the relevant regulated
      activity must arise out of, or be complementary to, the provision
      of a particular professional service to a particular client.
    \item
      Both the professional and regulated service must be supplied to
      the same person.
    \end{itemize}
  \item
    General test

    \begin{itemize}
    \tightlist
    \item
      Activities carries out by the firm which would otherwise be
      regulated cannot be a major part of the firm's activities.
      Consider

      \begin{itemize}
      \tightlist
      \item
        Scale of regulated activities as a proportion
      \item
        Whether regulated activities are held out as a separate service
      \item
        The impression given of how the firm provides those activities
      \end{itemize}
    \end{itemize}
  \end{itemize}
\item
  The firm must carry out only regulated activities permitted by the DPB
  (i.e., the SRA).

  \begin{itemize}
  \tightlist
  \item
    SRA has published the SRA Financial Services (Scope) Rules and the
    SRA Conduct of Business (COB) Rules for this purpose

    \begin{itemize}
    \tightlist
    \item
      These set out in great deal what is and is not prohibited.
    \end{itemize}
  \end{itemize}
\item
  The activities must not be prohibited by an order made by the
  Treasury, or any direction by the FCA under s 328 or 329.

  \begin{itemize}
  \tightlist
  \item
    Such prohibited activities are listed in FSMA 2000 (Professions)
    (Non-Exempt Activities) Order 2001
  \item
    FCA also has authority to prohibit stuff.
  \end{itemize}
\item
  The firm must not carry on any other regulated activities.

  \begin{itemize}
  \tightlist
  \item
    i.e., exemption cannot be used by FCA-authorised firms.
  \end{itemize}
\end{enumerate}

{[}{[}s 327 exemption flowchart.png{]}{]}

\hypertarget{sra-financial-services-conduct-of-business-rules}{%
\subsection{SRA Financial Services (Conduct of Business)
Rules}\label{sra-financial-services-conduct-of-business-rules}}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{COB Rules apply only when the firm is carrying out an exempt regulated activity. }
\end{Highlighting}
\end{Shaded}

\hypertarget{status-disclosure}{%
\subsubsection{Status Disclosure}\label{status-disclosure}}

r 2.1 and 2.2: A firm must provide clients with certain information
concerning the status of the firm (e.g., this firm is not authorised by
the FCA, complaint and redress mechanisms provided through SRA and Legal
Ombudsman).

\hypertarget{best-execution}{%
\subsubsection{Best Execution}\label{best-execution}}

r 3.1: A solicitor must act in the best interests of their clients
(Principle 7). The firm must carry out transactions for clients as soon
as possible unless it reasonably believes it is in the client's best
interest not to.

\hypertarget{transactions}{%
\subsubsection{Transactions}\label{transactions}}

r 4.1 and 4.2: The firm must keep records of:

\begin{itemize}
\tightlist
\item
  Instructions from clients to carry out transactions
\item
  Instructions to third parties to carry them out.
\end{itemize}

\hypertarget{commissions}{%
\subsection{Commissions}\label{commissions}}

r 5.1: The firm must keep records of commissions received in respect of
regulated activities and how these were dealt with.

\hypertarget{execution-only-clients}{%
\subsubsection{Execution-only Clients}\label{execution-only-clients}}

r 7.1: Where a firm acts for an execution-only client and the investment
concerned is a retail investment product, it must send a letter to the
client confirming that the client is not relying on the advice of the
solicitor.

\hypertarget{insurance-distribution-activities}{%
\subsubsection{Insurance Distribution
Activities}\label{insurance-distribution-activities}}

COB Rules, Part 3: Rules which must be complied with whenever any
insurance distribution activities are carried out. Detailed.

\hypertarget{consumer-credit-activity}{%
\subsection{Consumer Credit Activity}\label{consumer-credit-activity}}

From 01/04/14, consumer credit activity became a regulated activity for
the purposes of s 22 SMA 2000. These activities are set out in Pt 2 and
Pt 3A RAO 2001.

Solicitors carrying out such activities need FCA authorisation, or need
to ensure the activities fall within the s 327 exemption. Representing a
client in a litigation matter surrounding a consumer credit agreement is
not a credit-related matter for these purposes.

Allowing a client time to pay/ to pay in instalments will be regarded as
exempt so long as:

\begin{itemize}
\tightlist
\item
  Number of repayments does not exceed 12
\item
  Payment term does not exceed 12 months
\item
  Credit is provided without interest or other charge.
\end{itemize}

\hypertarget{insurance-distribution}{%
\subsubsection{Insurance Distribution}\label{insurance-distribution}}

Regulation extended to all general insurance selling and administration
in order to comply with the Insurance Mediation Directive (2002/92/EC)
and the Insurance Distribution Directive (2016/97/EC) - EU legislation.

Rights under contracts of insurance are specified investments. There are
other specified activities set out in RAO 2001 which specifically cover
insurance contracts.

If a solicitor assists a client in obtaining a contract of insurance,
even if they just introduce the client to an insurance broker, the
solicitor will be carrying out a specified activity.

If a solicitor is involved in an insurance claim against an insurance
company, this will also be caught. All these activities involve
insurance distribution.

So for insurance distribution, firm will have to:

\begin{itemize}
\tightlist
\item
  Rely on s,327 exemption, or
\item
  Seek authorisation from the FCA
\item
  Or rely on limited exemptions for insurance distribution activities
  (outside scope).
\end{itemize}

\hypertarget{financial-promotions}{%
\subsection{Financial Promotions}\label{financial-promotions}}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{title: FSMA 2000, s 21}
\NormalTok{A solicitor who is not authorised by the FCA will be unable to make a financial promotion (i.e., communicate an invitation or inducement to engage in investment activity) unless its contents are approved by an authorised person. }
\end{Highlighting}
\end{Shaded}

Use the same 4-stage test as for regulated activities. But note that the
exclusion test is not applicable. So although you may be carrying out an
activity which is not regulated, it may be controlled and subject to
restrictions on financial promotions.

Almost anything you say or write in connection with many transactions
could be construed as financial promotion.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{A solicitor advising PRs to sell the deceased\textquotesingle{}s assets could be an invitation to deal in investments if the estate includes shares or gilts.}
\end{Highlighting}
\end{Shaded}

Both real-time and non-real-time communications are caught.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{title: Solicited communication}
\NormalTok{A real{-}time communication is solicited where it is made in the course of a personal visit, phone call or other interactive dialogue, and that}
\NormalTok{{-} was initiated by the recipient, or}
\NormalTok{{-} takes place as a result of an express request from the recipient of the communication. }
\end{Highlighting}
\end{Shaded}

\hypertarget{exemptions}{%
\subsubsection{Exemptions}\label{exemptions}}

Some exemptions are set out in FPO 2005. These include:

\begin{itemize}
\tightlist
\item
  Trustees, PRs
\item
  Takeover of body corporate.
\end{itemize}

\hypertarget{exemptions-for-exempt-professional-firms}{%
\subsubsection{Exemptions for Exempt Professional
Firms}\label{exemptions-for-exempt-professional-firms}}

Two exemptions for firms claiming the special exemption under FSMA 2000
for exempt regulated activities.

\hypertarget{real-time-promotions-fpo-2005-art-55}{%
\paragraph{Real-time Promotions (FPO 2005, Art
55)}\label{real-time-promotions-fpo-2005-art-55}}

A solicitor who carries out exempt regulated activities may make a
real-time promotion:

\begin{itemize}
\tightlist
\item
  If made to a client who has engaged the solicitor to provide
  professional services; and
\item
  Where the controlled activity to which the communication relates is
  exempt because of the exemption for professional firms, or is exempted
  by the `necessary' exemption; and
\item
  Where the controlled activity would be undertaken for the purposes of,
  and incidental to, the provision of professional services at the
  request of the client.
\end{itemize}

\hypertarget{non-real-time-promotions-fpo-2005-art-55a}{%
\paragraph{Non-real-time Promotions (FPO 2005, Art
55A)}\label{non-real-time-promotions-fpo-2005-art-55a}}

Promotion must contain certain specific statements.

\hypertarget{one-off-promotions-fpo-2005-arts-2828a}{%
\subsubsection{One-off Promotions (FPO 2005 Arts
28/28A)}\label{one-off-promotions-fpo-2005-arts-2828a}}

One-off non-real-time communications and solicited real-time
communications are exempt if certain conditions are met; essentially if
they are personal to the client.

One-off, unsolicited real-time communications are exempt provided the
solicitor believes on reasonable grounds, that:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  The client understands the risks associated with engaging in the
  investment activity, and
\item
  At the time of communication, the client would expect to be contacted
  by the solicitor in relation to that investment activity.
\end{enumerate}

\hypertarget{introducers}{%
\subsubsection{Introducers}\label{introducers}}

A solicitor may make any real-time communication in order to introduce a
client to an ATP, provided (art 15 FPO 2005):

\begin{itemize}
\tightlist
\item
  Not connected to the ATP (e.g., a close relative)
\item
  Do not receive from anyone but the client any money/advantage from
  making the introduction
\item
  The client is not seeking/ has not sought advice from you as to the
  merits of engaging in investment activity.
\end{itemize}

\end{document}
