% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{article}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage[margin=1in]{geometry}
\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\newenvironment{Shaded}{}{}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{1.00,0.00,0.00}{\textbf{#1}}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textbf{\textit{#1}}}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.49,0.56,0.16}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.25,0.63,0.44}{#1}}
\newcommand{\BuiltInTok}[1]{#1}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{#1}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textit{#1}}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textbf{\textit{#1}}}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.53,0.00,0.00}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.00,0.44,0.13}{\textbf{#1}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.56,0.13,0.00}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.25,0.63,0.44}{#1}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.73,0.13,0.13}{\textit{#1}}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{1.00,0.00,0.00}{\textbf{#1}}}
\newcommand{\ExtensionTok}[1]{#1}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.25,0.63,0.44}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.02,0.16,0.49}{#1}}
\newcommand{\ImportTok}[1]{#1}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textbf{\textit{#1}}}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.00,0.44,0.13}{\textbf{#1}}}
\newcommand{\NormalTok}[1]{#1}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.40,0.40,0.40}{#1}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.00,0.44,0.13}{#1}}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.74,0.48,0.00}{#1}}
\newcommand{\RegionMarkerTok}[1]{#1}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{#1}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.73,0.40,0.53}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.10,0.09,0.49}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{#1}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textbf{\textit{#1}}}}
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\usepackage{xcolor}
\definecolor{aliceblue}{HTML}{F0F8FF}
\definecolor{antiquewhite}{HTML}{FAEBD7}
\definecolor{aqua}{HTML}{00FFFF}
\definecolor{aquamarine}{HTML}{7FFFD4}
\definecolor{azure}{HTML}{F0FFFF}
\definecolor{beige}{HTML}{F5F5DC}
\definecolor{bisque}{HTML}{FFE4C4}
\definecolor{black}{HTML}{000000}
\definecolor{blanchedalmond}{HTML}{FFEBCD}
\definecolor{blue}{HTML}{0000FF}
\definecolor{blueviolet}{HTML}{8A2BE2}
\definecolor{brown}{HTML}{A52A2A}
\definecolor{burlywood}{HTML}{DEB887}
\definecolor{cadetblue}{HTML}{5F9EA0}
\definecolor{chartreuse}{HTML}{7FFF00}
\definecolor{chocolate}{HTML}{D2691E}
\definecolor{coral}{HTML}{FF7F50}
\definecolor{cornflowerblue}{HTML}{6495ED}
\definecolor{cornsilk}{HTML}{FFF8DC}
\definecolor{crimson}{HTML}{DC143C}
\definecolor{cyan}{HTML}{00FFFF}
\definecolor{darkblue}{HTML}{00008B}
\definecolor{darkcyan}{HTML}{008B8B}
\definecolor{darkgoldenrod}{HTML}{B8860B}
\definecolor{darkgray}{HTML}{A9A9A9}
\definecolor{darkgreen}{HTML}{006400}
\definecolor{darkgrey}{HTML}{A9A9A9}
\definecolor{darkkhaki}{HTML}{BDB76B}
\definecolor{darkmagenta}{HTML}{8B008B}
\definecolor{darkolivegreen}{HTML}{556B2F}
\definecolor{darkorange}{HTML}{FF8C00}
\definecolor{darkorchid}{HTML}{9932CC}
\definecolor{darkred}{HTML}{8B0000}
\definecolor{darksalmon}{HTML}{E9967A}
\definecolor{darkseagreen}{HTML}{8FBC8F}
\definecolor{darkslateblue}{HTML}{483D8B}
\definecolor{darkslategray}{HTML}{2F4F4F}
\definecolor{darkslategrey}{HTML}{2F4F4F}
\definecolor{darkturquoise}{HTML}{00CED1}
\definecolor{darkviolet}{HTML}{9400D3}
\definecolor{deeppink}{HTML}{FF1493}
\definecolor{deepskyblue}{HTML}{00BFFF}
\definecolor{dimgray}{HTML}{696969}
\definecolor{dimgrey}{HTML}{696969}
\definecolor{dodgerblue}{HTML}{1E90FF}
\definecolor{firebrick}{HTML}{B22222}
\definecolor{floralwhite}{HTML}{FFFAF0}
\definecolor{forestgreen}{HTML}{228B22}
\definecolor{fuchsia}{HTML}{FF00FF}
\definecolor{gainsboro}{HTML}{DCDCDC}
\definecolor{ghostwhite}{HTML}{F8F8FF}
\definecolor{gold}{HTML}{FFD700}
\definecolor{goldenrod}{HTML}{DAA520}
\definecolor{gray}{HTML}{808080}
\definecolor{green}{HTML}{008000}
\definecolor{greenyellow}{HTML}{ADFF2F}
\definecolor{grey}{HTML}{808080}
\definecolor{honeydew}{HTML}{F0FFF0}
\definecolor{hotpink}{HTML}{FF69B4}
\definecolor{indianred}{HTML}{CD5C5C}
\definecolor{indigo}{HTML}{4B0082}
\definecolor{ivory}{HTML}{FFFFF0}
\definecolor{khaki}{HTML}{F0E68C}
\definecolor{lavender}{HTML}{E6E6FA}
\definecolor{lavenderblush}{HTML}{FFF0F5}
\definecolor{lawngreen}{HTML}{7CFC00}
\definecolor{lemonchiffon}{HTML}{FFFACD}
\definecolor{lightblue}{HTML}{ADD8E6}
\definecolor{lightcoral}{HTML}{F08080}
\definecolor{lightcyan}{HTML}{E0FFFF}
\definecolor{lightgoldenrodyellow}{HTML}{FAFAD2}
\definecolor{lightgray}{HTML}{D3D3D3}
\definecolor{lightgreen}{HTML}{90EE90}
\definecolor{lightgrey}{HTML}{D3D3D3}
\definecolor{lightpink}{HTML}{FFB6C1}
\definecolor{lightsalmon}{HTML}{FFA07A}
\definecolor{lightseagreen}{HTML}{20B2AA}
\definecolor{lightskyblue}{HTML}{87CEFA}
\definecolor{lightslategray}{HTML}{778899}
\definecolor{lightslategrey}{HTML}{778899}
\definecolor{lightsteelblue}{HTML}{B0C4DE}
\definecolor{lightyellow}{HTML}{FFFFE0}
\definecolor{lime}{HTML}{00FF00}
\definecolor{limegreen}{HTML}{32CD32}
\definecolor{linen}{HTML}{FAF0E6}
\definecolor{magenta}{HTML}{FF00FF}
\definecolor{maroon}{HTML}{800000}
\definecolor{mediumaquamarine}{HTML}{66CDAA}
\definecolor{mediumblue}{HTML}{0000CD}
\definecolor{mediumorchid}{HTML}{BA55D3}
\definecolor{mediumpurple}{HTML}{9370DB}
\definecolor{mediumseagreen}{HTML}{3CB371}
\definecolor{mediumslateblue}{HTML}{7B68EE}
\definecolor{mediumspringgreen}{HTML}{00FA9A}
\definecolor{mediumturquoise}{HTML}{48D1CC}
\definecolor{mediumvioletred}{HTML}{C71585}
\definecolor{midnightblue}{HTML}{191970}
\definecolor{mintcream}{HTML}{F5FFFA}
\definecolor{mistyrose}{HTML}{FFE4E1}
\definecolor{moccasin}{HTML}{FFE4B5}
\definecolor{navajowhite}{HTML}{FFDEAD}
\definecolor{navy}{HTML}{000080}
\definecolor{oldlace}{HTML}{FDF5E6}
\definecolor{olive}{HTML}{808000}
\definecolor{olivedrab}{HTML}{6B8E23}
\definecolor{orange}{HTML}{FFA500}
\definecolor{orangered}{HTML}{FF4500}
\definecolor{orchid}{HTML}{DA70D6}
\definecolor{palegoldenrod}{HTML}{EEE8AA}
\definecolor{palegreen}{HTML}{98FB98}
\definecolor{paleturquoise}{HTML}{AFEEEE}
\definecolor{palevioletred}{HTML}{DB7093}
\definecolor{papayawhip}{HTML}{FFEFD5}
\definecolor{peachpuff}{HTML}{FFDAB9}
\definecolor{peru}{HTML}{CD853F}
\definecolor{pink}{HTML}{FFC0CB}
\definecolor{plum}{HTML}{DDA0DD}
\definecolor{powderblue}{HTML}{B0E0E6}
\definecolor{purple}{HTML}{800080}
\definecolor{red}{HTML}{FF0000}
\definecolor{rosybrown}{HTML}{BC8F8F}
\definecolor{royalblue}{HTML}{4169E1}
\definecolor{saddlebrown}{HTML}{8B4513}
\definecolor{salmon}{HTML}{FA8072}
\definecolor{sandybrown}{HTML}{F4A460}
\definecolor{seagreen}{HTML}{2E8B57}
\definecolor{seashell}{HTML}{FFF5EE}
\definecolor{sienna}{HTML}{A0522D}
\definecolor{silver}{HTML}{C0C0C0}
\definecolor{skyblue}{HTML}{87CEEB}
\definecolor{slateblue}{HTML}{6A5ACD}
\definecolor{slategray}{HTML}{708090}
\definecolor{slategrey}{HTML}{708090}
\definecolor{snow}{HTML}{FFFAFA}
\definecolor{springgreen}{HTML}{00FF7F}
\definecolor{steelblue}{HTML}{4682B4}
\definecolor{tan}{HTML}{D2B48C}
\definecolor{teal}{HTML}{008080}
\definecolor{thistle}{HTML}{D8BFD8}
\definecolor{tomato}{HTML}{FF6347}
\definecolor{turquoise}{HTML}{40E0D0}
\definecolor{violet}{HTML}{EE82EE}
\definecolor{wheat}{HTML}{F5DEB3}
\definecolor{white}{HTML}{FFFFFF}
\definecolor{whitesmoke}{HTML}{F5F5F5}
\definecolor{yellow}{HTML}{FFFF00}
\definecolor{yellowgreen}{HTML}{9ACD32}
\usepackage[most]{tcolorbox}

\usepackage{ifthen}
\provideboolean{admonitiontwoside}
\makeatletter%
\if@twoside%
\setboolean{admonitiontwoside}{true}
\else%
\setboolean{admonitiontwoside}{false}
\fi%
\makeatother%

\newenvironment{env-31dc4002-571c-4c68-8eaf-69aecb086f1f}
{
    \savenotes\tcolorbox[blanker,breakable,left=5pt,borderline west={2pt}{-4pt}{firebrick}]
}
{
    \endtcolorbox\spewnotes
}
                

\newenvironment{env-5b1fa31a-cacb-478f-b341-687551f3a18e}
{
    \savenotes\tcolorbox[blanker,breakable,left=5pt,borderline west={2pt}{-4pt}{blue}]
}
{
    \endtcolorbox\spewnotes
}
                

\newenvironment{env-f2aa0ca7-c4b6-41cf-b821-e7cd2d42b933}
{
    \savenotes\tcolorbox[blanker,breakable,left=5pt,borderline west={2pt}{-4pt}{green}]
}
{
    \endtcolorbox\spewnotes
}
                

\newenvironment{env-dbab329c-1ff1-4e71-b7fc-dee98e4c23d8}
{
    \savenotes\tcolorbox[blanker,breakable,left=5pt,borderline west={2pt}{-4pt}{aquamarine}]
}
{
    \endtcolorbox\spewnotes
}
                

\newenvironment{env-7f6725cc-73a1-4a20-ac3d-bb8896c69a69}
{
    \savenotes\tcolorbox[blanker,breakable,left=5pt,borderline west={2pt}{-4pt}{orange}]
}
{
    \endtcolorbox\spewnotes
}
                

\newenvironment{env-2e27ec82-0030-432c-8af1-8472e68c2cba}
{
    \savenotes\tcolorbox[blanker,breakable,left=5pt,borderline west={2pt}{-4pt}{gold}]
}
{
    \endtcolorbox\spewnotes
}
                

\newenvironment{env-5ee27ddc-967a-4409-a565-00dfcc19c733}
{
    \savenotes\tcolorbox[blanker,breakable,left=5pt,borderline west={2pt}{-4pt}{darkred}]
}
{
    \endtcolorbox\spewnotes
}
                

\newenvironment{env-20735490-40bc-4e45-ac4a-32d20ddf055c}
{
    \savenotes\tcolorbox[blanker,breakable,left=5pt,borderline west={2pt}{-4pt}{pink}]
}
{
    \endtcolorbox\spewnotes
}
                

\newenvironment{env-b6566782-d32c-4f2f-b945-35ecab55fbd8}
{
    \savenotes\tcolorbox[blanker,breakable,left=5pt,borderline west={2pt}{-4pt}{cyan}]
}
{
    \endtcolorbox\spewnotes
}
                

\newenvironment{env-66f281dc-2740-483b-8cca-5e4df2d3aab2}
{
    \savenotes\tcolorbox[blanker,breakable,left=5pt,borderline west={2pt}{-4pt}{cyan}]
}
{
    \endtcolorbox\spewnotes
}
                

\newenvironment{env-8a18b9a7-7473-4f2b-8c80-f70982999ac8}
{
    \savenotes\tcolorbox[blanker,breakable,left=5pt,borderline west={2pt}{-4pt}{purple}]
}
{
    \endtcolorbox\spewnotes
}
                

\newenvironment{env-39e0192b-2d1c-4f2b-bae6-8c8d54728882}
{
    \savenotes\tcolorbox[blanker,breakable,left=5pt,borderline west={2pt}{-4pt}{darksalmon}]
}
{
    \endtcolorbox\spewnotes
}
                

\newenvironment{env-5ba29b22-f478-49ce-bbcf-17ec1b010fd6}
{
    \savenotes\tcolorbox[blanker,breakable,left=5pt,borderline west={2pt}{-4pt}{gray}]
}
{
    \endtcolorbox\spewnotes
}
                
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

\author{}
\date{}

\begin{document}

{
\setcounter{tocdepth}{3}
\tableofcontents
}
\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{min\_depth: 1}
\end{Highlighting}
\end{Shaded}

\hypertarget{consequences-of-exchange}{%
\section{Consequences of Exchange}\label{consequences-of-exchange}}

\hypertarget{effects}{%
\subsection{Effects}\label{effects}}

A binding contract exists from which normally neither party may withdraw
without incurring liability for breach. At common law, the beneficial
ownership in the property passes to the buyer, who becomes entitled to
any increase in value of the property but bears the risk of any loss
damage.

The seller retains legal title until completion, but holds the
beneficial interest on behalf of the buyer. Seller entitled to remain in
possession of the property and to any rental income, unless otherwise
agreed. Must pay any outgoings until completion. Owes a duty of care to
the buyer, and liable in damages if loss is caused through neglect or
destruction ({[}{[}Clarke v Ramuz {[}1891{]} 2 QB 456{]}{]};
{[}{[}Phillips v Lamdin {[}1949{]} 2 KB 33{]}{]}). Duty does not
terminate even if the seller vacates the property before completion.

\hypertarget{immediately-after}{%
\subsection{Immediately After}\label{immediately-after}}

\begin{itemize}
\tightlist
\item
  Seller's solicitor

  \begin{itemize}
  \tightlist
  \item
    Should inform the client that exchange has happened.
  \item
    If exchange telephonic, copy of the contract signed by seller
    immediately sent to buyer's solicitor (to fulfil any undertaking
    given during the phone call).
  \item
    Any received deposit immediately paid into client deposit account.
  \end{itemize}
\item
  Buyer's solicitor

  \begin{itemize}
  \tightlist
  \item
    Inform client and lender that exchange has happened
  \item
    If exchange telephonic, send seller's solicitor the signed contract
    and deposit cheque in accordance with the undertaking given, after
    checking the contract is dated and contains the agreed completion
    date.
  \item
    Protect the contract by registration, where appropriate.
  \end{itemize}
\end{itemize}

\hypertarget{registration-of-contract}{%
\subsection{Registration of Contract}\label{registration-of-contract}}

In practice, registration is uncommon since completion often takes place
shortly after exchange But should be registered if:

\begin{itemize}
\tightlist
\item
  There is a long delay (\textgreater2 months) before exchange and
  completion
\item
  Reason to doubt the seller's good faith
\item
  Dispute between seller and buyer
\item
  Seller delays completion beyond the contractual date.
\end{itemize}

Registered land: protect by entry of a notice on the register of title.

\hypertarget{buyer-in-possession}{%
\subsection{Buyer in Possession}\label{buyer-in-possession}}

\hypertarget{buyers-occupation-before-completion}{%
\subsubsection{Buyer's Occupation Before
Completion}\label{buyers-occupation-before-completion}}

Seller entitled to retain possession until completion, unless otherwise
agreed. Buyer's request to enter and occupy the premises before
completion should be regarded with caution (loss of incentive to
complete on time, hard to regain possession of property if required).

If the buyer is allowed to possess before completion, ensure the buyer's
occupation is as licensee and not a tenant, to avoid security of tenure
issues.

SCPCs contain no provisions allowing a buyer to take possession prior to
completion, nor any provisions regulating such occupation.

\hypertarget{tenanted-property}{%
\subsection{Tenanted Property}\label{tenanted-property}}

\begin{itemize}
\tightlist
\item
  SCPC 5.2.6: seller not to give or formally withhold any licence or
  approve under the lease without the buyer's consent.
\item
  SCPC 6.1.5: neither party is to agree a new rent figure unless the
  other agrees.
\end{itemize}

\hypertarget{death-of-a-contracting-party}{%
\subsection{Death of a Contracting
Party}\label{death-of-a-contracting-party}}

Death of one of the contracting parties between contract and completion
does not affect the validity of the contract. Benefit and burden of the
contract pass to the deceased's personal representatives, who are bound
to complete.

If completion does not take place by the contractual completion date,
there is a breach of contract. If time was not originally of the
essence, may be made so by service of a notice to complete addressed to
the deceased and executors named in the seller's will. A copy of the
notice must be sent to the Public Trustee.

\hypertarget{death-of-co-seller}{%
\subsubsection{Death of Co-seller}\label{death-of-co-seller}}

\begin{itemize}
\tightlist
\item
  Property owned by beneficial co-owners is held on a trust of land.
\item
  If following death there remain \(\geq 2\) trustees of the legal
  estate, the transaction can proceed to completion.
\item
  Else, (for a beneficial joint tenancy) appoint another trustee, so
  there are at least 2.
\end{itemize}

\hypertarget{death-of-buyer}{%
\subsubsection{Death of Buyer}\label{death-of-buyer}}

Personal representatives of the deceased bound to complete the contract.
Delay in completion may occur:

\begin{itemize}
\tightlist
\item
  Transfer deed will need redrafting
\item
  PRs cannot complete until they have obtained a grant of representation
\item
  Mortgage offer usually revoked.
\item
  Survivor of joint buyers can be forced to complete.
\end{itemize}

\hypertarget{pre-completion-steps}{%
\subsection{Pre-completion Steps}\label{pre-completion-steps}}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Preparation of the transfer deed
\item
  Redemption of seller's mortgage and creation of buyer's mortgage
\item
  Pre-completion searches by the buyer
\item
  Ensuring financial aspects in order.
\end{enumerate}

Conventionally 4 weeks between exchange and completion, but 2 weeks or
less not uncommon. Sometimes exchange and completion are simultaneous.

\hypertarget{request-for-completion-information}{%
\section{Request for Completion
Information}\label{request-for-completion-information}}

After exchange, the buyer's solicitor requests completion information
from the seller's solicitor (``requisitions on title''). Note that
although historically, title was deduced at this stage, evidence of
title is now almost always supplied to the buyer before exchange. The
contract usually contains a clause excluding the buyer's right to raise
requisitions once contracts have been exchanged (SCPC 7.2.1). If after
exchange an undisclosed incumbrance/ defect is discovered by the buyer's
solicitor, they will be able to require it to be remedied/ assert
remedies.

\begin{itemize}
\tightlist
\item
  SCPC 7.2.2: buyer must raise any further title requisitions within 6
  working days of a matter coming to her attention
\item
  SCPC 7.3.1: seller's solicitor must reply to requisitions 4 working
  days after receiving them from buyer's solicitor.
\end{itemize}

\hypertarget{standard-form-requests}{%
\subsection{Standard Form Requests}\label{standard-form-requests}}

Law stationers produce a standard form including common questions. Add
more to the bottom. Send 2 copies to seller's solicitor, who will return
one copy with their answers.

\hypertarget{discharge-of-sellers-existing-mortgage}{%
\subsection{Discharge of Seller's Existing
Mortgage}\label{discharge-of-sellers-existing-mortgage}}

Use TA13 Completion Information and Undertakings. Form contains a
request for confirmation that existing mortgages will be discharged on
completion. The answer to the question takes effect as an undertaking to
discharge mortgages referred to.

\hypertarget{transfer-deed}{%
\section{Transfer Deed}\label{transfer-deed}}

\hypertarget{preparation}{%
\subsection{Preparation}\label{preparation}}

\begin{itemize}
\tightlist
\item
  Usually the buyer's duty to prepare

  \begin{itemize}
  \tightlist
  \item
    Transfer prepared shortly after exchange and sent to seller for
    approval with buyer's Request for Completion Information
  \item
    Draft transfer be sent \(\geq 12\) working days before completion
    date (SCPC 7.3.2)
  \end{itemize}
\item
  Seller may reserve the right by contractual condition to prepare the
  transfer himself (s 48(1) LPA 1925).
\end{itemize}

SCPC: seller's solicitor must submit the draft transfer to seller's
solicitor when they approve the draft contract, which may be well before
exchange of contracts.

\hypertarget{form-of-transfer}{%
\subsection{Form of Transfer}\label{form-of-transfer}}

Transfer must be a deed to effect transfer of the legal estate to the
buyer (s 52 LPA 1925). Must reflect the terms of the contract.

\hypertarget{registered-land}{%
\subsubsection{Registered Land}\label{registered-land}}

Prescribed form of wording must be used. Standard Land Registry forms
are usually used. A traditional conveyance cannot be used.

\hypertarget{agreeing-the-transfer}{%
\subsection{Agreeing the Transfer}\label{agreeing-the-transfer}}

Buyer's solicitor must have access to:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  The contract
\item
  Official copy entries
\item
  Precedent on which to base the transfer under preparation.
\end{enumerate}

Once the draft has been prepared, 2 copies should be sent to the
seller's solicitor for their approval. Amendments should be restricted
to those necessary for the fulfilment of the document's legal purpose.
Must be returned 4 working days after delivery to the buyer's solicitor
(SCPC 7.3.2).

When amendments have been finalised, the buyer's solicitor should
prepare an engrossment (clean copy) of the transfer. Seller to execute
(sign). If the buyer also needs to execute, this is commonly done before
sending to the seller's solicitor.

SCPC 6.3.2: buyer must deliver engrossment of the transfer to the seller
at least 5 working days before completion.

\hypertarget{execution-and-delivery}{%
\subsection{Execution and Delivery}\label{execution-and-delivery}}

A valid deed must (s 1 LP(MP)A 1989):

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Indicate clearly that it is a deed
\item
  Be signed by the necessary parties in the presence of a witness; and
\item
  Be delivered.
\end{enumerate}

Signature of the seller is required to transfer the legal estate (s 52
LPA 1925). Buyer required to execute if the transfer contains a
covenant/ declaration on his behalf (e.g., indemnity covenant in respect
of existing covenants, or a declaration by buyers relating to trusts on
which they hold the property). If there is another party (e.g., lender
to release a mortgage), they should also sign.

Deed must be delivered: parties must be intended to be bound by it. This
takes effect on delivery. Common for the buyer to deliver the transfer
to the seller ``in escrow'' (conditionally) so the operation of the
transfer is postponed until completion.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{The letter accompanying the transfer should explain to the client: }
\NormalTok{{-} Contents and puropse}
\NormalTok{{-} Clear instructions relating to execution of the transfer}
\NormalTok{{-} When to return by}
\NormalTok{{-} Ask client not to date the transfer}
\end{Highlighting}
\end{Shaded}

\hypertarget{other}{%
\subsection{Other}\label{other}}

\hypertarget{plans}{%
\subsubsection{Plans}\label{plans}}

If the contract provides for the use of a plan, the transfer will also
refer to the plan. Not normally considered necessary for the sale of the
whole of a seller's property.

\hypertarget{transfer-of-whole}{%
\subsubsection{Transfer of Whole}\label{transfer-of-whole}}

A transfer of the whole of a registered title will follow TR1.

\hypertarget{pre-completion-searches}{%
\section{Pre-completion Searches}\label{pre-completion-searches}}

\hypertarget{bankruptcy-searches}{%
\subsection{Bankruptcy Searches}\label{bankruptcy-searches}}

\hypertarget{against-buyer-borrower}{%
\subsubsection{Against buyer/ Borrower}\label{against-buyer-borrower}}

Lender will require a clear bankruptcy search against the name of the
buyer/ borrower before releasing mortgage funds. Unless a full search of
the register has been made on Form K15, the lender's solicitor should
submit Form K16 electronically to Land Registry, completed with full and
correct names of the borrower. A search certificate will be returned. If
there is an adverse entry revealed, lender's instructions must be
immediately obtained.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{If the bankruptcy search reveals an adverse entry, the buyer\textquotesingle{}s solicitor can certify that the entry does not relate to the borrower (relates to someone else with the same name). But this is construed as an undertaking given by the solicitor. }
\end{Highlighting}
\end{Shaded}

Can also check Insolvency Service website, which will also reveal date
of birth.

\hypertarget{against-the-seller}{%
\subsubsection{Against the Seller}\label{against-the-seller}}

For registered land, most conveyancing solicitors do not consider a
bankruptcy check on an individual seller necessary. s 86(5) LRA 2002
provides that where no notice or restriction relating to bankruptcy
appears on the title, a buyer for value acting in good faith without
notice of the bankruptcy petition will take good title from the
bankruptcy seller. May be worth doing in a v high value transaction.

\hypertarget{company-search}{%
\subsection{Company Search}\label{company-search}}

May be done as a pre-contract search. Update at this stage of the
transaction.

\hypertarget{powers-of-attorney}{%
\subsection{Powers of Attorney}\label{powers-of-attorney}}

If the transfer deed is to be executed by a person acting under the
authority of an enduring or lasting power of attorney, do a search with
the Office of Public Guardian to check whether or not registration of
the power is effected/ pending.

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.0231}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.8150}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.1618}}@{}}
\toprule()
\begin{minipage}[b]{\linewidth}\raggedright
Power
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Registered
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Pending
\end{minipage} \\
\midrule()
\endhead
Enduring & Transaction may proceed to completion, provided the buyer has
no actual knowledge of any circumstances that would revoke the power.
Power no longer revocable without notification to the Public Guardian.
Office copy of the power should be obtained from the Office of Public
Guardian. & Do not proceed. Suggests mental incapacity of the donor. \\
Lasting & As above. & Do not proceed. Only comes into effect once
registered. \\
\bottomrule()
\end{longtable}

\hypertarget{local-land-searches-and-enquiries}{%
\subsection{Local Land Searches and
Enquiries}\label{local-land-searches-and-enquiries}}

Searches invariably made before exchange of contracts. Only repeat prior
to completion if:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  \(> 2\) months between exchange and completion and the search is not
  covered/ replaced by insurance.
\item
  Information received by the buyer's solicitor suggests a further
  search may be advisable.
\item
  Contract conditional on the satisfactory results of later searches.
\end{enumerate}

\hypertarget{inspection-of-the-property}{%
\subsection{Inspection of the
Property}\label{inspection-of-the-property}}

May be necessary just before completion.

\hypertarget{results-of-searches}{%
\subsection{Results of Searches}\label{results-of-searches}}

If the searches reveal something unexpected:

\begin{itemize}
\tightlist
\item
  Contact the seller's solicitor immediately to find out exactly what
  the issue is and if it can be rectified before completion.
\item
  For a Land Charges Department search, apply for an official copy of
  the entry using Form K19 (reveal name and address of the person with
  the benefit of the charge).
\item
  Keep the client, lender and other solicitors involved informed.
\end{itemize}

\hypertarget{removing-entry-from-the-register}{%
\subsubsection{Removing Entry From the
Register}\label{removing-entry-from-the-register}}

An application form for the removal of an entry from the register in
either registered or unregistered land will be accepted by the Chief
Land Registrar only if it is signed by the person with the benefit of
the charge or a person acting on his behalf.

An application form signed by the seller's solicitors, or an undertaking
given by them on completion to secure the removal of the charge may not,
therefore, suffice, unless the seller is the person with the benefit of
the charge (see Holmes v Kennard (H) and Son (A Firm) (1985) 49 P \& CR
202).

\hypertarget{irrelevant-land-charges-entries}{%
\subsubsection{Irrelevant Land Charges
Entries}\label{irrelevant-land-charges-entries}}

Charges registered at the Land Charges Department can only validly be
entered against the name of an estate owner in relation to the period in
which he owned the land in question. Sometimes the system throws up
irrelevant ones. Often seller's solicitor asked to certify that an entry
is irrelevant (undertaking).

\hypertarget{entries-that-cannot-be-removed}{%
\subsubsection{Entries That Cannot Be
Removed}\label{entries-that-cannot-be-removed}}

If a seller is unwilling/ unable to remove an adverse entry revealed by
a pre-completion search, check contract terms to see if seller is in
breach.

\hypertarget{official-certificates-of-search}{%
\subsubsection{Official Certificates of
Search}\label{official-certificates-of-search}}

An official certificate of search issued by Land Registry is not
conclusive in favour of the\\
searcher, who will thus take his interest in the land subject to
whatever entries are on the\\
register, irrespective of whether or not they were revealed by the
search certificate ({[}{[}Parkash v Irani Finance Ltd {[}1970{]} Ch
101{]}{]}).

\hypertarget{priority-periods}{%
\subsection{Priority Periods}\label{priority-periods}}

An official certificate of search issued by Land Registry following a
search made on Form OS1 or OS2 gives a priority period to the searcher
of 30 working days from the Certificate Date in the search. The searcher
will take priority over any entry made during the priority period,
provided that completion takes place and a correct application for
registration of the transaction is received by the appropriate Land
Registry Office by 12 noon on the day when the priority period given by
the search expires.

The expiry of the priority period is shown on the search certificate.
Cannot be extended.

Land Charge Searches:

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.2651}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.3614}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.3735}}@{}}
\toprule()
\begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Local
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Central
\end{minipage} \\
\midrule()
\endhead
Form: & LLC1 & K15 \\
When to make: & in every transaction & unregistered land only \\
Time to make: & before exchange & part of investigation of title \\
Send to: & local authority/Land Registry & Plymouth \\
Search against: & description of land & owners' names \\
Information revealed: & mainly public incumbrances & mainly private
incumbrances \\
Protection of search: & none & priority for 15 working days \\
\bottomrule()
\end{longtable}

Land Registry vs Land Charges searches:

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.1717}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.4949}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.3333}}@{}}
\toprule()
\begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Land Registry
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Land Charges Department
\end{minipage} \\
\midrule()
\endhead
Form: & OS1 or OS2 & K15 \\
When to use: & registered land & unregistered land \\
Search against: & title number & owners' names \\
Protectiongiven: & search not conclusive but compensation available &
conclusive in favour of searcher \\
Fee: & standard fee for each search & fee for each name searched \\
Priority period: & 30 working days & 15 working days \\
\bottomrule()
\end{longtable}

\hypertarget{preparing-for-completion}{%
\section{Preparing for Completion}\label{preparing-for-completion}}

\hypertarget{sellers-checklist}{%
\subsection{Seller's Checklist}\label{sellers-checklist}}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Check transfer deed approved and replies given to buyer's Request for
  Completion Information
\item
  Receive engrossed transfer deed from buyer
\item
  Obtain seller's signature and witness to transfer deed
\item
  Obtain redemption figure for seller's mortgage
\item
  Obtain receipts etc. where apportionments are to be made.
\item
  Prepare completion statement and send 2 copies to the buyer
\item
  Remind client to do final meter readings
\item
  Prepare undertakings to be given on completion
\item
  Contact lender to confirm arrangements for discharge of mortgage and
  payment.
\item
  Check file for queries
\item
  Prepare list of matters
\item
  Locate deeds and documents to be inspected/ handed over and prepare
  certified copies for the buyer
\item
  Prepare 2 copies of schedule of deeds to be handed to the buyer.
\item
  Prepare inventory and receipt for chattels
\item
  Check vacant possession and arrangement for handing over keys
\item
  Receive instructions from buyer's solicitor to act as their agent on
  completion
\item
  Inform estate agents of completion arrangements
\item
  Prepare bill.
\end{enumerate}

\hypertarget{buyers-checklist}{%
\subsection{Buyer's Checklist}\label{buyers-checklist}}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Ensure transfer deed approved and Request for Completion Information
  answered
\item
  Engross transfer and mortgage deeds
\item
  Get buyer to execute mortgage deed, transfer deed (and plan) and
  return
\item
  Send (executed) transfer deed to seller's solicitor for client's
  execution.
\item
  Pre-completion searches
\item
  Report on title to lender in format specified by lender and request
  mortgage advance
\item
  Receive and check completion statement and copies of last receipts.
\item
  Prepare and agree any undertaking to be given or received on
  completion
\item
  Confirm final arrangements with lender
\item
  Ensure any life policy required by the lender has been obtained.
\item
  Check file for queries
\item
  Prepare account and bill for client, submit together with a copy of
  the completion statement.
\item
  Receive mortgage advance from lender and balance of funds from buyer.
  Pay into client account and clear funds.
\item
  Arrange for final inspection by lender's valuer, if necessary.
\item
  Prepare matters to deal with on completion.
\item
  Check arrangements for vacant possession and handing over keys.
\item
  Instruct seller's solicitor to act as agent on completion.
\item
  Arrange time/ place of completion.
\item
  Ensure estate agents aware.
\item
  Arrange to send completion money to seller's solicitor.
\item
  SDLT forms.
\end{enumerate}

\hypertarget{client-mortgage}{%
\subsection{Client Mortgage}\label{client-mortgage}}

\hypertarget{seller}{%
\subsubsection{Seller}\label{seller}}

Seller must obtain from the seller's lender a ``redemption statement''
showing the exact amount to discharge the seller's mortgage. Daily rate
usually included.

Lender must inform Land Registry of the discharge of mortgage: Form DS1
or electronic ED system.

\hypertarget{buyer}{%
\subsubsection{Buyer}\label{buyer}}

After completing investigation of title, the solicitor acting for the
buyer's lender reports to the lender that that title to the property is
safe, marketable and acceptable security. City of London Law Society's
form of certificate often used. Solicitor acting for buyer's lender must
also make a formal request for the advance money from the lender.
Mortgage deed should be prepared for signature by the borrowers.

\hypertarget{apportionment}{%
\subsection{Apportionment}\label{apportionment}}

Outgoings attaching to the land can be split between parties on
completion. Calculations shown on completion statement. SCPC 9.3
includes apportionment provisions. SCPC 9.3.3: buyer is deemed to own
the property for the entirety of the day of completion.

\hypertarget{completion-statement}{%
\subsection{Completion Statement}\label{completion-statement}}

Shows the money required to complete the transaction. Usually requested
by the buyer when the Request for Completion Information is sent to the
seller. Usually only necessary when sum due includes apportionments/
other sums in excess of the balance of purchase price ({[}{[}Carne v De
Bono {[}1988{]} 3 All ER 485{]}{]}). 2 copies to be sent to the buyer.

\hypertarget{sdlt}{%
\subsection{SDLT}\label{sdlt}}

SDLT return should be made to HMRC, giving details of the transaction
and the buyer. Can be submitted online or by post using Form SDLT1. If
by post, seller must sign. If the buyer's solicitor is also acting for a
lender, a SDLT return should be completed by the solicitor and approved
by the buyer before the loan is used to finance the purchase. Solicitor
should check it has adequate cleared funds from the client to cover SDLT
payment.

If steps are not taken and the buyer refuses to pay SDLT, it will not be
possible to register the transaction at the Land Registry, so the lender
will have no security for its loan (and solicitor may be liable).

\hypertarget{money}{%
\subsection{Money}\label{money}}

Check figures sent by seller's solicitor. Think about if bridging
finance or a further loan is necessary. Solicitor may be guilty of a
criminal offence if they assist someone suspected of money laundering.
Be wary.

\hypertarget{completion-checklist}{%
\subsection{Completion Checklist}\label{completion-checklist}}

To be prepared and sent to any agents.

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Documents to be available at completion:

  \begin{enumerate}
  \def\labelenumii{(\alph{enumii})}
  \tightlist
  \item
    contract;\\
  \item
    evidence of title;\\
  \item
    copy transfer deed;\\
  \item
    replies to Request for Completion Information;\\
  \item
    completion statement.
  \end{enumerate}
\item
  Documents to be inspected by buyer:

  \begin{enumerate}
  \def\labelenumii{(\alph{enumii})}
  \tightlist
  \item
    title deeds, where in unregistered land these are not to be handed
    over on\\
    completion (e.g., on a sale of part);\\
  \item
    power(s) of attorney;\\
  \item
    grant of administration;\\
  \item
    receipts/demands for apportionments if not previously supplied.
  \end{enumerate}
\item
  Documents, etc, to be handed to buyer on completion:

  \begin{enumerate}
  \def\labelenumii{(\alph{enumii})}
  \tightlist
  \item
    where the property is unregistered land, title deeds and a schedule
    of deeds and a\\
    certified copy of any memorandum endorsed on retained deeds;\\
  \item
    original lease (on purchase of a lease);\\
  \item
    executed transfer deed;\\
  \item
    Form DS1/discharged mortgage or undertaking in respect of discharge
    of\\
    mortgage(s);\\
  \item
    receipt for money paid for chattels;\\
  \item
    keys of the property (if these are not available the seller's
    solicitor should be asked\\
    to telephone the keyholder to request the release of the keys).
  \end{enumerate}
\item
  Documents, etc, to be handed to seller on completion:

  \begin{enumerate}
  \def\labelenumii{(\alph{enumii})}
  \tightlist
  \item
    executed duplicate transfer deed/counterpart lease (where
    appropriate);\\
  \item
    receipted schedule of any deeds received from seller;\\
  \item
    release of deposit if held by third party in capacity of
    stakeholder.
  \end{enumerate}
\item
  Unregistered land only, endorsements on documents (if required by
  buyer):

  \begin{enumerate}
  \def\labelenumii{(\alph{enumii})}
  \tightlist
  \item
    endorsement of sale on most recently dated retained document of
    title (sale of\\
    part);\\
  \item
    abstract or epitome to be marked up as compared with the original
    deeds (in\\
    respect of any document the original of which is not handed over on
    completion).
  \end{enumerate}
\end{enumerate}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{{-}   In anticipation of completion, the TR1 which will transfer the legal title from the seller to the buyer will need to be executed.  The seller will always execute this document and the buyer will also do so if the buyer is "doing something" in the TR1 such as giving an indemnity covenant or declaring a beneficial interest under a trust. }
\NormalTok{{-}   On completion, the buyer\textquotesingle{}s solicitor will send the balance of the purchase price (as the seller will already have been given the deposit on exchange). }
\NormalTok{{-}   On receipt of the money, the seller\textquotesingle{}s solicitor will complete the transaction by dating the TR1.  }
\NormalTok{{-}   Once the TR1 has been dated, the buyer\textquotesingle{}s solicitor will need to date the mortgage deed.  This cannot be done until after the TR1 has been dated, as technically, the buyer does not own the property until then and so would not be in a position to grant a mortgage over it. }
\NormalTok{{-}   The seller\textquotesingle{}s solicitor will send the TR1 to the buyer\textquotesingle{}s solicitor.   SDLT must then be paid, if due.  The application to register the buyer as the new registered proprietor cannot take place until this has been done. }
\NormalTok{{-}   Once SDLT, if due, has been paid, an application to Land Registry must be made to have the buyer registered as the new registered proprietor of the land in the title and to register the new mortgage.  This must be done before the expiry of the priority period of the OS1 search that should have been carried out prior to completion. }
\NormalTok{{-}   Land Registry will process the application for registration and insert buyer as the new registered proprietor and the details of the new charge.  Once this has been done, Land Registry will forward an official copy of the entries on the register, known as a Title Information Document (or TID) to the buyer\textquotesingle{}s solicitor.  Although this is not strictly a document of title, it is for buyer to keep as evidence that it is the new owner. It should be checked to ensure that the transfer has been correctly carried out. }
\NormalTok{{-}   Once the TID has been checked, the solicitor acting for the buyer will need to deal with the safe custody of the TID and other papers.  Where the property has been bought with the aid of a mortgage, as here, it is usual for these papers to be lodged with the lender. }
\NormalTok{{-}   Finally, when all the above matters have been dealt with and the bill paid, it is possible to close the file.}
\end{Highlighting}
\end{Shaded}


\end{document}
