% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{article}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage[margin=1in]{geometry}
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\usepackage{xcolor}
\definecolor{aliceblue}{HTML}{F0F8FF}
\definecolor{antiquewhite}{HTML}{FAEBD7}
\definecolor{aqua}{HTML}{00FFFF}
\definecolor{aquamarine}{HTML}{7FFFD4}
\definecolor{azure}{HTML}{F0FFFF}
\definecolor{beige}{HTML}{F5F5DC}
\definecolor{bisque}{HTML}{FFE4C4}
\definecolor{black}{HTML}{000000}
\definecolor{blanchedalmond}{HTML}{FFEBCD}
\definecolor{blue}{HTML}{0000FF}
\definecolor{blueviolet}{HTML}{8A2BE2}
\definecolor{brown}{HTML}{A52A2A}
\definecolor{burlywood}{HTML}{DEB887}
\definecolor{cadetblue}{HTML}{5F9EA0}
\definecolor{chartreuse}{HTML}{7FFF00}
\definecolor{chocolate}{HTML}{D2691E}
\definecolor{coral}{HTML}{FF7F50}
\definecolor{cornflowerblue}{HTML}{6495ED}
\definecolor{cornsilk}{HTML}{FFF8DC}
\definecolor{crimson}{HTML}{DC143C}
\definecolor{cyan}{HTML}{00FFFF}
\definecolor{darkblue}{HTML}{00008B}
\definecolor{darkcyan}{HTML}{008B8B}
\definecolor{darkgoldenrod}{HTML}{B8860B}
\definecolor{darkgray}{HTML}{A9A9A9}
\definecolor{darkgreen}{HTML}{006400}
\definecolor{darkgrey}{HTML}{A9A9A9}
\definecolor{darkkhaki}{HTML}{BDB76B}
\definecolor{darkmagenta}{HTML}{8B008B}
\definecolor{darkolivegreen}{HTML}{556B2F}
\definecolor{darkorange}{HTML}{FF8C00}
\definecolor{darkorchid}{HTML}{9932CC}
\definecolor{darkred}{HTML}{8B0000}
\definecolor{darksalmon}{HTML}{E9967A}
\definecolor{darkseagreen}{HTML}{8FBC8F}
\definecolor{darkslateblue}{HTML}{483D8B}
\definecolor{darkslategray}{HTML}{2F4F4F}
\definecolor{darkslategrey}{HTML}{2F4F4F}
\definecolor{darkturquoise}{HTML}{00CED1}
\definecolor{darkviolet}{HTML}{9400D3}
\definecolor{deeppink}{HTML}{FF1493}
\definecolor{deepskyblue}{HTML}{00BFFF}
\definecolor{dimgray}{HTML}{696969}
\definecolor{dimgrey}{HTML}{696969}
\definecolor{dodgerblue}{HTML}{1E90FF}
\definecolor{firebrick}{HTML}{B22222}
\definecolor{floralwhite}{HTML}{FFFAF0}
\definecolor{forestgreen}{HTML}{228B22}
\definecolor{fuchsia}{HTML}{FF00FF}
\definecolor{gainsboro}{HTML}{DCDCDC}
\definecolor{ghostwhite}{HTML}{F8F8FF}
\definecolor{gold}{HTML}{FFD700}
\definecolor{goldenrod}{HTML}{DAA520}
\definecolor{gray}{HTML}{808080}
\definecolor{green}{HTML}{008000}
\definecolor{greenyellow}{HTML}{ADFF2F}
\definecolor{grey}{HTML}{808080}
\definecolor{honeydew}{HTML}{F0FFF0}
\definecolor{hotpink}{HTML}{FF69B4}
\definecolor{indianred}{HTML}{CD5C5C}
\definecolor{indigo}{HTML}{4B0082}
\definecolor{ivory}{HTML}{FFFFF0}
\definecolor{khaki}{HTML}{F0E68C}
\definecolor{lavender}{HTML}{E6E6FA}
\definecolor{lavenderblush}{HTML}{FFF0F5}
\definecolor{lawngreen}{HTML}{7CFC00}
\definecolor{lemonchiffon}{HTML}{FFFACD}
\definecolor{lightblue}{HTML}{ADD8E6}
\definecolor{lightcoral}{HTML}{F08080}
\definecolor{lightcyan}{HTML}{E0FFFF}
\definecolor{lightgoldenrodyellow}{HTML}{FAFAD2}
\definecolor{lightgray}{HTML}{D3D3D3}
\definecolor{lightgreen}{HTML}{90EE90}
\definecolor{lightgrey}{HTML}{D3D3D3}
\definecolor{lightpink}{HTML}{FFB6C1}
\definecolor{lightsalmon}{HTML}{FFA07A}
\definecolor{lightseagreen}{HTML}{20B2AA}
\definecolor{lightskyblue}{HTML}{87CEFA}
\definecolor{lightslategray}{HTML}{778899}
\definecolor{lightslategrey}{HTML}{778899}
\definecolor{lightsteelblue}{HTML}{B0C4DE}
\definecolor{lightyellow}{HTML}{FFFFE0}
\definecolor{lime}{HTML}{00FF00}
\definecolor{limegreen}{HTML}{32CD32}
\definecolor{linen}{HTML}{FAF0E6}
\definecolor{magenta}{HTML}{FF00FF}
\definecolor{maroon}{HTML}{800000}
\definecolor{mediumaquamarine}{HTML}{66CDAA}
\definecolor{mediumblue}{HTML}{0000CD}
\definecolor{mediumorchid}{HTML}{BA55D3}
\definecolor{mediumpurple}{HTML}{9370DB}
\definecolor{mediumseagreen}{HTML}{3CB371}
\definecolor{mediumslateblue}{HTML}{7B68EE}
\definecolor{mediumspringgreen}{HTML}{00FA9A}
\definecolor{mediumturquoise}{HTML}{48D1CC}
\definecolor{mediumvioletred}{HTML}{C71585}
\definecolor{midnightblue}{HTML}{191970}
\definecolor{mintcream}{HTML}{F5FFFA}
\definecolor{mistyrose}{HTML}{FFE4E1}
\definecolor{moccasin}{HTML}{FFE4B5}
\definecolor{navajowhite}{HTML}{FFDEAD}
\definecolor{navy}{HTML}{000080}
\definecolor{oldlace}{HTML}{FDF5E6}
\definecolor{olive}{HTML}{808000}
\definecolor{olivedrab}{HTML}{6B8E23}
\definecolor{orange}{HTML}{FFA500}
\definecolor{orangered}{HTML}{FF4500}
\definecolor{orchid}{HTML}{DA70D6}
\definecolor{palegoldenrod}{HTML}{EEE8AA}
\definecolor{palegreen}{HTML}{98FB98}
\definecolor{paleturquoise}{HTML}{AFEEEE}
\definecolor{palevioletred}{HTML}{DB7093}
\definecolor{papayawhip}{HTML}{FFEFD5}
\definecolor{peachpuff}{HTML}{FFDAB9}
\definecolor{peru}{HTML}{CD853F}
\definecolor{pink}{HTML}{FFC0CB}
\definecolor{plum}{HTML}{DDA0DD}
\definecolor{powderblue}{HTML}{B0E0E6}
\definecolor{purple}{HTML}{800080}
\definecolor{red}{HTML}{FF0000}
\definecolor{rosybrown}{HTML}{BC8F8F}
\definecolor{royalblue}{HTML}{4169E1}
\definecolor{saddlebrown}{HTML}{8B4513}
\definecolor{salmon}{HTML}{FA8072}
\definecolor{sandybrown}{HTML}{F4A460}
\definecolor{seagreen}{HTML}{2E8B57}
\definecolor{seashell}{HTML}{FFF5EE}
\definecolor{sienna}{HTML}{A0522D}
\definecolor{silver}{HTML}{C0C0C0}
\definecolor{skyblue}{HTML}{87CEEB}
\definecolor{slateblue}{HTML}{6A5ACD}
\definecolor{slategray}{HTML}{708090}
\definecolor{slategrey}{HTML}{708090}
\definecolor{snow}{HTML}{FFFAFA}
\definecolor{springgreen}{HTML}{00FF7F}
\definecolor{steelblue}{HTML}{4682B4}
\definecolor{tan}{HTML}{D2B48C}
\definecolor{teal}{HTML}{008080}
\definecolor{thistle}{HTML}{D8BFD8}
\definecolor{tomato}{HTML}{FF6347}
\definecolor{turquoise}{HTML}{40E0D0}
\definecolor{violet}{HTML}{EE82EE}
\definecolor{wheat}{HTML}{F5DEB3}
\definecolor{white}{HTML}{FFFFFF}
\definecolor{whitesmoke}{HTML}{F5F5F5}
\definecolor{yellow}{HTML}{FFFF00}
\definecolor{yellowgreen}{HTML}{9ACD32}
\usepackage[most]{tcolorbox}

\usepackage{ifthen}
\provideboolean{admonitiontwoside}
\makeatletter%
\if@twoside%
\setboolean{admonitiontwoside}{true}
\else%
\setboolean{admonitiontwoside}{false}
\fi%
\makeatother%

\newenvironment{env-4232b519-e861-40a4-8532-4fa98c1ba1f6}
{
    \savenotes\tcolorbox[blanker,breakable,left=5pt,borderline west={2pt}{-4pt}{firebrick}]
}
{
    \endtcolorbox\spewnotes
}
                

\newenvironment{env-b81a0bad-19ae-4a38-aa54-7c89ad4aa09e}
{
    \savenotes\tcolorbox[blanker,breakable,left=5pt,borderline west={2pt}{-4pt}{blue}]
}
{
    \endtcolorbox\spewnotes
}
                

\newenvironment{env-794f47f9-38df-4a52-b7f1-227d11939e2b}
{
    \savenotes\tcolorbox[blanker,breakable,left=5pt,borderline west={2pt}{-4pt}{green}]
}
{
    \endtcolorbox\spewnotes
}
                

\newenvironment{env-44715b9d-4e2d-4299-8f00-b7d1037cd4ad}
{
    \savenotes\tcolorbox[blanker,breakable,left=5pt,borderline west={2pt}{-4pt}{aquamarine}]
}
{
    \endtcolorbox\spewnotes
}
                

\newenvironment{env-9a8b2d5a-e80b-4017-bd60-2afb8ee62329}
{
    \savenotes\tcolorbox[blanker,breakable,left=5pt,borderline west={2pt}{-4pt}{orange}]
}
{
    \endtcolorbox\spewnotes
}
                

\newenvironment{env-30d64c79-3a29-45f7-aad4-8833488b1ec1}
{
    \savenotes\tcolorbox[blanker,breakable,left=5pt,borderline west={2pt}{-4pt}{blue}]
}
{
    \endtcolorbox\spewnotes
}
                

\newenvironment{env-f09e27d9-67b8-4931-ab1f-ca6031966775}
{
    \savenotes\tcolorbox[blanker,breakable,left=5pt,borderline west={2pt}{-4pt}{yellow}]
}
{
    \endtcolorbox\spewnotes
}
                

\newenvironment{env-15034df6-36d4-4909-a9e5-98bc9083e742}
{
    \savenotes\tcolorbox[blanker,breakable,left=5pt,borderline west={2pt}{-4pt}{darkred}]
}
{
    \endtcolorbox\spewnotes
}
                

\newenvironment{env-793b6877-d305-4048-810e-53d49b281cd7}
{
    \savenotes\tcolorbox[blanker,breakable,left=5pt,borderline west={2pt}{-4pt}{pink}]
}
{
    \endtcolorbox\spewnotes
}
                

\newenvironment{env-a0fe8824-15a9-47d5-bf01-f54598533eae}
{
    \savenotes\tcolorbox[blanker,breakable,left=5pt,borderline west={2pt}{-4pt}{cyan}]
}
{
    \endtcolorbox\spewnotes
}
                

\newenvironment{env-02a65e52-3e2f-4e42-a1ad-a9c88e6d3852}
{
    \savenotes\tcolorbox[blanker,breakable,left=5pt,borderline west={2pt}{-4pt}{cyan}]
}
{
    \endtcolorbox\spewnotes
}
                

\newenvironment{env-fa538958-d8dd-4e54-9729-545967e84e8a}
{
    \savenotes\tcolorbox[blanker,breakable,left=5pt,borderline west={2pt}{-4pt}{purple}]
}
{
    \endtcolorbox\spewnotes
}
                

\newenvironment{env-07aa4a8b-d1af-4595-9861-b5f1e8010637}
{
    \savenotes\tcolorbox[blanker,breakable,left=5pt,borderline west={2pt}{-4pt}{darksalmon}]
}
{
    \endtcolorbox\spewnotes
}
                

\newenvironment{env-74ec9a83-00fa-4ee6-a586-5d505a5676c9}
{
    \savenotes\tcolorbox[blanker,breakable,left=5pt,borderline west={2pt}{-4pt}{gray}]
}
{
    \endtcolorbox\spewnotes
}
                
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Facility Agreements},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

\title{Facility Agreements}
\author{}
\date{}

\begin{document}
\maketitle

{
\setcounter{tocdepth}{3}
\tableofcontents
}
\hypertarget{raising-finance}{%
\section{Raising Finance}\label{raising-finance}}

\hypertarget{borrowers}{%
\subsection{Borrowers}\label{borrowers}}

3 main categories:

\begin{itemize}
\tightlist
\item
  Governments -- sovereign debt
\item
  Companies
\item
  Others (partnerships, sole traders, supranational organisations,
  educational institutions).
\end{itemize}

4 methods of raising money:

\begin{enumerate}
\tightlist
\item
  Share capital;

  \begin{itemize}
  \tightlist
  \item
    Ordinary shareholders invest money in return for a share of the
    business.
  \item
    Money non-returnable as far as the company is concerned.
  \item
    But profits are usually distributed to shareholders.
  \item
    Gearing = debt/ equity.
  \end{itemize}
\item
  Retained profits;

  \begin{itemize}
  \tightlist
  \item
    Directors decide how to use any profit.
  \item
    Paid to shareholders/ working capital
  \end{itemize}
\item
  Loan facilities; and

  \begin{itemize}
  \tightlist
  \item
    Company borrows with a flexible and reliable source of funds.
  \item
    Overdraft/ term loan etc.
  \end{itemize}
\item
  Capital markets instruments.

  \begin{itemize}
  \tightlist
  \item
    Debt securities
  \item
    Most common: bond

    \begin{itemize}
    \tightlist
    \item
      Issuer repays borrowed capital on maturity
    \item
      Interest paid at intervals
    \end{itemize}
  \item
    Tradeable asset
  \item
    Bonds, notes, commercial paper, certificates of deposit and loan
    stock.
  \end{itemize}
\end{enumerate}

\hypertarget{lenders}{%
\subsection{Lenders}\label{lenders}}

Two main types of commercial lender:

\begin{enumerate}
\tightlist
\item
  Banks

  \begin{itemize}
  \tightlist
  \item
    Raise money primarily from depositors, from issuing debt securities
    and by borrowing from each other.
  \item
    Individuals can lend indirectly through their contributions to
    pension funds and insurance policies.
  \end{itemize}
\item
  Institutional lenders

  \begin{itemize}
  \tightlist
  \item
    Create funds in which investors invests
  \end{itemize}
\end{enumerate}

\hypertarget{banks}{%
\subsection{Banks}\label{banks}}

\begin{longtable}[]{@{}ll@{}}
\toprule()
Type & Details \\
\midrule()
\endhead
Commercial banks & Raise money by attracting deposits and then lend that
money to individuals, companies governments and others. \\
Investment banks & Sell advice and underwrite securities issues. \\
\bottomrule()
\end{longtable}

\hypertarget{glass-steagall}{%
\subsubsection{Glass-Steagall}\label{glass-steagall}}

Banking Act 1933 (Glass-Steagall Act) regulated the investments of many
commercial banks and effectively prohibited commercial banks and
investment banks from combining their businesses.

European banks didn't have Glass-Steagall equivalent so {\(\exists\)}
universal banks. Intended to benefit from economies of scale and
synergies. Capital and customer base of the commercial business fed into
the investment and advisory activities. Markets knew that the government
would not allow the retail arms to fail, benefitting the investment
banking operations.

1999: Glass-Steagall was eventually repealed, and US banks followed suit
with universal banks (Citi, JP Morgan).

UK Independent Commission on Banking (``Vickers Report'') was set up to
consider reforms of the UK banking sector to promote financial
stability. Recommended forcing banks to separate their retail and
investment bank businesses, and was adopted in HM Treasury White Paper.
US authorities adopted the Volcker Rule, which is similar.

\hypertarget{central-banks}{%
\subsubsection{Central Banks}\label{central-banks}}

Bank of England acts as the Government's banker and adviser on monetary
policy. Regulates the money markets by lending or borrowing and setting
interest rates. Lender of last resort for other banks. Linked to UK bank
supervision at the FCA and PRA.

\hypertarget{financial-solicitor}{%
\subsection{Financial Solicitor}\label{financial-solicitor}}

Most city law firms have the following finance teams:

\begin{enumerate}
\tightlist
\item
  General banking
\item
  Capital markets and securitisation
\item
  Project finance
\item
  Asset finance
\item
  Derivatives and regulatory practices.
\end{enumerate}

\hypertarget{due-diligence-to-term-sheet}{%
\section{Due Diligence to Term
Sheet}\label{due-diligence-to-term-sheet}}

\hypertarget{context}{%
\subsection{Context}\label{context}}

There are 3 contexts for commercial bank lending:

\begin{enumerate}
\tightlist
\item
  Investment grade loans/ corporate credit
\item
  Leveraged loans -- debt substantial compared to borrower's earnings/
  share capital. Usually used for acquisition finance.
\item
  Cross-over credit -- borrower borderline investment grade.
\end{enumerate}

The process and terms relating to an investment grade loan are
discussed. The borrower usually approaches the bank, but when the
markets are flush with cash, banks may approach potential borrowers.

Bank must obtain approval from internal credit procedures before a loan
is offered. Once loan approved, the term sheet is drafted.

\hypertarget{due-diligence}{%
\subsection{Due Diligence}\label{due-diligence}}

Fact finding exercise. Scope of investigation depends on risk of
transaction, size of loan, market liquidity, etc. Typical features:

\begin{itemize}
\tightlist
\item
  Account officer

  \begin{itemize}
  \tightlist
  \item
    Individual assigned to each borrower, provides a point of contact.
    Account officer will put together a very basic package for agreed
    loan
  \end{itemize}
\item
  Credit analysis

  \begin{itemize}
  \tightlist
  \item
    Account officer's proposal undergoes credit analysis
  \item
    May be done by a different team/ by the account officer.
  \item
    Internal procedures may stipulate that there must be a material
    adverse change clause or cross-default clause (it is an event of
    default to default on another loan).
  \item
    Private company has 9 months after year-end to publish accounts,
    public company 6 months, London company has 4 months.
  \item
    Smaller borrowers may be asked to produce interim or management
    accounts.
  \end{itemize}
\item
  Credit clearance

  \begin{itemize}
  \tightlist
  \item
    Credit department sanctions the analysis.
  \item
    Internal limits and policies checked.
  \item
    Proposal may be rejected.
  \end{itemize}
\item
  Final term sheet drafted by account officer.
\item
  Borrower may immediately accept bank's draft, or negotiate (requiring
  re-approval by credit department). Bank may involve advisers.
\item
  Solicitors instructed to put together initial draft facility agreement
  on the basis of the term sheet.
\end{itemize}

\begin{env-b81a0bad-19ae-4a38-aa54-7c89ad4aa09e}

Note

A solicitor can obtain copies of a corporate borrower's registered
documents to get information. But if the borrower has not filed returns
propely, this would result in a fine for the company and its
officers.But there is no provision for compensating a 3rd party who
suffers loss.

\end{env-b81a0bad-19ae-4a38-aa54-7c89ad4aa09e}

\hypertarget{acquisition-finance}{%
\subsubsection{Acquisition Finance}\label{acquisition-finance}}

Traditionally \textgreater DD, since more leveraged and risky. But
disclosure could be limited in a liquid market with multiple bidders/
during a hostile takeover.

\hypertarget{term-sheet}{%
\subsection{Term Sheet}\label{term-sheet}}

Term sheet records the principal terms of a transaction. Signed by the
parties, but not intended to be commercially binding. Term sheets are
often attached as an Appendix to a short letter -- the commitment
letter/ mandate letter -- which contains any legally binding terms
required at the outset.

The point of the term sheet is to ``focus minds''. If the facility is to
be syndicated, helpful for the arranging bank to have an accurate
summary of the main terms. Acts as an outline from which the bank's
solicitor can extract key information.

\hypertarget{commitment-letter}{%
\subsection{Commitment Letter}\label{commitment-letter}}

Contains:

\begin{enumerate}
\tightlist
\item
  Whether the bank's obligation to arrange the facility is ``best
  efforts'' or underwritten
\item
  Any general conditionality to the offer to arrange (e.g., time limits,
  subject to credit approval, due diligence).
\item
  Material adverse change provision allowing the bank to withdraw the
  offer if there is a market change (``market mac'') or to the
  borrower's situation (``business mac'')
\item
  Clear market clause: borrower agrees not to raise other finance while
  facility is being arranged
\item
  ``Market flex'' clause; allowing the bank to change aspects if
  necessary to attract other banks to participate.
\item
  Provisions for recovery of fees, costs and expenses if the deal
  doesn't go ahead, and an indemnity for any losses the bank might
  suffer.
\end{enumerate}

\hypertarget{basic-term-sheet}{%
\subsection{Basic Term Sheet}\label{basic-term-sheet}}

A term sheet is not a facility agreement:

\begin{itemize}
\tightlist
\item
  May specify the type of undertakings, but will not include a draft
  clause
\item
  "Guarantees may be given by all material subsidiaries of the
  borrower".
\item
  Use inclusive lists; preface open lists by "including, but not limited
  to".
\item
  Specify unusual provisions.
\end{itemize}

\hypertarget{legal-effect}{%
\subsection{Legal Effect}\label{legal-effect}}

Term sheet not intended to be legally binding. Term sheets have
considerable ``moral authority''. Usually not binding under English law.
An agreement to agree is not legally enforceable if there is a material
term of the future contract which has not been expressly or impliedly
agreed. But usual to use language specifying that parties do not intend
it to be legally binding.

Private equity houses raising finance for an acquisition usually produce
their own term sheet from which to negotiate the loan. These are long,
detailed and favour the borrower.

There is no implied duty of good faith under English law. So either
party may withdraw at any time (Walford v Miles {[}1992{]} 2 WLR 174,
with the \emph{Petromec} exception).

\hypertarget{facility-agreement-options}{%
\section{Facility Agreement Options}\label{facility-agreement-options}}

\hypertarget{types-of-facility}{%
\subsection{Types of Facility}\label{types-of-facility}}

3 types of general purpose loan facility:

\begin{enumerate}
\tightlist
\item
  Overdraft
\item
  Term loan
\item
  Revolving credit facility.
\end{enumerate}

A loan facility may be committed or uncommitted.

\begin{env-f09e27d9-67b8-4931-ab1f-ca6031966775}

Definition

A facility is committed if the facility agreement, once executed,
obliges the bank to advance monies at the borrower's request. If the
bank has some discretion, the facility is uncommitted.

\end{env-f09e27d9-67b8-4931-ab1f-ca6031966775}

\hypertarget{overdraft}{%
\subsubsection{Overdraft}\label{overdraft}}

\begin{itemize}
\tightlist
\item
  A tool to aid cash flow to aid any shortfall in working capital.
\item
  Uncommitted facility
\item
  ``On demand'' {\(\rightarrow\)} must be repaid whenever the bank
  demands, even f the borrower has not defaulted.
\item
  Usually provided on the bank's standard terms, with little room for
  negotiation.
\item
  Subject to a maximum aggregate amount which may be borrowed at any one
  time
\item
  May be subject to a ``clean-down'' provision.
\item
  Relatively expensive way to borrow.
\end{itemize}

\hypertarget{term-loan}{%
\subsubsection{Term Loan}\label{term-loan}}

\begin{itemize}
\tightlist
\item
  Provides a specified capital sum over a set period (term) with an
  agreed repayment schedule.
\item
  Usually a lump sum borrowed for a specific purpose.
\item
  Committed facility
\item
  Borrower allowed availability period/ commitment period after
  executing the facility agreement during which it can draw down a lump
  sum up to a specified amount.
\item
  May be an extended availability period during which the money can be
  drawn down in portions (tranches, advances, loans).
\item
  Schedule of repayment:

  \begin{itemize}
  \tightlist
  \item
    Amortisation - repayment in equal amounts at regular intervals over
    the term of the loan. Repayment ``holiday'' at the start of the
    loan.
  \item
    Balloon repayment -- instalment amount increases towards maturity
  \item
    Bullet repayment -- at maturity.
  \end{itemize}
\item
  Once capital has been repaid, it cannot be re-borrowed.
\item
  Refinancing might be agreed.
\item
  Generally term is {\(\leq 5\)} years
\item
  Not usually ``on demand'' - the bank cannot withdraw the facility
  unless the borrower defaults.
\item
  Might be a multicurrency term loan; drawn amounts may be
  redenominated.
\item
  Interest on a term loan usually floats at a fixed percentage above
  LIBOR/ SONIA.
\end{itemize}

\hypertarget{revolving-credit-facility}{%
\subsubsection{Revolving Credit
Facility}\label{revolving-credit-facility}}

\begin{itemize}
\tightlist
\item
  Provides a maximum aggregate amount of capital available over a
  specified period.
\item
  Allows a borrower to draw down and repay tranches of the available
  capital.
\item
  Combines flexibility of an overdraft with certainty of a term loan.
\item
  Borrower can save money by not drawing the whole loan at once.
\item
  Each tranche usually borrowed for a short period of time (e.g., 1-6
  months) after which it is technically repayable. But if the borrower
  has not defaulted, the tranche to be repaid can immediately re-drawn.
  Known as a ``rollover''.
\item
  With a revolving credit facility, the utilisation and repayment will
  be subject to more restrictions than an overdraft:

  \begin{itemize}
  \tightlist
  \item
    Minimum notice periods before each utilisation
  \item
    Max/ min amounts which may be drawn down in one tranche.
  \item
    Minimum period for which an amount must be borrowed before repaid.
  \item
    Maximum number of tranches which may be drawn at any time.
  \item
    Control over the frequency and timing of repayment.
  \end{itemize}
\item
  Bank will usually charge a ``commitment fee'', calculated as a \% of
  the undrawn facility from time to time.
\item
  Clean down provision may be included

  \begin{itemize}
  \tightlist
  \item
    Ensures the money borrowed under an RCF is used for working capital
    rather than long-term capital.
  \item
    e.g., borrower has to reduce facilities to £0 for 5 consecutive
    business days a year.
  \end{itemize}
\end{itemize}

\hypertarget{ancillary-facilities}{%
\subsubsection{Ancillary Facilities}\label{ancillary-facilities}}

Syndicated facility agreements may include both a term loan and an RCF
in one document.

\hypertarget{syndicated-facilities}{%
\subsection{Syndicated Facilities}\label{syndicated-facilities}}

Large or sophisticated facilities are usually provided by a group of
banks known as a syndicate. Not unusual to find syndicates of
\textgreater30 banks.

\hypertarget{structure}{%
\subsubsection{Structure}\label{structure}}

\begin{itemize}
\tightlist
\item
  Usually, one or more banks act as the ``arranger(s)''.
\item
  This structure is known as primary syndication.
\item
  Syndicate members participate in the syndicate on common terms and
  with common documentation.
\item
  Each syndicate member only liable for its own lending obligations;
  will not undertake any responsibility for other syndicate members'
  commitments.
\end{itemize}

\hypertarget{arranger}{%
\paragraph{Arranger}\label{arranger}}

The arranger/ mandated lead arranger/ MLA is the bank responsible for:

\begin{enumerate}
\tightlist
\item
  advising the borrower as to the type of facility required;
\item
  finding other banks to form the initial syndicate; and
\item
  negotiating the broad terms of the loan.
\end{enumerate}

\hypertarget{secondary-syndication}{%
\paragraph{Secondary Syndication}\label{secondary-syndication}}

Once the facility has been made available, the participating banks might
sell part or all of their commitment to other banks or institutional
lenders -- secondary syndication.

But a borrower will usually require the arranger to retain a substantial
part of the facility, known as the ``minimum hold'', so the arranger has
a stake in organising the facility and ensuring the syndicate runs
smoothly.

\begin{env-44715b9d-4e2d-4299-8f00-b7d1037cd4ad}

Action

The facility agreement should specifically state that the arranger does
not owe any other syndicate members any common law fiduciary duties.

\end{env-44715b9d-4e2d-4299-8f00-b7d1037cd4ad}

\hypertarget{administrative-agent}{%
\paragraph{Administrative Agent}\label{administrative-agent}}

Common for one bank, known as the administrative agent, to act as agent
for the others in the day-to-day running of the loan. Must act as the:

\begin{enumerate}
\tightlist
\item
  Interpreter (communicating between borrower and syndicate)
\item
  Counsellor (dealing with any practical problems)
\item
  Policeman (monitoring borrower's compliance)
\item
  Postman
\item
  Record keeper
\item
  Conduit for payments.
\end{enumerate}

Carries out the duties and responsibilities of an agent. But the
facility agreement will contain provisions to limit the relationship,
known as ``exculpatory provisions'':

\begin{enumerate}
\tightlist
\item
  Agent will not want any obligation to act unless instructed by the
  syndicate
\item
  Agent will want the ability to delegate its functions and to take
  professional advice if necessary
\item
  Agent will want syndicate members to take responsibility for their own
  credit assessment of the borrower.
\item
  Syndicate members to complete their own KYC (Know Your Customer)
  checks to comply with money laundering obligations.
\item
  Agent permitted to take an agency fee and indemnity for costs,
  liabilities and losses, from the borrower
\item
  To the extent that costs, losses and liabilities are not recovered
  from the borrower under (e) above, the agent will want an indemnity
  from the syndicate;
\item
  Agent will always want a right to resign, and the syndicate will want
  the right to replace it. This will entail provisions dealing with
  notice periods and the procedure for appointing a successor (either by
  the agent or by the majority lenders depending on the circumstances).
  Borrower given consultation rights on any replacement to the agent.
\end{enumerate}

\hypertarget{security-trustee}{%
\paragraph{Security Trustee}\label{security-trustee}}

\begin{env-f09e27d9-67b8-4931-ab1f-ca6031966775}

Definition

If a syndicated loan is secured, one entity (usually one of the
syndicate banks) will usually hold the security on trust for the whole
syndicate. This entity is known as the `security trustee'.

\end{env-f09e27d9-67b8-4931-ab1f-ca6031966775}

Fiduciary duties imposed are more extensive than those of a simple
agent.

\hypertarget{taking-a-role}{%
\paragraph{Taking a Role}\label{taking-a-role}}

Several advantages;

\begin{itemize}
\tightlist
\item
  Prestige
\item
  Ability to charge fees
\item
  Relationship with the borrower
\end{itemize}

\hypertarget{tombstones}{%
\subsubsection{Tombstones}\label{tombstones}}

Once a large deal is closed, announced by way of an advert in the
financial press with details. Cast in Perspex block and given to parties
and advisers. Now less common.

\hypertarget{liability-of-arrangers}{%
\subsubsection{Liability of Arrangers}\label{liability-of-arrangers}}

Arranger will provide information about the borrower in an ``information
memorandum'' (info memo). Usually contains financial information,
general information and market information.

The arranger is potentially liable for any wrong information the info
memo contains which is relied on by the lenders deciding to joint the
syndicate. Claims may be brought for misrepresentation or negligence.
There is also a tortious duty of care (Sumiomo Bnk Ltd v Banque
Bruxelles Lambers SA, 1996).

So it is standard for a clause in the info memo to state:

\begin{itemize}
\tightlist
\item
  the borrower has supplied the information and confirms it is true,
  complete and accurate; has been prepared with reasonable care; is not
  misleading; and does not omit anything material;
\item
  the borrower confirms that any assumptions or projections it makes are
  reasonable and in good faith;
\item
  the arranger has not independently verified the information; does not
  make any representations or undertakings with respect to it; does not
  take responsibility for it; and accepts no liability for loss caused
  by reliance on it; and
\item
  participants in the facility acknowledge the info memo is not a
  valuation of the borrower's credit; that it is not a recommendation to
  participate; and that they should make their own independent
  assessment of the documentation and the borrower.
\end{itemize}

Effectiveness of such disclaimers:

\begin{itemize}
\tightlist
\item
  Courts tend to interpret limitation clauses strictly.
\item
  UCTA 1977\textasciitilde{} disclaimer of liability for negligence or
  misrepresentation must be reasonable.
\item
  Arranger is almost certainly under duties anyway not to withhold
  relevant information and to answer fully any specific questions.
\item
  In IFE Fund SA v Goldman Sachs International {[}2007{]} EWCA Civ 811,
  Court of Appeal ruled that disclaimer language commonly used in
  information memoranda is effective to deny an arranger's liability for
  the accuracy of information and to deny responsibility for updating
  it.
\end{itemize}

\hypertarget{leveraged-acquisition-facilities}{%
\subsection{Leveraged Acquisition
Facilities}\label{leveraged-acquisition-facilities}}

Usually involves more than one source of lending to raise the maximum
amount of debt. Consists of a senior facility providing a term loan and
revolving facility ("revolver") and a mezzanine facility providing a
second term loan. Senior lenders get first right to repayment and
control the enforcement process on default. Mezzanine lenders get a
higher interest rate. But mezzanine facilities rare recently - there is
usually a senior facility and a high yield bond.

\hypertarget{short-term-lending}{%
\subsection{Short-term Lending}\label{short-term-lending}}

Other forms:

\begin{itemize}
\tightlist
\item
  Standby facility -- backup facility, may include a utilisation fee.
\item
  Bankers' acceptance -- bill of exchange which the bank has agreed to
  be responsible for paying which it falls due. Borrower must eventually
  pay back the bank.
\item
  Swing-line facility -- made available to a company which has a
  commercial paper programme (IOU). Can be used on very short notice to
  allow the borrower to pay back any maturing commercial paper which it
  is unable to reissue (rollover).
\end{itemize}

\hypertarget{anatomy-of-facility-agreement}{%
\subsection{Anatomy of Facility
Agreement}\label{anatomy-of-facility-agreement}}

Provisions in a loan document almost all designed to protect the bank's
position.

\hypertarget{bank-concerns}{%
\subsubsection{Bank Concerns}\label{bank-concerns}}

Will want to ensure:

\begin{enumerate}
\tightlist
\item
  Fees are paid
\item
  Interest received on the loan
\item
  Loan capital returned.
\end{enumerate}

Aims best served by ensuring that:

\begin{itemize}
\tightlist
\item
  Loan monies only used for a specific purpose
\item
  Borrower doesn't do anything financially risky
\item
  Borrower monitored to give early warning of any possibility of default
\item
  All monies due to the bank secured over the borrower's assets.
\end{itemize}

4 main types of clause:

\begin{enumerate}
\tightlist
\item
  Mechanics clauses
\item
  Margin protection clauses
\item
  General protection clauses
\item
  Boilerplate clauses.
\end{enumerate}

\hypertarget{borrowers-concerns}{%
\paragraph{Borrower's Concerns}\label{borrowers-concerns}}

Primary concern: flexibility. Main areas to negotiate relate to the
monitoring of and restrictions on the borrower.

\hypertarget{matched-funding}{%
\subsection{Matched Funding}\label{matched-funding}}

Most facility agreements assume that when the borrower requests a
utilisation, the facility banks will fund their commitment by borrowing
the sum from other banks on the London interbank market. Loans on
interbank market typically short term (\textasciitilde6 months).
Interest payable on successive interbank loans recovered from the
borrower.

This assumption drives lots of provisions, such as interest periods,
repetition of representations, repayment, prepayment and break costs.

\hypertarget{representations}{%
\section{Representations}\label{representations}}

\hypertarget{facility-agreement}{%
\subsection{Facility Agreement}\label{facility-agreement}}

Through the due diligence process, the borrower will provide facts,
records, forecasts, business plans, documents of title etc. Most
important pieces of information are recorded in the facility agreement
as representations. Reasons:

\begin{enumerate}
\tightlist
\item
  To elicit information
\item
  To provide a drawstop if they are untrue
\item
  To allow the bank to call a default if they are untrue.
\end{enumerate}

\hypertarget{legal-representations}{%
\subsubsection{Legal Representations}\label{legal-representations}}

Relate to the validity of the borrower, any guarantor and the loan
documentation. All 6 are usually repeated.

\begin{itemize}
\tightlist
\item
  Status

  \begin{itemize}
  \tightlist
  \item
    Borrower required to state it is duly incorporated under the laws of
    the original jurisdiction and the representation repeated to ensure
    the jurisdiction has not changed.
  \item
    Borrower represents that it has the power to own its assets and
    carry on its business.
  \end{itemize}
\item
  Binding obligations

  \begin{itemize}
  \tightlist
  \item
    Borrower confirms that obligations it undertakes in the facility
    agreement and associated documents are legal, valid, binding, and
    enforceable. Common to qualify this representation by general
    principles limiting the borrower's obligations.
  \end{itemize}
\item
  Non-conflict

  \begin{itemize}
  \tightlist
  \item
    Borrower must state that entering into the facility agreement will
    not conflict with any laws or regulations, their constitutional
    documents, or any other agreement or instrument to which they are a
    party.

    \begin{itemize}
    \tightlist
    \item
      Practically, the bank does not want the borrower to immediately
      default on another contract.
    \item
      Legally, if entering the loan causes the borrower to breach an
      existing contract, the other party to that contract might sue the
      bank in tort for `procuring a breach of contractual relations' (or
      `interference with contractual relations').
    \item
      So the bank should ensure this does not just protect material
      breaches.
    \end{itemize}
  \end{itemize}
\item
  Power and authority

  \begin{itemize}
  \tightlist
  \item
    Borrower represents that it has the power and authority to enter
    into the loan and perform its obligations. Breach will lead to
    default, which allows the bank to refuse advances or demand
    repayment.
  \end{itemize}
\item
  Validity and admissibility in evidence

  \begin{itemize}
  \tightlist
  \item
    Borrower represents that it has done everything required to ensure
    the facility agreement is admissible in evidence in its jurisdiction
    of incorporation.
  \end{itemize}
\item
  Governing law and enforcement.

  \begin{itemize}
  \tightlist
  \item
    Representation that the law governing the facility agreement and any
    judgment obtained will be recognised and enforced in the borrower's
    jurisdiction of incorporation.
  \end{itemize}
\end{itemize}

\hypertarget{commercial-representations}{%
\subsubsection{Commercial
Representations}\label{commercial-representations}}

\begin{itemize}
\tightlist
\item
  Deduction of tax

  \begin{itemize}
  \tightlist
  \item
    Borrower represents that it is not required to deduct tax from
    payments made to banks define as ``Qualifying Lenders'' (those
    exempt from withholding tax). Not repeated: ongoing tax issues dealt
    with separately.
  \end{itemize}
\item
  No filing or stamp duties

  \begin{itemize}
  \tightlist
  \item
    Stamp duty is imposed on the transfer of `stock', `marketable
    securities' and `interests in partnership', with rates ranging from
    0.5\% to 1.5\% of the consideration or market value of the property
    being transferred (known as \emph{ad valorem} stamp duty).
  \item
    Failure to pay stamp duty within 30 days of execution will usually
    mean that the document is not admissible in evidence or accepted for
    registration purposes.
  \item
    Not usually payable on a loan facility, but represented to protect
    against possible foreign duties.
  \end{itemize}
\item
  No event of default

  \begin{itemize}
  \tightlist
  \item
    Borrower represents that there is no \textbf{event of default}
    continuing, or which might reasonably be expected to occur as a
    result of utilising the facility.
  \item
    Borrower represents that there is no \textbf{event of default} under
    any other of its agreements which would have a material adverse
    effect.

    \begin{itemize}
    \tightlist
    \item
      Negotiation will focus on narrowing the definition of material
      adverse effect.
    \item
      Repeated representation.
    \end{itemize}
  \end{itemize}
\item
  No misleading information

  \begin{itemize}
  \tightlist
  \item
    Borrower represents that it has not supplied information to the bank
    which is misleading.

    \begin{itemize}
    \tightlist
    \item
      Information provided for the information memorandum true and
      accurate
    \item
      Financial projections prepared by the company's directors based on
      recent information and contain reasonable assumptions
    \item
      All other information provided to banks was true, complete, and
      accurate.
    \end{itemize}
  \item
    Borrower will want the representation to apply only to written
    information provided.
  \item
    Borrower will want representation qualified to information that was
    ``true and accurate in all material respects''.
  \item
    Not necessary to repeat.
  \end{itemize}
\item
  Accounting principles

  \begin{itemize}
  \tightlist
  \item
    Borrower represents that the original financial statements it
    provided to the bank were prepared in accordance with the relevant
    GAAP, consistently applied, and give a `true and fair view'.
  \item
    Borrower undertakes to deliver future accounts on the same basis,
    with a proviso to deal with any change in policies.
  \item
    Borrower represents that there has been no material adverse change
    in its business or financial condition since a set of financial
    statements.

    \begin{itemize}
    \tightlist
    \item
      Usually not repeated if the facility has a material adverse change
      event of default.
    \item
      If repeated, does it refer to a material adverse change from the
      original financial statements or the latest financial statements?
      Negotiate.
    \end{itemize}
  \end{itemize}
\item
  Ranking

  \begin{itemize}
  \tightlist
  \item
    Borrower represents that any claims the bank may have under the loan
    documentation will rank at least \emph{pari passu} (equally) with
    any other unsecured creditors, other than as a result of any
    preference through insolvency regulations.
  \item
    Under English law, this is the default legal position. Protects
    banks against a change in the law and protects against foreign
    jurisdiction practices.
  \end{itemize}
\item
  Litigation

  \begin{itemize}
  \tightlist
  \item
    Borrower represents that it faces no current or threatened
    litigation -- though this is likely to be qualified by materiality/
    de minimis/ excluding frivolous and vexatious claims or to refer to
    the ``reasonably likely outcome''.
  \item
    Litigation representations likely to include claims against the
    borrower group's directors (derivative claims).
  \end{itemize}
\item
  Winding-up proceedings

  \begin{itemize}
  \tightlist
  \item
    Borrower represents that no insolvency proceedings in any form have
    been started or are threatened against the borrower.
  \item
    Borrower should press for de minimis provision and wording to
    exclude debts which it disputes in good faith.
  \item
    Usually repeated
  \end{itemize}
\item
  Encumbrances

  \begin{itemize}
  \tightlist
  \item
    Borrower represents that there are no encumbrances other than those
    disclosed before execution.

    \begin{itemize}
    \tightlist
    \item
      Even though the bank will have done a company search on the
      borrower and examined its register of mortgages and charges.
    \end{itemize}
  \item
    Bank represents that it has valid title in its own name to all
    property over which security is granted.
  \item
    Not usually repeated
  \end{itemize}
\item
  Environmental

  \begin{itemize}
  \tightlist
  \item
    If the borrower breaches environmental laws, its lender may also
    incur direct liability (`lender liability') if it is found to have
    `knowingly permitted' contamination or if it takes possession of
    contaminated land.
  \item
    Borrower represents that it is in compliance with all environmental
    laws and has obtained and complies with all environmental licences
    necessary for its business.
  \item
    Often subject to materiality and not typically repeated
    (environmental undertaking used instead).
  \end{itemize}
\item
  Catch all

  \begin{itemize}
  \tightlist
  \item
    Banks may try to include such a clause; borrower typically resists.
  \end{itemize}
\end{itemize}

\hypertarget{disclosure-letter}{%
\subsubsection{Disclosure Letter}\label{disclosure-letter}}

\begin{itemize}
\tightlist
\item
  The bank's solicitor will draft the representations as widely as
  possible.
\item
  Borrower forced to disclose instances in which the representation is
  not true.

  \begin{itemize}
  \tightlist
  \item
    These could be included as a specific exception.
  \item
    Neater to keep the representation ``clean'' and use a separate
    document to deal with anything which would otherwise be a breach --
    the disclosure letter. List of agreed specific items which the bank
    has agreed to waive.
  \end{itemize}
\item
  Disclosures should be worded to modify all representations.
\end{itemize}

\hypertarget{repetition-of-representations}{%
\subsubsection{Repetition of
Representations}\label{repetition-of-representations}}

Representations may be ``evergreen'' - deemed repeated every day
throughout the life of the loan. More common for representations to be
repeated immediately prior to each utilisation and on the first day of
each interest period.

To prevent the borrower arguing that the representations were merely
initial pre-conditions of lending:

\begin{itemize}
\tightlist
\item
  Loan document may specifically state that the representations are to
  survive the execution and first utilisation.
\item
  A certificate by the borrower to be periodically provided confirming
  that the representations are still correct.
\end{itemize}

This also helps prove reliance on a future misrepresentation claim.
Reliance will be harder to show from a deemed representation.

\hypertarget{banks-perspective}{%
\subsection{Bank's Perspective}\label{banks-perspective}}

Some representations are best cast in general terms due to caveat
emptor. Representations are about risk allocation, allowing the bank to
take action. So don't give in to blanket qualifications for most terms.

\hypertarget{borrowers-perspective}{%
\subsection{Borrower's Perspective}\label{borrowers-perspective}}

Borrower will want to negotiate wide margins of error.

\begin{env-9a8b2d5a-e80b-4017-bd60-2afb8ee62329}

Warning

\begin{itemize}
\tightlist
\item
  A borrower should not repeat a representation that there is no
  "potential event of default", since breach would turn a potential into
  a full default.
\item
  The disclosure letter is effective only on the first instance. If a
  representation is repeated, the disclosure letter is no longer
  effective.
\end{itemize}

\end{env-9a8b2d5a-e80b-4017-bd60-2afb8ee62329}

The bank will strongly resist wording that representations are qualified
with ``to the best of our knowledge and belief'', since this defeats the
point.

Borrowers may want to negotiate a materiality qualification, should
ensure that the qualification is certain enough to actually be used
(e.g., a \emph{de minimus} amount of money). This clause can also be
used to restrict the companies in the borrower's group to which a
representation applies (e.g., material subsidiaries).

\hypertarget{contract-law}{%
\subsection{Contract Law}\label{contract-law}}

A representation is a statement of fact, made to induce the party to
whom it is addressed to enter into a contract. The statement may be oral
or written, and if it is drafted into the contract, it also becomes a
term of that contract. If the statement is untrue, it constitutes
misrepresentation.

\begin{longtable}[]{@{}ll@{}}
\toprule()
Term & Definition \\
\midrule()
\endhead
Representation & A statement asserting the truth of a given state of
facts \\
Representor & The party who allegedly made the representation \\
Representee & The party who allegedly received the representation \\
Actionable representation & An unambiguous false statement of fact made
to the claimant and which induces the claimant to enter into the
contract with the statement maker \\
\bottomrule()
\end{longtable}

\hypertarget{misrepresentation-sequence}{%
\subsubsection{Misrepresentation
Sequence}\label{misrepresentation-sequence}}

\begin{enumerate}
\tightlist
\item
  Identify potential misrepresentation (common sense)
\item
  Identify if the potential misrepresentation is an Actionable
  representation
\item
  Categorise into one of the Misrepresentation categories
\item
  Consider which of the Misrepresentation remedies to apply
\item
  Consider if any Misrepresentation alternative claims apply
\end{enumerate}

\begin{longtable}[]{@{}llllll@{}}
\toprule()
Category & Legal basis & Test & Rescission? & Indemnity? & Damages \\
\midrule()
\endhead
Fraudulent & Tort of deceit & Representee to prove that representor made
the statement knowing it was untrue/ without believing it to be true/
reckless as to the truth & Yes & Unlikely to be needed & Yes, generous
Tort Law/Remoteness rules, no reduction for contributory negligence \\
Negligent & s 2(1) MA 1967 & Representor fails to show reasonable
grounds to believe true/ not believing true up until contract made & Yes
& Unlikely to be needed & As for fraud, but potential reduction for
contributory negligence. Can also be given in lieu of rescission. \\
Innocent & s 2(1) MA 1967 & Representor has shown reasonable grounds to
believe, and actual belief up until contract made & Yes & Yes, as part
of rescission & Can be given in lieu of rescission. \\
\bottomrule()
\end{longtable}

\hypertarget{actionable-representation}{%
\subsubsection{Actionable
Representation}\label{actionable-representation}}

Examining each component in more detail:

\hypertarget{unambiguous}{%
\subsubsection{Unambiguous}\label{unambiguous}}

The representation must be clear and will only form the basis of a claim
in misrepresentation if it unambiguously has the meaning put forward by
the representee.

Representor will not be liable if the representee has placed its own
unreasonable construction on the representation (McInerny v Lloyd's Bank
Ltd {[}1974{]} 1 Lloyd's Rep 246).

\hypertarget{false}{%
\subsubsection{False}\label{false}}

The statement must be false; it will not be false if it is substantially
correct

\begin{quote}
``{[}A{]} representation may be true without being entirely correct,
provided it is substantially correct and the difference between what is
represented and what is actually correct would not have been likely to
induce a reasonable person in the position of the claimants to enter
into the contracts''\\
Rix J, Avon Insurance Plc v Swire Fraser Ltd {[}2000{]} 1 All ER (Comm)
573.
\end{quote}

\hypertarget{statement-of-fact}{%
\subsubsection{Statement of Fact}\label{statement-of-fact}}

\begin{quote}
'A representation is not an undertaking to do, or not to do, something.
It is a statement asserting a given state of affairs'\\
Kleinwort Benson Ltd v Malaysia Mining Corp {[}1989{]} 1 WLR 379
\end{quote}

\begin{longtable}[]{@{}ll@{}}
\toprule()
Includes & Does not include \\
\midrule()
\endhead
Statements of fact & Statements of opinion \\
Some conduct & Statements of intention \\
Statements of law & Silence \\
\bottomrule()
\end{longtable}

\hypertarget{facts-vs-puff}{%
\paragraph{Facts Vs Puff}\label{facts-vs-puff}}

Representations distinguished from mere 'puff' (Dimmock v Hallett (1866)
LR 2 Ch App 21).

\hypertarget{conduct}{%
\paragraph{Conduct}\label{conduct}}

Statements usually made by words, but can also be made by conduct
(Gordon v Selico (1986) 278 EG 53 concealment of dry rot).

\hypertarget{statement-of-law}{%
\paragraph{Statement of Law}\label{statement-of-law}}

Traditional rule was that a statement of law could not give rise to an
actionable misrepresentation. But the distinction between statements of
fact and law has been abolished. It was held that a false statement as
to the existence of an Act of Parliament is a misrepresentation of fact
(West London Commercial Bank v Kitson (1884) LR 13 QBD 360).

\hypertarget{statement-of-opinion}{%
\paragraph{Statement of Opinion}\label{statement-of-opinion}}

\hypertarget{general-rule}{%
\subparagraph{General Rule}\label{general-rule}}

A statement of opinion is not a statement of fact, so cannot usually
form the basis of a claim in misrepresentation (e.g., Bisset v Wilkinson
{[}1927{]} AC 1977).

\hypertarget{exceptions}{%
\subparagraph{Exceptions}\label{exceptions}}

When someone expresses an opinion, they impliedly state they know the
facts justifying that opinion. When the same facts are known to each
party, the implicit statement is unlikely to have much impact on the
representee. But if the representor has greater knowledge than the
representee, the implied statement that there are facts justifying the
opinion can significantly mislead the representee (Smith v Land and
House Property Corporation (1885) LR 28 Ch D 7 (CA)).

Where the representor is in a position of superior knowledge or
experience, a statement of opinion can be held to involve a statement of
fact that there are reasonable grounds for their opinion. So if there
are no reasonable grounds for this opinion (i.e., the opinion is one
which someone with the knowledge of the representor could not reasonably
have held) then a false statement has been made (Esso v Mardon
{[}1976{]} QB 801).

There is also an implied statement in an expression of opinion that the
representor believes the opinion that they express. If this is not the
case, such a representation will be a misrepresentation.

\hypertarget{statement-of-future-intention}{%
\paragraph{Statement of Future
Intention}\label{statement-of-future-intention}}

Representation is a statement of fact, so can have no reference to
future events or promises.

\begin{quote}
``There is a clear difference between a representation of fact and a
representation that something will be done in the future. A
representation that something will be done in the future cannot be true
or false at the moment it is made; and although you may call it a
representation, if anything it is a contract or promise''\\
Beattie v Ebury (1872) LR 7 Ch App 777, per Mellish LJ.
\end{quote}

Consequently, it is not a misrepresentation if the representor makes a
promise regarding a future intention, but is prevented from doing so or
if circumstances alter so that they change their mind about that
intention (Wales v Wadham {[}1977{]} 1 WLR 199).

However, if the representor states they intend to do something, they are
making a limited statement of fact: stating they do have that intention.
So if, at that time, they know they cannot do what they state, or do not
intend to do it, they are misrepresenting their existing intention
(Edgington v Fitzmaurice (1885) 29 Ch D 459).

Bowen LJ:

\begin{quote}
``The state of a man's mind is as much a fact as the state of his
digestion\ldots''
\end{quote}

\hypertarget{silence}{%
\paragraph{Silence}\label{silence}}

Silence will generally not give rise to an action for misrepresentation.
No general duty to disclose facts which, if known, might affect the
other party's decision to enter the contract (Keates v The Earl of
Cadogan (1851) 10 CB 591).

But there are some exceptions to the rule in Keates:

\hypertarget{half-truths}{%
\subparagraph{Half-truths}\label{half-truths}}

It is a misrepresentation to make statements which are technically true
but misleading (Nottingham Patent Brick \& Tile Co v Butler (1866)).
e.g., to describe property which is the subject of negotiations for sale
as 'fully let' without disclosing that, although the property is indeed
fully let at that time, the tenants have given notice to quit, is a
misrepresentation (Dimmock v Hallett (1866) LR 2 Ch App 21).

\hypertarget{continuing-representation}{%
\subparagraph{Continuing
Representation}\label{continuing-representation}}

If, at the beginning of negotiations, a statement is made which is true
but which prior to entering into the contract becomes false, the
representor is under an obligation to correct the representation. If
they fail to do so and allow the other party to enter into the contract
still believing that the representation is true, then they will be
liable for misrepresentation (With v O'Flanagan {[}1936{]} Ch 575). This
is the principle of 'continuing representations'.

\hypertarget{contracts-uberrimae-fidei-utmost-good-faith}{%
\subparagraph{Contracts Uberrimae Fidei (utmost Good
faith)}\label{contracts-uberrimae-fidei-utmost-good-faith}}

There is a duty to disclose material facts in some types of contracts in
which one party is in a particularly strong position to know the
material facts which form the basis of the contract. Most common is
contract of insurance: at common law, disclosure of all material facts
must be made to the insurer. Also, often this obligation to disclose
information exists between parties which are in a fiduciary relationship
(e.g., between a company and its directors, between a trustee and
beneficiaries of a trust).

\hypertarget{addressed-to-the-claimant}{%
\subsubsection{Addressed to the
Claimant}\label{addressed-to-the-claimant}}

The misrepresentation must be addressed by the representor to the
claimant.

\hypertarget{induces-the-claimant-to-enter-into-the-contract-with-statement-maker}{%
\subsubsection{Induces the Claimant to Enter into the Contract with
Statement
Maker}\label{induces-the-claimant-to-enter-into-the-contract-with-statement-maker}}

Requirement not satisfied in JEB Fasteners v Mark Bloom {[}1983{]} 1 All
ER 583, where the Court of Appeal held that the defendants'
representation did not play a 'real and substantial' part in inducing
the claimants to act.

Court asks:

\begin{quote}
Was the representation material?
\end{quote}

Objective test for this: did the statement relate to an issue that would
have influenced a reasonable person (Pan Atlantic Co Ltd v Pine Top
Insurance Co Ltd {[}1995{]} 1 AC 501)?

\begin{itemize}
\tightlist
\item
  If a statement is found to be material, then inducement will generally
  be inferred as a matter of fact (Smith v Chadwick (1884) 9 App Cas
  187). Then burden shift to the defendant to rebut the inference that
  the claimant was induced. This is done by proving that the claimant
  was not subjectively induced.
\item
  If a statement is not found to be material, then inducement of the
  claimant cannot be inferred as a matter of fact. The claimant must
  prove that they were subjectively induced. If they do this
  successfully, it will be held that they have been induced by the
  misrepresentation (Museprime Properties Ltd v Adhill Properties Ltd
  (1990) 61 P. \& C.R. 111).
\end{itemize}

MERMAID1

Since the representation must have induced the representee to enter the
contract, there is no actionable misrepresentation where:

\begin{itemize}
\tightlist
\item
  the statement was not actually communicated to the representee;
\item
  the statement did not affect the representee's decision to enter the
  contract
\item
  the statement was known to be untrue by the representee.
\end{itemize}

The misrepresentation need not be the only reason the claimant entered
the contract (Edgington v Fitzmaurice (1885) 29 Ch D 459).

A representor may seek to argue that the representee was not induced
where the representee chooses to test the validity of the representor's
statement by making its own investigations (Attwood v Small (1838) 6 CL
\& F 232).

The key point in Attwood is not that separate enquiries were made or
could have been made --- the crucial point is that the separate
enquiries showed that the vendor's statements were not relied upon
(Redgrave v Hurd (1881) 20 Ch D 1). This case also established that
there is no general duty to check the misrepresentor's statement.

If a representee does not check, where the court considers it reasonable
for them to have done so, or carries out a negligent investigation, this
would open up the possibility of a defence of Contributory negligence
against the representee. This cannot be pleaded where the
misrepresentation is fraudulent. The more commercial the representee is,
the more likely it is that the court will consider it reasonable for the
representee to have investigated (by analogy to Smith v Eric Bush
{[}1990{]} 1 AC 831).

\hypertarget{categories-of-misrepresentation}{%
\subsubsection{Categories of
Misrepresentation}\label{categories-of-misrepresentation}}

Once the elements of an actionable misrepresentation have been
established, the remedies available depend on the category of the
misrepresentation in question.

Under
\href{https://www.legislation.gov.uk/ukpga/1967/7}{Misrepresentation Act
1967}, the categories of misrepresentation available are:

\begin{longtable}[]{@{}lll@{}}
\toprule()
Category & Legal basis & Test \\
\midrule()
\endhead
Fraudulent & Tort of deceit & Representee to prove that representor made
the statement knowing it was untrue/ without believing it to be true/
reckless as to the truth \\
Negligent & s 2(1) MA 1967 & Representor fails to show reasonable
grounds to believe true/ not believing true up until contract made \\
Innocent & s 2(1) MA 1967 & Representor has shown reasonable grounds to
believe, and actual belief up until contract made \\
\bottomrule()
\end{longtable}

\hypertarget{fraudulent-misrepresentation}{%
\paragraph{Fraudulent
Misrepresentation}\label{fraudulent-misrepresentation}}

Defined by Lord Herschell in Derry v Peek (1889) 14 App Cas 337:

\begin{quote}
"First, in order to sustain an action of deceit, there must be proof of
fraud, and nothing short of that will suffice. Secondly, fraud is proved
when it is shown that a false representation has been made (a)
knowingly, or (b) without belief in its truth, or (c) recklessly,
careless whether it be true or false. Although I treated the second and
third as distinct cases, I think the third is but an instance of the
second, for one who makes a statement under such circumstances can have
no real belief in the truth of what he states\ldots Thirdly, if fraud be
proved, the motive of the person guilty of it is immaterial. It matters
not that there was no intention to cheat or injure the person to whom
the statement was made."
\end{quote}

'Recklessness' was defined in Thomas Witter Ltd v TBP Industries Ltd
{[}1996{]} 2 All ER 573, where it was established that to find fraud,
the level of recklessness required was 'flagrant disregard for the
truth'.

\hypertarget{change-in-circumstances}{%
\subparagraph{Change in Circumstances}\label{change-in-circumstances}}

It can be difficult categorising a misrepresentation based on change of
circumstances. Lord Wright MR in With v O'Flanagan {[}1936{]} Ch 575:

\begin{quote}
``\ldots the Court is more reluctant to use the word ``fraud'' and would
not generally use the word ``fraud'' in that connection because the
failure to disclose, though wrong and a breach of duty, may be due to
inadvertence or a failure to realise that the duty rests upon the party
who has made the representation not to leave the other party under an
error when the representation has become falsified by a change of
circumstances''
\end{quote}

So the Court would need to be satisfied that failure to disclose was
deliberate/dishonest.

\hypertarget{burden-of-proof}{%
\subparagraph{Burden of Proof}\label{burden-of-proof}}

The burden of proof on the claimant to prove fraudulent representation
is a difficult burden to discharge: allegations of fraud taken very
seriously.

\hypertarget{negligent-misrepresentation}{%
\paragraph{Negligent
Misrepresentation}\label{negligent-misrepresentation}}

Section 2(1) of the MA 1967 provides:

\begin{quote}
``Where a person has entered into a contract after a misrepresentation
has been made to him by another party thereto and as a result thereof he
has suffered loss, then, if the person making the misrepresentation
would be liable to damages in respect thereof had the misrepresentation
been made fraudulently, that person shall be so liable notwithstanding
that the misrepresentation was not made fraudulently, unless he proves
that he had reasonable grounds to believe and did believe up to the time
the contract was made that the facts represented were true.''
\end{quote}

So defendant liable for negligent misrepresentation unless they can
prove they had reasonable grounds to believe, and did believe up until
the time of the contract, that the statement was true.

Where a representor is found liable for a negligent misrepresentation,
they will be treated to all intents and purposes as if he had made a
fraudulent misrepresentation (Royscot Trust Ltd v Rogerson {[}1991{]} 2
QB 297 but widely criticised). This has ramifications for damages.

See also Howard Marine and Dredging Co. Ltd. v A. Ogden \& Sons
(Excavations) Ltd. {[}1978{]} 2 WLR 515.

Important to note that the statutory right to damages for negligent
misrepresentation reverses the normal burden of proof by requiring the
representor (defendant) to prove that they had reasonable grounds to
believe their statement, and did believe their statement.

Note that it is difficult for a defendant to get their misrepresentation
downgraded from negligent to innocent, because the courts want to
discourage people from making statements without checking/ thinking
about their validity.

\hypertarget{innocent-misrepresentation}{%
\paragraph{Innocent
Misrepresentation}\label{innocent-misrepresentation}}

An innocent misrepresentation is one which is not made fraudulently or
negligently. It is a statement made where the representor:

\begin{itemize}
\tightlist
\item
  Proves they had reasonable grounds for belief in the truth of their
  statement
\item
  Proves that they believed up to the time of contract that what they
  were saying is true.
\end{itemize}

It is defined in s
2(1)\href{https://www.legislation.gov.uk/ukpga/1967/7}{Misrepresentation
Act 1967}.

\hypertarget{remedies-for-misrepresentation}{%
\subsubsection{Remedies for
Misrepresentation}\label{remedies-for-misrepresentation}}

\hypertarget{rescission}{%
\paragraph{Rescission}\label{rescission}}

Misrepresentation renders the contract voidable. So the contract is
still valid and subsisting until the representee decides to rescind it.
Available for any type of representation.

General rule: in order to rescind, the representee must communicate the
intention to do so to the representor. Or may initiate proceedings for
the rescission of the contract.

This is an equitable remedy, so given or withheld at the discretion of
the court. Generally only awarded where parties can be restored to their
original position. Reasons why potentially not awarded:

\begin{itemize}
\tightlist
\item
  Affirmation

  \begin{itemize}
  \tightlist
  \item
    Representee declaring intention to affirm, or doing some act from
    which such intention may reasonably be inferred.
  \end{itemize}
\item
  Lapse of time

  \begin{itemize}
  \tightlist
  \item
    Delay defeats equity
  \item
    Lapse of time without attempt to effect rescission not itself
    affirmation, but could be treated as evidence for such an intention
  \end{itemize}
\item
  Restitution impossible

  \begin{itemize}
  \tightlist
  \item
    \emph{restitutio in integrum} no longer possible (Clarke v Dickson
    (1858) EB \& E 148)
  \end{itemize}
\item
  Third party rights accrue

  \begin{itemize}
  \tightlist
  \item
    The contract remains valid up until the time notice is given of the
    intention to rescind. So at any time prior to rescission, a person
    acquiring goods under such a contract is able to pass good title to
    those goods to an innocent third party, who purchases the goods
    without notice of the misrepresentation.
  \item
    This would prevent restitution
  \end{itemize}
\end{itemize}

\hypertarget{indemnity}{%
\paragraph{Indemnity}\label{indemnity}}

As part of the equitable process of rescission, an indemnity may be
awarded to cover expenses for obligations assumed as a direct result of
the contract. The obligations must have been created by the contract.

Generally, where an action for misrepresentation will give a right to
damages, an indemnity will not be awarded. No damages are available for
an innocent misrepresentation, so here an action of indemnity is more
likely to be awarded. Note the distinction between an indemnity and the
common law right to damages (Whittington v Seale‑Hayne (1900) 82 LT 49).

\hypertarget{damages}{%
\paragraph{Damages}\label{damages}}

Damages are potentially the greatest where the misrepresentation is
fraudulent. But damages for an action for negligent misrepresentation
will under s 2(1) MA 1967 often match those available for fraudulent
misrepresentation. Damages in lieu of rescission may be available under
s 2(2) MA 1967.

Innocent misrepresentation does not afford any damages per se, but
damages in lieu of rescission may be available under s 2(2) MA 1967.

Negligent misrepresentation and fraudulent misrepresentation both
entitle the innocent party to damages, under the Misrepresentation Act
1967, s 2(1) and the tort of deceit respectively. In both cases,
however, any damages will be on a tortious basis and, therefore, a bank
would get no compensation for loss of bargain or profit (see Smith New
Court Securities Ltd v Scrimgeour Vickers (Asset Management) Ltd
{[}1997{]} AC 254, HL).

\hypertarget{damages-for-fraudulent-misrepresentation}{%
\paragraph{Damages for Fraudulent
Misrepresentation}\label{damages-for-fraudulent-misrepresentation}}

Fraudulent misrepresentation is part of the tort of deceit, so damages
will be calculated on a tortious basis. The measure of damages is that
which is necessary to place the innocent injured party in the position
they would have been in had the misrepresentation not been made.

Held in Doyle v Olby (Ironmongers) Ltd {[}1969{]} 2 QB 158 that claimant
can recover

\begin{quote}
``\ldots all the damage directly flowing from the tortious act of
fraudulent inducement which was not rendered too remote by the
plaintiff's own conduct, whether or not the defendant could have
foreseen the loss.''
\end{quote}

Requirements have since been added:

\begin{itemize}
\tightlist
\item
  That the claimant must mitigate as soon as the fraud is discovered
\item
  That any damages awarded to the claimant will be reduced by the value
  of any benefit the claimant has acquired as a result of the contract
\end{itemize}

Contributory negligence is \textbf{not} available as a defence to a
claim for fraudulent misrepresentation (Standard Chartered Bank v
Pakistan National Shipping Corporation (No.2) {[}2003{]} 1 AC 959).

\hypertarget{damages-for-negligent-misrepresentation}{%
\paragraph{Damages for Negligent
Misrepresentation}\label{damages-for-negligent-misrepresentation}}

\hypertarget{s-21-damages}{%
\subparagraph{S 2(1) Damages}\label{s-21-damages}}

s 2(1) CA 1967:

\begin{quote}
``Where a person has entered into a contract after a misrepresentation
has been made to him by another party thereto and as a result thereof he
has suffered loss, then, if the person making the misrepresentation
would be liable to damages in respect thereof had the misrepresentation
been made fraudulently, that person shall be so liable notwithstanding
that the misrepresentation was not made fraudulently, unless he proves
that he had reasonable grounds to believe and did believe up to the time
the contract was made that the facts represented were true.''
\end{quote}

In Royscot Trust Ltd v Rogerson {[}1991{]} 2 QB 297 the court confirmed
that the effect of this statute is that the correct measure of damages
for negligent misrepresentation must be based on the tort of deceit (not
the tort of negligence). This implies that the claimant is entitled to
recover all losses, even if those losses are unforeseeable. Note that
this draws no distinction between the honest and careless representor,
and the fraudulent representor (which is criticised). But for negligent
misrepresentation, damages may be reduced for Contributory negligence.

\hypertarget{s-22-damages}{%
\subparagraph{S 2(2) Damages}\label{s-22-damages}}

s 2(2) CA 1967:

\begin{quote}
``Where a person has entered into a contract after a misrepresentation
has been made to him otherwise than fraudulently, and he would be
entitled by reason of the misrepresentation, to rescind the contract,
then, if it is claimed, in any proceedings arising out of the contract,
that the contract ought to be or has been rescinded, the court or
arbitrator may declare the contract subsisting and award damages in lieu
of rescission, if of the opinion that it would be equitable to do so,
having regard to the nature of the misrepresentation and the loss that
would be caused by it if the contract were upheld, as well as to the
loss that rescission would cause to the other party.''
\end{quote}

Remedy of damages in lieu of rescission only available at the discretion
of the court, and only available for negligent and innocent
misrepresentation.

The court must consider various factors (UCB Corporate Services v
Thomason {[}2005{]} 1 All ER (Comm) 601):

\begin{itemize}
\tightlist
\item
  The nature and seriousness of the misrepresentation
\item
  The loss that would be caused if the contract were upheld
\item
  The loss that rescission would cause the other party
\end{itemize}

\hypertarget{distinction}{%
\subparagraph{Distinction}\label{distinction}}

Note that s 2(2) damages are a separate award of damages to those
awarded under s 2(1)

\begin{longtable}[]{@{}ll@{}}
\toprule()
Damages awarded & Rationale \\
\midrule()
\endhead
s 2(1) damages & Intended to compensate the claimant for the loss
directly flowing from the negligent misrepresentation. \\
s 2(2) damages & Intended to compensate the representee where the court
has decided, at its discretion, not to award rescission. \\
\bottomrule()
\end{longtable}

Where damages in lieu of rescission are awarded, the damages awarded
under s 2(1) will be reduced to reflect those awarded under s 2(2) (as
explained by s 2(3))

\hypertarget{damages-for-innocent-misrepresentation}{%
\paragraph{Damages for Innocent
Misrepresentation}\label{damages-for-innocent-misrepresentation}}

Where an innocent misrepresentation has been made, the representee is
only entitled to the remedy of rescission and, if applicable, an
indemnity to cover the cost of the legal obligations arising from the
contract entered into. There is no automatic right to damages for an
innocent misrepresentation. But, as with negligent misrepresentation,
the court has the discretion under s 2(2) to award damages in lieu of
rescission. Note Contributory negligence cannot be claimed for innocent
misrepresentation.

\hypertarget{misrepresentation-and-exemption-clauses}{%
\paragraph{Misrepresentation and Exemption
Clauses}\label{misrepresentation-and-exemption-clauses}}

If a contract contains Exemption Clauses purporting to protect a party
from liability for misrepresenting, or purporting to exclude or restrict
any remedy available to the other party, the clause will be of no effect
except in so far as it satisfies the requirement of reasonableness (s 3
MA 1967 as substituted by s 8 Unfair Contract Terms Act 1977). The
requirement of reasonableness comes from
\href{https://www.legislation.gov.uk/ukpga/1977/50/section/11}{s 11(1)
of the Unfair Contract Terms Act 1977}.

\hypertarget{non-reliance-clauses}{%
\paragraph{Non-reliance Clauses}\label{non-reliance-clauses}}

It is doubtful whether an 'entire agreement' clause would be sufficient
to exclude liability for any misrepresentation. A possible boilerplate
clause excluding this liability could be

\begin{quote}
The parties agree and confirm that neither party has made any
representations to the other in relation to this contract or its subject
matter, and neither has any party relied on any representation from the
other in entering into this contract.
\end{quote}

The court will not allow liability for fraudulent misrepresentation to
be avoided in this way.

\hypertarget{hedley-byrne-v-heller}{%
\subsubsection{Hedley Byrne V Heller}\label{hedley-byrne-v-heller}}

There is a claim for negligent misstatement in the law of tort, based on
the judgment in Hedley Byrne v Heller {[}1964{]} AC 465. This depends on
there being a relationship between two parties that gives rise to a duty
of care. The rule in Hedley Byrne v Heller is, however, eclipsed by the
Misrepresentation Act 1967 in a situation of inducement to contract,
because the onus of proof in the latter is more favourable to a
claimant.

\hypertarget{facility-agreement-1}{%
\subsubsection{Facility Agreement}\label{facility-agreement-1}}

The most important remedies for misleading and inaccurate information
are built into the facility agreement, typically as events of default.
The general law on misrepresentations is therefore only relevant in the
unlikely event that a material representation was not caught by the
facility agreement, or the facility agreement was unenforceable.

\begin{env-a0fe8824-15a9-47d5-bf01-f54598533eae}

Tip

The ultimate remedy appropriate for breach of representation in a
Facilities Agreement is the immediate repayment of the Loan, together
with interest and any other amounts outstanding under the Agreement.
Damages aren't really relevant.

\end{env-a0fe8824-15a9-47d5-bf01-f54598533eae}

\hypertarget{undertakings}{%
\section{Undertakings}\label{undertakings}}

Aim: to give the lender control over the borrower once the agreement is
executed, to protect the loan monies.

\hypertarget{form-and-purpose}{%
\subsection{Form and Purpose}\label{form-and-purpose}}

Promise given by the bank to the borrower. May involve a promise to do/
not to something, or to procure that something is done/ not done by a
third party within the borrower's control.

Categories:

\begin{enumerate}
\tightlist
\item
  Financial covenants

  \begin{itemize}
  \tightlist
  \item
    Financial targets which the borrower is required to meet during the
    life of the loan
  \item
    Relate to profit, cash flow or net worth.
  \end{itemize}
\item
  Information undertakings

  \begin{itemize}
  \tightlist
  \item
    Promises the provide the bank with certain information.
  \end{itemize}
\item
  General undertakings

  \begin{itemize}
  \tightlist
  \item
    e.g., by prohibiting disposals, maintaining insurance and sticking
    to the key business.
  \end{itemize}
\end{enumerate}

\begin{env-9a8b2d5a-e80b-4017-bd60-2afb8ee62329}

Warning

Breach of an undertaking is made an event of default, allowing the bnak
to cancel its commitment and demand repayment.

\end{env-9a8b2d5a-e80b-4017-bd60-2afb8ee62329}

\hypertarget{undertaking-vs-representation}{%
\subsection{Undertaking Vs
Representation}\label{undertaking-vs-representation}}

Undertakings can often be recast as repeat representations (just stating
the fact).

In either case, the likely result of the breach is an event of default.
But preferable to get a borrower to represent the position at signing
and also undertake to maintain the position throughout the term of the
loan.

\begin{env-a0fe8824-15a9-47d5-bf01-f54598533eae}

Tip

This results in a "tighter" constraint: the borrower is in default
immediately as the undertaking is broken, rather than a repeating
representation being tested at the time of repetition (typically at the
start of every interest period).

\end{env-a0fe8824-15a9-47d5-bf01-f54598533eae}

Also gives more flexibility: can make the representation more strict
than the subsequent undertaking.

\begin{itemize}
\tightlist
\item
  Breach of representation may give remedies under Misrepresentation Act
  1967;
\item
  Breach of an undertaking may give remedies of specific performance/
  injunction.
\item
  Courts will not order specific performance where damages are an
  adequate remedy (almost always for a facility agreement).
\end{itemize}

\hypertarget{financial-covenants}{%
\subsection{Financial Covenants}\label{financial-covenants}}

Undertakings which impose financial targets/ limitations on the
borrower. Common for leveraged finance, unusual for investment grade
borrowers. Two broad types: maintenance covenants and incurrence
covenants.

\hypertarget{maintenance-covenants}{%
\subsubsection{Maintenance Covenants}\label{maintenance-covenants}}

They require the borrower to maintain a specified level of financial
performance and are tested regularly over the life of the loan facility.
Measure past performance, not future prospects.

\hypertarget{incurrence-covenants}{%
\subsubsection{Incurrence Covenants}\label{incurrence-covenants}}

Also set financial targets, but borrower must show compliance only if it
wants to do specific things (e.g., dispose of large assets, acquire new
companies, pay dividends).

If these are not done, the covenants are not incurred (not tested).
Originated from high yield bonds.

\hypertarget{limits-and-ratios}{%
\subsubsection{Limits and Ratios}\label{limits-and-ratios}}

Some financial covenants simply set a financial limit which a borrower
must attain. More common to consider ratios.

\hypertarget{leverage-ratio}{%
\paragraph{Leverage Ratio}\label{leverage-ratio}}

\[\text{Net\ Debt}:\text{EBITDA}\]

\begin{itemize}
\tightlist
\item
  Compares debt to earnings (profit) over a period (e.g., a year).
\item
  Net debt

  \begin{itemize}
  \tightlist
  \item
    Debt defined as commercial borrowing, not short-term trade
    creditors.
  \item
    Net debt is gross debt -- freely available cash held
  \item
    May be consolidated across a group.
  \end{itemize}
\item
  EBITDA

  \begin{itemize}
  \tightlist
  \item
    Earnings before interest, tax, depreciation and amortisation
  \item
    Measure of profit.
  \item
    Exceptional items typically excluded
  \item
    Typical acquisition financing have a ratio of 4:1, 5:1.
  \item
    Leverage ratio also called a "debt cover ratio". Different from
    gearing, which is {\(\text{Net\ Debt}:\text{Net\ Worth}\)}.
  \end{itemize}
\end{itemize}

\hypertarget{interest-cover-ratio}{%
\paragraph{Interest Cover Ratio}\label{interest-cover-ratio}}

\[\text{EBITDA}:\text{Net\ Finance\ Charges}\]

\begin{itemize}
\tightlist
\item
  Borrowers ability to cover the interest payment and fee obligations
  under its borrowing.
\item
  Net Finance Charges exclude capital repayments.
\item
  Any interest earned by the borrower from its investments is usually
  deducted from interest payable.
\item
  \textasciitilde{} 2 : 1
\end{itemize}

\hypertarget{cashflow-cover-ratio}{%
\paragraph{Cashflow Cover Ratio}\label{cashflow-cover-ratio}}

\[\text{Cashflow}:\text{Net\ Debt\ Service}\]

\begin{itemize}
\tightlist
\item
  Also known as ``fixed charge cover''/ ``debt service cover''
\item
  Ability of a borrower to generate enough cash to pay all amounts due
  under its borrowing.
\item
  Cashflow

  \begin{itemize}
  \tightlist
  \item
    Definition starts with EBITDA and then adjusts items included and
    excluded to represent cash movements more closely.
  \item
    Starting point for whether the bank thinks the borrower has more
    cash than needed to run its business. Excess cash is often taken and
    used to prepay the loan in a ``cash sweep''.
  \end{itemize}
\item
  Net debt service

  \begin{itemize}
  \tightlist
  \item
    Includes fees and interest payments due under commercial borrowing,
    but also any capital payments due
  \end{itemize}
\item
  \textasciitilde{} 1:1
\end{itemize}

\hypertarget{minimum-net-worth}{%
\paragraph{Minimum Net Worth}\label{minimum-net-worth}}

\begin{itemize}
\tightlist
\item
  A minimum net worth financial covenant aims to ensure the company
  maintains a minimum value of assets in the business.
\item
  The definition of net worth will vary.
\item
  Intended to represent the difference between assets and liabilities of
  the borrower.
\item
  Starting point: paid-up share capital + any reserves (shareholder
  funds in the balance sheet)
\item
  Most calculations exclude intangible assets. These include:

  \begin{itemize}
  \tightlist
  \item
    Goodwill
  \item
    Capitalised R\&D costs

    \begin{itemize}
    \tightlist
    \item
      R\&D usually entered as an expense
    \item
      May be lumped together and treated as an asset if it is reasonably
      certain that they will be recovered once the product is developed.
    \item
      Capitalised amount amortised to spread the cost over time as a
      form of accrual.
    \end{itemize}
  \item
    Patents, copyrights, licences etc.

    \begin{itemize}
    \tightlist
    \item
      Often not transferrable, worth little on insolvency
    \end{itemize}
  \item
    Payments on account

    \begin{itemize}
    \tightlist
    \item
      Excluded since they wold need to be repaid if the borrower did not
      perform.
    \end{itemize}
  \item
    Unaudited revaluation

    \begin{itemize}
    \tightlist
    \item
      To avoid the borrower over-inflating its value.
    \end{itemize}
  \item
    Tax provisions

    \begin{itemize}
    \tightlist
    \item
      Money set aside to pay taxes.
    \end{itemize}
  \end{itemize}
\end{itemize}

\hypertarget{capital-expenditure}{%
\paragraph{Capital Expenditure}\label{capital-expenditure}}

Puts a limit on the amount which the borrower can spend on ``long-term''
business assets over a given period. Helps to retain cash to service the
debt. Colloquially, ``capex'' provision.

\hypertarget{borrowing}{%
\paragraph{Borrowing}\label{borrowing}}

Bank limits a borrower's exposure to other creditors. May be achieved
through a financial covenant capping total borrowings. Might be linked
to tangible net worth. Transactions likely to be included in the
expanded definition of ``borrowing'':

\begin{itemize}
\tightlist
\item
  Financial leases and instalment credit options.

  \begin{itemize}
  \tightlist
  \item
    Lesee will usually be required over the life of the lease to pay
    instalments to the lessor equivalent to the full value of the leased
    asset, together with a return on the capital value.
  \item
    Lesee usually responsible for maintaining the asset.
  \item
    Commonly used to fund computer hardware/ heavy machinery/ aircraft.
  \end{itemize}
\item
  Deferred purchase agreements

  \begin{itemize}
  \tightlist
  \item
    Buy now pay later
  \end{itemize}
\item
  Capital instruments

  \begin{itemize}
  \tightlist
  \item
    Bonds, notes, commercial paper.
  \end{itemize}
\item
  Guarantees and indemnities

  \begin{itemize}
  \tightlist
  \item
    Financial assurance given by the borrower in respect of any
    transaction.
  \end{itemize}
\end{itemize}

When defining ratios, define terms specifically rather than relying on
standard terminology.

\hypertarget{information-for-testing}{%
\subsubsection{Information for Testing}\label{information-for-testing}}

\begin{itemize}
\tightlist
\item
  All companies must produce year end accounts
\item
  Listed companies must produce interim accounts (unaudited) for first 6
  months of their financial year.
\item
  To test more than twice a year: bank can require B to supply a
  ``compliance certificate'' signed by 2 directors, confirming the
  financial covenants have been met on the requisite date. Usually
  accompanies by financial statements covering the period.
\end{itemize}

\hypertarget{gaap-and-accounting}{%
\subsubsection{GAAP and Accounting}\label{gaap-and-accounting}}

B usually allowed to follow changes to GAAP as long as auditors provide
a description of changes necessary to ``convert'' the statements to the
original policies (``frozen GAAP'' provision). Usually a provision for
parties to negotiate re-setting the covenants, to avoid frozen GAAP
becoming a big burden.

\hypertarget{equity-cure}{%
\subsubsection{Equity Cure}\label{equity-cure}}

Borrowers traditionally could not ``cure'' a breach of financial
covenants. Strong equity sponsors (PE firms etc.) have negotiated a
right to inject cash into the borrower to help it meet financial
covenants, either by prepayment or putting money into an account.

Equity sponsors would prefer to treat the new cash as EBITDA, but rarely
given. Usually treated as positive cash flow.

\hypertarget{information-undertakings}{%
\subsection{Information Undertakings}\label{information-undertakings}}

Effectiveness of most provisions depends on the bank receiving a regular
supply of information from the borrower.

\hypertarget{financial-information}{%
\subsubsection{Financial Information}\label{financial-information}}

Used to monitor the borrower's financial wellbeing and check for any
breaches of financial covenants. Main sources:

\begin{enumerate}
\tightlist
\item
  Year end accounts (and possibly interims/ pro forma quarterlies),
  included consolidated accounts.
\item
  Management accounts (unpublished pro forma accounts for internal use).
\item
  Forecast of borrower performance, usually in the form of a projected
  balance statement.
\item
  Certificate showing compliance with all financial covenants
\item
  "Sweep-up" provision, any other information relating to financial
  condition which the bank requests. Borrower should negotiate this
  subject to no material cost (at least prior to default).
\end{enumerate}

\hypertarget{shareholder-documents}{%
\subsubsection{Shareholder Documents}\label{shareholder-documents}}

Often restricted to information which borrower is required by statute to
send to shareholders.

\hypertarget{kyc}{%
\subsubsection{KYC}\label{kyc}}

Undertaking by borrower to supply information reasonably requested by
any lender in order to fulfil ongoing KYC obligations -- this is
standard. Borrower may ask to limit this to information within their
possession and control, or which they can obtain with reasonable
endeavours.

\hypertarget{insurance}{%
\subsubsection{Insurance}\label{insurance}}

Maintaining adequate insurance for major assets -- even if loan
unsecured. Often coupled with undertakings to notify bank of material
changes to insurance cover and produce copies of policies if requested.

\hypertarget{notification-of-default}{%
\subsubsection{Notification of Default}\label{notification-of-default}}

Borrower to notify the bank if it defaults under the loan, and possibly
also any ``potential'' events of default which might occur. This should
be carefully defined.

Written confirmation to be given by directors at bank's request,
confirming there has been no default or potential event of default.

\hypertarget{general-undertakings}{%
\subsection{General Undertakings}\label{general-undertakings}}

Can be positive or negative promises.

\hypertarget{positive-undertakings}{%
\subsubsection{Positive Undertakings}\label{positive-undertakings}}

\begin{itemize}
\tightlist
\item
  Insurance

  \begin{itemize}
  \tightlist
  \item
    Keep major assets insured.
  \item
    Can also include "keyman" insurance (insurance for if key personnel
    leave the company, where the business is small). If the key man
    dies, lump sum paid to borrower.
  \end{itemize}
\item
  Consents, licences and laws

  \begin{itemize}
  \tightlist
  \item
    Undertaking that borrower maintain all necessary consents,
    authorisations and licences to perform its obligations under the
    loan document.
  \end{itemize}
\item
  Access to information and assistance

  \begin{itemize}
  \tightlist
  \item
    Very occasionally, bank allowed to inspect books/ records/ assets.
    Should be at reasonable notice and within office hours.
  \end{itemize}
\end{itemize}

\hypertarget{negative-undertakings}{%
\subsubsection{Negative Undertakings}\label{negative-undertakings}}

Prohibit borrower from doing something.

\hypertarget{negative-pledge}{%
\paragraph{Negative Pledge}\label{negative-pledge}}

\begin{itemize}
\tightlist
\item
  Undertaking by the borrower not to create any security over its
  assets.
\item
  Standard if borrower is unsecured, often insisted upon even if the
  bank has security, to protect the bank's floating charge security.
\item
  Borrower may want to negotiate exceptions:

  \begin{itemize}
  \tightlist
  \item
    Security necessary under equipment leasing/ hire purchase agreements
  \item
    Liens arising in the ordinary course of business (e.g., asset
    retained by third party until payment for services). Bank may
    restrict to a time period.
  \item
    Cash collateralisation necessary for business
  \item
    Security pre-dating the facility
  \item
    Security acquired through purchasing asset/ business which is
    already subject to a security
  \item
    Netting (``set-off'') arrangements between group companies
  \item
    Retention of title clauses.
  \item
    Project finance. Banks will usually want security over the asset
    being funded.
  \item
    De minimis amount (``basket'') allowing a minimum amount of debt for
    which the borrower may create security.
  \end{itemize}
\item
  Often wording attempts to create security for the ban automatically in
  the event of a breach. Doubts over whether this is effective in
  English law.
\end{itemize}

\hypertarget{pari-passu-ranking}{%
\paragraph{Pari Passu Ranking}\label{pari-passu-ranking}}

\begin{itemize}
\tightlist
\item
  The claims of the bank rank at least pari passu with the claims of all
  other present or future unsecured and subordinated creditors, except
  for those preferred by bankruptcy, insolvency, liquidation or other
  similar laws.
\item
  Tries to ensure bank is not subordinated to unsecured creditors.
  Repeated representation often used here.
\end{itemize}

\hypertarget{disposals-and-acquisitions}{%
\paragraph{Disposals and
Acquisitions}\label{disposals-and-acquisitions}}

\begin{itemize}
\tightlist
\item
  Restriction/ prohibition on disposal of assets other than those in the
  ordinary course of trading/ in exchange for comparable assets.
\item
  May allow disposal of obsolete assets.
\item
  The clause might allow disposals at arm's length up to a maximum
  aggregate annual value (a `basket'), or be tied in to mandatory
  prepayment.
\item
  Restriction from substantial acquisitions/ mergers
\item
  Undertaking that there will not be a substantial change to the general
  nature of the business.
\end{itemize}

\hypertarget{lending}{%
\paragraph{Lending}\label{lending}}

\begin{itemize}
\tightlist
\item
  Carries the risk of non-repayment and a depletion of assets.
\item
  Typically, an undertaking not to lend must be given, other than in the
  ordinary course of business/ within the group.
\end{itemize}

\hypertarget{shares-and-dividends}{%
\paragraph{Shares and Dividends}\label{shares-and-dividends}}

\begin{itemize}
\tightlist
\item
  Undertaking not to make any distribution to shareholders without the
  bank's consent/ meeting of a financial test.
\item
  The bank may agree that the undertaking will not apply to the
  borrower's subsidiaries.
\item
  Take care if the borrower has issued any preference shares -- giving
  such an undertaking may result in a breach of contract.
\item
  If the borrower has subsidiaries, the bank may require the borrower to
  undertake not to allow subsidiaries to issue shares to third parties.
\end{itemize}

\hypertarget{auditors}{%
\paragraph{Auditors}\label{auditors}}

\begin{itemize}
\tightlist
\item
  Any contractual clause that restricts an audited person's choice of
  auditor has no legal effect.
\item
  Auditors and Third Country Auditors Regulations 2016
\end{itemize}

\hypertarget{persons-with-significant-control}{%
\paragraph{Persons with Significant
Control}\label{persons-with-significant-control}}

\begin{itemize}
\tightlist
\item
  Register of individuals and entities must be kept.
\item
  If a company fails to comply with this `people with significant
  control' regime, the sanctions may prevent security being taken or
  enforced over the company's shares. Secured facilities will therefore
  usually include an undertaking that the group companies will comply
  with the relevant requirements.
\end{itemize}

\hypertarget{banks-perspective-1}{%
\subsection{Bank's Perspective}\label{banks-perspective-1}}

Choice of undertakings depends on the context of the loan.

When drafting, take care of ``shadow director'' provisions. If a bank
uses the undertakings in its facility agreement to force a borrower to
take certain action, it seems possible that the bank might be a `shadow
director'. The time of greatest risk is usually when a borrower is in
financial difficulties.

A shadow director may be held liable for wrongful trading under the IA
1986, s 214(7), and may have to contribute to a company's assets on its
liquidation.

\hypertarget{borrowers-perspective-1}{%
\subsubsection{Borrower's Perspective}\label{borrowers-perspective-1}}

Provision of financial projections and management accounts can be v
time-consuming if not already part of the directors' remit. Borrower's
solicitor will need to ensure undertakings are not too tight. Insert
``reasonable'' wherever possible. Try to negotiate ``cure'' periods for
as many breaches as possible. Try to get references to subsidiaries
changed to ``material subsidiaries''.

\hypertarget{events-of-default}{%
\section{Events of Default}\label{events-of-default}}

Dictate the circumstances in which a bank can terminate its loan early
or take other action. Usually very heavily negotiated.

\hypertarget{purpose}{%
\subsection{Purpose}\label{purpose}}

If a provision of an agreement is breached, the injured party may seek
remedy for breach of contract. The remedy available will depend on the
time of term breached:

\begin{enumerate}
\tightlist
\item
  \textbf{Conditions}. Terms which are fundamental to a contract. Breach
  of a condition may allow the injured party to terminate the contract
  and claim damages, or to affirm the contract and claim damages.
\item
  \textbf{Warranties}. Warranties are terms which are not considered
  fundamental to the contract. Breach of a warranty will not allow the
  injured party to terminate the contract, but merely to claim damages
  for any loss arising from the breach.
\item
  \textbf{Intermediate (or innominate) terms}. Intermediate terms which
  are capable of being breached either in a trivial, or in a fundamental
  way. The remedy available on the breach of an intermediate term will
  be either damages, or damages and termination, depending on the effect
  of the breach on the contract.
\end{enumerate}

MERMAID2

If a party makes a false representation to induce the other party into
contract, the remedies for misrepresentation may be available.

Events of default provide the bank with a contractual right suspend/
cancel a commitment, put the loan on demand or immediately demand
repayment (acceleration). Means the bank does not have to rely on
statutory/ common law remedies.

\hypertarget{common-events-of-default}{%
\subsection{Common Events of Default}\label{common-events-of-default}}

\hypertarget{late-payment}{%
\subsubsection{Late Payment}\label{late-payment}}

\begin{itemize}
\tightlist
\item
  Most important event of default.
\item
  Banks are highly sensitive about \textbf{punctual payment} by the
  borrower.
\item
  A bank will require default \textbf{interest} to be paid on the
  overdue amount.
\item
  General requirement to \textbf{indemnify} the bank against any costs
  associated with the default.
\item
  Missed payment probably signals cash flow problems in the business.
\item
  Default for late payment will be \textbf{non-negotiable}. Will cover
  late payment as well as interest and repayments of principal.
\item
  Bank may allow the borrower a short \textbf{grace period} to rectify
  any accidental late payment/ late payment due to technical issues.
\item
  Bank will ensure that any default interest and indemnity provisions
  run \textbf{from the date from which payment is overdue}, not from the
  end of any grace period.
\end{itemize}

\hypertarget{breach-of-representation}{%
\subsubsection{Breach of
Representation}\label{breach-of-representation}}

\begin{itemize}
\tightlist
\item
  Default for an incorrect/ misleading written representation often
  \textbf{drafted widely}.
\item
  Bank may concede that the requirement should be \textbf{material}/
  have material adverse effect.
\end{itemize}

\hypertarget{breach-of-undertaking}{%
\subsubsection{Breach of Undertaking}\label{breach-of-undertaking}}

\begin{itemize}
\tightlist
\item
  Expressed as an explicit event of default
\item
  May be a \textbf{grace period/ materiality} qualification.
\end{itemize}

\hypertarget{breach-of-financial-covenants}{%
\subsubsection{Breach of Financial
Covenants}\label{breach-of-financial-covenants}}

\begin{itemize}
\tightlist
\item
  Some equity sponsors manage to negotiate ``\textbf{equity cure}''
  rights to avoid a default on breach of a financial covenant.
\item
  Another concession is a \textbf{``Mulligan'' provision}. Works to
  either

  \begin{itemize}
  \tightlist
  \item
    Prevent the bank calling an event of default unless the borrower
    breaches a financial covenant on \textbf{two consecutive testing
    dates} (true Mulligan); or
  \item
    Allow the bank to call a default on the first breach of a financial
    covenant, but to \textbf{lose that right} if it fails to call a
    default \textbf{before} the borrower meets the financial covenant on
    the \textbf{subsequent test date} (not a true Mulligan).
  \end{itemize}
\end{itemize}

\hypertarget{cross-default}{%
\subsubsection{Cross-default}\label{cross-default}}

\begin{itemize}
\tightlist
\item
  Key event of default, heavily negotiated.
\item
  If a borrower \textbf{defaults on another loan or contract}, the other
  party may force terms on the borrower {\(\rightarrow\)} prejudices the
  lending bank.
\item
  A cross-default clause operates by \textbf{automatically defaulting}
  the borrower under the loan in the event that it defaults under
  certain other types of agreement.
\item
  Ensures the bank can protect its position and possibly accelerate the
  loan.
\item
  Effectively gives a bank the benefit of the default provisions in
  other agreements to which the borrower is a party.
\item
  Usually limited by

  \begin{itemize}
  \tightlist
  \item
    being confined to defaults which occur under other documents
    providing `\textbf{financial indebtedness}'; and
  \item
    including a \textbf{\emph{de minimus} threshold}.
  \end{itemize}
\item
  The borrower may also negotiate:

  \begin{itemize}
  \tightlist
  \item
    That the cross-default will operate only if the other default will
    have a \textbf{material} effect on its ability to perform its
    obligations under the facility agreement.
  \item
    That the cross-default will \textbf{trigger at a specific point} in
    the initial default (e.g. failure to pay amount due, creditor
    demands early repayment, creditor exercises a right to demand early
    repayment -- "\textbf{cross-acceleration}").

    \begin{itemize}
    \tightlist
    \item
      Cross-acceleration is the most desirable for the borrower.
    \item
      But the bank will be concerned that the borrower will favour the
      party with the benefit of default to persuade them not to
      accelerate.
    \end{itemize}
  \item
    Carve out for defaults which the borrower is \textbf{disputing in
    good faith}. Strongly resisted by most banks.
  \end{itemize}
\item
  The bank may want cross-default provision to extend to \textbf{other
  companies in the group}.

  \begin{itemize}
  \tightlist
  \item
    Borrower can try to limit this to \textbf{material or principal
    subsidiaries}.
  \end{itemize}
\end{itemize}

The bank has an \textbf{evidential problem} in finding out when a
borrower has defaulted under another agreement. The borrower breaching a
representation will then be useless (just means they've breached twice
rather than once). Furthermore, a bank is subject to an implied duty of
confidentiality with regard to a customer's affairs (see Tournier v
National Provincial and Union Bank of England {[}1924{]} 1 KB 461). Bank
will refer to its own knowledge and monitoring.

\hypertarget{change-of-business}{%
\subsubsection{Change of Business}\label{change-of-business}}

It will therefore usually be an event of default if the borrower makes
any \textbf{material change to its business}, ceases to carry on any
material part of its business, or disposes of a material part of its
business.

\hypertarget{change-of-control}{%
\subsubsection{Change of Control}\label{change-of-control}}

Change of control usually triggers \textbf{mandatory prepayments},
rather than an event of default.

\hypertarget{insolvency}{%
\subsubsection{Insolvency}\label{insolvency}}

\begin{itemize}
\tightlist
\item
  The insolvency of a borrower will, without question, be an
  \textbf{event of default}.
\item
  Any \textbf{negotiation} of this clause will simply relate to the
  \textbf{timing} of default.
\item
  The bank will want \textbf{time to take action} (e.g., enforce
  security) at the first sign of a problem.
\item
  Some clauses may make reference to the \textbf{two insolvency tests}
  in s 123 of the IA 1986.

  \begin{itemize}
  \tightlist
  \item
    Section 123(1) deems a company unable to pay its debts if it has not
    paid a debt \textbf{three weeks} after service of a
    \textbf{statutory demand} or court order (if the debt is over £750).
  \item
    Section 123(2) deems a company unable to pay its debts if the value
    of its \textbf{assets is less than its liabilities}.

    \begin{itemize}
    \tightlist
    \item
      BNY Corporate Trustee Services Ltd v Eurosail-UK-2007-3BL Plc \&
      Others {[}2011{]} EWCA Civ 227: s 123(2) `applies to a company
      whose assets and liabilities (including contingent and future
      liabilities) are such that it had reached a point of no return'.
    \end{itemize}
  \end{itemize}
\item
  Usually also triggers relating to \textbf{insolvency proceedings}.

  \begin{itemize}
  \tightlist
  \item
    Typically, this will include any corporate action, legal proceedings
    or other procedure or step.
  \item
    e.g., moratorium, appointment of liquidator, enforcement of any
    security over the borrower's assets.
  \end{itemize}
\item
  Borrower will want the event of default to trigger as late as
  possible, e.g., by negotiating a remedy period, a carve out for
  disputed debts, or a default limited to actual appointment of
  administrator/ receiver or similar.
\end{itemize}

\hypertarget{proceedings}{%
\subsubsection{Proceedings}\label{proceedings}}

\begin{itemize}
\tightlist
\item
  Event of default if the borrower becomes involved in any
  \textbf{material litigation, arbitration or similar} proceedings.
\item
  Borrower will try to negotiate a \emph{de minimus} limit or
  proceedings which will/ might have a material adverse effect on the
  borrower's ability to perform obligations.
\end{itemize}

\hypertarget{unlawful-performance}{%
\subsubsection{Unlawful Performance}\label{unlawful-performance}}

\begin{itemize}
\tightlist
\item
  Event of default if it becomes \textbf{unlawful} for a borrower to
  continue to perform its obligations under a facility agreement.
\item
  Event of default if the borrower should lose \textbf{any consents/
  authorisations} necessary to perform obligations under the loan.
\item
  Borrower will want \textbf{materiality} and may ask for
  \textbf{mandatory pre-payment} for materiality.
\end{itemize}

\hypertarget{general-default}{%
\subsubsection{General Default}\label{general-default}}

\begin{itemize}
\tightlist
\item
  Event of default if the borrower fails to comply with ``other
  obligations''.
\item
  Invariably drafted with a relatively long `\textbf{cure period}' of
  perhaps 30 days.
\item
  \textbf{Sweep up provision} for any minor defaults.
\end{itemize}

\hypertarget{qualified-accounts}{%
\subsubsection{Qualified Accounts}\label{qualified-accounts}}

\begin{itemize}
\tightlist
\item
  Event of default which is triggered if the \textbf{borrower's accounts
  are qualified} by the auditors.
\item
  Common in leveraged transactions.
\item
  Borrower will want to limit any default to a qualification which has a
  \textbf{material adverse effect} on its ability to perform its
  obligations under the loan facility.
\end{itemize}

\hypertarget{repudiation}{%
\subsubsection{Repudiation}\label{repudiation}}

\textbf{Actual or attempted repudiation} by the borrower is an event of
default. Rarely negotiated.

\hypertarget{material-adverse-change}{%
\subsubsection{Material Adverse Change}\label{material-adverse-change}}

\begin{itemize}
\tightlist
\item
  Colloquially, ``MAC''.
\item
  Acts as a \textbf{sweep up provision} allowing the bank to call a
  default if there is a material adverse change in the borrower's
  position or circumstances which might prevent it from complying with
  any provisions in the facility agreement.
\item
  Many borrowers will try to exclude this provision, and it is always
  heavily negotiated.
\item
  Strong clause: Any event or circumstance occurs which the
  \textbf{Majority Lenders} believe might have a \textbf{material
  adverse effect} on the ability of the Borrower to perform or comply
  with its obligations under this Agreement.
\item
  Borrower may negotiate:

  \begin{itemize}
  \tightlist
  \item
    Threshold for Majority Lenders (commonly {\(\frac{2}{3}\)})
  \item
    Insertion of "reasonably".
  \item
    Definition of material adverse effect.
  \end{itemize}
\item
  Most common use for the MAC is to threaten default and persuade a
  borrower to start a dialogue with the bank.
\item
  In BNP Paribas v Yukos Oil Company {[}2005{]} EWHC 1321 (Ch), the
  \textbf{court upheld the validity} of accelerating a loan based on a
  MAC event of default.
\end{itemize}

\hypertarget{interaction-with-other-clauses}{%
\subsection{Interaction with Other
Clauses}\label{interaction-with-other-clauses}}

Events of default should not overlap with any representations or
undertakings. If there are overlaps, check they do not contradict.

\hypertarget{action-on-default}{%
\subsection{Action on Default}\label{action-on-default}}

A bank may:

\begin{enumerate}
\tightlist
\item
  \textbf{Cancel} any undrawn commitments
\item
  Put all or part of the \textbf{facility on demand}, and take over the
  selection of interest periods.
\item
  \textbf{Demand immediate repayment} of part or all outstanding capital
  and interest, as well as any cash collateral.
\item
  Enforce any \textbf{security}.
\item
  Do \textbf{nothing}.
\end{enumerate}

A bank will want these remedies available `on or at any time after an
Event of Default has occurred and until it is waived \ldots'; a borrower
will want to add `\ldots{} \textbf{or remedied}'. Banks commonly allow
this. Implications:

\begin{enumerate}
\tightlist
\item
  If the borrower can \textbf{unilaterally remedy} an event of default,
  it will have \textbf{little incentive to disclose} when it occurs.
\item
  What constitutes a \textbf{remedy} of certain types of default is
  \textbf{not always clear}. Might say ``remedied to the satisfaction of
  the bank''.
\end{enumerate}

In a syndicated loan, the \textbf{agent} will often be given
\textbf{discretion} as to what action to take on an event of default,
with an overriding obligation to act in accordance with the instructions
of the majority banks.

\begin{itemize}
\tightlist
\item
  Enables a \textbf{fast} decision.
\item
  But agent takes potential \textbf{liability} for wrong decisions.
\end{itemize}

If the loan is terminated and outstanding amounts are immediately
repayable, default interest applies until the amounts are settled in
full.

\hypertarget{grace-periods-defaults-and-potential-defaults}{%
\subsection{Grace Periods, Defaults and Potential
Defaults}\label{grace-periods-defaults-and-potential-defaults}}

Some events of default include a \textbf{specified period} during which
the borrower is permitted to \textbf{remedy} the event or apply for a
\textbf{waiver}. During the grace period, the borrower is prevented from
borrowing new money (``draw stop'').

\begin{env-f09e27d9-67b8-4931-ab1f-ca6031966775}

Definition

An event of default with an in-built grace period is sometimes referred
to as a `potential event of default'.

\end{env-f09e27d9-67b8-4931-ab1f-ca6031966775}

\begin{env-b81a0bad-19ae-4a38-aa54-7c89ad4aa09e}

Variation

LMA facilities define ``default'' to mean

\begin{quote}
an Event of Default or any event or circumstance specified in Clause {[}
{]} (Events of Default) which would (with the expiry of a grace period,
the giving of notice, the making of any determination under the Finance
Documents or any combination of any of the foregoing) be an Event of
Default.
\end{quote}

The equivalent of a `potential event of default' is then described as `a
Default other than an Event of Default'.

\end{env-b81a0bad-19ae-4a38-aa54-7c89ad4aa09e}

Borrowers must undertake to tell the banks as soon as they are aware of
a `potential event of default', but

\begin{itemize}
\tightlist
\item
  Borrowers under a revolving credit facility (RCF) will argue that the
  draw stop within a grace period should not prevent them rolling over a
  maturing RCF loan.
\item
  Borrowers should not agree to give a repeating representation that
  there are no `potential events of default' (or equivalent).
\end{itemize}

\hypertarget{banks-perspective-2}{%
\subsection{Bank's Perspective}\label{banks-perspective-2}}

\textbf{Demanding early repayment} will be considered a last resort, but
a useful \textbf{negotiating tool}. The bank may demand increased fees/
margin, tighter financial covenants or other terms.

A bank will not accept arguments that the borrower should not be
defaulted on the basis of events beyond its control. The rationale is
\textbf{risk management}, not fairness.

\hypertarget{borrowers-perspective-2}{%
\subsection{Borrower's Perspective}\label{borrowers-perspective-2}}

Borrower must be very cautious when agreeing to the events of default.
In reality, though, a bank is unlikely to withdraw a loan without
considerable provocation.

\hypertarget{boiler-plate-and-mechanics-clauses}{%
\section{Boiler Plate and Mechanics
Clauses}\label{boiler-plate-and-mechanics-clauses}}

\hypertarget{repayment-prepayment-and-cancellation}{%
\subsection{Repayment, Prepayment and
Cancellation}\label{repayment-prepayment-and-cancellation}}

\hypertarget{repayment-clause}{%
\subsubsection{Repayment Clause}\label{repayment-clause}}

Arguably the most important provision in the agreement. Form of
repayment depends on type of facility. Overdraft usually on demand and
does not require a repayment date.

\hypertarget{prepayment}{%
\subsubsection{Prepayment}\label{prepayment}}

Most loans will allow a borrower to repay capital early. Without a
contractual provision, borrower probably has no right to prepay a loan.
Prepaying will save interest payments. Borrowers may refinance existing
loans partway through their term.

\hypertarget{broken-funding}{%
\paragraph{Broken Funding}\label{broken-funding}}

If B repays, the bank may be left to service an interbank loan without
an equivalent interest stream from the borrower. Or bank may incur
breakage fees for terminating the interbank loan early. So usually there
is a "break costs" clause. Paying break costs could be avoided if B
prepays at the end of an interest period.

\hypertarget{prepayment-fee}{%
\subsubsection{Prepayment Fee}\label{prepayment-fee}}

Banks will ask for a prepayment fee (``prepayment premium'') to cover
administration costs of early repayment. Loans for leveraged
acquisitions may feature call protection: an extra amount payable by the
borrower if it prepays within e.g., 12 months of utilisation. Unlikely
to be drafted into agreements for senior facilities.

\hypertarget{order-of-application}{%
\subsubsection{Order of Application}\label{order-of-application}}

Amortised loans: banks resist treating a partial pre-payment as an early
settlement of the next repayment due. The bank wants a borrower to
demonstrate the financial control necessary to maintain a steady
repayment regime. May insist that any prepayment is treated as
satisfying the final repayment first, and then in inverse order (reduces
life of loan) / spread over the scheduled amounts rateably (reduces
exposure but maintains relationship).

\hypertarget{preference}{%
\subsubsection{Preference}\label{preference}}

If borrower becomes insolvent {\(\leq 6\)} months of prepayment (or 2
years for connected party), risk of liquidator/ administrator applying
to set aside the prepayment as a preference under s 239 IA 1986.
Typically, a remote risk.

\hypertarget{mandatory-prepayment}{%
\subsubsection{Mandatory Prepayment}\label{mandatory-prepayment}}

For an investment grade facility, there are 2 circumstances in which a
borrower will be required to make a mandatory prepayment:

\begin{enumerate}
\tightlist
\item
  If it becomes unlawful for a bank to lend to the borrower (e.g.,
  sanctions).
\item
  If there is a change in control of the borrower. Control is typically
  defined as the ability to remove half or more of the directors, the
  ability to dictate how the company is run, the holding of a majority
  of the shares in the borrower, or the ability to cast more than half
  the shareholder votes.
\end{enumerate}

Leveraged facility usually includes additional mandatory prepayment
events:

\begin{enumerate}
\tightlist
\item
  If the borrower disposes of certain fundamental assets other than in
  the ordinary course of business (and assuming the proceeds are not
  required to purchase a replacement asset);
\item
  If the borrower receives insurance monies for an asset it is not
  replacing;
\item
  If the borrower generates more cash than it is deemed to require to
  run its business (this is known as an `excess cash sweep');
\item
  If the facility is required to fund an acquisition and the purchase
  price is subsequently adjusted downwards due to a revaluation event;
  and
\item
  If the borrower subsequently raises funds through an issue of shares
  or debt securities.
\end{enumerate}

Repayment is usually allocated pro rata across banks.

\hypertarget{cancellation}{%
\subsubsection{Cancellation}\label{cancellation}}

If a loan provides a commitment period over which the borrower is
allowed to utilise the facility, the bank most likely charge a
commitment fee for making the facility available. Many loans allow the
borrower to cancel during the commitment period and save on commitment
fees. May be mandatorily cancelled in the event of default.

\hypertarget{margin-protection}{%
\subsection{Margin Protection}\label{margin-protection}}

The interest rate charged by a bank on a syndicated loan is typically
calculated by adding the margin to the estimated cost of lending. If the
cost of lending rises, the bank will want to have mechanisms to pass
this onto the lender.

\hypertarget{regulatory-capital}{%
\subsubsection{Regulatory Capital}\label{regulatory-capital}}

If a bank cannot recover money it has lent, there is a danger it may be
unable to repay depositors and other lenders. To reduce this risk, banks
required to maintain regulatory capital. This is a minimum level of
assets which must be maintained as a ``buffer''.

\hypertarget{origins}{%
\paragraph{Origins}\label{origins}}

\begin{itemize}
\tightlist
\item
  The Basel Committee (1980s) came up with Basel Capital Accord, adopted
  by all major banking jurisdictions.
\item
  A more sophisticated version, Basel II (2008) was adopted by most
  countries (except US).
\item
  Basel III implemented 2013-23.
\end{itemize}

\hypertarget{rationale}{%
\paragraph{Rationale}\label{rationale}}

\begin{env-02a65e52-3e2f-4e42-a1ad-a9c88e6d3852}

Important

The key proposition of Basel regime is that a bank needs to recover its
assets (primarily loans) to repay its depositors: if there is a risk
that an asset cannot be recovered in full, it must be backed by
`capital' to create a `buffer'. Capital buffer known as "regulatory
capital".

\end{env-02a65e52-3e2f-4e42-a1ad-a9c88e6d3852}

Starting point: regulatory capital is 8\% of the value of each asset
(but this is a minimum). Assets must be risk weighted.

\hypertarget{risk-weightings}{%
\paragraph{Risk Weightings}\label{risk-weightings}}

\begin{itemize}
\tightlist
\item
  Basel I approach: calculation of counterparty credit risk was
  relatively crude, grouping diverse borrower types into single risk
  categories.
\item
  Basel II better: banks used a "standardised approach" or "internal
  ratings based" approach.
\end{itemize}

\begin{longtable}[]{@{}ll@{}}
\toprule()
Approach & Details \\
\midrule()
\endhead
Standardised approach & Banks apply counterparty risk weightings as
pre-determined in the regulations. \\
IRB approach & Allows a bank to use its own internal model to calculate
the risk of recovery from different counterparties. \\
\bottomrule()
\end{longtable}

\hypertarget{types-of-capital}{%
\paragraph{Types of Capital}\label{types-of-capital}}

What is capital? Not just cash in an account. Basel II splits into 3
types. At least half a bank's regulatory capital had to be core Tier 1
capital, and Tier 3 capital could only be used to cover limited
regulatory capital requirements.

\begin{longtable}[]{@{}ll@{}}
\toprule()
Tier of capital & Details \\
\midrule()
\endhead
Tier 1 & Includes ordinary share capital, verified net profits and
certain reserves. \\
Tier 2 & Less permanence than Tier 1 and might incur fixed costs. \\
Tier 3 & Least stable, e.g., short-term subordinated loans, non-verified
profits. \\
\bottomrule()
\end{longtable}

\hypertarget{basel-iii-iv}{%
\paragraph{Basel III \& IV}\label{basel-iii-iv}}

Even more stringent requirements, including new buffers, increase in
capital requirement for systemically important banks, liquidity
requirements.

\hypertarget{increased-costs-clause}{%
\paragraph{Increased Costs Clause}\label{increased-costs-clause}}

Cost of maintaining regulatory capital passed to the borrower as part of
the margin on a loan. If the cost of capital increases, this eats into
the margin and reduces profit. Such a provision can be resented by the
borrower.

\hypertarget{mandatory-cost}{%
\subsubsection{Mandatory Cost}\label{mandatory-cost}}

PRA funds itself by leveraging fees against regulated fees. Facility
agreements give banks the opportunity to recover these costs from the
borrower. But LMA withdrew the schedule from market documentation in
2013, and now not used much.

\hypertarget{costs-and-expenses}{%
\subsection{Costs and Expenses}\label{costs-and-expenses}}

Bank will require a borrower to indemnify it against the incidental
costs it incurs setting up and running the loan. Includes:

\begin{itemize}
\tightlist
\item
  Costs of negotiation, preparation and execution
\item
  Registration of loan and ancillary documents
\item
  Preservation and enforcement of the bank's rights under the loan.
\end{itemize}

Usually covered by an indemnity from the borrower in the mandate letter.

\hypertarget{fees}{%
\subsection{Fees}\label{fees}}

Types include

\begin{itemize}
\tightlist
\item
  Front end fees

  \begin{itemize}
  \tightlist
  \item
    For initial work done to put the loan together.
  \item
    For a syndicated loan, known as the ``arrangement fee''.
  \item
    Arranger may have used participation fee to attract other banks to
    the syndicate.
  \item
    Amount of fees remaining once the arranger has paid any
    participation fees is called "praecipium".
  \end{itemize}
\item
  Underwriting fee

  \begin{itemize}
  \tightlist
  \item
    Initial banks underwrite syndication.
  \end{itemize}
\item
  Commitment fee

  \begin{itemize}
  \tightlist
  \item
    Charged on committed but undrawn money.
  \item
    Usually 25-50\% of the margin.
  \end{itemize}
\item
  Agent's fee

  \begin{itemize}
  \tightlist
  \item
    Bank which takes on agency role (administrative or security agent)
  \end{itemize}
\item
  Cancellation fee
\item
  Utilisation fee

  \begin{itemize}
  \tightlist
  \item
    For facilities intended to be used only rarely, e.g., standby or
    swing-line facilities.
  \end{itemize}
\end{itemize}

\hypertarget{fee-letter}{%
\subsubsection{Fee Letter}\label{fee-letter}}

Fees usually put in a separate fee letter. Provides confidentiality from
syndicate members, other borrowers and the market. The facility
agreement will specifically refer to the agent's fees - but not the
amount -- to avoid argument that the arrangement is a secret profit in
breach of fiduciary duties.

\hypertarget{remedies}{%
\subsection{Remedies}\label{remedies}}

\hypertarget{remedies-and-waivers}{%
\subsubsection{Remedies and Waivers}\label{remedies-and-waivers}}

Most facilities contain the following boilerplate:

\begin{quote}
The rights and remedies of the Bank under this agreement:

\begin{enumerate}
\tightlist
\item
  shall not be waived by any failure, or delay, in their exercise;
\item
  are cumulative, and shall not exclude or restrict its rights and
  remedies under any other agreement or the general law; and
\item
  may be exercised in whole, in part, and as often as necessary.
\end{enumerate}
\end{quote}

(1) tries to avoid the borrower using the equitable doctrine of
``laches'' (form of estoppel) to preclude the bank exercising a right/
remedy after unreasonable delay. But by case law, continued performance
of a contract inconsistent with termination and will result in a loss of
the right to terminate (Tele2 International Card Company SA v Post
Office Limited {[}2009{]} EWCA Civ 9).

(2) says the bank can use more than remedy at once. Tries to avoid any
merger of rights between facility agreement and security document (about
not losing an inferior security, which may happen under case law if you
also have a superior security).

\hypertarget{partial-invalidity}{%
\subsubsection{Partial Invalidity}\label{partial-invalidity}}

Include:

\begin{quote}
If any provision of this agreement is prohibited, unenforceable, void or
invalid, that shall not invalidate, or otherwise affect the
enforceability of, the remaining provisions hereof.
\end{quote}

Two underlying principles in relation to severance:

\begin{enumerate}
\tightlist
\item
  a court will not construct a new contract: any severance must leave
  the original contract standing and not alter its basic scope or
  intention; and
\item
  a court will not sever if to do so is against public policy.
\end{enumerate}

Goldsoll v Goldman {[}1915{]} 1 Ch 292 blue pencil test (can you just
cross out the severed provisions in blue pen).

\hypertarget{amendments-and-waivers}{%
\subsubsection{Amendments and Waivers}\label{amendments-and-waivers}}

No alterations to the provisions of an executed document will be
effective without the agreement of all parties. Most syndicated
facilities allow amendments and waivers to most of their clauses to be
agreed by a majority of the syndicate (defined in the agreement as
either {\(\frac{2}{3}\)} or "simple majority" - {\(\frac{1}{2}\)}).

Common provisions:

\begin{longtable}[]{@{}ll@{}}
\toprule()
Provision & Details \\
\midrule()
\endhead
Yank the bank & Clause allowing the borrower to take a bank out of the
syndicate if the bank votes against an amendment/ waiver which the
majority of banks voted to allow. \\
Use or lose & If the bank does not vote on an amendment/ waiver within a
time period, the bank's commitment is disregarded in calculating whether
the borrower gets the required majority. \\
\bottomrule()
\end{longtable}

\hypertarget{force-majeure}{%
\subsection{Force Majeure}\label{force-majeure}}

Suppose there are circumstances in which the contracting parties
consider performance of their obligations is, through no fault of their
own, no longer possible. The clause may provide that obligations under
the contract are suspended until circumstances have passed or that the
contract is immediately terminated. Usually no sanctions attached.

\hypertarget{governing-law-and-jurisdiction}{%
\subsection{Governing Law and
Jurisdiction}\label{governing-law-and-jurisdiction}}

\hypertarget{common-law}{%
\subsubsection{Common Law}\label{common-law}}

Under common law, parties are free to choose governing law provided that
their choice is bona fide and legal (Vita Food Products Inc v Unus
Shipping Co Ltd (in Liquidation {[}1939{]} AC 277). This determines
issues including the validity of the document and any amendments,
interpretation of clauses and remoteness of damages.

\hypertarget{rome-i}{%
\subsubsection{Rome I}\label{rome-i}}

Allows complete freedom to choose the governing law of a
\textbf{contract} (expressly or impliedly), although public policy and
some mandatory rules may supersede the governing law if necessary. The
Regulation is retained EU law.

\hypertarget{rome-ii}{%
\subsubsection{Rome II}\label{rome-ii}}

Allows parties to choose the governing law of \textbf{non-contractual}
obligations. The choice will be upheld if:

\begin{enumerate}
\tightlist
\item
  The claim falls within Rome II
\item
  The context is a commercial activity
\item
  The choice of law was freely negotiated.
\end{enumerate}

\hypertarget{choosing}{%
\subsubsection{Choosing}\label{choosing}}

May be bank's country, borrower's country, or English law (for stability
and for being creditor friendly). If parties do not choose, courts check
for an implied choice (e.g., choice of jurisdiction).

\hypertarget{jurisdiction}{%
\subsubsection{Jurisdiction}\label{jurisdiction}}

Specifies the jurisdiction of enforcement.

The three regimes that have usually applied to a contract giving
jurisdiction to English courts are:

\begin{enumerate}
\tightlist
\item
  the Recast Brussels Regulation which applies to proceedings commenced
  on or after 10 January 2015 binding all EU members.
\item
  the 2007 Lugano Convention, which binds EU members, Denmark, Norway,
  Iceland and Switzerland; and
\item
  English domestic law.
\end{enumerate}

Complicated by Brexit.

Some facility agreements contain one-sided jurisdiction clauses,
allowing the bank to sue in the named jurisdiction or any other
competent jurisdiction, but only allowing the borrower to sue in the
named jurisdiction. Though these have been challenged in court, still
included in most facility agreements.

\hypertarget{agent-for-service-of-process}{%
\subsubsection{Agent for Service of
Process}\label{agent-for-service-of-process}}

If any of the obligors in the facility is not an English registered
company and does not have a branch or other place of business in Great
Britain, the service of any English court documents on that obligor
would require service `out of jurisdiction'.

\begin{env-794f47f9-38df-4a52-b7f1-227d11939e2b}

CPR 6.11(1) - Service of the claim form by contractually agreed method

Where --

\begin{itemize}
\tightlist
\item
  (a) a contract contains a term providing that, in the event of a claim
  being started in relation to the contract, the claim form may be
  served by a method or at a place specified in the contract; and
\item
  (b) a claim solely in respect of that contract is started,
\end{itemize}

the claim form may, subject to paragraph (2), be served on the defendant
by the method or at the place specified in the contract.

\end{env-794f47f9-38df-4a52-b7f1-227d11939e2b}

The usual contractually agreed method is to appoint a 3rd party to
receive documents on the obligors' behalf. This avoids a bank having to
apply to court for leave to serve a foreign corporation. Several
companies offer services for a fee.

\hypertarget{miscellaneous}{%
\subsection{Miscellaneous}\label{miscellaneous}}

Other provisions include:

\begin{itemize}
\tightlist
\item
  Inter-bank provisions

  \begin{itemize}
  \tightlist
  \item
    Monitoring of defaults
  \item
    Liability for actions and information
  \item
    Resignation of agent banks
  \item
    Indemnities
  \item
    Sharing of money if some banks receive money or recover money and
    others do not
  \end{itemize}
\item
  Set-off

  \begin{itemize}
  \tightlist
  \item
    Authority from borrower giving the bank a right (but not an
    obligation) to set off credit balances in favour of the borrower
    against any amounts due but unpaid.
  \item
    Set-off is the cancellation of mutual obligations outside of a court
    proceeding. It allows a debtor to reduce the amount it owes to a
    creditor by the amount the creditor owes to the debtor. Set-off is a
    self-help remedy that is governed by state statutory and common law.
  \end{itemize}
\item
  Notices

  \begin{itemize}
  \tightlist
  \item
    Communications must be written and may be sent by post or by fax. By
    email if agreed by both parties.
  \item
    Banks might want deemed service provisions for borrower, but will
    never accept a reciprocal arrangement.
  \end{itemize}
\item
  Defaulting lender provisions --

  \begin{itemize}
  \tightlist
  \item
    If a lender becomes insolvent, B may be able to

    \begin{itemize}
    \tightlist
    \item
      Cancel any undrawn commitment of a defaulting lender.
    \item
      No commitment fee payable on any retained commitment
    \item
      If defaulting lender retains a commitment, it will not be allowed
      to vote on any decisions under the facility (it is
      disenfranchised).
    \item
      Defaulting lender may be forced to transfer participation in the
      facility to another lender at par.
    \item
      Any amount lent under revolving credit facility automatically
      turned into a term loan.
    \item
      Identity of a defaulting lender disclosed by the agent to the
      borrower.
    \item
      If facility agent defaults, agent may be bypassed and can be
      removed by a majority vote of the remaining banks.
    \end{itemize}
  \end{itemize}
\end{itemize}

\hypertarget{execution}{%
\subsection{Execution}\label{execution}}

Where a syndicated loan requires a party to perform several defined
roles, conventional for a bank to execute separately for each of the
roles. s 44(6) CA 2006 requires that if a director is signing a document
as director for more than one company (quite common for large groups),
they must sign separately for each company.

\begin{env-74ec9a83-00fa-4ee6-a586-5d505a5676c9}

Question

Whhat if the loan involves a large syndicate and it is not possible to
find a signing date which works for everyone?

\end{env-74ec9a83-00fa-4ee6-a586-5d505a5676c9}

\begin{enumerate}
\tightlist
\item
  Documents can be sent to each signatory in turn (cumbersome)
\item
  Absent party may appoint someone to sign on their behalf under a power
  of attorney, created through a deed. The attorney can sign in their
  own name or that of the principal (s 7(1) Powers of Attorney Act
  1971).
\item
  Document may be executed in counterparts -- identical copies
  simultaneously signed. Methods:

  \begin{enumerate}
  \tightlist
  \item
    Printed documents sent to each party by post/ courier. After
    signing, parties return their documents to the solicitor. The
    original agreement is a composite of the counterparts.
  \item
    Documents circulated by email/ made accessible via an online
    platform. Signed document signed as PDF and returned.
  \item
    Documents circulated by email, instructions to print and sign
    signature pages, and then scan into PDF and return with the
    electronic copy of the unsigned original. The true original document
    is the electronic version.
  \item
    Signature pages printed, signed and returned. This means there is no
    ``original document''. Cannot be used for executing deeds or
    agreements with land clauses.
  \item
    Just signature pages circulated. Solicitor authorised to affix
    signature pages and document treated as executed. Cannot be used for
    deeds/ land agreements.
  \end{enumerate}
\end{enumerate}

\hypertarget{us-vs-english-loan-facilities}{%
\subsection{US Vs English Loan
Facilities}\label{us-vs-english-loan-facilities}}

Factors:

\begin{enumerate}
\tightlist
\item
  Similar insolvency regimes.
\item
  Generally easier to take security over all the borrower's assets under
  US law and registration/ enforcement regimes are more standardised.
\item
  US more litigious: unconnected 3rd parties can purchase and pursue
  legal claims
\item
  US has more liquidity -- convergence between loan and bond market
  terms.
\end{enumerate}

Differences:

\begin{itemize}
\tightlist
\item
  Call protection

  \begin{itemize}
  \tightlist
  \item
    US phrase for early prepayment fee.
  \end{itemize}
\item
  Limited amortisation

  \begin{itemize}
  \tightlist
  \item
    Annual amortisation \textasciitilde1\% on the term loan
  \end{itemize}
\item
  Covenant lite

  \begin{itemize}
  \tightlist
  \item
    English leveraged loan facilities oblige the borrower to show
    regularly (e.g., 3 months) that it is maintaining financial
    performance predicted in business plan.
  \item
    US facilities just have a single financial covenant, tested if the
    borrower wants to take specified actions. Cash flow covenants rare.
    Much less onerous.
  \end{itemize}
\item
  MAC event of default.

  \begin{itemize}
  \tightlist
  \item
    Not included in US law facilities.
  \end{itemize}
\item
  Majority lender votes

  \begin{itemize}
  \tightlist
  \item
    US typically requires 1/2, English usually requires 2/3.
  \end{itemize}
\item
  Loan transfer

  \begin{itemize}
  \tightlist
  \item
    US facilities will allow the term loan to be freely transferred
    without borrower consent.
  \item
    English more restrictive; may be a white list of acceptable
    transferees/ borrower consent required.
  \end{itemize}
\end{itemize}

\hypertarget{novation-assignment-sub-participation}{%
\section{Novation, Assignment,
Sub-participation}\label{novation-assignment-sub-participation}}

A bank may use loans like fixed assets: purchasing them, keeping them
for a while to earn income, and then disposing of them.

\hypertarget{why-sell-a-loan}{%
\subsection{Why Sell a Loan}\label{why-sell-a-loan}}

\begin{enumerate}
\tightlist
\item
  Realising capital

  \begin{enumerate}
  \tightlist
  \item
    Improve liquidity to take advantage of new/ better lending
    opportunities.
  \item
    Focus on a particular market.
  \end{enumerate}
\item
  Risk management

  \begin{enumerate}
  \tightlist
  \item
    Sell loans to diversify away from one type of loan/ borrower.
  \end{enumerate}
\item
  Balance sheet

  \begin{enumerate}
  \tightlist
  \item
    Bank must maintain a \% of its capital as cover for loans
    (regulatory capital).
  \item
    Sell loans to raise capital for others
  \end{enumerate}
\item
  Profit

  \begin{enumerate}
  \tightlist
  \item
    Short-term profit by selling a loan (e.g., if interest rates fall)
  \item
    If a borrower has defaulted/ is performing badly, a lender may
    choose to sell participation at a discount to crystallise loss
  \end{enumerate}
\item
  Prestige

  \begin{enumerate}
  \tightlist
  \item
    Bank may want initial involvement in a facility which is
    high-profile/ important, but may sell this after.
  \end{enumerate}
\item
  Syndication

  \begin{enumerate}
  \tightlist
  \item
    Some loans provide for syndication post-closing.
  \end{enumerate}
\end{enumerate}

\hypertarget{moving-rights-and-obligations}{%
\subsection{Moving Rights and
Obligations}\label{moving-rights-and-obligations}}

\begin{itemize}
\tightlist
\item
  Most contracts operate by giving rights and obligations (liabilities)
  to each party.
\item
  Under most facility agreements, a bank's fundamental obligation is to
  lend money (the borrower has a right to borrow), whilst its
  fundamental rights are usually to receive interest and eventual return
  of capital (the borrower is obliged to service and repay the loan).
\item
  In order to fully dispose of a loan, a bank must dispose of both its
  rights and its obligations.
\item
  Under English law, contractual obligations can be moved only with the
  consent of the party to whom the obligation is owed (i.e., borrower's
  consent is required).
\end{itemize}

\hypertarget{selling-assets}{%
\subsection{Selling Assets}\label{selling-assets}}

Methods of disposing of an asset:

\begin{enumerate}
\tightlist
\item
  Novation
\item
  Legal (``statutory''/ ``disclosed'') assignment
\item
  Equitable (``undisclosed'') assignment
\item
  Sub-participation
\item
  Risk participation.
\end{enumerate}

Main aim of the bank: to remove the risk associated with the asset.

\hypertarget{novation}{%
\subsubsection{Novation}\label{novation}}

The only way in which a party can effectively `transfer' all its rights
and obligations under a contract is with the consent of all the parties
involved.

\begin{env-02a65e52-3e2f-4e42-a1ad-a9c88e6d3852}

Important

Novation involves one party's rights and obligations under a contract
being cancelled and discharged, whilst a third party assumes identical
new rights and obligations in their place.

\end{env-02a65e52-3e2f-4e42-a1ad-a9c88e6d3852}

The new bank assumes identical rights and obligations with respect to
the borrower as applied to the existing bank.

\begin{env-9a8b2d5a-e80b-4017-bd60-2afb8ee62329}

Warning

Strictly, a novation does not transfer rights and liabilities: it
cancels an existing contract and replaces it with another. The
borrower's promise to perform its obligations in favour of a new bank is
consideration for the existing bank releasing the old debt.

\end{env-9a8b2d5a-e80b-4017-bd60-2afb8ee62329}

\hypertarget{advantages}{%
\paragraph{Advantages}\label{advantages}}

\begin{itemize}
\tightlist
\item
  Moving obligations

  \begin{itemize}
  \tightlist
  \item
    Only proven method of moving contractual obligations as well as
    rights.
  \item
    Allows the existing bank to dispose of a loan which has an
    unutilised commitment (e.g., under an RCF).
  \item
    The new bank achieves a relationship with the borrower as if it were
    a party to the facility agreement.
  \end{itemize}
\item
  Risk transfer

  \begin{itemize}
  \tightlist
  \item
    Can fully remove a loan (including undrawn commitments) from the
    existing bank's balance sheet and exclude it from regulatory capital
    requirements.
  \end{itemize}
\item
  Easy syndication

  \begin{itemize}
  \tightlist
  \item
    If the original facility agreement included transfer certificates,
    the existing bank can take a large commitment without the delay of
    putting together an underwriting syndicate.
  \end{itemize}
\end{itemize}

\hypertarget{disadvantages}{%
\paragraph{Disadvantages}\label{disadvantages}}

\begin{itemize}
\tightlist
\item
  Consent

  \begin{itemize}
  \tightlist
  \item
    Consent of all parties involved in the original loan document is
    required, including any parties which guaranteed the borrower's
    performance.
  \item
    Puts lots of power in the borrower's hands.
  \item
    Solution: use of transfer certificates

    \begin{itemize}
    \tightlist
    \item
      Facility agreement will contain a clause under which the parties
      agree that a bank may dispose by novation any or all of its
      commitment -- open offer to take a loan (though borrower's consent
      may be required in certain circumstances).
    \item
      Pro-forma certificate usually included as a schedule.
    \end{itemize}
  \item
    Controversy

    \begin{itemize}
    \tightlist
    \item
      Australian case Goodridge v Macquarie Bank Limited {[}2010{]} FCA
      67 held that a prior agreement to novation in a loan agreement was
      invalid and simply an `agreement to agree'.
    \item
      Habibsons Bank Ltd v Standard Chartered Bank (Hong Kong) Limited
      {[}2010{]} EWCA Civ 1335: CoA obiter criticised the decision.
    \end{itemize}
  \item
    A system similar to the transfer certificate is the transferable
    loan certificate (TLC). The only significant difference is that the
    TLC operates through the agent bank issuing certificates and keeping
    a register of transfers.
  \end{itemize}
\item
  Secured loans

  \begin{itemize}
  \tightlist
  \item
    Novation (probably) restarts hardening periods during which security
    may be set aside as a transaction at an undervalue/ preferred
    transaction/ avoidance of floating charges.
  \item
    Security being re-dated may lose priority over other security.

    \begin{itemize}
    \tightlist
    \item
      Solution: appoint a security trustee to hold any security granted
      under the loan on trust for the banks.
    \item
      Banks participating in the facility from time to time are
      beneficiaries.
    \end{itemize}
  \end{itemize}
\item
  Disclosure

  \begin{itemize}
  \tightlist
  \item
    Difficult to hide the identity of a transferee bank.
  \end{itemize}
\end{itemize}

\hypertarget{legal-assignment}{%
\subsubsection{Legal Assignment}\label{legal-assignment}}

Assignment is the transfer of \textbf{rights}. \textbf{s 136 LPA 1925}
provides that an assignment of debts and other choses in action will
only be recognised at common law if it is:

\begin{enumerate}
\tightlist
\item
  In writing and signed by the assignor;
\item
  Absolute (i.e., unconditional); and

  \begin{itemize}
  \tightlist
  \item
    To be `absolute' in the context of loan transfers, an assignment
    must transfer the whole of the debt owed to the existing bank
    (Walter and Sullivan Ltd v Murphy (J) and Sons Ltd {[}1955{]} 2 QB
    584) (and cannot be `by way of charge')
  \end{itemize}
\item
  Notified in writing to any person(s) against whom the assignor could
  enforce the assigned rights.
\end{enumerate}

\begin{env-a0fe8824-15a9-47d5-bf01-f54598533eae}

Tip

\begin{itemize}
\tightlist
\item
  If any one of the conditions for legal (or "statutory") assignment is
  not fulfilled, the assignment will be an equitable one.
\item
  An assignment can transfer only rights, not obligations
\end{itemize}

\end{env-a0fe8824-15a9-47d5-bf01-f54598533eae}

\hypertarget{advantages-1}{%
\paragraph{Advantages}\label{advantages-1}}

\begin{itemize}
\tightlist
\item
  Rights transferred

  \begin{itemize}
  \tightlist
  \item
    Rights under a facility agreement transferred without requiring the
    borrower's consent.
  \end{itemize}
\item
  Borrower's payments

  \begin{itemize}
  \tightlist
  \item
    On receiving notice of an assignment, the borrower is obliged to pay
    any monies due under the assigned loan to the new bank.
  \end{itemize}
\item
  Secured loan

  \begin{itemize}
  \tightlist
  \item
    Any security, or a bank's rights as beneficiary of any security
    sharing agreement, may be assigned along with the debt, but is
    usually held by a Security Trustee or under a parallel debt
    structure.
  \end{itemize}
\end{itemize}

\hypertarget{disadvantages-1}{%
\paragraph{Disadvantages}\label{disadvantages-1}}

\begin{itemize}
\tightlist
\item
  Obligations not transferred

  \begin{itemize}
  \tightlist
  \item
    Existing bank cannot assign any undrawn commitments (makes assigning
    a revolving credit facility problematic)
  \item
    Obligations to other syndicate banks not transferred.
  \item
    Some facility documents therefore include assignment mechanics which
    purport to assign the existing bank's rights, release the existing
    bank from any obligations relating to that part of its commitment,
    and require the new bank to assume equivalent obligations.
  \end{itemize}
\item
  Rights not transferred

  \begin{itemize}
  \tightlist
  \item
    The rights under a contract can be assigned without consent only in
    `cases where it can make no difference to the person on whom the
    obligation lies to which of two persons he is to discharge it'
    (Tolhurst v Associated Portland Cement Manufacturers (1900) Ltd
    {[}1902{]} 2 KB 660).
  \item
    Objective test -- character of the assignee will b ignored.
  \item
    Danger that any indemnity provisions for increased costs may be
    unassignable without consent because they are personal.
  \end{itemize}
\end{itemize}

\hypertarget{equitable-assignment}{%
\subsubsection{Equitable Assignment}\label{equitable-assignment}}

An assignment which does not meet any of the provisions required for
legal assignment under the LPA 1925, s 136 will be an equitable
assignment.

Procedurally, an equitable assignee must join the assignor (the existing
bank) in any action on the debt (Three Rivers DC v Bank of England
{[}1995{]} 4 All ER 312).

\hypertarget{advantages-2}{%
\paragraph{Advantages}\label{advantages-2}}

\begin{itemize}
\tightlist
\item
  Partial assignment

  \begin{itemize}
  \tightlist
  \item
    If a bank wants to assign part of an outstanding loan, it must use
    an equitable assignment.
  \end{itemize}
\item
  Non-disclosure

  \begin{itemize}
  \tightlist
  \item
    If a bank wants to assign an outstanding loan without disclosing the
    identity of the new bank to the borrower, it must use an equitable
    assignment.
  \end{itemize}
\end{itemize}

\hypertarget{disadvantages-2}{%
\paragraph{Disadvantages}\label{disadvantages-2}}

\begin{itemize}
\tightlist
\item
  Payments

  \begin{itemize}
  \tightlist
  \item
    If the equitable assignment is not notified to the borrower, the
    borrower is entitled to continue making payments through the
    existing bank.

    \begin{itemize}
    \tightlist
    \item
      Old bank might be happy to skim interest
    \item
      New bank taking a double credit risk.
    \end{itemize}
  \end{itemize}
\item
  Subject to equities

  \begin{itemize}
  \tightlist
  \item
    If the equitable assignment is unnotified, a new bank will be
    subject to all `equities' (e.g., mutual rights of set-off) which
    arise between the existing bank and the borrower, until notification
    of assignment.
  \end{itemize}
\end{itemize}

\hypertarget{form-of-notice}{%
\paragraph{Form of Notice}\label{form-of-notice}}

A notice of assignment (legal or equitable) should require a borrower to
confirm:

\begin{enumerate}
\tightlist
\item
  the amount of the existing bank's debt;
\item
  that it has no rights of set-off or counter-claim against the existing
  bank;
\item
  that it has no notice of other assignments of the same debt;
\item
  the facility agreement has not been varied; and
\item
  that it will pay amounts due and payable under the underlying contract
  to the new bank's order.
\end{enumerate}

\hypertarget{sub-participation}{%
\subsubsection{Sub-participation}\label{sub-participation}}

Novation and assignment are attractive to banks because they enable it
to remove loans from its balance sheet. Possible to achieve the same
effect while leaving the original loan in place: sub-participation
(``funded participation'').

\begin{env-f09e27d9-67b8-4931-ab1f-ca6031966775}

Sub-participation

An arrangement under which an existing bank matches part or all of its
loan to a borrower with a deposit it takes from a new bank (the
`sub-participant'). The new bank agrees that its deposit will be
serviced and repaid only when the borrower services and repays the loan
from the existing bank. The sub-participant has effectively taken on the
risk of the first loan.

\end{env-f09e27d9-67b8-4931-ab1f-ca6031966775}

\begin{itemize}
\tightlist
\item
  The transaction does not involve any transfer of rights or
  obligations: the existing bank's contractual relationship with the
  borrower is unaffected.
\item
  Since sub-participation is an entirely separate agreement, the
  sub-participant bank has no-recourse against the borrower
  (``non-recourse transaction'').
\item
  Security provided remains with the existing bank.
\end{itemize}

\hypertarget{advantages-3}{%
\paragraph{Advantages}\label{advantages-3}}

\begin{itemize}
\tightlist
\item
  Risk transfer

  \begin{itemize}
  \tightlist
  \item
    Sub-participation removes a loan from inclusion in the existing
    bank's regulatory capital requirements, other than in respect of
    undrawn commitments.
  \end{itemize}
\item
  Consent

  \begin{itemize}
  \tightlist
  \item
    Unless there is a prohibition in the original facility agreement,
    existing bank can sub-participate without the consent of the
    borrower.
  \item
    All banks owe a duty of confidentiality to clients (Tournier v
    National Provincial and Union Bank of England {[}1924{]} 1 KB 461).

    \begin{itemize}
    \tightlist
    \item
      Existing bank intending to sub-participate will want to include a
      clause in the original facility agreement will want a clause
      allowing it to release information about the borrower and the
      loan.
    \end{itemize}
  \end{itemize}
\item
  Non-disclosure

  \begin{itemize}
  \tightlist
  \item
    Existing bank does not have to disclose sub-participation to the
    borrower, and most sub-participations will be silent.
  \end{itemize}
\end{itemize}

\hypertarget{disadvantages-3}{%
\paragraph{Disadvantages}\label{disadvantages-3}}

\begin{itemize}
\tightlist
\item
  Sub-participant's credit risk

  \begin{itemize}
  \tightlist
  \item
    Double credit risk for the sub-participation.
  \end{itemize}
\item
  Risk transfer

  \begin{itemize}
  \tightlist
  \item
    May not off-load the entire risk of undrawn commitments if it just
    results in the existing bank taking a risk on the new bank instead
    of the borrower.
  \item
    The sub-participation agreement must ensure that the existing bank
    is put in funds in time to meet the borrower's demands for
    utilisation.
  \end{itemize}
\end{itemize}

\hypertarget{risk-participation}{%
\subsubsection{Risk Participation}\label{risk-participation}}

\begin{itemize}
\tightlist
\item
  Form of participation which acts like a guarantee.
\item
  Risk participant agrees for a fee to put the existing bank in funds in
  certain circumstances (e.g., payment default by borrower).
\item
  Common use: to cover "unfunded" assets (guarantees, letters of credit,
  swaps).

  \begin{itemize}
  \tightlist
  \item
    If existing bank's customer fails to meet reimbursement obligations,
    the new bank will put the existing bank in funds.
  \end{itemize}
\end{itemize}

\hypertarget{global-transfer-agreement}{%
\subsection{Global Transfer Agreement}\label{global-transfer-agreement}}

\begin{env-f09e27d9-67b8-4931-ab1f-ca6031966775}

Definition

A new contract which amends (and restates) the facility agreement to add
a new schedule of commitments.

\end{env-f09e27d9-67b8-4931-ab1f-ca6031966775}

\begin{itemize}
\tightlist
\item
  A deal may require a large amount of simultaneous transfers.
\item
  To avoid preparing a large number of transfer certificates, a Global
  Transfer Agreement may be used.
\item
  But transparency may not suit all deals.
\end{itemize}

\hypertarget{loan-market-association}{%
\subsection{Loan Market Association}\label{loan-market-association}}

\begin{itemize}
\tightlist
\item
  Founded 1996.
\item
  Key objective is `improving liquidity, efficiency and transparency in
  the primary and secondary syndicated loan markets in EMEA'.
\item
  Creates widely used loan documentation.
\end{itemize}

\hypertarget{tax-implications}{%
\subsection{Tax Implications}\label{tax-implications}}

2 main types of tax which might affect asset sales: stamp duty (or
sometimes, stamp duty reserve tax) and withholding tax.

\begin{itemize}
\tightlist
\item
  Sub-participation does not involve any transfer of rights, and so will
  not attract stamp duty.
\item
  Novation will not usually involve a transfer of property for stamp
  duty purposes and so will not normally attract the tax (unless the
  asset includes, e.g., rights to acquire property).
\item
  Assignment may attract stamp duty or SDRT, but exemptions can often by
  used.
\end{itemize}

\begin{env-f09e27d9-67b8-4931-ab1f-ca6031966775}

Definition

Income tax in respect of recurring payments at source is sometimes
collected by requiring the payer to make a deduction on account of
income tax before making the payment. This deduction is generally known
as withholding tax. It frequently applies to dividends (although not
those paid by UK companies), interest and royalties.

\end{env-f09e27d9-67b8-4931-ab1f-ca6031966775}

In all asset sales between banks, and involving a UK resident/domiciled
borrower, the relevance of withholding tax on interest payments will
depend largely on whether the exemption under s 879 of the ITA 2007
applies.

The s 879 exemption will apply to loans which have been assigned as long
as the advance was initially made by a `bank' and the assignee is
beneficially entitled to the interest and within the charge to UK
corporation tax at the time interest is paid. If a loan is novated, this
probably creates a new advance and so must be made to a `bank' to ensure
s 879 still applies.

\begin{env-794f47f9-38df-4a52-b7f1-227d11939e2b}

s 879 ITA 2007 - interest paid on advances from banks

(1) The duty to deduct a sum representing income tax under section 874
does not apply to a payment of interest on an advance from a bank if, at
the time when the payment is made, the person beneficially entitled to
the interest is within the charge to corporation tax as respects the
interest {[}F1or is a bank that would be within the charge to
corporation tax as respects the interest apart from section 18A of CTA
2009{]}.

(2) Section 991 (meaning of ``bank'') applies for the purposes of this
section.

\end{env-794f47f9-38df-4a52-b7f1-227d11939e2b}

Summary -- withholding tax not payable if:

\begin{itemize}
\tightlist
\item
  The bank is a UK corporation taxpayer
\item
  The bank is international, but there is a double-tax treaty between
  the bank's country and the UK
\end{itemize}

\hypertarget{fsma-implications}{%
\subsection{FSMA Implications}\label{fsma-implications}}

\begin{env-794f47f9-38df-4a52-b7f1-227d11939e2b}

s 19 FSMA 2000

No person may carry on a regulated activity in the United Kingdom or
purport to do so unless he is an authorised person or an exempt person.

\end{env-794f47f9-38df-4a52-b7f1-227d11939e2b}

When selling participations, the bank should ensure that it is
authorised to do so under the Act.

\hypertarget{securitisation}{%
\subsection{Securitisation}\label{securitisation}}

Securitisation involves the owner of a pool of assets (the `originator')
transferring them to a specially formed company (a `special purpose
vehicle' or `SPV'). The SPV funds its purchase of the assets by issuing
debt securities (it is therefore known as the `issuer'). Coupon and
redemption of the debt securities are met with, and secured over, the
asset pool.

\hypertarget{banks-perspective-3}{%
\subsection{Bank's Perspective}\label{banks-perspective-3}}

\begin{itemize}
\tightlist
\item
  Most banks will require the flexibility to sell a loan if they decide
  it is necessary or desirable, but will accept some restrictions.
\item
  A bank's solicitor should ensure the facility transfer language allows
  transfer not just to other banks but also to trusts, funds, or other
  entities engaged in buying loans.
\end{itemize}

\hypertarget{borrowers-perspective-3}{%
\subsection{Borrower's Perspective}\label{borrowers-perspective-3}}

However, a borrower's relationship with its lenders will be crucial if
it requires waivers to actual or potential defaults, if it wants to make
changes to the loan documents going forward, or needs tolerance in
interpreting provisions.

Most facilities allow transfer subject to the borrower's consent, such
consent not to be unreasonably withheld or delayed.

Commonly negotiated conditions:

\begin{itemize}
\tightlist
\item
  Borrower consent deemed given if not given within a certain timeframe
\item
  Transfers to borrower's competitors prohibited
\item
  Transfers to certain fund types prohibited (those more activist ones)
\item
  List of named institutions to which banks can transfer without consent
  pre-agreed.
\item
  Freedom to transfer upon the borrower's event of default?
\end{itemize}

\begin{env-74ec9a83-00fa-4ee6-a586-5d505a5676c9}

Can a borrower buy back portions of their own loan debt?

\begin{itemize}
\tightlist
\item
  This is a very cheap way to repay a debt.
\item
  May be issues with prepayment restrictions, pro-rata sharing between
  lenders, restrictions on cash usage, and issues with novation.
\item
  Even if buy-back is allowed, the buying entity is usually required to
  allow all the syndicate banks a chance to sell.
\end{itemize}

\end{env-74ec9a83-00fa-4ee6-a586-5d505a5676c9}

\hypertarget{drafting-tips}{%
\section{Drafting Tips}\label{drafting-tips}}

\hypertarget{creating-document}{%
\subsection{Creating Document}\label{creating-document}}

\begin{itemize}
\tightlist
\item
  Written facility agreement primarily used to benefit the bank.
\item
  Drafting the document traditionally remains in the control of the
  bank's solicitor.
\item
  LMA precedents for investment grade and leveraged deals widely used as
  a starting point. Can eliminate some basic negotiation.
\end{itemize}

\hypertarget{aims}{%
\subsection{Aims}\label{aims}}

\begin{itemize}
\tightlist
\item
  Certainty

  \begin{itemize}
  \tightlist
  \item
    Document must reflect what has been agreed.
  \item
    Can help avoid litigation.
  \end{itemize}
\item
  Ambiguity

  \begin{itemize}
  \tightlist
  \item
    Often desired by a party, e.g., ``reasonableness''.
  \item
    Uncertainty usually resolved in favour of the borrower.
  \end{itemize}
\item
  Workable document

  \begin{itemize}
  \tightlist
  \item
    Document must be reasonably user-friendly.
  \item
    Use headings and contents table.
  \item
    Drafting must work in practice, e.g., references to business days.
  \end{itemize}
\end{itemize}

\begin{env-9a8b2d5a-e80b-4017-bd60-2afb8ee62329}

Warning

If using precedents, ensure conformity of terminology and check if any
standard clauses have been removed/ amended.

\end{env-9a8b2d5a-e80b-4017-bd60-2afb8ee62329}

\hypertarget{techniques}{%
\subsection{Techniques}\label{techniques}}

\hypertarget{clause-construction}{%
\subsubsection{Clause Construction}\label{clause-construction}}

Coode's principle of clause construction:

\begin{enumerate}
\tightlist
\item
  Refer to any exceptions first: e.g. `Subject to \ldots'.
\item
  Next, set out the circumstances or conditions upon which the legal
  right or obligation depends, using the present tense: e.g. `If X does
  {[}\ldots{]} and Y does {[}\ldots{]}'.
\item
  Next, set out the right or obligation using the active and avoiding
  the passive: e.g. `X must do 1, 2, 3, 4 and 5'.
\item
  Finally, put the provision into paragraphs and give it a heading.
\end{enumerate}

\hypertarget{qualifying-an-obligation}{%
\subsubsection{Qualifying an
Obligation}\label{qualifying-an-obligation}}

Many obligations will be drafted as absolute. A borrower may be able to
agree a qualifier being added to an obligation. The usual qualifiers are
`reasonable', `all reasonable' or `best' endeavours.

\begin{longtable}[]{@{}ll@{}}
\toprule()
Term & Definition \\
\midrule()
\endhead
Reasonable endeavours & Requires a party to adopt and pursue one
reasonable course of action to achieve the result, bearing in mind its
own commercial interests and the likelihood of success. \\
All reasonable endeavours & Requires a party to explore all avenues
reasonably open to it, and explore them to the extent reasonable. \\
Best endeavours & Most onerous, requires a party to take steps which a
prudent, determined and reasonable obligee, acting in its own interests
and desiring to achieve that result, would take. \\
\bottomrule()
\end{longtable}

It is worth noting that `all reasonable' and `best' endeavours are very
similar, and some judges have held there is no practical difference
(Rhodia International Holdings v Huntsman International LLC {[}2007{]}
All ER (D) 264, Jet2.com Ltd v Blackpool Airport Ltd {[}2011{]} EWHC
1529 (Comm)). The onus is on the obligor to show it has met the required
level of `endeavour'.

Parameters can be specified, but then this leads to yet more negotiation
(so usual not to bother).

\hypertarget{materiality}{%
\subsubsection{Materiality}\label{materiality}}

The borrower may argue that a term constrains it too tightly given
commercial realities. So ``material'' may be drafted in. This holds some
risk for a bank; a bank will be wary of calling a default only later to
be proven to have acted negligently.

Options:

\begin{itemize}
\tightlist
\item
  Quantify ``material''

  \begin{itemize}
  \tightlist
  \item
    Complicated/ impossible
  \end{itemize}
\item
  Entitle the bank to decide what is material

  \begin{itemize}
  \tightlist
  \item
    Often by a majority of syndicated banks.
  \end{itemize}
\end{itemize}

\hypertarget{would-or-might}{%
\subsubsection{Would or Might}\label{would-or-might}}

Some clauses (e.g., certain events of default) will be drafted so that
they are triggered by events which `might' have particular consequences
(e.g., litigation which might affect B's ability to perform
obligations). Borrower may try to argue that ``will'' or ``would''
should be used instead.

The parties will usually agree to some form of compromise, such as `is
reasonably likely to \ldots', or `will in the bank's opinion'.

\hypertarget{de-minimis}{%
\subsubsection{De Minimis}\label{de-minimis}}

\emph{De minimis non curat lex} is the general legal principle that some
breaches of duty or mistakes are too trivial to attract a legal remedy.

The level at which a de minimis is set is crucial; it must allow the
borrower sufficient `leeway' to run its day-to-day business, whilst not
being so high as to prevent the provision which it modifies from
operating effectively. Some de minimis provisions will set an aggregate
outstanding maximum to avoid lots of minor breaches being permissible,
which together could signal a serious issue.

\hypertarget{reasonableness}{%
\subsubsection{Reasonableness}\label{reasonableness}}

Powerful tool for the borrower to reduce the severity of a provision.
Difficult for the bank to objectively decide what is reasonable.

\hypertarget{key-words}{%
\subsection{Key Words}\label{key-words}}

\hypertarget{subsidiaries}{%
\subsubsection{Subsidiaries}\label{subsidiaries}}

"Subsidiary" will usually be defined in the facility agreement as
referring to one of the two definitions in CA 2006:

\hypertarget{s-1159-as-supplemented-by-sch-6}{%
\paragraph{S 1159 (as Supplemented by Sch
6)}\label{s-1159-as-supplemented-by-sch-6}}

For a company to be its subsidiary, the `parent' company must:

\begin{enumerate}
\tightlist
\item
  hold a majority of the voting rights which can be exercised at a
  general meeting of the `subsidiary'; or
\item
  be a member of the `subsidiary' and have the right to appoint or
  remove its directors who together control more than half the voting
  rights at board meetings; or
\item
  be a member of the `subsidiary' and control, pursuant to an agreement
  with other shareholders, a majority of the voting rights which can be
  exercised at a general meeting of the `subsidiary'; or
\item
  be a holding company of another subsidiary which is itself the holding
  company of the `subsidiary'.\\
  In determining the voting rights, etc, of the `holding company', the
  voting rights of its subsidiaries and certain nominees and trustees
  are taken into account.
\end{enumerate}

\hypertarget{s-1162-and-sch-7}{%
\paragraph{S 1162 and Sch 7}\label{s-1162-and-sch-7}}

Definition of subsidiaries in relation to financial statements. The
important difference from s 1159 is the concept of `subsidiary
undertaking', whereby the ability to exercise a `dominant influence'
over a company (or partnership, or other unincorporated association)
makes it a `subsidiary undertaking' for accounting purposes. This might
include joint-venture arrangements which a company would not usually
consider to be a subsidiary.

A bank will favour the wider s 1162 definition. The borrower will ask
that provisions only apply to its material subsidiaries (those worth
over a certain amount). The bank will want ``material subsidiaries from
time to time'' to catch any that grow.

In Enviroco Ltd v Farstad Supply A/S {[}2009{]} EWCA Civ 1399, the Court
of Appeal decided that a company ceased to be a subsidiary of its
holding company when the subsidiary's shares were secured to a bank by
way of legal mortgage.

\hypertarget{notwithstanding}{%
\subsubsection{Notwithstanding}\label{notwithstanding}}

Means ``in spite of''. Antonym: ``subject to''.

\hypertarget{definitions}{%
\subsubsection{Definitions}\label{definitions}}

\begin{itemize}
\tightlist
\item
  ``Means'' gives a precise definition.
\item
  ``Includes'' makes the definition broader and less certain.
\end{itemize}

\hypertarget{s-61-lpa-1925}{%
\subsubsection{S 61 LPA 1925}\label{s-61-lpa-1925}}

Section 61 of the LPA 1925 applies, amongst other things, to all deeds
and contracts made after 1925 and, unless the context otherwise
requires, will deem:

\begin{enumerate}
\tightlist
\item
  `month' to mean a calendar month;
\item
  `person' to include a corporation;
\item
  the singular to include the plural (and vice versa);
\item
  the masculine to include the feminine (and vice versa).
\end{enumerate}

Under the Interpretation Act 1978, any reference to statutory provisions
in a document will be deemed (subject to contrary intention) to be a
reference to those provisions as amended or re-enacted from time to
time.

\hypertarget{eiusdem-generis-rule}{%
\subsubsection{Eiusdem Generis Rule}\label{eiusdem-generis-rule}}

Where general words follow a list of specific words, they are taken to
be limited to categories of the same kind as the specific words. To
avoid this rule, use words like ``whatsoever'' to emphasise generality.

\hypertarget{time-computations}{%
\subsection{Time Computations}\label{time-computations}}

Unless there is a contrary intention, courts will hold that:

\begin{enumerate}
\tightlist
\item
  The day on which an event occurs is not included when calculating a
  period which runs from the occurrence of that event (Lester v Garland
  (1808) 15 Ves 248).
\item
  The period of a `month' (or `months') will end on the day which has
  the same date in the subsequent month (Dodds v Walker {[}1981{]} 2 All
  ER 609). For example, two months from 20 March will be 20 May. If this
  rule would not yield a date (e.g., one month from 31 August), the
  period is taken to end on the last day of the month in which it
  expires (e.g., 30 September). The period will end at midnight on the
  appropriate day.
\item
  If a period is stated as `not less than {\(x\)} months nor more than
  {\(y\)} months', the rule outlined in (b) above will apply. For
  example, if a notice is to be given `not more than 3 months nor less
  than 2 months before cancellation', then cancellation on 30 November
  may be notified from 31 August to 30 September (inclusive);
\item
  notice of `at least' or `no fewer than' {\(x\)} days will usually
  exclude the day on which notice is given and the day on which it
  expires. The same effect is obtained by the expression `{\(x\)} clear
  days';
\item
  `day' is a calendar day (not just business days).
\end{enumerate}

Best to be very specific about time periods.

\end{document}
